### 2017-04-24 Manuscript/EngineBalance.tex
Latest update. Nothing was checked, but I supposedly finished the data processing for the main engines

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{verbatim}

%opening
\title{Demonstration of the solution of the system of mass and energy balances for MS Birka Main Engines}
\author{Francesco Baldi}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Initial system}

\subsection{Mass balances}

Mass balance on the split after the compressor, with one flow going to the engine inlet valves and the other flow going to the bypass valve
\begin{equation}
\dot{m}_{air,comp} = \dot{m}_{air,cyl} + \dot{m}_{air,bypass} \\
\end{equation}

Mass balance on the mixer before the turbine, with one flow coming from the engine exhaust valves and the other flow coming from the bypass valve
\begin{equation}
\dot{m}_{eg,turb} = \dot{m}_{eg,cyl} + \dot{m}_{air,bypass} \\
\end{equation}

Note that the exhaust gas mass flow from the cylinders is fully defined as follows:
\begin{equation}
\dot{m}_{eg,cyl} = \dot{m}_{air,cyl} + \dot{m}_{fuel,cyl} \\
\end{equation}

\subsection{Energy balances}

Energy balance on the turbocharger: the power generated by the turbine must be equal to the power absorbed by the compressor
\begin{equation}
\dot{m}_{air,comp} \Delta h_{comp} = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
\end{equation}

Energy balance on the mixer between the air flow after the bypass valve and the exhaust gas flow after the exhaust valve
\begin{equation}
\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0)
\end{equation}

The system of four equations is to be solved in the four following unknowns:
\begin{itemize}
	\item $ \dot{m}_{eg,turb} $
	\item $ \dot{m}_{air,bypass} $
	\item $ \dot{m}_{air,comp} $
	\item $ T_{turb,in} $
\end{itemize}

\section{Calculating the explicit system}

First, we make all equations explicit in $ \dot{m}_{air,bypass} $ thus eliminating both 
$ \dot{m}_{eg,turb} $ and $ \dot{m}_{air,comp} $

$$
(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech} 
$$
$$
\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_0)
$$

At this point, we use the first equation to simplify the system and make it explicit in the $ \dot{m}_{air,bypass} $ variable alone:

\begin{eqnarray*}
(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}  & = & (\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} \\
T_{turb,in} - T_{turb,out} & = & \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} \\
T_{turb,in} & = & T_{turb,out} + \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}}  \\
\end{eqnarray*}

We can now substitute this expression in the second equation:

\begin{multline}
	\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\ 
	(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,out} + \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} - T_0)
\end{multline}

We now have to simplify this equation in order to make the term $ \dot{m}_{air,bypass} $ explicit. 
\begin{comment}
One first thing we can do is noticing that the fraction:
$$
\frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}}
$$
Can be rewritten as:
$$
\frac{1}{1 + \frac{\dot{m}_{fuel,cyl}}{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}}
$$
Since the numerator of the "subfraction" is much smaller than the denominator, we can make the assumption that this fraction behaves as:
$$
\frac{1}{1 + \epsilon}
$$
For which we know that:
$$
\frac{1}{1 + \epsilon} \approx 1 - \epsilon 
$$
That corresponds to:
$$
1 - \frac{\dot{m}_{fuel,cyl}}{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}
$$
\begin{multline}
\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\ 
(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} \left[T_{turb,out} + \left( 1 - \frac{\dot{m}_{fuel,cyl}}{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}} \right) \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} - T_0\right]
\end{multline}
\end{comment}

We first execute the multiplication in the term on the right of the equal:
\begin{multline}
\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\
\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_0) + \\
\dot{m}_{air,bypass} c_{p,eg} (T_{turb,out} - T_0) + \\ 
\frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}}(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) c_{p,eg}
\end{multline}

Now, we can group together the elements in the variable of interest:
\begin{multline}
\dot{m}_{air,bypass} \left[ c_{p,air} (T_{comp,out} - T_0) - c_{p,eg} (T_{turb,out} - T_0) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_0) - \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
\end{multline}

This can be further simplified into:
\begin{multline}
\dot{m}_{air,bypass} \left[ c_{p,air} (T_{comp,out} - T_0) - c_{p,eg} (T_{turb,out} - T_0) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
\end{multline}

Finally, we can assume the specific heat at constant pressure of air and exhaust gas to be sufficiently similar for considering them equal in the term at the numerator (we ignored changes in the properties of the mixture in other points as well):
\begin{multline}
\dot{m}_{air,bypass} \left[ c_{p,eg} (T_{comp,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
\end{multline}

From which we can derive the final form:
\begin{equation}
\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}}{c_{p,eg} (T_{comp,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}} 
\end{equation}

Which can be rewritten, in order to have both numerator and denominator positive, as follows:
\begin{equation}
\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{cyl,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}}{c_{p,eg} (T_{turb,out} - T_{comp,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}} 
\end{equation}

\end{document}

### 2017-04-24 Python files/Constants.py

@@ -10,7 +10,7 @@ def general():
     output["CP_AIR"] = 1.02   # Air specific heat, in [kJ/kgK] 
     output["CP_EG"] = 1.08   # EG specific heat, in [kJ/kgK] 
     output["CP_LO"] = 2.1   # Lubricating oil specific heat, in [kW/kgK]
-    output["CP_W"] = 4.187   # Water specifi heat, in [kW/kgK] 
+    output["CP_WATER"] = 4.187   # Water specifi heat, in [kW/kgK]
     output["RHO_W"] = 1000.0   # Water density, in [kg/m^3]
     output["RHO_LO"] = 850.0   # Lubricating oil density, in [kg/m^3]
     output["LHV_HFO"] = 41170.0   # Heavy Fuel Oil lower heating value, in [kJ/kg]
@@ -20,7 +20,10 @@ def general():
     output["CP_HFO"] = 1.8   # Fuel specific heat, [kJ/kg]
     output["RHO_HFO"] = numpy.mean([890, 919, 924, 926, 925, 921, 924, 918, 920, 919, 933])  # HFO density, in [kg/m^3]
     output["AIR_STOIC"] = 14.7  # Stoichiometric ratio fuel/air for Diesel-type fuels
-    output["ETA_VOL"] = 0.97 # Assumption about volumetric efficiency 
+    output["ETA_VOL"] = 0.97 # Assumption about volumetric efficiency
+    output["P_ATM"] = 1.10325 # Assumption on atmospheric pressure
+    output["ISO"] = {"LHV": 42700, "T_CA": 298, "T_LT": 298, "ETA_MECH": 0.8} # Reference values for ISO conditions
+    output["NAMES"] = {"MainEngines": ["ME1", "ME2", "ME3", "ME4"], "AuxiliaryEngines": ["AE1", "AE2", "AE3", "AE4"]}
     return output
 
 def steamProperties():

### 2017-04-24 Python files/Energyanalysis.py

@@ -1,207 +1,1 @@
-## BirkaEA_main_engines_efficiency ##
-
-# Initialising variables for speed
-BirkaEA_variable_initialisation_energy 
-
-# Correction factor for main / auxiliary engines power and load based on
-# the comparison with the fuel meter
-# BirkaEA_fuel_consumption_correction_AE3onDiesel
-## NOTE: THE CORRECTION IS FIXED TO 1.15 TO AVOID STRANGE THINGS
-## THE ANALYSIS IS BASED ON THE ABOVE SCRIPT
-# See the constant ME_ETA_CORR, AE_ETA_CORR
-
-# NOTE: THIS SHOULD BE MOVED TO ANOTHER PYTHON FILE
-def engineFuelFlowCalculation(dataframe,CONSTANTS):
-    # This function is used to calculate the fuel flow of an engine given its fuel rack position and its rotating speed.
-    # Inputs are free, as long as the same unit is used for the fuel rack position and the engine speed in both the input and the "design" values in the "CONSTANTS" dictionary
-    output = CONSTANTS["MainEngines"]["MFR_FUEL_DES"] * ( (CONSTANTS["MainEngines"]["POLY_FRP_2_MF"][0] + CONSTANTS["MainEngines"]["POLY_FRP_2_MF"][1] * fuel_rack_position) / (CONSTANTS["MainEngines"]["POLY_FRP_2_MF"][0] + CONSTANTS["MainEngines"]["POLY_FRP_2_MF"][1] * CONSTANTS["MainEngines"]["FRP_DES"])) * (engine_speed / CONSTANTS["MainEngines"]["RPM_DES"])
-    return output
-
-def powerAndFuel(dataframe,CONSTANTS):
-    engine_fuel_flow = dataframe["ME"]
-    dataframe
-
-for i = 1 : 4
-    ## Main engines
-    ## Power - fuel flow - efficiency
-    ##ME_load_fuel.(char(ME_names(i))) = ME_pos_fuel_rack.(char(ME_names(i))) .* ME_rpm.(char(ME_names(i))) ./ polyval(polyfit([0.295 0.515 0.735 0.835 1],[315 397 454 474 500],2),ME_pos_fuel_rack.(char(ME_names(i)))) .* ME_on.(char(ME_names(i))) ;
[…]

### 2017-04-24 Python files/Main.py

@@ -1,5 +1,4 @@
 
-
 ###########################################################################
 #
 ###########		ENERGY AND EXERGY ANALYSIS OF A CRUISE SHIP		###########
@@ -18,7 +17,7 @@
 
 
 
-# The Main.py script calls other scripts and functions. It is divided in the following sections:
+# The main.py script calls other scripts and functions. It is divided in the following sections:
 # - INPUT
 # - DATA READING
 # - DATA CLEANING
@@ -37,7 +36,7 @@
 
 # Loading appropriate modules
 import pandas as pd
-import input 
+import input
 
 filenames = input.filenames() # Note: this is just a test
 
@@ -53,21 +52,20 @@ filenames = input.filenames() # Note: this is just a test
 # Responsible: FA
 
 data_path = 'C:\\Users\\FrancescoBaldi\\Dropbox\\Energy and exergy analysis of a cruise ship - Journal Extension\\'
-df_load = pd.read_hdf(data_path + 'birka_all_data.h5','table')
+df_load = pd.read_hdf(data_path + 'birka_all_data.h5' ,'table')
 
-import Constants
+import constants
 CONSTANTS = {}
-CONSTANTS["General"] = Constants.general() # loading dictionary with general, physical constants
-CONSTANTS["Steam"] = Constants.steamProperties() # loading dictionary with steam properties constants
-CONSTANTS["MainEngines"] = Constants.mainEngines(CONSTANTS) # loading dictionary with main-engine related constants
-CONSTANTS["AuxEngines"] = Constants.auxiliaryEngines(CONSTANTS) # loading dictionary with auxiliary-engine related constants
-CONSTANTS["OtherUnits"] = Constants.otherUnits()
-N_POINTS = 319*4*24
-temp = Constants.monthLimits(N_POINTS)
+CONSTANTS["General"] = constants.general() # loading dictionary with general, physical constants
+CONSTANTS["Steam"] = constants.steamProperties() # loading dictionary with steam properties constants
+CONSTANTS["MainEngines"] = constants.mainEngines(CONSTANTS) # loading dictionary with main-engine related constants
+CONSTANTS["AuxEngines"] = constants.auxiliaryEngines \
+    (CONSTANTS) # loading dictionary with auxiliary-engine related constants
+CONSTANTS["OtherUnits"] = constants.otherUnits()
+N_POINTS = 319* 4 * 24
[…]

### 2017-04-24 Python files/preprocessing.py

@@ -1,0 +1,462 @@
+# This module includes the functions that perform the "pre-processing": In short, they prepare the data structure
+# that is then processed by the "Energy analysis" and "Exergy analysis" functions, all in one go.
+#
+# The module is made of :
+# - One function that simply reads in the data that is already available from the original dataset
+
+
+import numpy as np
+import pandas as pd
+
+
+
+def generalProcessing(raw, CONSTANTS):
+    CONSTANTS["General"]["T_0"] = raw["T_SW"]
+    return CONSTANTS
+
+
+
+
+def mainEngineProcessing(raw, processed, CONSTANTS, status):
+    # This script summarizes all the functions that calculate the required data for the Main Engines different flows
+
+    # Reading existing values
+    processed = readMainEnginesExistingValues(raw,processed,CONSTANTS)
+    # Adding various assumptions on speific values
+    processed = mainEnginesAssumptions(processed, CONSTANTS)
+    # Calculating the main engines fuel flows
+    processed = mainEngineFuelFlowCalculation(raw, processed, CONSTANTS)
+    # Calculating the main engines power output
+    processed = mainEnginePowerCalculation(processed, CONSTANTS)
+    # Calculating engine load, that is used many times later on
+    status = mainEngineStatusCalculation(processed,CONSTANTS,status)
+    # Calculate air and exhaust gas flows in the main engines
+    processed = mainEngineAirFlowCalculation(raw, processed, CONSTANTS)
+
+    return processed
+
+
+
+def readMainEnginesExistingValues(raw,processed,CONSTANTS):
+    # This function only reads existing series. It does not do any pre-processing action.
+    for name in CONSTANTS["NAMES"]["MainEngines"]:
+        # Reading main engines exhaust gas temperature, TC inlet and outlet
+        processed[name]["Cyl"]["EG_out"]["T"] = raw[name + "_TC_EG_T_in"]  # Measured before mixer with flow form bypass
+        processed[name]["TC"]["EG_out"]["T"] = raw[name + "_TC_EG_T_out"]  # Measured after mixer with waste gate
+        # Reading main engines exhaust gas temperature, after HRSG
+        if idx==1 or idx==2:
[…]

### 2017-04-24 Python files/unitstructures.py

@@ -39,26 +39,42 @@ def unitStructure():
     for idx in structure.keys():
         if idx[1] == "E":  # This basically means that this operation is only done if the system is an engine
             structure[idx] = {"TC": {}, "CAC_HT": {}, "CAC_LT": {}, "LOC": {}, "JWC": {}, "Cyl": {}}
-            structure[idx]["TC"] = {"Air": {"type": "PF"}, "EG": {"type": "PF"}}  # Turbocharger
-            structure[idx]["BPvalve"] = {"Air": {"type": "PF"}}  # Bypass valve
-            structure[idx]["WasteGate"] = {"EG": {"type": "PF"}}  # Waste gate
-            structure[idx]["CAC_HT"] = {"Air": {"type": "PF"}, "HTwater": {"type": "PF"}}  # Charge air cooler, HT stage
-            structure[idx]["CAC_LT"] = {"Air": {"type": "PF"}, "LTwater": {"type": "PF"}}  # Charge air cooler, LT stage
-            structure[idx]["LOC"] = {"LubOil": {"type": "PF"}, "LTwater": {"type": "PF"}}  # Lubricating oil cooler
-            structure[idx]["JWC_HT"] = {"QdotJW": {"type": "Qdot"}, "HTwater": {"type": "PF"}}  # Jacket water cooler
-            # Cylinders
-            structure[idx]["Cyl"] = {"Air": {"type": "PF"}, "HTwater": {"type": "PF"}, "Power": {"type": "Wdot"}}
-
+            structure[idx]["TC"] = {"Air_in": {"type": "CPF"}, "EG_in": {"type": "CPF"},
+                                    "Air_out": {"type": "CPF"}, "EG_out": {"type": "CPF"}}  # Turbocharger
+            structure[idx]["Comp"] = {"Air_in": {"type": "CPF"}, "Air_out": {"type": "CPF"}} # TC compressor
+            structure[idx]["BPvalve"] = {"Air_in": {"type": "CPF"}, "Air_out": {"type": "CPF"}}  # Bypass valve
+            structure[idx]["Turbine"] = {"EG_in": {"type": "CPF"}, "EG_out": {"type": "CPF"}}  # Turbocharger turbine
+            structure[idx]["WasteGate"] = {"EG_in": {"type": "CPF"},"EG_out": {"type": "CPF"}}  # Waste gate
[…]

### 2017-05-16 Manuscript/Manuscript.tex
Initiation of the manuscript. Included introduction and description of the heat demand estimation

%% 
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01

\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
\def\bibsection{\section*{References}}
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Energy}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Energy and exergy analysis of a cruise ship}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author[EPFL]{Francesco Baldi}
\author[Linneus]{Fredrik Ahlgren}
\author[DTU]{Tuong-Van Nguyen}
\author[CTU]{Cecilia Gabrielii}
\author[CTU]{Karin Andersson}

\address[EPFL]{Industrial Process Energy Systems Engineering (IPESE), \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, 1950, Sion, Switzerland}
\address[Linneus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
\address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
\address[CTU]{Department of Mechanics and Maritime Sciences, Chalmers University of technology, Gothenburg, Sweden}

\begin{abstract}
%% Text of abstract

\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

low carbon shipping \sep energy analysis \sep exergy analysis \sep energy efficiency

\end{keyword}

\end{frontmatter}

%% \linenumbers

%% main text

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%%  \bibliographystyle{elsarticle-num} 
%%  \bibliography{<your bibdatabase>}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

\section{Introduction} \label{sec:introduction}

\subsection{Background}

According to the third IMO GHG Study, in 2012 CO$_2$ emissions from shipping amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions [1]. Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios [1]. As an example, global transport demand has increased by 3.8\% in 2013, compared to a global GBP growth of 2.3\% the same year, which shows how shipping tends to rise even faster than global economy [2].

International Energy Agency data show that the OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3]. The fact that shipping needs to even further reduce its CO$_2$ emissions in the near future is essential for being able to achieve the goals of maintaining the climate below a 2-degree level in 2050 [4]. Finally, in the Baltic Sea an emission control area is enforced by the International Maritime Organisation since January 2015 which stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels.

Altogether, these conditions present a challenge to the shipping companies who attempt to reduce their fuel consumption, environmental impact, and operative costs. A wide range of fuel saving solutions for shipping are available and partially implemented in the existing fleet, both from the design and operational perspective; several specific studies have been conducted on these technologies, and a more detailed treatise would be out of the scope of this work. In this context, it has been acknowledged that the world fleet is heterogeneous, and measures need to be evaluated on a ship-to-ship basis [5,6]. In this process, a deeper understanding of energy use on board of the specific ship is vital.

\subsection{Previous work}

The idea of analyzing in detail the energy flows of ship systems is not new. The studies published in these regards can generally be divided in two main categories: those based on operational measurements, and those based on mathematical modeling the ship systems.

The work presented by \cite{Thomas2010}, \cite{Basurko2013} (fishing vessels) and \cite{Baldi2015} (product tanker) falls in the first category. Although the results differ from study to study, all agree in putting propulsion at the top of the list of energy requirement (from 76\% to 87.3\% in the case of fishing vessels \cite{Thomas2010,Basurko2013}, 68\% in the case of the chemical tanker \cite{Baldi2015}, although it should be noted that in the former two cases heat demand is not taken into account).

A number of studies concerning ship energy systems can be found in literature. Thomas et al. [7] and Basurko et al. [8] worked on energy auditing fishing vessels; Shi et al. [11, 12] proposed models for predicting ship fuel consumption for some specific vessel types; Balaji and Yaakob [9] analysed ship heat availability for use in ballast water treatment technologies. These studies have been of particular interest in their relative fields, but a more comprehensive approach of the totality of the ship energy system is missing. In addition, an analysis purely based on the First law of thermodynamics does not account for the irreversibilities of the systems and for the different quality of heat flows [16]. Exergy analysis, which is based on both the First and the Second laws of thermodynamics, can help addressing this shortcoming. Widely used in other industrial sectors, exergy analysis in not commonly employed in maritime technology studies, and is mostly related to waste heat recovery systems [17, 18] and refrigeration plants [19, 20]. The application of exergy analysis in shipping is still limited; Zaili and Zhaofeng [10] proposed the energy and exergy analysis of the propulsion system of an existing vessel showing that there is potential in improving ship power plant efficiency by recovering the exergy in the exhaust gas and by improving operations of the main engines.
In a previous study of the energy and exergy analysis of a product tanker [11] the dominance of propulsion as main consumer on board was highlighted, together with the substantial availability of waste heat for recover. On cruise vessels, the number of different uses of energy is larger and a complex system of different energy carriers (chemical, thermal, electrical or mechanical) is present in order to fulfill the needs for transport combined with passenger services and comfort, such as cooking and cooling in restaurants, air conditioning, and passenger entertainment facilities.
The complexity of the energy system of a ship where the energy required by propulsion is no longer the trivial main contributor to the whole energy consumption thus makes this kind of system of particular interest for the analysis of how energy is produced, transformed, and used on board. The complexity of such systems was modelled and investigated previously by Marty et al. [12,13], but to the best of our knowledge there is no study in literature describing cruise ships’ energy and exergy analyses based on actual measurements.

To add:
- REF: Marty, Exergy analysis of complex ship energy systems
- the fact that many people talk about ORCs, without focusing on the larger picture
- REF: Bouman, State-of-the-art technologies, measures and potential for reducing GHG emissions from shipping, A review
- REF: Andersson, Shipping and the Environment (book)
- REF: Koroglu, Advanced exergy analysis of an ORC WHR of a marine power plant
- REF: A comparative life cycle assessment of marine power systems

\subsection{Aim}

The aim of this paper is to provide a better understanding of how energy is used on board of a cruise ship and where the largest potential for improvement is located by applying energy and exergy analysis to the a case study. The combination of a method rarely applied in the shipping sector to a ship type featuring a complex energy system is considered as the main contribution of this work to the existing literature in the field.

\section{Method} \label{sec:method}

\subsection{Method overview} \label{sec:met:overview}

\subsection{Energy analysis} \label{sec:met:energy}

\subsection{Exergy analysis} \label{sec:met:exergy}

\subsection{Case study vessel} \label{sec:met:case}

The ship under study is a cruise ship operating on a daily basis in the Baltic Sea between Stockholm and the island of Åland. The ship is 176.9 m long and has a beam of 28.6 m, and has a design speed of 21 knots. The ship was built in Aker Finnyards, Raumo Finland in 2004. The ship has a capacity of 1800 passengers and has several restaurants, night clubs and bars, as well as saunas and pools. This means that the energy system regarding the heat and electricity demand is more complex than a regular cargo vessel in the same size. Typical ship operations, although they can vary slightly between different days, are represented in Figure 1. It should be noted that the ship stops and drifts in open sea during night hours before mooring at its destination in the morning, if allowed by weather conditions.

The ship systems are summarized in Figure 2. The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller. The main engines are four Wärtsilä 4-stroke Diesel engines (ME) rated 5850 kW each All engines are equipped with selective catalytic reactors (SCR) for NOX emissions abatement. Propulsion power is needed whenever the ship is sailing; however, it should be noted that the ship rarely sails at full speed, and most of the time it only needs one or two engines operated simultaneously.

Auxiliary power is provided by four auxiliary engines (AE) rated 2760 kW each. Auxiliary power is needed on board for a number of alternative functions, from pumps in the engine room to lights, restaurants, ventilation and entertainment for the passengers.

Auxiliary heat needs are fulfilled by the exhaust gas steam generators (HRSG) located on all four AEs and on two of the four MEs or by oil-fired auxiliary boilers (mainly when in port, or during winter), by the heat recovery on the HT cooling water systems (HRHT), and by the auxiliary, oil fired boilers (AB). The heat is needed for passenger and crew accommodation, as well as for the heating of the highly viscous heavy fuel oil used for engines and boilers. This last part, however, is drastically reduced since the 1st of January 2015, as new regulations entering into force require the use of low-sulphur fuels, which require a much more limited heating.

\subsection{Data gathering} \label{sec:met:gathering}

\subsection{Data preprocessing} \label{sec:met:processing}

\subsection{Ship energy system modeling} \label{sec:met:modeling}

\subsection{Estimation of the heat demand} \label{sec:met:heat}

\subsubsection*{Production-side}

In a top-down model, the demand of an energy system is modeled based on the production side of the energy balance. According to the top-down approach, the total heat consumption is calculated based on the sum of the heat provided by the three main heat sources available on board:
\begin{itemize}
	\item Exhaust gas boilers
	\item Auxiliary oil-fired boilers
	\item Heat recovery on the HT cooling systems
\end{itemize}

For the case of the EGBs, the heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
\begin{equation}
\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
\end{equation}\label{eq:egb}

Where measurements of the exhaust gas temperature before and after the EGBs (respectively $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$) are available from the DLS; the mass flow of exhaust gas ($\dot{m}_{eg}$) is calculated as described in appendices \ref{app:me} and \ref{app:ae}; and the specific heat at constant pressure of the exhaust gas ($c_{p,eg}$) is assumed to be constant and equal to 1.08 $\frac{kJ}{kgK}$.

For the case of the ABs, the heat transferred to the steam ($\dot{Q}_{AB}$) is calculated according to equation \ref{eq:ab}:
\begin{equation}
\dot{Q}_{AB} = \dot{m}_{fuel,AB} \eta_{AB}
\end{equation} \label{eq:ab}

Where the fuel consumption of the ABs ($\dot{m}_{fuel,AB}$) is available on a monthly basis. No information was available concerning the first-law efficiency of the boilers ($ \eta_{AB}$). \cite{Cohen1962} estimated it to vary between 0.83 and 0.89 depending on the load of the boiler, while according to a more recent experimental campaign presented in \cite{Mrzljak2017} the efficiency varies between 0.7 and 0.79. In the absence of more specific available data, a value of 0.8 was selected for $ \eta_{AB}$

The contribution from the heat recovery on the high temperature cooling systems of the engines ($\dot{Q}_{HRHT}$) represents the main uncertainty related to the top-down estimations. This contribution cannot be measured, neither directly or indirectly, and hence needs to be estimated.

For this reason, two calculated values for the top-down heat demand are provided:
\begin{itemize}
	\item A \textbf{high boundary} case, where it is assumed that almost all of the heat available from the HT cooling systems (an utilization factor of 0.9 is assumed to account for losses) is used. This corresponds to the assumption that the use of the waste heat from the HT cooling systems is prioritized over the use of waste heat from the exhaust gas of the engines.
	\begin{equation}
	\dot{Q}_{HRHT} = f_{HRHT} \dot{Q}_{HT,tot} 
	\end{equation}
	\item a \textbf{low boundary} case, where it is assumed that the waste heat from the cooling systems is only used when the ship is in port (again assuming a utilization factor of 0.9). 
\end{itemize}

In both cases, the available heat from the HT cooling systems is estimated as explained in Section \ref{eq:met:modeling}.

\subsubsection*{Consumers-side}
Following, the assumptions that we have made related to the bottom-up estimations of ship heat demand:
\begin{itemize}
	\item The maximum demand for \textbf{spacial heating} is connected to a design minimum outer air temperature of -5$^o$C.
	\item The minimum demand for \textbf{spacial heating} is 0 for an external temperature of 20$^o$C (NOTE: Maybe 15 is more reasonable?)
	\item \textbf{Spacial heating} demand is linearly dependent on the external air temperature
	\item In the HVAC unit, the "\textbf{Preheater}" is only used when heating (i.e., in winter) while the "\textbf{Reheater}" is only used when cooling (i.e., in summer)
	\item Every passenger contributes with a total of 150 W of free heat during the day (8-23, ref. to "walking, seated" activity from \cite{Wang2000}) and 100 W during night (23-8, ref. to "seated at rest" activity from \cite{Wang2000}). For the crew, as it is assumed that they perform a more intense activity than the passengers, the daily contribution is considered equal to 160 (ref. to "moderate work" activity from \cite{Wang2000}) 
	\item The heat required for \textbf{tank heating} is linearly dependent on the external sea water temperature. The maximum heat is required for an external seawater temperature of 0$^o$C
	\item The heat required from the \textbf{galley} depends on two parameters:
	\begin{itemize}
		\item The \textbf{number of passengers} (linear dependence)
		\item the \textbf{hour of the day}. The hour-dependent factor is 0 at night (21-7), 0.1 during non-eating times (10-11 and 14-18) and 1 during eating times
	\end{itemize}
	\item The heat required from the \textbf{hot water} depends on two parameters:
	\begin{itemize}
		\item The \textbf{number of passengers} (linear dependence)
		\item the \textbf{hour of the day}. The hour-dependent factor is assumed based on the estimations for a land-based hotel reported in \cite{Chung2015}.
	\end{itemize}   
\end{itemize}

In Table \ref{tab:heatConsumers} we summarize all heat consumers on board the case study vessel. For each entry, we provide the design heat demand and the variables we assumed the actual heat demand to depend on:
\begin{table}
	\begin{tabular}{lrccccc}
		\hline 
		Name & $\dot{Q}_{des}$ [kW] & \multicolumn{5}{c}{Depends on...} \\ 
		\hline 
		 & & $T_{air}$ & t & $N_{pax}$ & $f_{TL}$ & $\lambda$ \\ 
		\hline 
		Hot water heater & 1200 &  & x & x &  &  \\ 
		HVAC Preheater & 3500 & x & x & x &  &  \\ 
		HVAC Reheater & 1780 & x & x & x &  &  \\ 
		Tank heating & 208 &  &  &  & x &  \\ 
		Other tanks & 138 &  &  &  & x &  \\ 
		HFO Tank heating & 271 &  &  &  & x &  \\ 
		Machinery space heaters & 281 & x &  &  &  &  \\ 
		Bilge water separator & 26 &  &  &  &  &  \\ 
		Hot water calorifier & 366 &  & x &  &  &  \\ 
		Fuel oil heater & 103 &  &  &  &  & x \\ 
		HFO Separator & 37 &  &  &  &  &  \\ 
		Galley & 602 &  & x & x &  &  \\
		\hline 
	\end{tabular} \label{tab:heatConsumers}
\end{table}
\section{Energy analysis}
\label{sec:energy}

\subsection{Exploratory data analysis}

\subsection{Mechanical and electric energy demand}

\subsection{Heat demand}

\section{Exergy Analysis}
\label{sec:exergy}

\section{Discussion}
\label{sec:discussion}

\section{Conclusion}
\label{sec:conclusion}

\appendix

\section{Estimation of air and exhaust gas flows in the main engines}

\section{Estimation of air and exhaust gas flows in the auxiliary engines}

\bibliographystyle{plain}
\bibliography{bibliography}

%% \begin{thebibliography}{widestlabel}
%% \bibitem{label}
%% Text of bibliographic item
%%\end{thebibliography}

\end{document}
\endinput
%%
%% End of file `elsarticle-template-num.tex'.

### 2017-05-18 Manuscript/Manuscript.tex
Manuscript update

@@
 
-According to the third IMO GHG Study, in 2012 CO$_2$ emissions from shipping amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions [1].
-Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios [1].
-As an example, global transport demand has increased by 3.8\% in 2013, compared to a global GBP growth of 2.3\% the same year, which shows how shipping tends to rise even faster than global economy [2].
+According to the third IMO GHG Study, in 2012 CO$_2$ emissions from shipping amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions \cite{Smith2014}.
+Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios \cite{Smith2014}.
+As an example, global transport demand has increased by 3.4\% in 2014, compared to a global GBP growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}.
 
@@
 Finally, in the Baltic Sea an emission control area is enforced by the International Maritime Organisation since January 2015 which stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels.
+More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}
 
@@
 A wide range of fuel saving solutions for shipping are available and partially implemented in the existing fleet, both from the design and operational perspective; several specific studies have been conducted on these technologies, and a more detailed treatise would be out of the scope of this work.
-In this context, it has been acknowledged that the world fleet is heterogeneous, and measures need to be evaluated on a ship-to-ship basis [5,6].
+In this context, it has been acknowledged that the world fleet is heterogeneous, and measures need to be evaluated on a ship-to-ship basis \cite{Bouman2017}.
 In this process, a deeper understanding of energy use on board of the specific ship is vital.
@@
 
-The idea of analyzing in detail the energy flows of ship systems is not new.
-The studies published in these regards can generally be divided in two main categories: those based on operational measurements, and those based on mathematical modeling the ship systems.
+The idea of improving the understanding of the behavior of the energy system of a ship is not new.
+Most of the work published around this subject relates to the use  of mathematical models of the ship systems.
 
-The work presented by \cite{Thomas2010}, \cite{Basurko2013} (fishing vessels) and \cite{Baldi2015} (product tanker) falls in the first category.
-Although the results differ from study to study, all agree in putting propulsion at the top of the list of energy requirement (from 76\% to 87.3\% in the case of fishing vessels \cite{Thomas2010,Basurko2013}, 68\% in the case of the chemical tanker \cite{Baldi2015}, although it should be noted that in the former two cases heat demand is not taken into account).
+Most authors focused on the propulsion part of the problem, as this is often the most relevant energy demand on board.
+Shi et al.
+proposed a modeling approach for predicting ship fuel consumption of a cargo/passenger ferry \cite{Shi2009}; Theotokoats and Tzelepis applied a similar procedure to the case of a Handymax product carrier \cite{Theotokatos2015}, similarly to Tillig et al., who also added the dynamic element to their predictive model \cite{Tillig2016}.
 
-A number of studies concerning ship energy systems can be found in literature.
-Thomas et al.
-[7] and Basurko et al.
-[8] worked on energy auditing fishing vessels; Shi et al.
-[11, 12] proposed models for predicting ship fuel consumption for some specific vessel types; Balaji and Yaakob [9] analysed ship heat availability for use in ballast water treatment technologies.
-These studies have been of particular interest in their relative fields, but a more comprehensive approach of the totality of the ship energy system is missing.
-In addition, an analysis purely based on the First law of thermodynamics does not account for the irreversibilities of the systems and for the different quality of heat flows [16].
-Exergy analysis, which is based on both the First and the Second laws of thermodynamics, can help addressing this shortcoming.
-Widely used in other industrial sectors, exergy analysis in not commonly employed in maritime technology studies, and is mostly related to waste heat recovery systems [17, 18] and refrigeration plants [19, 20].
-The application of exergy analysis in shipping is still limited; Zaili and Zhaofeng [10] proposed the energy and exergy analysis of the propulsion system of an existing vessel showing that there is potential in improving ship power plant efficiency by recovering the exergy in the exhaust gas and by improving operations of the main engines.
-In a previous study of the energy and exergy analysis of a product tanker [11] the dominance of propulsion as main consumer on board was highlighted, together with the substantial availability of waste heat for recover.
-On cruise vessels, the number of different uses of energy is larger and a complex system of different energy carriers (chemical, thermal, electrical or mechanical) is present in order to fulfill the needs for transport combined with passenger services and comfort, such as cooking and cooling in restaurants, air conditioning, and passenger entertainment facilities.
-The complexity of the energy system of a ship where the energy required by propulsion is no longer the trivial main contributor to the whole energy consumption thus makes this kind of system of particular interest for the analysis of how energy is produced, transformed, and used on board.
-The complexity of such systems was modelled and investigated previously by Marty et al.
-[12,13], but to the best of our knowledge there is no study in literature describing cruise ships’ energy and exergy analyses based on actual measurements.
+The work referenced above allowed improving the understanding of how many operational (such as the ship's speed) and environmental (such as wave height and wind speed) parameters influence the ship's energy performance.
+These studies, however, are not based on actual measurements of the ship's operations.
+Also, they focus entirely on the ship's power demand for propulsion.
+This is a very reasonable practice for most ship types, given that propulsion represent the largest part of the total energy demand, but does not allow improving the knowledge of the remaining part of the system.
 
-To add:
-- REF: Marty, Exergy analysis of complex ship energy systems
-- the fact that many people talk about ORCs, without focusing on the larger picture
-- REF: Bouman, State-of-the-art technologies, measures and potential for reducing GHG emissions from shipping, A review
-- REF: Andersson, Shipping and the Environment (book)
-- REF: Koroglu, Advanced exergy analysis of an ORC WHR of a marine power plant
-- REF: A comparative life cycle assessment of marine power systems
+% Data-driven analysis
+Other authors filled the gap by both including electric energy demand, and by basing their work on measurements from actual ship operations.
+The work presented by Thomas et al.
+\cite{Thomas2010} and Basurko et al.
+\cite{Basurko2013} shows the application of energy auditing methods to fishing vessels.
+This approach represents a step forward towards improving the understanding of how energy is used on board during actual ship operations.
+
+More generally, several authors have highlighted the importance of a detailed knowledge of the ship's operational profile in order to appropriately assess and optimize possible alternatives for improving ship energy efficiency \cite{Banks2013}.
+Coraddu et al., started from a statistical analysis of measured ship operations and used the aggregated data, together with a computational model of the vessel, to provide a better prediction of the actual ship's operational efficiency \cite{Coraddu2014}.
+The importance of considering the operational profile was also showed in the case of the optimization of engine-propeller interaction \cite{Baldi2015c}, in the process of retrofitting existing systems \cite{Baldi2015b,Choi2013}, and in ship design \cite{Ghassemi2017,Solem2015}.
+
+% Heat availability
+Heat demand is rarely a subject of concern on board ships, with some notable exceptions.
+This is due to a combination of generally low demand and high availability from the waste heat of the engines.
+In a previous publications by the authors, for instance, it is shown that in the case of a product tanker, although the heat demand was estimated to account for roughly 20\% of the total energy demand of the ship on a yearly basis, it only contributes to 4.1\% of the total fuel consumption (contribution of the auxiliary boilers), while the rest of the demand is fulfilled using waste heat \cite{Baldi2015a}.
+
+%% WHR and heat availability
+It should be noted that, however, much research effort has been devoted, especially in recent times, to the improvement of the efficiency of ship energy systems by recovering the waste heat available from the engines.
+With reference to different types of technologies, case studies, and designs,the several authors showed the existance of a quite significant potential for energy saving when WHR systems are employed, ranging from around 1\% for single-pressure steam cycles applied to two-stroke engines \cite{Theotokatos2012} to more complex systems based on ORCs (up to 10\%, \cite{Hountalas2012}) or including the cooling systems as a source of waste heat (over10\%, \cite{Dimopoulos2012}).
+The case of the installation of an ORC on board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2015}, showing a similar potential.
+
+The potential uses for waste heat on board are not limited to improving the efficiency of the power plant.
+Waste heat is commonly used for fulfilling on board heat demand for spatial heating and freshwater generation \cite{Molland2011,Baldi2015a,Mccarthy1990};  Balaji et al.
+proposed the use of waste heat for ballast-water systems \cite{Balaji2012}; Salmi et al.
+suggested its use for adsorption refrigeration systems \cite{Salmi2017}.
+A detailed review of potential uses for waste heat from marine engines is presented by Shu et al \cite{Shu2013}.
+
+% Cruise ships
+Some ship types constitute notable examples.
+On cruise ships the heat demand is significantly higher compared to standard cargo ships.
+Referring to winter conditions, Marty et al.
+estimated the instantaneous heat demand of a selected cruise ship to reach roughly 23 MW, compared to an estimated peak of 49 MW for propulsion and electric auxiliaries combined \cite{Marty2012}.
+
+This shows how the heat from the engines should be considered as a potential resource, rather than a waste, and that there is potential for improvement based on the optimal use of these heat sources.
+The work presented by the authors in \cite{Baldi2016} represent a step in this direction; however, there is the need for an increased detail in the estimation of the heat demand.
+
+% Exergy analysis
+Exergy analysis provides a more accurate estimation of the potential for energy recovery on board.
+The application of exergy analysis to the case of ship energy systems is, however, still limited.
+Dimopoulos et al.
+showed how the process optimizing the WHR system of a container ship can be more efficient if exergy efficiency, rather than energy efficiency, is used as the target of the optimization \cite{Dimopoulos2012}; Baldi et al.
+also analyzed the exergy flows on board of a product tanker, showing that this allows having a more accurate understanding of what parts of the system show potential for improvement \cite{Baldi2015a}; similar results were obtained by Marty et al., who focused on the power plant of a cruise ship \cite{Marty2016}.
+Koroglu et al.
+made a step futher also including advanced exergy analysis in their study \cite{Koroglu2017}.
 
@@
 
-The aim of this paper is to provide a better understanding of how energy is used on board of a cruise ship and where the largest potential for improvement is located by applying energy and exergy analysis to the a case study.
-The combination of a method rarely applied in the shipping sector to a ship type featuring a complex energy system is considered as the main contribution of this work to the existing literature in the field.
+Given the existing limitations of the current available information in scientific literature highlighted in the previous section, the objective of this work is the following:
+\begin{itemize}
+	\item Analyse the demand of a cruise ship in terms of propulsion power, electric power and heat, based on operational measurements.
+	\item Analyse the current efficiency of the system, and potential ways to improve it, by means of applying exergy analysis
+	\item Provide a more detailed analysis of the heat demand of a cruise ship, as a basis for further studies on how to optimize the whole efficiency of the propulsion plant.
+\end{itemize}
 
@@
 
-\subsection{Method overview} \label{sec:met:overview}
+Here an overview of the method is provided: 
+\begin{itemize}
+	\item Data gathering
+	\item Data filtering
+	\item Data pre-processing
+	\item Heat demand estimation
+	\item Statistical analysis
+	\item Energy and exergy analysis
+\end{itemize}
+
+In this paper, we present the application of energy and exergy analysis (described in sections \ref{set:met:energy} and \ref{sec:met:exergy} to a cruise ship.
+This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; these are used to generate the required information for calculating energy and exergy flows (Section \ref{sec:met:processing}.
+In these regards, the estimation of the heat demand is treated in a separate section (Sec.
+\ref{sec:met:heat} as it constituted a particularly challenging task and it is one of the central aspects of this article.
 
@@
 
+This section should be written by TV :-)
+
 \subsection{Exergy analysis} \label{sec:met:exergy}
+
+This section should be written by TV :-)
 
@@
 
-The ship under study is a cruise ship operating on a daily basis in the Baltic Sea between Stockholm and the island of Åland.
+The energy and exergy analysis are here applied to a specific  cruise ship operating daily cruises in the Baltic Sea between Stockholm and the island of \.{A}land.
 The ship is 176.9 m long and has a beam of 28.6 m, and has a design speed of 21 knots.
 The ship was built in Aker Finnyards, Raumo Finland in 2004.
+
 The ship has a capacity of 1800 passengers and has several restaurants, night clubs and bars, as well as saunas and pools.
 This means that the energy system regarding the heat and electricity demand is more complex than a regular cargo vessel in the same size.
-Typical ship operations, although they can vary slightly between different days, are represented in Figure 1.
+Typical ship operations, although they can vary slightly between different days, are represented in Figure \ref{fig:typicalShipOperations}.
 It should be noted that the ship stops and drifts in open sea during night hours before mooring at its destination in the morning, if allowed by weather conditions.
 
-The ship systems are summarized in Figure 2.
+The ship systems are summarized in Figure \ref{fig:shipSystems}.
 The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller.
@@
 
-\subsection{Data gathering} \label{sec:met:gathering}
+\subsection{Data gathering and pre-processing} \label{sec:met:gathering}
 
-\subsection{Data preprocessing} \label{sec:met:processing}
+This section should be written by FA :-)
 
-\subsection{Ship energy system modeling} \label{sec:met:modeling}
+\subsection{Data processing: Ship energy system modeling} \label{sec:met:modeling}
+
+Not all of the variables required to perform a full energy and exergy analysis of the system are available from measurements.
+In some cases, they are measurable, but not measured (e.g.
+some temperatures, mass flows, etc.).
+In other cases, they are simply impossible, or impractical, to measure (e.g.
+specific enthalpy, specific entropy).
+For this reason, part of the ship systems needed to be modeled in order to derive the unknown variables.
+
+\subsubsection*{Main engines}
+
+Make a figure representing the main engines, with specified the variables for which we have measurements of.
+
+List the main assumptions
+
+\subsubsection*{Auxiliary engines}
+
+Make a figure representing the main engines, with specified the variables for which we have measurements of.
+
+List the main assumptions
+
+\subsubsection*{Auxiliary boilers}
+
+\subsubsection*{Cooling systems}
 
@@
 	\item a \textbf{low boundary} case, where it is assumed that the waste heat from the cooling systems is only used when the ship is in port (again assuming a utilization factor of 0.9). 
+	\item a reference case, (NOTE: We should discuss what the assumption could be here. The one we had in the ECOS paper? Or a value halfway between the highest and lowest estimate?)
 \end{itemize}
@@
 \end{table}
-\section{Energy analysis}
-\label{sec:energy}
 
-\subsection{Exploratory data analysis}
+\section{Results (and discussion)}
 
-\subsection{Mechanical and electric energy demand}
+\subsection{Exploratory analysis} \label{sec:res:exploratory}
 
-\subsection{Heat demand}
+Here we show and discuss about some interesting values. 
+\begin{itemize}
+	\item External temperature
+	\item Sea water temperature
+	\item Ship speed
+	\item Other
+\end{itemize}
 
-\section{Exergy Analysis}
-\label{sec:exergy}
+The results are presented in two main forms:
+\begin{itemize}
+	\item Time series (comparing some relevant trips with a daily time scale)
+	\item Distribution plot (or box plot...box plots are easier to compare, distribution plots are easier to interpret)
+\end{itemize}
+
+\subsection{Energy analysis} \label{sec:res:energy}
+
+The results are presented in three main forms:
+\begin{itemize}
+	\item Time series (comparing some relevant trips with a daily time scale)
+	\item Distribution plot (or box plot...box plots are easier to compare, distribution plots are easier to interpret)
+	\item Sankey diagrams
+\end{itemize}
+
+The results should be presented both for the full database and for each operational mode (port-sailing-drifting)
+
+\subsection{Heat demand} \label{sec:res:heat}
+
+\subsection{Exergy Analysis} \label{sec:res:exergy}
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{unsrt}
 \bibliography{bibliography}

### 2017-05-18 Python files/unitstructures.py

@@ -34,7 +34,7 @@
 import pandas as pd
 
 
-def unitStructure():
+def flowStructure():
     structure = {"ME1": {}, "ME2": {}, "ME3": {}, "ME4": {}, "AE1": {}, "AE2": {}, "AE3": {}, "AE4": {}, "Other": {}}
     for idx in structure.keys():
         if idx[1] == "E":  # This basically means that this operation is only done if the system is an engine
@@ -84,26 +84,31 @@ def flowPreparation(structure, database_index):
     for system in structure:
         for unit in structure[system]:
             for flow in structure[system][unit]:
-                if structure[system][unit][flow]["type"] == "IPF": # Incompressible physical energy flow
+                if flow["type"] == "IPF": # Incompressible physical energy flow
                     structure[system][unit][flow]["mdot"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["T"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["cp"] = 0  # Note that the CP is a fixed, individual value
-                elif structure[system][unit][flow]["type"] == "CPF": # Compressible physical energy flow
+                elif flow["type"] == "CPF": # Compressible physical energy flow
                     structure[system][unit][flow]["mdot"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["T"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["h"] = pd.Series(index=database_index)
+                    structure[system][unit][flow]["h0"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["p"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["s"] = pd.Series(index=database_index)
+                    structure[system][unit][flow]["s0"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["cp"] = 0  # Note that the CP is a fixed, individual value
-                elif structure[system][unit][flow]["type"] == "Qdot": # Heat flow
+                elif flow["type"] == "Qdot": # Heat flow
                     structure[system][unit][flow]["Qdot"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["T"] = pd.Series(index=database_index)
-                elif structure[system][unit][flow]["type"] == "Wdot": # Work flow
+                elif flow["type"] == "Wdot": # Work flow
                     structure[system][unit][flow]["Wdot"] = pd.Series(index=database_index) # in KW
[…]

### 2017-05-23 Manuscript/Manuscript.tex
Some important code updates

@@
 %% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
+\usepackage{amsthm}
+%% Tthe amsmath package provides the "align" environment for nicer equation formatting
+\usepackage{amsmath}
 \def\bibsection{\section*{References}}
@@
 
-\subsection{Energy analysis} \label{sec:met:energy}
+\subsection{Energy analysis}
 
-This section should be written by TV :-)
+\emph{Energy} can be stored, transformed from one form to another (e.g.
+heat to power) and transferred between systems, but can neither be created nor destroyed (conservation law)~\textbf{Bejan et al.}.
+The system under study is the \emph{ship energy system} and is thus taken as control volume.
+The energy balance can be expressed as:
 
-\subsection{Exergy analysis} \label{sec:met:exergy}
+\begin{align}
+	\sum_{\mathrm{in}} \dot{H}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{H}_{\mathrm{out}} \\
+	\dot{H}_{\mathrm{fuel}} + \dot{H}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}+\dot{W}_{\mathrm{el}}+\dot{Q}_{\mathrm{heating}}
+\end{align}
 
-This section should be written by TV :-)
+The left-hand side term represents, on a time rate basis, the energy associated with the fuel consumed in the boilers and engines ($\dot{H}_{\mathrm{fuel}}$) and the air used in the combustion processes ($\dot{H}_{\mathrm{air}}$).
+The right-hand side term denotes the power ($\dot{W}_{\mathrm{el}}$) and heat ($\dot{Q}_{\mathrm{heating}}$) required on-site (e.g.
+propulsion and fuel heating), and the heat discharged into the environment ($\sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}$) with, for instance, the exhaust gases.
+
+The energy flow associated with a material stream is calculated as the sum of the physical and chemical enthalpies, and kinetic and potential energies are neglected.
+The physical energy is taken as the relative enthalpy, as underlined in~\textbf{Kotas et al.}, and the chemical energy is taken as the lower/higher heating value.
+The environmental conditions taken for the present analysis are the ambient pressure and seawater temperature.
+
+\subsection{Exergy analysis}
+
+\emph{Exergy} may be defined as the `maximum theoretical useful work (shaft work or electrical work) as the system is brought into complete thermodynamic equilibrium with the thermodynamic environment while the system interacts with it only'~\citep{Moran1989}.
+Unlike energy, exergy is not conserved but some is destroyed because of the irreversible phenomena taking place in real processes (e.g.
+chemical reactions like combustion).
+The exergy balance for the system under study can be expressed as:
+
+\begin{align}
+	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} = \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \\
+	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d} 
+\end{align}
+
+The left-hand side term represents, on a time rate basis, the exergy associated with the fuel and air.
+The right-hand side term denotes the exergy of the waste streams, the exergy transfers with heat and power, and the exergy destroyed in the ship system.
+Kinetic and potential exergies are neglected, and the exergy of a given material stream is derived as the sum of the physical and chemical exergies.
+The chemical exergy is calculated based on the reference environment of \textbf{Szargut (1989)}, and its value is approximatively equal to the higher heating value for hydrocarbon fuels.
+The exergy destruction can be calculated from the Gouy-Stodola theorem~\textbf:{Kotas (1995)}.
+
+The exergy balance may alternatively be formulated as:
+\begin{equation}
+\dot{E}_{\mathrm{p}} = \dot{E}_{\mathrm{f}} - \dot{E}_{\mathrm{d}} - \dot{E}_{\mathrm{l}} 
+\end{equation}
+
+where $\dot{E}_{\mathrm{p}}$ is called the exergy product, and corresponds to the desired output of the system, in exergy terms (for example, the power produced in an engine).
+$\dot{E}_{\mathrm{f}}$ denotes the exergy fuel, and represents the resources spent to drive the studied process (for instance, the fuel used in a combustion process).
+The last term $\dot{E}_{\mathrm{l}}$ corresponds to the losses of a system, such as the heat discharged into the environment with cooling water.
+
+\subsection{System performance}
+
+The following indicators are used to evaluate the system performance:
+\begin{itemize}
+	\item the exergy efficiency $\varepsilon$, defined as the ratio between the exergy product and fuel of a given component or system
+	\begin{equation} \varepsilon = \frac{\dot{E}_{\mathrm{p}}}{\dot{E}_{\mathrm{f}}} \end{equation}
+	\item the efficiency defect $\lambda$, presented in the work of Kotas, defined as the fraction of the total exergy input destroyed in the successive irreversible processes
+	\begin{equation} \lambda = \frac{\dot{E}_{\mathrm{d}}}{\dot{E}_{\mathrm{in,tot}}} \end{equation}
+	\item  the irreversibility share $\delta$, suggested in the work of Tsatsaronis, defined as the ratio between the exergy destroyed in the $i$-th component in relation to the exergy destroyed in the entire system
+	\begin{equation} \delta_i = \frac{\dot{E}_{\mathrm{d},i}}{\dot{E}_{\mathrm{d,tot}}} \end{equation}
+
+\end{itemize}
 

### 2017-05-23 Python files/Energyanalysis.py

@@ -1,15 +1,23 @@
 # Temporarily empty
+import CoolProp.CoolProp as cp
+import numpy as np
+import pandas as pd
 
-
-def eYergyAnalzsis(structure):
+def eYergyAnalzsis(structure,T0):
     for system in structure:
         for unit in structure[system]:
             for flow in structure[system][unit]:
-                if flow["Type"] == "CPF":
+                if structure[system][unit][flow]["Type"] == "CPF":
                     # If the specific enthalpy is available already, it needs to be calculated
-                    if flow["h"]
-
-
+                    p = structure[system][unit][flow]["p"].values()  # It's not really needed to do it in two steps..
+                    T = structure[system][unit][flow]["T"].values()
+                    if structure[system][unit][flow]["h"].sum() == 0:
+                        structure[system][unit][flow]["h"] = pd.Series(cp.PropSI('H','T',T,'P',p))
+                    if structure[system][unit][flow]["s"].sum() == 0:
+                        structure[system][unit][flow]["s"] = pd.Series(cp.PropSI('S', 'T', T, 'P', p))
+                    structure[system][unit][flow]["h0"] = pd.Series(cp.PropSI('H', 'T', T0, 'P', p))
+                    structure[system][unit][flow]["s0"] = pd.Series(cp.PropSI('S', 'T', T0, 'P', p))
+                    del p,T  # I am not sure if deleting variables serves a purpose in Python, but it doesn't cost
                     # Calculating the energy flows
                     structure[system][unit][flow]["Edot"] = structure[system][unit][flow]["mdot"] * (
                         structure[system][unit][flow]["h"] - structure[system][unit][flow]["h0"])
@@ -17,3 +25,17 @@ def eYergyAnalzsis(structure):
                     structure[system][unit][flow]["Bdot"] = structure[system][unit][flow]["mdot"] * (
                         (structure[system][unit][flow]["h"] - structure[system][unit][flow]["h0"]) - T0 *
                         (structure[system][unit][flow]["s"] - structure[system][unit][flow]["s0"]))
+                elif structure[system][unit][flow]["Type"] == "IPF":
+                    # Calculating the energy flows
+                    structure[system][unit][flow]["Edot"] = structure[system][unit][flow]["mdot"] * structure[system][unit][flow]["cp"] * (
+                        structure[system][unit][flow]["T"] - T0)
+                    # Calculating the exergy flows
[…]

### 2017-05-23 Python files/Main.py

@@ -56,21 +56,17 @@ import datareading as dr
 data_path = 'C:\\Users\\FrancescoBaldi\\Dropbox\\Energy and exergy analysis of a cruise ship - Journal Extension\\'
 dataset_raw = pd.read_hdf(data_path + 'birka_all_data.h5' ,'table')
 
-
 #N_POINTS = 319* 4 * 24
 #temp = CONSTANTS.monthLimits(N_POINTS)
 #MONTH_LIMIT_IDX = temp[0]
 #DAY_LIMIT_IDX = temp[1]
-dataset_raw = dr.keysRenaming(dataset_raw)
+#dataset_raw = dr.keysRenaming(dataset_raw)
 
 ######################################
 ## DATA CLEANING			##
 ######################################
 
 # Responsible: FA
-
-
-
 
 
 ######################################

### 2017-05-23 Python files/plotting.py

@@ -1,0 +1,85 @@
+import matplotlib as mp
+
+def plotMain(type,filename,structure):
+    # This function plots the contents of the analysed files.
+    # Two input values for "type" are accepted:
+    # - "off" [default]: does not print anything
+    # - "prompt": prompts the user for what s/he wants to print
+    # - "csv": reads info on what is to be read from a csv file
+    if type == "propmpt":
+        plotPrompt(structure)
+    elif type == "csv":
+        plotFromCSV(filename)
+    else:
+        return
+
+
+def plotPrompt(structure):
+    # This function prompts the user for the required inputs to plot.
+    # Two alternatives are available:
+    # - The default is "automatic". It is the one that is set by default. The input needs to be provided using a
+    #   specific format
+    # - The alternative is "simple", that is activated either typing "S" or "s" or "Simple" or "simple". This helps the
+    #   user plotting by specifically asking for the required input
+    output = {}
+    output["variables"] = []
+    input1 = input("Provide input for advanced plotting. If you wish to switch to simple mode, type ""s"" ")
+    if input1 == "s" or input1 == "S" or input1 == "Simple" or input1 == "simple":
+        while True:
+            accepted_plot_mode = ["sankey", "hist", "timeSeries"]
+            plot_mode = input("Switched to simple mode. Please enter the plot type. Available choices are: ""sankey"", ""hist"", ""timeSeries"" ")
+            if plot_mode in accepted_plot_mode:
+                output["plot_mode"] = plot_mode
+                break
+            else:
+                print("The input you gave (" + plot_mode + ") for the plot type is not correct! Maybe there was a typo? Try again!")
+        while True:
+            # We start here a loop to allow for printing more things in the same figure
+            plot_info = {}
+            while True:
+                plot_system = input("Now input what system you like to plot for. Available are choices are ""ME#"" (#=1,2,3,4), ""AE#"" (#=1,2,3,4), and ""Other"" ")
+                if plot_system in structure.keys():
+                    plot_info["system"] = plot_system
+                    break
+                else:
+                    print("The input you gave (" + plot_system + ") for the system name is not correct! Maybe there was a typo? Try again!")
+            while True:
+                plot_component = input("Now input what component of the chosen system you like to plot for:")
+                if plot_component in structure[plot_info["system"]].keys():
+                    plot_info["component"] = plot_component
+                    break
+                else:
[…]

### 2017-05-23 Python files/unitstructures.py

@@ -84,11 +84,11 @@ def flowPreparation(structure, database_index):
     for system in structure:
         for unit in structure[system]:
             for flow in structure[system][unit]:
-                if flow["type"] == "IPF": # Incompressible physical energy flow
+                if structure[system][unit][flow]["type"] == "IPF": # Incompressible physical energy flow
                     structure[system][unit][flow]["mdot"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["T"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["cp"] = 0  # Note that the CP is a fixed, individual value
-                elif flow["type"] == "CPF": # Compressible physical energy flow
+                elif structure[system][unit][flow]["type"] == "CPF": # Compressible physical energy flow
                     structure[system][unit][flow]["mdot"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["T"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["h"] = pd.Series(index=database_index)
@@ -97,17 +97,17 @@ def flowPreparation(structure, database_index):
                     structure[system][unit][flow]["s"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["s0"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["cp"] = 0  # Note that the CP is a fixed, individual value
-                elif flow["type"] == "Qdot": # Heat flow
+                elif structure[system][unit][flow]["type"] == "Qdot": # Heat flow
                     structure[system][unit][flow]["Qdot"] = pd.Series(index=database_index)
                     structure[system][unit][flow]["T"] = pd.Series(index=database_index)
-                elif flow["type"] == "Wdot": # Work flow
+                elif structure[system][unit][flow]["type"] == "Wdot": # Work flow
                     structure[system][unit][flow]["Wdot"] = pd.Series(index=database_index) # in KW
                     structure[system][unit][flow]["omega"] = pd.Series(index=database_index) # In rpm
                     # Note that Wdot flows apply to chemical, electrical and mechanical power
                 else:
                     print("Error, input type not recognized")
-                structure[system][unit][flow]["Edot"] = pd.Series(index=database_index)
-                structure[system][unit][flow]["Hdot"] = pd.Series(index=database_index)
+                structure[system][unit][flow]["Edot"] = pd.Series(index=database_index) # Energy flow
+                structure[system][unit][flow]["Bdot"] = pd.Series(index=database_index) # Exergy flow
 
     return structure
 

### 2017-05-23 create_dict.py

@@ -1,5 +1,5 @@
 import csv
-headers_dict = '/Users/fredde/Dropbox/GitHub/Ecos2015PaperExtension/data_import/headers_dict.csv'
+headers_dict = 'C:\\Users\\FrancescoBaldi\\switchdrive\\Work in progress\\Paper 0\\Ecos2015PaperExtension\\data_import\\headers_dict.csv'
 
 #%%
 
@@ -14,7 +14,7 @@ with open(headers_dict, 'r') as f:
         d[k] = v
 
 
-d[headers[0]]
+##d[headers[0]]
 
 
 #%%

### 2017-06-04 Manuscript/Manuscript.tex
Modifications to headers file and manuscript

@@
 
+Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse.
+The assumption was validated using for the June-December period, for which onboard measurements were available.
+This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
+% We did the calculation, resampling the data from SHMI from a 1(3)h to 15min frequency using linear interpolation.
+
 \subsection{Data processing: Ship energy system modeling} \label{sec:met:modeling}

### 2017-06-05 Python files/Constants.py
Update on the Python files for data analysis

@@ -31,7 +31,7 @@ def general():
     output["RHO_HFO"] = numpy.mean([890, 919, 924, 926, 925, 921, 924, 918, 920, 919, 933])  # HFO density, in [kg/m^3]
     output["AIR_STOIC"] = 14.7  # Stoichiometric ratio fuel/air for Diesel-type fuels
     output["ETA_VOL"] = 0.97 # Assumption about volumetric efficiency
-    output["P_ATM"] = 1.10325 # Assumption on atmospheric pressure
+    output["P_ATM"] = 1.01325 # Assumption on atmospheric pressure
     output["ISO"] = {"LHV": 42700, "T_CA": 298, "T_LT": 298, "ETA_MECH": 0.8} # Reference values for ISO conditions
     output["NAMES"] = {"MainEngines": ["ME1", "ME2", "ME3", "ME4"], "AuxEngines": ["AE1", "AE2", "AE3", "AE4"]}
     return output
@@ -92,6 +92,7 @@ def mainEngines(CONSTANTS):
     output["ETA_GB"] = 0.985   # Mechanical efficiency of the gearbox
     output["ETA_SHAFT"] = 0.99  # Mechanical efficiency of the engine shaft
     output["FRP_DES"] = {"ME1": 51, "ME2": 47, "ME3": 47, "ME4": 46}  # Value of the fuel rack position at 100% load
+    output["BYPASS_FLOW"] = 1.1
     return output
     
     
@@ -125,6 +126,7 @@ def auxiliaryEngines(CONSTANTS):
     output["MFR_HT"] = 60.0 * CONSTANTS["General"]["RHO_W"] / 3600.0  # Mass flow rate of HT cooling water, in [kg/s]
     output["ETA_CORR"] = 1.15  # Used because one of the engines need correction, to be checked
     output["ETA_SG"] = 0.97
+    output["EPS_CAC_HTSTAGE"] = 0.85  # Effectiveness, as defined by the epsNTU method, of the High Temperature stage of the Charge Air Cooler, in [-]
     return output
 
 

### 2017-06-05 Python files/Energyanalysis.py

@@ -21,18 +21,22 @@ def eYergyAnalzsis(structure,T0):
                     # Calculating the energy flows
                     structure[system][unit][flow]["Edot"] = structure[system][unit][flow]["mdot"] * (
                         structure[system][unit][flow]["h"] - structure[system][unit][flow]["h0"])
+                    # Calculating the specific exergy
+                    structure[system][unit][flow]["b"] = (
+                        structure[system][unit][flow]["h"] - structure[system][unit][flow]["h0"]) - T0 * (
+                        structure[system][unit][flow]["s"] - structure[system][unit][flow]["s0"])
                     # Calculating the exergy flows
-                    structure[system][unit][flow]["Bdot"] = structure[system][unit][flow]["mdot"] * (
-                        (structure[system][unit][flow]["h"] - structure[system][unit][flow]["h0"]) - T0 *
-                        (structure[system][unit][flow]["s"] - structure[system][unit][flow]["s0"]))
+                    structure[system][unit][flow]["Bdot"] = structure[system][unit][flow]["mdot"] * (structure[system][unit][flow]["b"])
                 elif structure[system][unit][flow]["Type"] == "IPF":
+                    # Calculating the specific exergy
+                    structure[system][unit][flow]["b"] = structure[system][unit][flow]["cp"] * (
+                        (structure[system][unit][flow]["T"] - T0) - T0 * (
+                        np.log(structure[system][unit][flow]["T"] / T0)) )
                     # Calculating the energy flows
                     structure[system][unit][flow]["Edot"] = structure[system][unit][flow]["mdot"] * structure[system][unit][flow]["cp"] * (
                         structure[system][unit][flow]["T"] - T0)
                     # Calculating the exergy flows
-                    structure[system][unit][flow]["Bdot"] = structure[system][unit][flow]["mdot"] * structure[system][unit][flow]["cp"] * (
-                        (structure[system][unit][flow]["T"] - T0) - T0 *
-                        np.log(structure[system][unit][flow]["T"] / T0))
+                    structure[system][unit][flow]["Bdot"] = structure[system][unit][flow]["mdot"] * structure[system][unit][flow]["b"]
                 elif structure[system][unit][flow]["Type"] == "Qdot":
                     structure[system][unit][flow]["Edot"] = structure[system][unit][flow]["Qdot"]
                     structure[system][unit][flow]["Bdot"] = structure[system][unit][flow]["Qdot"] / structure[system][unit][flow]["T"]

### 2017-06-05 Python files/Main.py

@@ -1,4 +1,3 @@
-
 ###########################################################################
 #
 ###########		ENERGY AND EXERGY ANALYSIS OF A CRUISE SHIP		###########
@@ -15,8 +14,6 @@
 # - Process data so to generate the variables of interest: in particular, energy and exergy flows
 # - Statistically analyze the data so to produce appropriate results
 
-
-
 # The main.py script calls other scripts and functions. It is divided in the following sections:
 # - INPUT
 # - DATA READING
@@ -26,6 +23,9 @@
 # - ENERGY ANALYSIS
 # - EXERGY ANALYSIS
 
+
+input_run = "yes"
+datareading_run = "yes"
 
 
 
@@ -37,10 +37,11 @@
 # Loading appropriate modules
 import pandas as pd
 import input
+import os
+import plotting as plot
+
 
 filenames = input.filenames() # Note: this is just a test
-
-
 
 
 
@@ -53,10 +54,11 @@ filenames = input.filenames() # Note: this is just a test
 
 import datareading as dr
 
-data_path = 'C:\\Users\\FrancescoBaldi\\switchdrive\Work in progress\\Paper 0\\Ecos2015PaperExtension\\data_import\\'
-translate_filename = data_path + 'headers_dict_FB.xlsx'
+project_path = os.path.realpath('..')
+translate_filename = project_path + '\\General\\headers_dict.xlsx'
 
-dataset_raw = pd.read_hdf(data_path + 'selected_data_1year_comp.h5' ,'table')
+#dataset_raw = pd.read_hdf(project_path + '\\Database\\selected_data_1year_comp.h5' ,'table')
+dataset_raw = pd.read_hdf(project_path + '\\Database\\selected_df.h5' ,'table')
 header_names = dr.keysRenaming(dataset_raw, translate_filename)
 
 ######################################
@@ -84,7 +86,7 @@ dataset_status = us.generalStatus() # Here we simply initiate the "status" struc
 # Running the pre-processing required for filling in the data structures:
 import preprocessing as pp
 # First updating the "CONSTANTS" dictionary with the some additional information
-CONSTANTS = pp.assumptions(dataset_raw, dataset_processed, CONSTANTS, header_names)
+dataset_processed = pp.assumptions(dataset_raw, dataset_processed, CONSTANTS, header_names)
 # Updating the fields of the MainEngines and the auxiliary engines
[…]

### 2017-06-05 Python files/preprocessing.py

@@ -8,6 +8,17 @@ import numpy as np
 import pandas as pd
 
 
+def assumptions(raw, processed, CONSTANTS, hd):
+    # This function includes generic assumed values in the main structure
+    for name in CONSTANTS["General"]["NAMES"]["MainEngines"]:
+        # The pressure at the turbocharger inlet is equal to the atmospheric pressure
+        processed[name]["TC"]["Air_in"]["p"] = CONSTANTS["General"]["P_ATM"]
+    for name in CONSTANTS["General"]["NAMES"]["AuxEngines"]:
+        # The pressure at the turbocharger inlet is equal to the atmospheric pressure
+        processed[name]["TC"]["Air_in"]["p"] = CONSTANTS["General"]["P_ATM"]
+    CONSTANTS["General"]["T_0"] = raw[hd["ER13_SW_T_IN"]]
+    return processed
+
 
 def mainEngineProcessing(raw, processed, CONSTANTS, status, hd):
     # This script summarizes all the functions that calculate the required data for the Main Engines different flows
@@ -20,7 +31,7 @@ def mainEngineProcessing(raw, processed, CONSTANTS, status, hd):
     # Calculating engine load, that is used many times later on
     status = mainEngineStatusCalculation(processed, CONSTANTS, status)
     # Calculating air and exhaust gas flows in the main engines
-    processed = mainEngineAirFlowCalculation(raw, processed, CONSTANTS)
+    processed = mainEngineAirFlowCalculation(raw, processed, status, CONSTANTS)
     # Calculating cooling flows
     processed = engineCoolingSystemsCalculation(processed, CONSTANTS, status, "MainEngines")
     return (processed, status)
@@ -30,15 +41,15 @@ def auxEngineProcessing(raw, processed, CONSTANTS, status, hd):
     # This script summarizes all the functions that calculate the required data for the Main Engines different flows
     # Reading existing values
     processed = readAuxEnginesExistingValues(raw, processed, CONSTANTS, hd)
-    # Calculating the main engines fuel flows
-    processed = auxEngineFuelFlowCalculation(raw, processed, CONSTANTS)
[…]

### 2017-07-03 Manuscript/Manuscript.tex
Update on the Python files for data analysis
Various extensions to Python files

@@
 Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse.
-The assumption was validated using for the June-December period, for which onboard measurements were available.
+The assumption was validated based on  June-December period, for which onboard measurements were available.
 This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
@@
 \begin{itemize}
-	\item The maximum demand for \textbf{spacial heating} is connected to a design minimum outer air temperature of -5$^o$C.
-	\item The minimum demand for \textbf{spacial heating} is 0 for an external temperature of 20$^o$C (NOTE: Maybe 15 is more reasonable?)
-	\item \textbf{Spacial heating} demand is linearly dependent on the external air temperature
 	\item In the HVAC unit, the "\textbf{Preheater}" is only used when heating (i.e., in winter) while the "\textbf{Reheater}" is only used when cooling (i.e., in summer)
+	\item \textbf{Spacial heating} demand during \textbf{winter} is linearly dependent on the external air temperature, from a minimum of 0 for an external temperature of 20$^o$C to a maximum of 3500 kW (design power of the preheater)  for an outer air temperature of -20$^o$C.
+	\item \textbf{Spacial heating} demand during \textbf{summer} is related to the need for dehumidification, and harder to estimate based on available measurements.
+As a first assumption, it is estimated to be equal to the cooling demand (estimated as detailed in Sec.
+ref{sec:met:electric})
 	\item Every passenger contributes with a total of 150 W of free heat during the day (8-23, ref.
@@
 		\item The \textbf{number of passengers} (linear dependence)
-		\item the \textbf{hour of the day}. The hour-dependent factor is 0 at night (21-7), 0.1 during non-eating times (10-11 and 14-18) and 1 during eating times
+		\item the \textbf{hour of the day}. See Figure \ref{fig:hwhvstime}
 	\end{itemize}
@@
 \end{table}
+
+\subsection{Estimation of the electric power demand} \label{sec:met:electric}
+
+The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy
+\begin{equation}
+P_{el,tot} = \sum_{1}^{4} AE_i
+\end{equation}
+
+The determination of how different consumers contribute to the total electric power demand is less trivial.
+The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc.
+None of the individual contributions is directly measured, hence estimations need to be based on indirect measurements.
+
+\subsubsection*{HVAC systems}
+
+For what concerns the air conditioning unit demand (HVAC), the observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a clear period, corresponding to the summer months, during which there is a significant increase of the demand.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pel_vs_time}
+	\caption[Total electric power demand versus time (Daily average)]{Total electric power demand versus time (Daily average)}
+	\label{fig:pelvstime}
+\end{figure}
+
+Since the total demand seems to be rather constant otherwise, it can be concluded that the demand of the HVAC compressors related to the need for cooling is concentrated during the summer months.
+Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a random summer and winter day, shows that the daily variation is comparable.
+This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/PelRel_vs_time}
+	\caption{Instantaneous power demand over daily average, winter versus summer day}
+	\label{fig:pelrelvstime}
+\end{figure}
+
+Hence, the HVAC electric power demand was estimated as follows:
+\begin{equation}
+P_{el,HVAC} =
+\begin{cases}
+P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
+0, & \text{otherwise}
+\end{cases}
+\end{equation}
+where $P_{el,ref}(t)$ is calculated as:
+\begin{equation}
+	P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
+\end{equation}
+
+\subsubsection*{Thrusters}
+
+When entering port areas, the ship needs to use thrusters to maneuver and berth.
+In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.
+
+The observation of Figure XX shows how the power demand from thrusters can be identified as "spikes" in the total electric power demand.
+In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
+From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).
+
+By comparing the reference profile to the instantaneous one, the points where the former is more than 10\% higher than the latter are identified as "thruster-on" points and treated consequently.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Thruster_selection}
+	\caption{Example of the procedure of selection of thruster power demand}
+	\label{fig:thruster_selection}
+\end{figure}
 

### 2017-07-03 Main.py

@@ -41,6 +41,8 @@ do_energy_analysis = "yes"
 import sys
 import os
 import pandas as pd
+import matplotlib.pylab as plt
+import numpy as np
 project_path = os.getcwd()
 path_files = project_path + os.sep + 'Python files' + os.sep
 sys.path.append(path_files)
@@ -55,6 +57,7 @@ import preprocessingAE as ppa
 import preprocessingME as ppm
 import preprocessingO as ppo
 import energyanalysis as ea
+import auxiliaryDemand as aux
 import coolingsystems as cs
 
 #%%
@@ -80,8 +83,6 @@ CONSTANTS = kk.constantsSetting()
 CONSTANTS["filenames"] = filenames
 (dict_structure, processed) = us.structurePreparation(CONSTANTS, dataset_raw.index, CONSTANTS["filenames"]["dataset_output_empty"], do_processed_data_preparation)
 
-# Running the pre-processing required for filling in the data structures:
-
 # First updating the "CONSTANTS" dictionary with the some additional information
 processed = ppo.assumptions(dataset_raw, processed, CONSTANTS, header_names)
 # Updating the fields of the MainEngines and the auxiliary engines
@@ -90,14 +91,18 @@ if do_main_engines_analysis == "yes":
 if do_aux_engines_analysis == "yes":
     processed = ppa.auxEngineProcessing(dataset_raw, processed, dict_structure, CONSTANTS, header_names)
 processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-1")
+# Calculating the auxiliary power demands: heating and electric power
+processed = aux.auxPowerAnalysis(processed, CONSTANTS, dict_structure)
+# Calculating the central cooling systems
+processed = cs.centralCoolingSystems(processed, CONSTANTS)
+# Filling in
 processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-2")
-
-
-
+processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-3")
 # Assigning defined values to all flows for engines off
-
 processed = ea.energyAnalysisLauncher(processed, dict_structure, CONSTANTS, do_energy_analysis)
-
+# Re-doing the calculation of the connected points
+processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-4")
+processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-5")
 
 ######################################
 ## RESULTS CHECK                	##

### 2017-07-03 Python files/Constants.py

@@ -1,5 +1,4 @@
 import numpy as np
-import pdb
 import CoolProp.CoolProp as cp
 
 
@@ -16,8 +15,8 @@ def constantsSetting():
 def general():
     # GENERAL CONSTANTS
     output = {}
-    output["R_0"] = 8314.0 # Ideal gas constant, in [J/mol*K]
-    output["R_AIR"] = 8314.0 / 29 # Air gas constant, in [kJ/kg*K]
+    output["R_0"] = 8.3140 # Ideal gas constant, in [J/mol*K]
+    output["R_AIR"] = output["R_0"] * 1000 / 29 # Air gas constant, in [J/kg*K]
     output["K_AIR"] = 1.4   # Air specific heat ratio
     output["CP_AIR"] = 1.01   # Air specific heat, in [kJ/kgK]
     output["CP_EG"] = 1.10   # EG specific heat, in [kJ/kgK]
@@ -33,10 +32,12 @@ def general():
     output["ISO"] = {"LHV": 42700, "T_CA": 298, "T_LT": 298, "ETA_MECH": 0.8} # Reference values for ISO conditions
     output["NAMES"] = {"MainEngines": ["ME1", "ME2", "ME3", "ME4"], "AuxEngines": ["AE1", "AE2", "AE3", "AE4"]}
     output["PROPERTY_LIST"] = {}
-    output["PROPERTY_LIST"]["CPF"] =  ["mdot", "T", "p", "h", "b", "Edot", "Bdot"]
-    output["PROPERTY_LIST"]["IPF"] = ["mdot", "T", "p", "h", "b", "Edot", "Bdot"]
-    output["PROPERTY_LIST"]["Qdot"] = ["Qdot", "T", "Edot", "Bdot"]
-    output["PROPERTY_LIST"]["Wdot"] = ["Wdot", "omega", "Edot", "Bdot"]
+    output["PROPERTY_LIST"]["CPF"] =  ["mdot", "T", "p", "h", "b", "Edot", "Bdot"] # Compressible physical flow
+    output["PROPERTY_LIST"]["IPF"] =  ["mdot", "T", "h", "b", "Edot", "Bdot"] # Incompressible physical flow
[…]

### 2017-07-03 Python files/auxiliaryDemand.py

@@ -1,0 +1,161 @@
+import numpy as np
+import pandas as pd
+from helpers import d2df
+
+
+def auxPowerAnalysis(processed, CONSTANTS, dict_structure):
+    # Calculating the total auxiliary power demand
+    processed = auxPowerTotalDemand(processed)
+    # Calculating the contribution from the thrusters
+    #processed = thrusters(processed, CONSTANTS)  # The second method works better. But I still keep it
+    processed = thrusters2(processed, CONSTANTS)
+    processed = HVAC(processed, CONSTANTS)
+    processed = heatDemand(processed, CONSTANTS, dict_structure)
+    processed = HTHR(processed, CONSTANTS)
+    return processed
+
+
+
+def auxPowerTotalDemand(processed):
+    processed["Demands:Electricity:Total:Edot"] = processed["AE1:AG:Power_out:Edot"]+processed["AE2:AG:Power_out:Edot"]+processed["AE3:AG:Power_out:Edot"]+processed["AE4:AG:Power_out:Edot"]
+    # We also need to store (for further use) a series made of the daily averages
+    processed["Demands:Electricity:Total:Edot1D"] = processed["Demands:Electricity:Total:Edot"].resample('1D').mean().reindex(processed["Demands:Electricity:Total:Edot"].index, method='ffill')
+    processed["Demands:Electricity:Other:Edot"] = processed["Demands:Electricity:Total:Edot"]
+    return processed
+
+
+
+def HVAC(processed, CONSTANTS):
+    # The idea of the calculation of the HVAC demand, very much simplified, lies in saying that during the summer monhs
+    # the HVAC demand is the difference between the average consumption before summer AND the instananteous one.
+    mask = (processed.index > CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HVAC_START"]) & (processed.index <= CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HVAC_END"])
[…]

### 2017-07-03 Python files/consistencycheck.py

@@ -55,9 +55,8 @@ def missingValues(processed, dict_structure, CONSTANTS):
         for unit in dict_structure["systems"][system]["units"]:
             for flow in dict_structure["systems"][system]["units"][unit]["flows"]:
                 for property in dict_structure["systems"][system]["units"][unit]["flows"][flow]["properties"]:
-                    if property not in {"h", "b", "Edot", "Bdot"}:
-                        if processed[d2df(system,unit,flow,property)].isnull().sum() == len(processed[system + ":" + "on"]):
-                            text_file.write("The field {}:{}:{}:{} is still empty \n".format(system,unit,flow,property))
+                    if processed[d2df(system,unit,flow,property)].isnull().sum() == len(processed[system + ":" + "on"]):
+                        text_file.write("The field {}:{}:{}:{} is still empty \n".format(system,unit,flow,property))
     text_file.close()
     print("...done!")
 

### 2017-07-03 Python files/coolingsystems.py

@@ -9,8 +9,8 @@ def engineCoolingSystemsCalculation(processed, CONSTANTS, engine_type):
     # This function calculates the different flows related to the cooling systems of the main engines.
     for system in CONSTANTS["General"]["NAMES"][engine_type]:
         # Calculating the total energy flow going to the cooling systems, based on the energy balance on the engine
-        energy_2_cooling = (processed[d2df(system,"Cyl","FuelCh_in","Wdot")] -
-            processed[d2df(system,"Cyl","Power_out","Wdot")] +
+        energy_2_cooling = (processed[d2df(system,"Cyl","FuelCh_in","Edot")] -
+            processed[d2df(system,"Cyl","Power_out","Edot")] +
             CONSTANTS["General"]["HFO"]["CP"] * processed[d2df(system,"Cyl","FuelPh_in","mdot")] *
                 (processed[d2df(system,"Cyl","FuelPh_in","T")] - processed["T_0"]) -
             CONSTANTS["General"]["CP_EG"] * processed[d2df(system,"Turbine","Mix_in","mdot")] *
@@ -34,8 +34,8 @@ def engineCoolingSystemsCalculation(processed, CONSTANTS, engine_type):
         energy_2_cac_lt = energy_2_cac - energy_2_cac_ht
         # The energy to the JWC results as a balance over the HT cooling systems
         energy_2_jwc = energy_2_ht - energy_2_cac_ht
-        processed[d2df(system,"JWC","QdotJW_in","Qdot")] = energy_2_jwc
-        processed[d2df(system, "Cyl", "QdotJW_out", "Qdot")] = energy_2_jwc
+        processed[d2df(system,"JWC","QdotJW_in","Edot")] = energy_2_jwc
+        processed[d2df(system, "Cyl", "QdotJW_out", "Edot")] = energy_2_jwc
         # The energy to the LOC results as a balance over the LT cooling systems
         energy_2_loc = energy_2_lt - energy_2_cac_lt
         # Finally, the temperatures in the flows are calculated based on the calculated energy and mass flow values
@@ -55,6 +55,10 @@ def engineCoolingSystemsCalculation(processed, CONSTANTS, engine_type):
             processed[d2df(system,"HRSG","Steam_in","mdot")] = ((processed[d2df(system,"HRSG","Mix_in","T")] - processed[d2df(system,"HRSG","Mix_out","T")]) *
[…]

### 2017-07-03 Python files/heatdemand.py

@@ -1,0 +1,35 @@
+import numpy as np
+import pandas as pd
+
+
+def heatDemand(processed, CONSTANTS):
+    # This function calculates the heat demand from the on board systems
+    db_index = processed.index
+    processed[""] = CONSTANTS["Other"]["HEAT_DEMAND"]["HOT_WATER_HEATER"] * pd.Series(len(processed)/96*np.repeat(CONSTANTS["Other"]["HEAT_DEMAND"]["HWH_HOURLY"],4), index=db_index)
+    qdot_hvac_preheater = CONSTANTS["Other"]["HEAT_DEMAND"]["HVAC_PREHEATER"] * (CONSTANTS["Other"]["HEAT_DEMAND"]["T_INSIDE"] - processed["T_air"]) / (CONSTANTS["Other"]["HEAT_DEMAND"]["T_INSIDE"] - CONSTANTS["Other"]["HEAT_DEMAND"]["T_AIR_REF"])
+    qdot_hvac_preheater[qdot_hvac_preheater < 0.1 * CONSTANTS["Other"]["HEAT_DEMAND"]["HVAC_PREHEATER"]] = 0 # The pre-heater is not working during summer
+    # qdot_hvac_reheater = CONSTANTS["Other"]["HEAT_DEMAND"]["HVAC_PREHEATER"] * (CONSTANTS["Other"]["HEAT_DEMAND"]["T_INSIDE"] - processed["T_air"]) / (CONSTANTS["Other"]["HEAT_DEMAND"]["T_INSIDE"] - CONSTANTS["Other"]["HEAT_DEMAND"]["T_AIR_REF"])
+
+
+
+
+    return processed
+
+
+
+
+
+
+
+
+
+
+
+
+def heatDemandProducerSide(processed, CONSTANTS):
+
+
+
+
+
+    return processed

### 2017-07-03 Python files/preprocessingAE.py

@@ -49,7 +49,7 @@ def readAuxEnginesExistingValues(raw, processed,CONSTANTS,hd):
         # Reading charge air temperature.
         processed[d2df(system,"CAC_LT","Air_out","T")] = raw[hd[system + "-CAC_AIR_T_OUT"]] + 273.15
         # Reading power output
-        processed[d2df(system,"AG","Power_out","Wdot")] = raw[hd[system + "_POWER_Wdot_OUT"]]
+        processed[d2df(system,"AG","Power_out","Edot")] = raw[hd[system + "_POWER_Wdot_OUT"]]
         # Reading Engine rpm
         # processed[d2df(system, "Cyl", "Power_out", "omega")] = raw[hd[system + "__RPM_"]]
         # Reading the pressure in the cooling flows
@@ -61,6 +61,11 @@ def readAuxEnginesExistingValues(raw, processed,CONSTANTS,hd):
         # Reading the pressure in the cooling flows
         processed[d2df(system, "CAC_LT", "LTWater_in", "p")] = (raw[hd[system + "-LT-CAC_FW_P_IN"]] + 1.01325) * 100000
         processed[d2df(system, "JWC", "HTWater_in", "p")] = (raw[hd[system + "-HT-JWC_FW_P_IN"]] + 1.01325) * 100000
+        # Reading turbocharger speed
+        processed[d2df(system, "TCshaft", "Power_in", "omega")] = raw[hd[system + "-TC__RPM_"]]
+        processed[d2df(system, "TCshaft", "Power_out", "omega")] = raw[hd[system + "-TC__RPM_"]]
+        processed[d2df(system, "Turbine", "Power_out", "omega")] = raw[hd[system + "-TC__RPM_"]]
+        processed[d2df(system, "Compressor", "Power_in", "omega")] = raw[hd[system + "-TC__RPM_"]]
     print("...done!")
     return processed
 
@@ -69,12 +74,12 @@ def auxEnginePowerCalculation(processed, CONSTANTS):
     print("Started calculating auxiliary engine power...", end="", flush=True)
     # Calculating the power of the auxiliary engines
     for system in CONSTANTS["General"]["NAMES"]["AuxEngines"]:
-        load = processed[d2df(system,"AG","Power_out","Wdot")] / CONSTANTS["AuxEngines"]["MCR"]
+        load = processed[d2df(system,"AG","Power_out","Edot")] / CONSTANTS["AuxEngines"]["MCR"]
[…]

### 2017-07-03 Python files/preprocessingME.py

@@ -60,10 +60,15 @@ def readMainEnginesExistingValues(raw, processed, CONSTANTS, hd):
         # Pressure of the charge air, at the compressor outlet (and, hence, at the cylinder inlet)
         processed[d2df(system, "Comp", "Air_out", "p")] = (raw[hd[system + "-CAC_AIR_P_OUT"]] + 1.01325) * 100000
         # Reading the pressure of the lubricating oil
-        processed[d2df(system, "LOC", "LubOil_out", "p")] = (raw[hd[system + "-LOC_OIL_P_IN"]] + 1.01325) * 100000
+        # processed[d2df(system, "LOC", "LubOil_out", "p")] = (raw[hd[system + "-LOC_OIL_P_IN"]] + 1.01325) * 100000
         # Reading the pressure in the cooling flows
         processed[d2df(system, "CAC_LT", "LTWater_in", "p")] = (raw[hd[system + "-LT-CAC_FW_P_IN"]] + 1.01325) * 100000
         processed[d2df(system, "JWC", "HTWater_in", "p")] = (raw[hd[system + "-HT-JWC_FW_P_IN"]] + 1.01325) * 100000
+        # Reading turbocharger speed
+        processed[d2df(system, "TCshaft", "Power_in", "omega")] = raw[hd[system+"-TC__RPM_"]]
+        processed[d2df(system, "TCshaft", "Power_out", "omega")] = raw[hd[system + "-TC__RPM_"]]
+        processed[d2df(system, "Turbine", "Power_out", "omega")] = raw[hd[system + "-TC__RPM_"]]
+        processed[d2df(system, "Compressor", "Power_in", "omega")] = raw[hd[system + "-TC__RPM_"]]
 
 
     print("...done!")
@@ -105,9 +110,9 @@ def mainEnginePowerCalculation(processed, CONSTANTS):
         processed[d2df(system,"Cyl","FuelPh_in","mdot")] = processed[d2df(system,"Cyl","FuelPh_in","mdot")] * bsfc / bsfc_iso
         # Calculates the power of the engine as mfr/bsfc, with unit conversion to get the output in kW
         # Shaft energy out
-        processed[d2df(system,"Cyl","Power_out","Wdot")] = processed[d2df(system,"Cyl","FuelPh_in","mdot")] / bsfc * 1000 * 3600
+        processed[d2df(system,"Cyl","Power_out","Edot")] = processed[d2df(system,"Cyl","FuelPh_in","mdot")] / bsfc * 1000 * 3600
[…]

### 2017-07-03 Python files/preprocessingO.py

@@ -6,6 +6,55 @@ import CoolProp.CoolProp as cp
 
 
 
+def assumptions(raw, processed, CONSTANTS, hd):
+    # This function includes generic assumed values in the main structure
+    # ALL ENGINES
+    for system in {"ME1", "ME2", "ME3", "ME4", "AE1", "AE2", "AE3", "AE4"}:
+        # The pressure at the turbocharger air inlet and exhaust outlet is equal to the atmospheric pressure
+        processed[d2df(system,"Comp","Air_in","p")][:] = CONSTANTS["General"]["P_ATM"]
+        processed[d2df(system, "Turbine", "Mix_out", "p")][:] = CONSTANTS["General"]["P_ATM"]
+        # Temperature in the engine room, i.e. inlet to the compressor of the TC
+        processed[d2df(system, "Comp", "Air_in", "T")] = raw[hd["ER_AIR_T_"]] + 273.15
+        # Assuming that the pressure in the exhaust gas is 90% of the pressure in the inlet manifold. Somewhat reasonable
+        processed[d2df(system, "Cyl", "EG_out", "p")] = (0.9 * raw[hd[system + "-CAC_AIR_P_OUT"]] + 1.01325) * 100000
+        # Assuming the pressure of the fuel to be around 9 barg, based on measurements from ME4
+        processed[d2df(system, "Cyl", "FuelPh_in", "p")][:] = (9 + 1.01325) * 10e5
+        # Assuming the temperature of the cylinder wall to be 150 degC
+        processed[d2df(system, "Cyl", "QdotJW_out", "T")][:] = 150 + 273.15
+        processed[d2df(system, "JWC", "QdotJW_in", "T")][:] = 150 + 273.15
+        # Assuming a temperature of 100 degC for heat losses from the TC shaft
+        processed[d2df(system, "TCshaft", "Losses_out", "T")][:] = 100 + 273.15
+        # Assuming the steam pressure and temperature in the HRSG to be constant...
+        hrsg_pressure_assumption = (6 + 1.01325) * 100000
+        #if system in {"AE1", "AE2", "AE3", "AE4", "ME2", "ME3"}:
+        #    processed[d2df(system, "HRSG", "Steam_in", "T")][:] = cp.PropsSI('T', 'P', hrsg_pressure_assumption, 'Q', 0.5, "Water")
+        #    processed[d2df(system, "HRSG", "Steam_in", "p")][:] = hrsg_pressure_assumption
[…]

### 2017-07-03 Python files/unitstructures.py

@@ -56,15 +56,16 @@ def structurePreparation(CONSTANTS, index, empty_dataset_filesystem, data_struct
 
 def flowStructure():
     print("Started preparing the dataset_processed multi-level dictionary...", end="", flush=True)
-    structure = {"systems": {"ME1": {}, "ME2": {}, "ME3": {}, "ME4": {}, "AE1": {}, "AE2": {}, "AE3": {}, "AE4": {}, "Other": {}}}
+    structure = {"systems": {"ME1": {}, "ME2": {}, "ME3": {}, "ME4": {}, "AE1": {}, "AE2": {}, "AE3": {}, "AE4": {},
+                             "CoolingSystems": {}, "HTHR":{}, "Steam":{}, "Demands":{}}}
     for system in structure["systems"]:
         structure["systems"][system]["equations"] = {}
         if system[1] == "E":  # This basically means that this operation is only done if the system is an engine
             structure["systems"][system]["units"] = {"Comp": {}, "Turbine": {}, "BPsplit": {}, "BPmerge": {}, "BPvalve": {},
                                                      "CAC_HT": {}, "CAC_LT": {}, "LOC": {}, "JWC": {}, "Cyl": {},
-                                                     "HTsplit": {}, "HTmerge": {}, "LTsplit": {}}
+                                                     "HTsplit": {}, "HTmerge": {}, "LTsplit": {}, "TCshaft": {}}
             # Compressor
-            structure["systems"][system]["units"]["Comp"]["flows"] = {"Air_in": {"type": "CPF"}, "Air_out": {"type": "CPF"}} # TC compressor
+            structure["systems"][system]["units"]["Comp"]["flows"] = {"Air_in": {"type": "CPF"}, "Air_out": {"type": "CPF"}, "Power_in": {"type": "Wdot"}} # TC compressor
             structure["systems"][system]["units"]["Comp"]["equations"] = ["MassBalance"]
             # Bypass Split
             structure["systems"][system]["units"]["BPsplit"]["flows"] = {"Air_in": {"type": "CPF"}, "Air_out": {"type": "CPF"}, "BP_out": {"type": "CPF"}}  # Bypass valve
@@ -76,8 +77,11 @@ def flowStructure():
             structure["systems"][system]["units"]["BPvalve"]["flows"] = {"BP_in": {"type": "CPF"}, "BP_out": {"type": "CPF"}}
             structure["systems"][system]["units"]["BPvalve"]["equations"] = ["MassBalance", "ConstantTemperature"]
[…]

### 2017-07-05 Manuscript/Manuscript.tex
More updates on the python files

@@
 
-Here we show and discuss about some interesting values. 
+In this section, we observe and analyze direct measurements from ship systems, with no pre-processing, that are expected to have an influence on the energy analysis.
+
+Figure XX represents the distribution and the evolution over the year of the ambient air and sea water temperature.
+The relatively low temperatures that are experienced by the case study ship during its operations suggest that heating demand can be expected to be higher than cooling demand.
+On the other hand, the fact that, in summer, air temperatures reach up to 26 degrees Celsius justify the existence of an HVAC unit that can also operate in cooling mode.
+
+Figure XX represents the distribution of the ship speed, both in terms of statistical distribution and in terms of a reference 24h period.
+The ship operates almost for the entire year according to a fix schedule.
+
+NOTE: It would be interesting to find a way to categorize all periods in each phase (sailing in the archipelago - sailing at open sea) and trying to analyze statistically their duration / mean power / etc.
+
 \begin{itemize}
@@
 	\item Ship speed
-	\item Other
 \end{itemize}
@@
 \subsection{Heat demand} \label{sec:res:heat}
+
+Estimated heat demand
+Maximum heat demand 
+Minimum heat demand
 

### 2017-07-05 Main.py

@@ -24,10 +24,8 @@
 # - EXERGY ANALYSIS
 
 
-do_processed_data_preparation = "yes"
-do_main_engines_analysis = "yes"
-do_aux_engines_analysis = "yes"
-do_energy_analysis = "yes"
+do_processed_data_preparation = "no"
+do_data_processing = "yes"
 
 
 
@@ -83,34 +81,49 @@ CONSTANTS = kk.constantsSetting()
 CONSTANTS["filenames"] = filenames
 (dict_structure, processed) = us.structurePreparation(CONSTANTS, dataset_raw.index, CONSTANTS["filenames"]["dataset_output_empty"], do_processed_data_preparation)
 
-# First updating the "CONSTANTS" dictionary with the some additional information
-processed = ppo.assumptions(dataset_raw, processed, CONSTANTS, header_names)
-# Updating the fields of the MainEngines and the auxiliary engines
-if do_main_engines_analysis == "yes":
+erase_old_file = "yes"
+try:
+    processed_temp = pd.read_hdf(CONSTANTS["filenames"]["dataset_output"], 'processed')
+except FileNotFoundError:
+    if do_data_processing == "no":
+        do_data_processing = "yes"
+    erase_old_file = "no"
+
+# Running the code
+if do_data_processing == "no":
+    # If the file exists AND we want to use it, we just assign the "processed" variable to it
+    processed = processed_temp
+elif do_data_processing == "yes":
+    # Otherwise, we simply do as if it didn't exist
+    if erase_old_file != "no":
+        os.remove(CONSTANTS["filenames"]["dataset_output"])
+    # First updating the "CONSTANTS" dictionary with the some additional information
+    processed = ppo.assumptions(dataset_raw, processed, CONSTANTS, header_names)
+    # Updating the fields of the MainEngines and the auxiliary engines
     processed = ppm.mainEngineProcessing(dataset_raw, processed, dict_structure, CONSTANTS, header_names)
-if do_aux_engines_analysis == "yes":
     processed = ppa.auxEngineProcessing(dataset_raw, processed, dict_structure, CONSTANTS, header_names)
-processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-1")
-# Calculating the auxiliary power demands: heating and electric power
-processed = aux.auxPowerAnalysis(processed, CONSTANTS, dict_structure)
-# Calculating the central cooling systems
[…]

### 2017-07-05 Python files/Energyanalysis.py

@@ -8,26 +8,9 @@ import os
 
 
 
-def energyAnalysisLauncher(processed, dict_structure, CONSTANTS, run_decision):
-    try:
-        # Checking first if the file exists
-        processed_temp = pd.read_hdf(CONSTANTS["filenames"]["dataset_output"], 'processed')
-        if run_decision == "no":
-            # If the file exists AND we want to use it, we just assign the "processed" variable to it
-            processed = processed_temp
-        elif run_decision == "yes":
-            # Otherwise, we simply do as if it didn't exist
-            del processed_temp
-            os.remove(CONSTANTS["filenames"]["dataset_output"])
-            processed = propertyCalculator(processed, dict_structure, CONSTANTS)
-            processed.to_hdf(CONSTANTS["filenames"]["dataset_output"], "processed", format='fixed', mode='w')
-        else:
-            print("The data_structure_preparation value should either be yes or no")
-    except FileNotFoundError:
-        processed = propertyCalculator(processed, dict_structure, CONSTANTS)
-        processed.to_hdf(CONSTANTS["filenames"]["dataset_output"], "processed", format='fixed', mode='w')
+def energyAnalysisLauncher(processed, dict_structure, CONSTANTS):
+    processed = propertyCalculator(processed, dict_structure, CONSTANTS)
     return processed
-
 
 def propertyCalculator(processed, dict_structure, CONSTANTS):
     print("Started calculating flow properties...", end="", flush=True)
@@ -42,7 +25,7 @@ def propertyCalculator(processed, dict_structure, CONSTANTS):
                             temp = processed[d2df(system,unit,flow,"T")]  # .values()
                             # If the specific enthalpy is available already, it needs to be calculated
                             if ("Air" in flow) or ("BP" in flow):
-                                print("Calculating properties for " + system + "_" + unit + "_" + flow)
+                                # print("Calculating properties for " + system + "_" + unit + "_" + flow)
                                 # dh = pd.Series(cp.PropsSI('H','T',np.array(temp),'P',np.array(press), 'Air.mix'), index=df_index) - pd.Series(cp.PropsSI('H', 'T', np.array(processed["T_0"]), 'P', CONSTANTS["General"]["P_ATM"], 'Air.mix'), index=df_index)
[…]

### 2017-07-05 Python files/auxiliaryDemand.py

@@ -12,6 +12,7 @@ def auxPowerAnalysis(processed, CONSTANTS, dict_structure):
     processed = HVAC(processed, CONSTANTS)
     processed = heatDemand(processed, CONSTANTS, dict_structure)
     processed = HTHR(processed, CONSTANTS)
+    processed = steamSystems(processed, CONSTANTS)
     return processed
 
 
@@ -97,7 +98,7 @@ def heatDemand(processed, CONSTANTS, dict_structure):
     processed["Demands:Heat:HVACreheater:Edot"] = processed["Demands:Electricity:HVAC:Edot"]
     processed["Demands:Heat:TankHeating:Edot"] = 0.5 * CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["TANK_HEATING"]
     processed.loc[:,"Demands:Heat:OtherTanks:Edot"] = 0.5 * CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["OTHER_TANKS"]
-    processed.loc[:,"Demands:Heat:HFOTankHeating:Edot"] = 0.5 * CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HFO_TANK_HEATING"]
+    processed.loc[:,"Demands:Heat:HFOtankHeating:Edot"] = 0.5 * CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HFO_TANK_HEATING"]
     processed["Demands:Heat:MachinerySpaceHeaters:Edot"] = CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["MACHINERY_SPACE_HEATERS"] * (CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["T_INSIDE"] - processed["T_air"]) / (CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["T_INSIDE"] - CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["T_AIR_REF"])
     processed.loc[processed["Demands:Heat:HVACpreheater:Edot"] < 0.1 * CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HVAC_PREHEATER"], "Demands:Heat:Machinery_space_heaters:Edot"] = 0  # The pre-heater is not working during summer
     processed["Demands:Heat:HFOheater:Edot"] = CONSTANTS["General"]["HFO"]["CP"] * (
[…]

### 2017-07-05 Python files/consistencycheck.py

@@ -2,6 +2,16 @@ import pandas as pd
 import numpy as np
 from helpers import d2df
 
+def systemCheck(processed, CONSTANTS, dict_structure):
+    enginesCheck(processed, CONSTANTS)
+    HTHRcheck(processed, CONSTANTS)
+    SteamCheck(processed, CONSTANTS, dict_structure)
+    missingValues(processed, dict_structure, CONSTANTS)
+    massBalance(processed, dict_structure, CONSTANTS)
+    energyBalance(processed, dict_structure, CONSTANTS)
+
+
+
 def enginesCheck(processed, CONSTANTS):
     # This function checks that all relative values are consistent
     text_file = open(CONSTANTS["filenames"]["consistency_check_report"], "a")
@@ -46,6 +56,65 @@ def enginesCheck(processed, CONSTANTS):
     text_file.close()
     print("...done!")
 
+
+def HTHRcheck(processed, CONSTANTS):
+    # This function checks that all relative values are consistent
+    tot = len(processed.index)
+    text_file = open(CONSTANTS["filenames"]["consistency_check_report"], "a")
+    print("Started consistency check...")
+    text_file.write("\n *** CONSISTENTCY CHECK FOR THE HTHR SYSTEM *** \n")
+    # Checking that the temperature at the end of the circuit is lower than 90 degC
+    temp = sum(processed["HTHR:HTHR13:HTWater_in:T"] <= 90 + 273.15) / tot * 100
+    warn = "WARNING " if temp < 95 else ""
+    text_file.write(warn + " HTHR-13 inlet temperatures on the HR side are consistent for " + str(temp) + " % of the datapoints \n")
+    # HTHR 13 Temperature consistency (HT side)
+    temp = sum(processed["HTHR:HTHR13:HTWater_in:T"] >= processed["HTHR:HTHR13:HTWater_out:T"]) / tot * 100
+    warn = "WARNING " if temp < 95 else ""
+    text_file.write(warn + " HTHR-13 (HT side) temperatures are consistent for " + str(temp) + " % of the datapoints \n")
+    # HTHR 13 Temperature consistency (HR side)
+    temp = sum(processed["HTHR:HTHR13:HRWater_in:T"] <= processed["HTHR:HTHR13:HRWater_out:T"]) / tot * 100
+    warn = "WARNING " if temp < 95 else ""
[…]

### 2017-07-05 Python files/coolingsystems.py

@@ -68,9 +68,11 @@ def centralCoolingSystems(processed, CONSTANTS):
     ER13set = {"AE1", "AE3", "ME1", "ME3"}
     # Calculating the temperature at the LT collector outlet
     processed[d2df("CoolingSystems", "LTcollector13", "LTWater_out", "T")] = (
-        sum(processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "T")] *
-            processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "mdot")] for idx in ER13set) /
-        sum(processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "mdot")] for idx in ER13set))
+        (sum(processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "T")] *
+            processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "mdot")] for idx in ER13set) +
+        processed["HTHR:HTHR13:HTWater_out:mdot"] * processed["HTHR:HTHR13:HTWater_out:T"]) / (
+        sum(processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "mdot")] for idx in ER13set) +
+        processed["HTHR:HTHR13:HTWater_out:mdot"]))
     # Note that the temperature at the LT distribution inlet is calculated based on the average temperature at the LT collector outlets
     processed[d2df("CoolingSystems", "LTdistribution13", "LTWater_in", "T")] = (
         sum(processed[d2df("CoolingSystems", "LTdistribution13", "LTWater_" + idx + "_out", "T")] *
@@ -83,9 +85,11 @@ def centralCoolingSystems(processed, CONSTANTS):
     ER24set = {"AE2", "AE4", "ME2", "ME4"}
     # Calculating the temperature at the LT collector outlet
     processed[d2df("CoolingSystems", "LTcollector24", "LTWater_out", "T")] = (
-        sum(processed[d2df("CoolingSystems", "LTcollector24", "LTWater_" + idx + "_in", "T")] *
-            processed[d2df("CoolingSystems", "LTcollector24", "LTWater_" + idx + "_in", "mdot")] for idx in ER24set) /
-        sum(processed[d2df("CoolingSystems", "LTcollector24", "LTWater_" + idx + "_in", "mdot")] for idx in ER24set))
[…]

### 2017-07-05 Python files/preprocessingAE.py

@@ -78,7 +78,7 @@ def auxEnginePowerCalculation(processed, CONSTANTS):
         eta_AG =  CONSTANTS["AuxEngines"]["AG"]["ETA_DES"] - CONSTANTS["AuxEngines"]["AG"]["A"] * np.exp(
             -CONSTANTS["AuxEngines"]["AG"]["k"] * (load))
         processed[d2df(system,"AG","Power_in","Edot")] = processed[d2df(system,"AG","Power_out","Edot")] / eta_AG
-        processed[d2df(system,"AG","Losses","Qdot")] = processed[d2df(system,"AG","Power_in","Edot")] - processed[d2df(system,"AG","Power_out","Edot")]
+        processed[d2df(system,"AG","Losses_out","Edot")] = processed[d2df(system,"AG","Power_in","Edot")] - processed[d2df(system,"AG","Power_out","Edot")]
         processed[d2df(system,"Cyl","Power_out","Edot")] = processed[d2df(system,"AG","Power_in","Edot")]
     print("...done!")
     return processed
@@ -153,7 +153,7 @@ def auxEngineAirFlowCalculation(raw, processed, CONSTANTS):
         processed[d2df(system, "Comp", "Power_in", "Edot")] = CONSTANTS["General"]["CP_AIR"] * processed[d2df(system, "BPsplit", "Air_in", "mdot")] * (
             processed[d2df(system, "Comp", "Air_out", "T")] - processed[d2df(system, "Comp", "Air_in", "T")])
         # Losses at the TC shaft level are calculated
-        processed[d2df(system, "TCshaft", "Losses_out", "Qdot")] = processed[d2df(system, "Turbine", "Power_out", "Edot")] - processed[d2df(system, "Comp", "Power_in", "Edot")]
+        processed[d2df(system, "TCshaft", "Losses_out", "Edot")] = processed[d2df(system, "Turbine", "Power_out", "Edot")] - processed[d2df(system, "Comp", "Power_in", "Edot")]
 
         # processed[system + ":EG_Composition"] = ppo.mixtureComposition(
         #     processed[d2df(system, "Cyl", "Air_in", "mdot")],

### 2017-07-05 Python files/preprocessingO.py

@@ -11,19 +11,19 @@ def assumptions(raw, processed, CONSTANTS, hd):
     # ALL ENGINES
     for system in {"ME1", "ME2", "ME3", "ME4", "AE1", "AE2", "AE3", "AE4"}:
         # The pressure at the turbocharger air inlet and exhaust outlet is equal to the atmospheric pressure
-        processed[d2df(system,"Comp","Air_in","p")][:] = CONSTANTS["General"]["P_ATM"]
-        processed[d2df(system, "Turbine", "Mix_out", "p")][:] = CONSTANTS["General"]["P_ATM"]
+        processed.loc[:,d2df(system,"Comp","Air_in","p")] = CONSTANTS["General"]["P_ATM"]
+        processed.loc[:,d2df(system, "Turbine", "Mix_out", "p")] = CONSTANTS["General"]["P_ATM"]
         # Temperature in the engine room, i.e. inlet to the compressor of the TC
         processed[d2df(system, "Comp", "Air_in", "T")] = raw[hd["ER_AIR_T_"]] + 273.15
         # Assuming that the pressure in the exhaust gas is 90% of the pressure in the inlet manifold. Somewhat reasonable
         processed[d2df(system, "Cyl", "EG_out", "p")] = (0.9 * raw[hd[system + "-CAC_AIR_P_OUT"]] + 1.01325) * 100000
         # Assuming the pressure of the fuel to be around 9 barg, based on measurements from ME4
-        processed[d2df(system, "Cyl", "FuelPh_in", "p")][:] = (9 + 1.01325) * 10e5
+        processed.loc[:,d2df(system, "Cyl", "FuelPh_in", "p")] = (9 + 1.01325) * 10e5
         # Assuming the temperature of the cylinder wall to be 150 degC
-        processed[d2df(system, "Cyl", "QdotJW_out", "T")][:] = 150 + 273.15
-        processed[d2df(system, "JWC", "QdotJW_in", "T")][:] = 150 + 273.15
+        processed.loc[:,d2df(system, "Cyl", "QdotJW_out", "T")] = 150 + 273.15
+        processed.loc[:,d2df(system, "JWC", "QdotJW_in", "T")] = 150 + 273.15
         # Assuming a temperature of 100 degC for heat losses from the TC shaft
-        processed[d2df(system, "TCshaft", "Losses_out", "T")][:] = 100 + 273.15
+        processed.loc[:,d2df(system, "TCshaft", "Losses_out", "T")] = 100 + 273.15
[…]

### 2017-07-05 Python files/unitstructures.py

@@ -124,6 +124,8 @@ def flowStructure():
                 structure["systems"][system]["units"].update({"AG": {}})
                 structure["systems"][system]["units"]["AG"]["flows"] = {"Power_in": {"type": "Wdot"}, "Power_out": {"type": "CEF"}, "Losses_out": {"type": "Qdot"}}
                 structure["systems"][system]["units"]["AG"]["equations"] = []
+
+
         elif system == "CoolingSystems":
             structure["systems"][system]["units"] = { "SWC13": {}, "SWC24": {}, "HTcollector13": {}, "HTcollector24": {},
                                                      "LTcollector13": {}, "LTcollector24": {}, "LTdistribution13": {}, "LTdistribution24": {}}
@@ -165,6 +167,8 @@ def flowStructure():
                 "HTWater_ME2_in": {"type": "IPF"}, "HTWater_ME4_in": {"type": "IPF"},
                 "HTWater_out": {"type": "IPF"}}
             structure["systems"][system]["units"]["HTcollector24"]["equations"] = ["MassBalance"]
+
+
         elif system == "HTHR":
             structure["systems"][system]["units"] = {"HTHR13": {}, "HTHR24": {}, "HTHRsplit": {}, "HTHRmerge": {},
                             "HVACpreheater": {}, "HVACreheater": {}, "SteamHeater": {}, "HotWaterHeater": {}}
@@ -178,43 +182,47 @@ def flowStructure():
                                                                              "HRWater_in": {"type": "IPF"}, "HRWater_out": {"type": "IPF"}}
             structure["systems"][system]["units"]["SteamHeater"]["equations"] = ["MassBalance", "ConstantPressure"]
             structure["systems"][system]["units"]["HotWaterHeater"]["flows"] = {"HRWater_in": {"type": "IPF"}, "HRWater_out": {"type": "IPF"},
-                                                                                "Qdot_out": {"type": "IPF"}}
+                                                                                "Qdot_out": {"type": "Qdot"}}
             structure["systems"][system]["units"]["HotWaterHeater"]["equations"] = ["MassBalance"]
             structure["systems"][system]["units"]["HVACpreheater"]["flows"] = {"HRWater_in": {"type": "IPF"}, "HRWater_out": {"type": "IPF"},
[…]

### 2017-07-05 Manuscript/Manuscript.tex
Other updates to the py files...I am getting bored of writing this :-)

@@
 
-Figure XX represents the distribution and the evolution over the year of the ambient air and sea water temperature.
+Figure \ref{fig:Tsea_vs_time} represents the distribution and the evolution over the year of the ambient air and sea water temperature.
 The relatively low temperatures that are experienced by the case study ship during its operations suggest that heating demand can be expected to be higher than cooling demand.
@@
 
-Figure XX represents the distribution of the ship speed, both in terms of statistical distribution and in terms of a reference 24h period.
+The corresponding ditributions are presented in Figure \ref{fig:Tsea_hist}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Tsea_vs_time}
+	\caption{Yearly evolution of sea water and ambient air temperatures. Landsort, 2014}
+	\label{fig:Tsea_vs_time}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Tsea_hist}
+	\caption{Yearly distribution of sea water and ambient air temperatures. Landsort, 2014}
+	\label{fig:Tsea_hist}
+\end{figure}
+
+Figure \ref{fig:vship_hist} represents the distribution of the ship speed, both in terms of statistical distribution and in terms of a reference 24h period.
 The ship operates almost for the entire year according to a fix schedule.
@@
 NOTE: It would be interesting to find a way to categorize all periods in each phase (sailing in the archipelago - sailing at open sea) and trying to analyze statistically their duration / mean power / etc.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/vship_hist}
+	\caption{Yearly distribution of ship speed}
+	\label{fig:vship_hist}
+\end{figure}
 
@@
 \subsection{Heat demand} \label{sec:res:heat}
+
+In Figure \ref{fig:Heat_vs_time} the heat generated by different hot utilities, as estimated by the model, is shown.
+It should be noticed that, while the heat generated by the HRSGs represents a rather accurate estimation, all other data series (auxiliary boilers, HTHR) are the consequence of rather strong assumptions.
+
+The figure also shows the operation of validation of the assumptions by comparison of the measured values for the boilers fuel consumption (here translated into an energy-per-day basis to ease the comparison).
+It should be noted that, although the absolute value is still quite different compared to the calculated one (mean error is 255\%, minimum 3\%), there is evidence that the order of magnitude is similar, and that some trends are followed
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/Heat_vs_time}
+	\caption{Yearly evolution of heat generated by different sources. "Measured" and "Calculated" relate to the auxiliary oil-fired boilers}
+	\label{fig:Heat_vs_time}
+\end{figure}
 

### 2017-07-05 Python files/auxiliaryDemand.py

@@ -131,14 +131,14 @@ def HTHR(processed, CONSTANTS):
             processed[d2df("CoolingSystems", "HTcollector13", "HTWater_" + idx + "_in", "mdot")] for idx in ER13set) /
         sum(processed[d2df("CoolingSystems", "HTcollector13", "HTWater_" + idx + "_in", "mdot")] for idx in ER13set))
     # When all engines are off, there is no flow, so the temperature calculated above is NaN. Hence, we set it to T0
-    processed.loc[ER13off,"CoolingSystems:HTcollector24:HTWater_out:T"] = processed["T_0"][ER13on]
+    processed.loc[ER13off,"CoolingSystems:HTcollector13:HTWater_out:T"] = processed["T_0"][ER13off]
 
     # Same thing for the HT collector, ER 2/4
     processed["CoolingSystems:HTcollector24:HTWater_out:T"] = (
         sum(processed[d2df("CoolingSystems", "HTcollector24", "HTWater_" + idx + "_in", "T")] *
             processed[d2df("CoolingSystems", "HTcollector24", "HTWater_" + idx + "_in", "mdot")] for idx in ER24set) /
         sum(processed[d2df("CoolingSystems", "HTcollector24", "HTWater_" + idx + "_in", "mdot")] for idx in ER24set))
-    processed.loc[ER24off, "CoolingSystems:HTcollector24:HTWater_out:T"] = processed["T_0"][ER24on]
+    processed.loc[ER24off, "CoolingSystems:HTcollector24:HTWater_out:T"] = processed["T_0"][ER24off]
 
     # Now we need to calculate the conditions after the heat is given to the hot water heater, preheater, reheater
     processed["HTHR:HTHR24:HRWater_in:T"] = processed["HTHR:SteamHeater:HRWater_out:T"] - (
@@ -150,6 +150,7 @@ def HTHR(processed, CONSTANTS):
         processed["CoolingSystems:HTcollector24:HTWater_out:T"] - processed[d2df("HTHR", "HTHR24", "HRWater_in", "T")])
     # Once again, to avoid risks, when all engines of the ER 24 are off, the heat flow here is 0
     qdot_HTHR24[ER24off] = 0
[…]

### 2017-07-05 Python files/consistencycheck.py

@@ -2,13 +2,18 @@ import pandas as pd
 import numpy as np
 from helpers import d2df
 
-def systemCheck(processed, CONSTANTS, dict_structure):
+def systemCheck(processed, CONSTANTS, dict_structure, dataset_raw):
     enginesCheck(processed, CONSTANTS)
     HTHRcheck(processed, CONSTANTS)
     SteamCheck(processed, CONSTANTS, dict_structure)
     missingValues(processed, dict_structure, CONSTANTS)
     massBalance(processed, dict_structure, CONSTANTS)
     energyBalance(processed, dict_structure, CONSTANTS)
+
+    # Checking the boilers
+    boilers_measured = (dataset_raw["Boiler_Port"].resample("D").mean() + dataset_raw["Boiler_starbord"].resample("D").mean()) * CONSTANTS["General"]["HFO"]["LHV"]
+    boilers_calculated = processed["Steam:Boiler1:FuelPh_in:mdot"].resample("D").sum() * 60 * 15 * CONSTANTS["General"]["HFO"]["LHV"]
+    ((boilers_calculated - boilers_measured) / boilers_measured * 100).describe()
 
 
 

### 2017-07-05 Python files/coolingsystems.py

@@ -55,10 +55,10 @@ def engineCoolingSystemsCalculation(processed, CONSTANTS, engine_type):
             processed[d2df(system,"HRSG","Steam_in","mdot")] = ((processed[d2df(system,"HRSG","Mix_in","T")] - processed[d2df(system,"HRSG","Mix_out","T")]) *
                 processed[d2df(system, "HRSG", "Mix_in", "mdot")] * processed[system+":CP_MIX"] /
                 (CONSTANTS["Steam"]["H_STEAM_VS"] - CONSTANTS["Steam"]["H_STEAM_LS"]))
-        # Calculation of the mass recirculating from the LT outlet to the HT inlet
-        processed[d2df(system,"HTmerge","LTWater_in","mdot")] = processed[d2df(system,"JWC","HTWater_out","mdot")] * (
+        # Calculation of the mass recirculating from the LT outlet to the HT inlet (not exactly...)
+        processed[d2df(system, "HTmerge", "LTWater_in", "mdot")] = processed[d2df(system,"JWC","HTWater_out","mdot")] * (
             processed[d2df(system, "CAC_HT", "HTWater_out", "T")] - processed[d2df(system,"JWC","HTWater_in","T")]) / (
-            processed[d2df(system, "CAC_HT", "HTWater_out", "T")] - processed[d2df(system,"LOC","LTWater_out","T")])
+            processed[d2df(system, "CAC_HT", "HTWater_out", "T")] - CONSTANTS[engine_type]["T_COOLING_MIX"])
     print("...done!")
     return processed
 
@@ -112,8 +112,10 @@ def coolingFlows(processed, CONSTANTS, engine_type):
         heat_LT = processed[d2df(system, "CAC_LT", "LTWater_in", "p")][~processed[system+":on"]].mean()
         processed[d2df(system, "CAC_LT", "LTWater_in", "mdot")] = pumpFlow(processed[d2df(system, "Cyl", "Power_out", "omega")],
                        processed[d2df(system, "CAC_LT", "LTWater_in", "p")], heat_LT, CONSTANTS, engine_type)
+        processed.loc[:,d2df(system, "CAC_LT", "LTWater_in", "mdot")] = processed[d2df(system, "CAC_LT", "LTWater_in", "mdot")] / max(processed[d2df(system, "CAC_LT", "LTWater_in", "mdot")]) * CONSTANTS[engine_type]["MFR_LT"]
[…]

### 2017-07-05 Python files/preprocessingO.py

@@ -35,10 +35,11 @@ def assumptions(raw, processed, CONSTANTS, hd):
             processed.loc[:,d2df(system, "AG", "Losses_out", "T")] = 100 + 273.15
             processed.loc[:,d2df(system, "Cyl", "Power_out", "omega")] = 750
     # Others
-    processed["T_0"] = raw[hd["SEA_SW_T_"]] + 273.15
-    processed["T_air"] = raw[hd["OUTSIDE_AIR_T_"]] + 273.15
+    processed.loc[:,"T_0"] = raw[hd["water_T_forsmark_smhi-opendata"]] + 273.15
+    processed.loc[:,"T_air"] = raw[hd["air_T_sv_hogarna_smhi-opendata"]] + 273.15
     processed[d2df("CoolingSystems","SWC13","SeaWater_out","T")] = raw[hd["SWC13_SW_T_OUT"]]  # CHECK IF IT IS IN OR OUT
     processed[d2df("CoolingSystems","SWC24","SeaWater_out","T")] = raw[hd["SWC24_SW_T_OUT"]]  # CHECK IF IT IS IN OR OUT
+    processed.loc[:,"CoolingSystems:LTHTmixer:HTWater_out:T"] = CONSTANTS["MainEngines"]["T_COOLING_MIX"]
     # HTHR system
     processed.loc[:,"HTHR:SteamHeater:HRWater_out:T"] = 90 + 273.15 # From the heat balance, the temperature needs to rise at 90 degrees
     processed.loc[:,"HTHR:SteamHeater:HRWater_out:mdot"] = 298 / 3600 * CONSTANTS["General"]["RHO_W"] # the original value is in m3/h

### 2017-07-05 Python files/unitstructures.py

@@ -63,7 +63,7 @@ def flowStructure():
         if system[1] == "E":  # This basically means that this operation is only done if the system is an engine
             structure["systems"][system]["units"] = {"Comp": {}, "Turbine": {}, "BPsplit": {}, "BPmerge": {}, "BPvalve": {},
                                                      "CAC_HT": {}, "CAC_LT": {}, "LOC": {}, "JWC": {}, "Cyl": {},
-                                                     "HTsplit": {}, "HTmerge": {}, "LTsplit": {}, "TCshaft": {}}
+                                                     "HTsplit": {}, "HTmerge": {}, "TCshaft": {}}
             # Compressor
             structure["systems"][system]["units"]["Comp"]["flows"] = {"Air_in": {"type": "CPF"}, "Air_out": {"type": "CPF"}, "Power_in": {"type": "Wdot"}} # TC compressor
             structure["systems"][system]["units"]["Comp"]["equations"] = ["MassBalance"]
@@ -109,9 +109,6 @@ def flowStructure():
             # HT cooling systems merge
             structure["systems"][system]["units"]["HTmerge"]["flows"] = {"HTWater_in": {"type": "IPF"}, "HTWater_out": {"type": "IPF"}, "LTWater_in": {"type": "IPF"}}
             structure["systems"][system]["units"]["HTmerge"]["equations"] = ["MassBalance"]
-            # LT cooling systems split
-            structure["systems"][system]["units"]["LTsplit"]["flows"] = {"LTWater_in": {"type": "IPF"}, "HTWater_out": {"type": "IPF"}, "LTWater_out": {"type": "IPF"}}  # Bypass valve
-            structure["systems"][system]["units"]["LTsplit"]["equations"] = ["MassBalance", "ConstantTemperature"]
             # Only auxiliary engines AND main engines 2/3 have the exhaust gas boiler
             if system[0] == "A" or system[2] == "2" or system[2] == "3":
                 # Heat recovery steam generator
@@ -128,24 +125,25 @@ def flowStructure():
 
         elif system == "CoolingSystems":
             structure["systems"][system]["units"] = { "SWC13": {}, "SWC24": {}, "HTcollector13": {}, "HTcollector24": {},
-                                                     "LTcollector13": {}, "LTcollector24": {}, "LTdistribution13": {}, "LTdistribution24": {}}
[…]

### 2017-07-11 Manuscript/Manuscript.tex
Modifications to the .py files - STABLE VERSION
Almost everything is included. Most bugs have been fixed. The whole thing kind of works!

@@
 
-The corresponding ditributions are presented in Figure \ref{fig:Tsea_hist}
+The corresponding distributions are presented in Figure \ref{fig:Tsea_hist}
 
@@
 Figure \ref{fig:vship_hist} represents the distribution of the ship speed, both in terms of statistical distribution and in terms of a reference 24h period.
-The ship operates almost for the entire year according to a fix schedule.
+The ship operates almost for the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
+This can also be seen from Figure \ref{fig:Pme_vs_vship}, where the main cluster of operations at middle-low speed can be clearly identified, as well as the smaller area of operations at high speed.
 
@@
 
-\begin{itemize}
-	\item External temperature
-	\item Sea water temperature
-	\item Ship speed
-\end{itemize}
-
-The results are presented in two main forms:
-\begin{itemize}
-	\item Time series (comparing some relevant trips with a daily time scale)
-	\item Distribution plot (or box plot...box plots are easier to compare, distribution plots are easier to interpret)
-\end{itemize}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pme_vs_vship}
+	\caption{Scatter plot, Propulsion power versus ship speed}
+	\label{fig:Pme_vs_vship}
+\end{figure}
 
@@
 
+Here we can see the typical example (note: the day was not selected as "typical", but simply randomly) of the energy demand during a winter and a summer day (Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}).
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_W}
+	\caption{Energy demand during a day of operations, Winter}
+	\label{fig:Demand_vs_time_W}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_S}
+	\caption{Energy demand during a day of operations, Summer}
+	\label{fig:Demand_vs_time_S}
+\end{figure}
+
 \subsection{Heat demand} \label{sec:res:heat}
@@
 The figure also shows the operation of validation of the assumptions by comparison of the measured values for the boilers fuel consumption (here translated into an energy-per-day basis to ease the comparison).
-It should be noted that, although the absolute value is still quite different compared to the calculated one (mean error is 255\%, minimum 3\%), there is evidence that the order of magnitude is similar, and that some trends are followed
+It should be noted that, although the absolute value is still quite different compared to the calculated one (mean error is 255\%, minimum 3\%), there is evidence that the order of magnitude is similar, and that some trends are followed.
 

### 2017-07-11 Main.py

@@ -126,10 +126,11 @@ cc.systemCheck(processed, CONSTANTS, dict_structure, dataset_raw)
 
 #%%
 ######################################
-## EXPLORATORY DATA ANALYSIS	##
+## PLOTTING	##
 ######################################
 
-# Responsible: FB
+plot.predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure, ["TimeSeries:Heat_vs_time", "Pie:TotalEnergy", "Pie:Heat"])
+# plot.plotMain("prompt", dict_structure, processed)
 
 
 

### 2017-07-11 Python files/Constants.py

@@ -165,6 +165,9 @@ def otherUnits ():
     output["BOILER"]["ETA_REGR_Y"] = [0.8787, 0.8830, 0.8864, 0.8889, 0.8910, 0.8897, 0.8870, 0.8842, 0.8810, 0.8777, 0.8740, 0.8692, 0.86486, 0.8613, 0.8570, 0.8528, 0.8491, 0.8462, 0.8427, 0.8390, 0.8356, 0.8317]
     for idx in range(len(output["BOILER"]["ETA_REGR_Y"])):
         output["BOILER"]["ETA_REGR_Y"][idx] = output["BOILER"]["ETA_REGR_Y"][idx] / max(output["BOILER"]["ETA_REGR_Y"])
+    output["PROPULSION"] = {}
+    output["PROPULSION"]["ETA_GB"] = 0.98
+    output["PROPULSION"]["ETA_SH"] = 0.99
     output["HEAT_DEMAND"] = {}
     output["HEAT_DEMAND"]["HOT_WATER_HEATER"] = 1200
     output["HEAT_DEMAND"]["HVAC_PREHEATER"] = 3500

### 2017-07-11 Python files/Energyanalysis.py

@@ -12,6 +12,13 @@ def energyAnalysisLauncher(processed, dict_structure, CONSTANTS):
     processed = propertyCalculator(processed, dict_structure, CONSTANTS)
     return processed
 
+
+def summaryEnergyFlows(processed, CONSTANTS):
+
+    return processed
+
+
+
 def propertyCalculator(processed, dict_structure, CONSTANTS):
     print("Started calculating flow properties...", end="", flush=True)
     df_index = processed.index

### 2017-07-11 Python files/auxiliaryDemand.py

@@ -6,6 +6,7 @@ from helpers import d2df
 def auxPowerAnalysis(processed, CONSTANTS, dict_structure):
     # Calculating the total auxiliary power demand
     processed = auxPowerTotalDemand(processed)
+    processed = propPowerDemand(processed, CONSTANTS)
     # Calculating the contribution from the thrusters
     #processed = thrusters(processed, CONSTANTS)  # The second method works better. But I still keep it
     processed = thrusters2(processed, CONSTANTS)
@@ -22,6 +23,32 @@ def auxPowerTotalDemand(processed):
     # We also need to store (for further use) a series made of the daily averages
     processed["Demands:Electricity:Total:Edot1D"] = processed["Demands:Electricity:Total:Edot"].resample('1D').mean().reindex(processed["Demands:Electricity:Total:Edot"].index, method='ffill')
     processed["Demands:Electricity:Other:Edot"] = processed["Demands:Electricity:Total:Edot"]
+    return processed
+
+
+def propPowerDemand(processed, CONSTANTS):
+    # Looking at the first propulsion line, for engines 1 and 2
+    processed["Propulsion:Gearbox1:Power_in1:Edot"] = processed["ME1:Cyl:Power_out:Edot"]
+    processed["Propulsion:Gearbox1:Power_in2:Edot"] = processed["ME2:Cyl:Power_out:Edot"]
+    processed["Propulsion:Gearbox1:Power_out:Edot"] = (processed["Propulsion:Gearbox1:Power_in1:Edot"] + processed["Propulsion:Gearbox1:Power_in2:Edot"]) * CONSTANTS["OtherUnits"]["PROPULSION"]["ETA_GB"]
+    processed["Propulsion:Gearbox1:Losses:Edot"] = (processed["Propulsion:Gearbox1:Power_in1:Edot"] + processed["Propulsion:Gearbox1:Power_in2:Edot"]) - processed["Propulsion:Gearbox1:Power_out:Edot"]
+    processed["Propulsion:Shaft1:Power_in:Edot"] = processed["Propulsion:Gearbox1:Power_out:Edot"]
[…]

### 2017-07-11 Python files/coolingsystems.py

@@ -73,6 +73,8 @@ def centralCoolingSystems(processed, CONSTANTS):
         processed["HTHR:HTHR13:HTWater_out:mdot"] * processed["HTHR:HTHR13:HTWater_out:T"]) / (
         sum(processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "mdot")] for idx in ER13set) +
         processed["HTHR:HTHR13:HTWater_out:mdot"]))
+    # The temperature at the other output of the collector is the same
+    processed["CoolingSystems:LTcollector13:HTWater_out:T"] = processed["CoolingSystems:LTcollector13:LTWater_out:T"]
     # Note that the temperature at the LT distribution inlet is calculated based on the average temperature at the LT collector outlets
     processed[d2df("CoolingSystems", "LTdistribution13", "LTWater_in", "T")] = (
         sum(processed[d2df("CoolingSystems", "LTdistribution13", "LTWater_" + idx + "_out", "T")] *
@@ -90,14 +92,34 @@ def centralCoolingSystems(processed, CONSTANTS):
          processed["HTHR:HTHR24:HTWater_out:mdot"] * processed["HTHR:HTHR24:HTWater_out:T"]) / (
             sum(processed[d2df("CoolingSystems", "LTcollector24", "LTWater_" + idx + "_in", "mdot")] for idx in ER24set) +
             processed["HTHR:HTHR24:HTWater_out:mdot"]))
+    processed["CoolingSystems:LTcollector24:HTWater_out:T"] = processed["CoolingSystems:LTcollector24:LTWater_out:T"]
     # Note that the temperature at the LT distribution inlet is calculated based on the average temperature at the LT collector outlets
     processed[d2df("CoolingSystems", "LTdistribution24", "LTWater_in", "T")] = (
         sum(processed[d2df("CoolingSystems", "LTdistribution24", "LTWater_" + idx + "_out", "T")] *
             processed[d2df("CoolingSystems", "LTdistribution24", "LTWater_" + idx + "_out", "mdot")] for idx in ER24set) /
         sum(processed[d2df("CoolingSystems", "LTdistribution24", "LTWater_" + idx + "_out", "mdot")] for idx in ER24set))
[…]

### 2017-07-11 Python files/heatdemand.py

@@ -1,35 +1,0 @@
-import numpy as np
-import pandas as pd
-
-
-def heatDemand(processed, CONSTANTS):
-    # This function calculates the heat demand from the on board systems
-    db_index = processed.index
-    processed[""] = CONSTANTS["Other"]["HEAT_DEMAND"]["HOT_WATER_HEATER"] * pd.Series(len(processed)/96*np.repeat(CONSTANTS["Other"]["HEAT_DEMAND"]["HWH_HOURLY"],4), index=db_index)
-    qdot_hvac_preheater = CONSTANTS["Other"]["HEAT_DEMAND"]["HVAC_PREHEATER"] * (CONSTANTS["Other"]["HEAT_DEMAND"]["T_INSIDE"] - processed["T_air"]) / (CONSTANTS["Other"]["HEAT_DEMAND"]["T_INSIDE"] - CONSTANTS["Other"]["HEAT_DEMAND"]["T_AIR_REF"])
-    qdot_hvac_preheater[qdot_hvac_preheater < 0.1 * CONSTANTS["Other"]["HEAT_DEMAND"]["HVAC_PREHEATER"]] = 0 # The pre-heater is not working during summer
-    # qdot_hvac_reheater = CONSTANTS["Other"]["HEAT_DEMAND"]["HVAC_PREHEATER"] * (CONSTANTS["Other"]["HEAT_DEMAND"]["T_INSIDE"] - processed["T_air"]) / (CONSTANTS["Other"]["HEAT_DEMAND"]["T_INSIDE"] - CONSTANTS["Other"]["HEAT_DEMAND"]["T_AIR_REF"])
-
-
-
-
-    return processed
-
-
-
-
-
-
-
-
-
-
-
-
-def heatDemandProducerSide(processed, CONSTANTS):
-
-
-
-
-
-    return processed

### 2017-07-11 Python files/plotting.py

@@ -1,31 +1,30 @@
-#import numpy as np
-#import matplotlib.mlab as mlab
 import matplotlib.pyplot as plt
+from helpers import d2df
 
-def plotMain(type, structure, *args):
+def plotMain(type, dict_structure, processed, *args):
     # This function plots the contents of the analysed files.
     # Two input values for "type" are accepted:
     # - "off" [default]: does not print anything
     # - "prompt": prompts the user for what s/he wants to print
     # - "csv": reads info on what is to be read from a csv file
     if type == "prompt":
-        plot_info = plotPrompt(structure)
+        plot_info = plotPrompt(dict_structure)
         data_type = "processed"
     elif type == "prompt_raw":
         hd = args[0]
-        plot_info = plotPromptRaw(structure,hd)
+        plot_info = plotPromptRaw(dict_structure,hd)
         data_type = "raw"
     elif type == "csv":
         filename = args[0]
-        plot_info = plotFromCSV(filename)
+        plot_info = plotFromCSV(dict_structure)
         data_type = "processed"
     else:
         print("Either the -type- input was not correct, or you don't wish to print anything. No printout")
         return
-    plottingFunction(plot_info, structure, data_type)
+    plottingFunction(plot_info, processed, data_type)
 
 
-def plotPrompt(structure):
+def plotPrompt(dict_structure):
     # This function prompts the user for the required inputs to plot.
     # Two alternatives are available:
     # - The default is "automatic". It is the one that is set by default. The input needs to be provided using a
@@ -61,28 +60,28 @@ def plotPrompt(structure):
             line_info = {}
             while True:
                 plot_system = input("Now input what system you like to plot for. Available are choices are ""ME#"" (#=1,2,3,4), ""AE#"" (#=1,2,3,4), and ""Other"" ")
-                if plot_system in structure.keys():
+                if plot_system in dict_structure["systems"]:
                     line_info["system"] = plot_system
                     break
                 else:
                     print("The input you gave (" + plot_system + ") for the system name is not correct! Maybe there was a typo? Try again!")
             while True:
-                plot_component = input("Now input what component of the chosen system you like to plot for:")
-                if plot_component in structure[line_info["system"]].keys():
-                    line_info["component"] = plot_component
[…]

### 2017-07-11 Python files/preprocessingO.py

@@ -24,6 +24,8 @@ def assumptions(raw, processed, CONSTANTS, hd):
         processed.loc[:,d2df(system, "JWC", "QdotJW_in", "T")] = 150 + 273.15
         # Assuming a temperature of 100 degC for heat losses from the TC shaft
         processed.loc[:,d2df(system, "TCshaft", "Losses_out", "T")] = 100 + 273.15
+        # We assume the mixgin temperature at the LT inlet of the HT mixer to be constant
+        processed.loc[:,d2df(system, "HTmerge", "LTWater_in", "T")] = CONSTANTS["MainEngines"]["T_COOLING_MIX"]
         # Assuming the steam pressure and temperature in the HRSG to be constant...
         hrsg_pressure_assumption = (6 + 1.01325) * 100000
         #if system in {"AE1", "AE2", "AE3", "AE4", "ME2", "ME3"}:
@@ -39,7 +41,6 @@ def assumptions(raw, processed, CONSTANTS, hd):
     processed.loc[:,"T_air"] = raw[hd["air_T_sv_hogarna_smhi-opendata"]] + 273.15
     processed[d2df("CoolingSystems","SWC13","SeaWater_out","T")] = raw[hd["SWC13_SW_T_OUT"]]  # CHECK IF IT IS IN OR OUT
     processed[d2df("CoolingSystems","SWC24","SeaWater_out","T")] = raw[hd["SWC24_SW_T_OUT"]]  # CHECK IF IT IS IN OR OUT
-    processed.loc[:,"CoolingSystems:LTHTmixer:HTWater_out:T"] = CONSTANTS["MainEngines"]["T_COOLING_MIX"]
     # HTHR system
     processed.loc[:,"HTHR:SteamHeater:HRWater_out:T"] = 90 + 273.15 # From the heat balance, the temperature needs to rise at 90 degrees
     processed.loc[:,"HTHR:SteamHeater:HRWater_out:mdot"] = 298 / 3600 * CONSTANTS["General"]["RHO_W"] # the original value is in m3/h
@@ -133,7 +134,7 @@ def connectionAssignment(processed, dict_structure, CONSTANTS, system, unit, cal
     text_file = open(CONSTANTS["filenames"]["consistency_check_report"], "a")
     counter = 0
     for flow in dict_structure["systems"][system]["units"][unit]["flows"]:
-        if system+":"+unit+":"+flow == "CoolingSystems:HTcollector13:HTWater_out":
[…]

### 2017-07-11 Python files/unitstructures.py

@@ -138,12 +138,12 @@ def flowStructure():
             structure["systems"][system]["units"]["LTcollector13"]["flows"] = {
                 "LTWater_AE1_in": {"type": "IPF"}, "LTWater_AE3_in": {"type": "IPF"},
                 "LTWater_ME1_in": {"type": "IPF"}, "LTWater_ME3_in": {"type": "IPF"},
-                "LTWater_out": {"type": "IPF"}, "HTWater_in": {"type": "IPF"}, "HTWater_out": {"type": "IPF"}}
+                "LTWater_out": {"type": "IPF"}, "HTWater_out": {"type": "IPF"}}
             structure["systems"][system]["units"]["LTcollector13"]["equations"] = ["MassBalance"]
             structure["systems"][system]["units"]["LTcollector24"]["flows"] = {
                 "LTWater_AE2_in": {"type": "IPF"}, "LTWater_AE4_in": {"type": "IPF"},
                 "LTWater_ME2_in": {"type": "IPF"}, "LTWater_ME4_in": {"type": "IPF"},
-                "LTWater_out": {"type": "IPF"}, "HTWater_in": {"type": "IPF"}, "HTWater_out": {"type": "IPF"}}
+                "LTWater_out": {"type": "IPF"}, "HTWater_out": {"type": "IPF"}}
             structure["systems"][system]["units"]["LTcollector24"]["equations"] = ["MassBalance"]
             structure["systems"][system]["units"]["LTdistribution13"]["flows"] = {
                 "LTWater_in": {"type": "IPF"},
@@ -180,16 +180,16 @@ def flowStructure():
             structure["systems"][system]["units"]["HTsplit13"]["equations"] = ["MassBalance", "ConstantTemperature"]
             structure["systems"][system]["units"]["HTsplit24"]["flows"] = {
                 "HTWater_in": {"type": "IPF"}, "HTWater_out": {"type": "IPF"}, "LTWater_out": {"type": "IPF"}}
-            structure["systems"][system]["units"]["HTsplit24"]["equations"] = ["MassBalance"]
+            structure["systems"][system]["units"]["HTsplit24"]["equations"] = ["MassBalance", "ConstantTemperature"]
 
         elif system == "HTHR":
             structure["systems"][system]["units"] = {"HTHR13": {}, "HTHR24": {}, "HTHRsplit": {}, "HTHRmerge": {},
[…]

### 2017-07-19 Manuscript/Manuscript.tex
pre-processing manuscript

@@
 
-This section should be written by FA :-)
+The operational data was collected on board from the ships' machine logging and surveillance system.
+The on board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB.
+The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average.
+The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python [**INSERT REF TO PANDAS**].
+A new structured naming of headers and a consistent time frequency over all data points were created in a HDF5 table database which is the base for the analysis.
 

### 2017-07-19 Manuscript/Manuscript.tex
some updates.

@@
 \author[EPFL]{Francesco Baldi}
-\author[Linneus]{Fredrik Ahlgren}
+\author[Linnaeus]{Fredrik Ahlgren}
 \author[DTU]{Tuong-Van Nguyen}
@@
 \address[EPFL]{Industrial Process Energy Systems Engineering (IPESE), \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, 1950, Sion, Switzerland}
-\address[Linneus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
+\address[Linnaeus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
 \address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
@@
 A new structured naming of headers and a consistent time frequency over all data points were created in a HDF5 table database which is the base for the analysis.
+
+All data-points were checked individually by creating a descriptive statistic and a histogram of each data-point.
+A filter was applied by setting up a pre-defined maximum and minimum value.
+As we do not know, or had any meaningful way of deriving, each individual sensors placement, sensor type or calibration status all data must be filtered and checked accordingly for outliers or bad data.
 

### 2017-07-19 Manuscript/Manuscript.tex
smhi-data manuscript

@@
 
-Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse.
+Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska Högarna for the ambient temperature.
+The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska Högarna lighthouse is along the ship route, in between the Swedish archipelago and Åland.
 The assumption was validated based on  June-December period, for which onboard measurements were available.
 This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
-% We did the calculation, resampling the data from SHMI from a 1(3)h to 15min frequency using linear interpolation.
+The fit the SMHI-data with the rest of the data-set the SMHI data was resampled from 1h for the seawater and 3h for the ambient temperature, to 15min frequency using a linear interpolation.
 

### 2017-07-19 Manuscript/Manuscript.tex
updated bibtex

@@
 The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average.
-The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python [**INSERT REF TO PANDAS**].
-A new structured naming of headers and a consistent time frequency over all data points were created in a HDF5 table database which is the base for the analysis.
+The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python \cite{mckinney-proc-scipy-2010}.
+A new structured naming of headers and a consistent time frequency over all data points were created, and 245 selected data points were saved in a HDF5 table time series database which is the base for the analysis.
 

### 2017-07-19 Manuscript/Manuscript.tex
Changes to manuscript

--- 2017-07-19 Manuscript/Manuscript.tex
@@
 \author[EPFL]{Francesco Baldi}
-\author[Linnaeus]{Fredrik Ahlgren}
+\author[Linneus]{Fredrik Ahlgren}
 \author[DTU]{Tuong-Van Nguyen}
@@
 \address[EPFL]{Industrial Process Energy Systems Engineering (IPESE), \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, 1950, Sion, Switzerland}
-\address[Linnaeus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
+\address[Linneus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
 \address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
@@
 \begin{align}
-	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} = \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \\
+	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \\
 	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d} 
@@
 
-The operational data was collected on board from the ships' machine logging and surveillance system.
-The on board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB.
-The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average.
-The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python [**INSERT REF TO PANDAS**].
-A new structured naming of headers and a consistent time frequency over all data points were created in a HDF5 table database which is the base for the analysis.
-
-All data-points were checked individually by creating a descriptive statistic and a histogram of each data-point.
-A filter was applied by setting up a pre-defined maximum and minimum value.
-As we do not know, or had any meaningful way of deriving, each individual sensors placement, sensor type or calibration status all data must be filtered and checked accordingly for outliers or bad data.
+This section should be written by FA :-)
 
@@
 
-The results are presented in three main forms:
-\begin{itemize}
-	\item Time series (comparing some relevant trips with a daily time scale)
-	\item Distribution plot (or box plot...box plots are easier to compare, distribution plots are easier to interpret)
-	\item Sankey diagrams
-\end{itemize}
-
-The results should be presented both for the full database and for each operational mode (port-sailing-drifting)
-
 Here we can see the typical example (note: the day was not selected as "typical", but simply randomly) of the energy demand during a winter and a summer day (Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}).
+It can be seen clearly how, as expected, the heating demand in winter is much higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduce need for compartment heating.
+On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors.
 
@@
 	\label{fig:Demand_vs_time_S}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pie_TotalEnergy}
+	\caption{Total energy demand over one year of opeartions}
+	\label{fig:Pie_TotalEnergy}
+\end{figure}
+
+The average HVAC daily demand, together with the average total electric power demand, is represented in Figure \ref{fig:PelHVAC_vs_time}.
+As it can be seen, the HVAC demand represents only a minor contribution to the total electric energy demand, as it is only present for a few months during summer, and at rather low power demand.
+This is not surprising, since it can be observed that (see Figure \ref{fig:Tair_cum}) the air is always below 27 degC, and 90\% of the time below 17 degC.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/PelHVAC_vs_time}
+	\caption{HVAC power demand versus total demand. Daily averages for one year of operations}
+	\label{fig:PelHVAC_vs_time}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Tair_cum}
+	\caption{Cumulative distribution of measured ambient air temperatures}
+	\label{fig:Tair_cum}
 \end{figure}
@@
 
-Estimated heat demand
-Maximum heat demand 
-Minimum heat demand
+Looking at Figure \ref{fig:HeatGen_hist} it is possible to appreciate the amount of heat that is recovered from on board available waste heat (cooling water and exhaust gas).
+It can be observed that the amount of heat recovered from the HT cooling systems is much lower than expected, with an average of around 350 kW.
+This value is quite unreasonable compared to the values found in the technical specifications from the yard, that go from 700 kW (when only one auxiliary engine is operted) to 4800 kW (when all main engines and two auxiliary engines are operated).
+\\
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/HeatGen_hist}
+	\caption{Distribution of thermal power generated from on board waste heat recovery}
+	\label{fig:HeatGen_hist}
+\end{figure}
+
+Looking at the assumptions, this is mostly related to the assumption that the flow of heat recovery water is fixed to 297 m3/h.
+When the heat demand is medium or low, since the temperature BEFORE the consumers is fixed to 90 degrees, this limits the temperature drop on the consumers, making the maximum heat exchangeable in the HTHR heat exchangers to be very low.
+This can be seen in Figure \ref{fig:HTHR_T_hist}, where the histograms of the inlet temperature of the HT water and the HR water in the HTHR24 heat exchanger are plotted together.
+Since the maximum amount of heat that can be exchanged depends directly on this difference, the explanation is clear.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/HTHR_T_hist}
+	\caption{Distribution of temperatures at the HTHR (24) inlet}
+	\label{fig:HTHR_T_hist}
+\end{figure}
+
+The alternative solution would be to assume a constant temperature at the HTHR24 HR water inlet instead.
+This would, on the one hand, dramatically increase the contribution of the HT water to the heat demand of the ship.
+On the other hand, there is no information of what this temperature is, and little certainty that this is how the system is regulated.
+
+Figure \ref{fig:Heat_comparison} shows the heat demand calculated based on different values assumed for the HTHR inlet temperature and of the reference outside temperature for the design demand of the HVAC preheater.
+The values of the average error (calculated both as absolute value and not) are given in Table \ref{tab:Heat_comparison}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/Heat_comparison}
+	\caption{Comparison of the calculated steam demand for different values of the fixed HTHR inlet temperature and of the reference temperature for the design HVAC preheater load.
+The shaded green area represents the time frame when no heating occurs}
+	\label{fig:Heat_comparison}
+\end{figure}
+
+\begin{table}
+	\centering
+	\begin{tabular}{cccc}
+		\hline 
+		\rule[-1ex]{0pt}{2.5ex} $T_{HTHR,in} $ & $T_{OUT,des} $ & $\epsilon$ & $\epsilon_{abs}$ \\ 
+		\hline 
+		\rule[-1ex]{0pt}{2.5ex} 70 & -20 & 100\% & 80\% \\ 
+		\hline 
+		\rule[-1ex]{0pt}{2.5ex} 65 & -20 & 76\% & 67\% \\ 
+		\hline 
+		\rule[-1ex]{0pt}{2.5ex} 60 & -20 & 57\% & 57\% \\ 
+		\hline 
+		\rule[-1ex]{0pt}{2.5ex} 55 & -20 & 42\% & 49\% \\ 
+		\hline 
+		\rule[-1ex]{0pt}{2.5ex} 50 & -20 & 28\% & 42\% \\ 
+		\hline 
+		\rule[-1ex]{0pt}{2.5ex} 70 & -30 & 70\% & 69\% \\ 
+		\hline 
+		\rule[-1ex]{0pt}{2.5ex} 50 & -30 & 8\% & 35\% \\ 
+		\hline 
+		\rule[-1ex]{0pt}{2.5ex} $\dot{m}=K$ = k & -30 & 207\% & 147\% \\ 
+		\hline 
+	\end{tabular} 
+	\label{tab:Heat_comparison}
+\end{table}
 

### 2017-07-19 Main.py

@@ -50,6 +50,7 @@ import plotting as plot
 import datareading as dr
 import unitstructures as us
 import constants as kk
+import fillerfunctions as ff
 import consistencycheck as cc
 import preprocessingAE as ppa
 import preprocessingME as ppm
@@ -57,6 +58,7 @@ import preprocessingO as ppo
 import energyanalysis as ea
 import auxiliaryDemand as aux
 import coolingsystems as cs
+from helpers import d2df
 
 #%%
 ######################################
@@ -98,23 +100,23 @@ elif do_data_processing == "yes":
     if erase_old_file != "no":
         os.remove(CONSTANTS["filenames"]["dataset_output"])
     # First updating the "CONSTANTS" dictionary with the some additional information
-    processed = ppo.assumptions(dataset_raw, processed, CONSTANTS, header_names)
+    processed = input.assumptions(dataset_raw, processed, CONSTANTS, header_names)
     # Updating the fields of the MainEngines and the auxiliary engines
     processed = ppm.mainEngineProcessing(dataset_raw, processed, dict_structure, CONSTANTS, header_names)
     processed = ppa.auxEngineProcessing(dataset_raw, processed, dict_structure, CONSTANTS, header_names)
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-1")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-1")
     # Calculating the auxiliary power demands: heating and electric power
     processed = aux.auxPowerAnalysis(processed, CONSTANTS, dict_structure)
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-2")
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Demands", "Demands-1")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-2")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Demands", "Demands-1")
     # Calculating the central cooling systems
     processed = cs.centralCoolingSystems(processed, CONSTANTS)
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-3")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-3")
[…]

### 2017-07-19 Python files/Constants.py

@@ -8,7 +8,7 @@ def constantsSetting():
     output["Steam"] = steamProperties()  # loading dictionary with steam properties constants
     output["MainEngines"] = mainEngines(output)  # loading dictionary with main-engine related constants
     output["AuxEngines"] = auxiliaryEngines(output)  # loading dictionary with auxiliary-engine related constants
-    output["OtherUnits"] = otherUnits()
+    output["OtherUnits"] = otherUnits(output)
     return output
 
 
@@ -18,8 +18,8 @@ def general():
     output["R_0"] = 8.3140 # Ideal gas constant, in [J/mol*K]
     output["R_AIR"] = output["R_0"] * 1000 / 29 # Air gas constant, in [J/kg*K]
     output["K_AIR"] = 1.4   # Air specific heat ratio
-    output["CP_AIR"] = 1.01   # Air specific heat, in [kJ/kgK]
-    output["CP_EG"] = 1.10   # EG specific heat, in [kJ/kgK]
+    output["CP_AIR"] = 1.015   # Air specific heat, in [kJ/kgK]
+    output["CP_EG"] = 1.11   # EG specific heat, in [kJ/kgK]
     output["CP_LO"] = 2.1   # Lubricating oil specific heat, in [kW/kgK]
     output["CP_WATER"] = 4.187   # Water specifi heat, in [kW/kgK]
     output["RHO_W"] = 1000.0   # Water density, in [kg/m^3]
@@ -30,16 +30,23 @@ def general():
     output["P_ATM"] = 101325 # Assumption on atmospheric pressure
     output["T_STANDARD"] = 323 # Standard temperature, for the calculation of the exergy in the exhaust gas
     output["ISO"] = {"LHV": 42700, "T_CA": 298, "T_LT": 298, "ETA_MECH": 0.8} # Reference values for ISO conditions
-    output["NAMES"] = {"MainEngines": ["ME1", "ME2", "ME3", "ME4"], "AuxEngines": ["AE1", "AE2", "AE3", "AE4"]}
[…]

### 2017-07-19 Python files/Energyanalysis.py

@@ -9,13 +9,10 @@ import os
 
 
 def energyAnalysisLauncher(processed, dict_structure, CONSTANTS):
-    processed = propertyCalculator(processed, dict_structure, CONSTANTS)
+    processed = propertyCalculator(processed, dict_structure, CONSTANTS)  # Calculates the energy and exergy flows
+    processed = efficiencyCalculator(processed, dict_structure, CONSTANTS)
     return processed
 
-
-def summaryEnergyFlows(processed, CONSTANTS):
-
-    return processed
 
 
 
@@ -116,6 +113,61 @@ def propertyCalculator(processed, dict_structure, CONSTANTS):
                     elif dict_structure["systems"][system]["units"][unit]["flows"][flow]["type"] in {"Wdot", "CEF"}:
                         processed[d2df(system, unit, flow, "Bdot")] = processed[d2df(system, unit, flow, "Edot")]
     print("...done!")
+    return processed
+
+
+
+def efficiencyCalculator(processed, dict_structure, CONSTANTS):
+    text_file = open(CONSTANTS["filenames"]["consistency_check_report"], "a")
+    text_file.write("\n *** CALCULATING EFFICIENCIES *** \n")
+    df_index = processed.index
+    for system in dict_structure["systems"]:
+        for unit in dict_structure["systems"][system]["units"]:
+            # Here I create eight (8) series: 4 Edot and 4 Bdot. 2 "Full" and 2 "useful".
+            temp_flow_list = {"Edot_in", "Edot_in_useful", "Edot_out", "Edot_out_useful", "Bdot_in", "Bdot_in_useful", "Bdot_out", "Bdot_out_useful"}
+            temp_df = pd.DataFrame(0, columns=temp_flow_list, index=df_index)
+            for flow in dict_structure["systems"][system]["units"][unit]["flows"]:
+                if flow[-2:] == "in":
+                    temp_df['Edot_in'] = temp_df['Edot_in'] + processed[d2df(system, unit, flow, "Edot")]
+                    temp_df['Bdot_in'] = temp_df['Bdot_in'] + processed[d2df(system, unit, flow, "Bdot")]
+                elif flow[-3:] == "out":
+                    temp_df['Edot_out'] = temp_df['Edot_out'] + processed[d2df(system, unit, flow, "Edot")]
+                    temp_df['Bdot_out'] = temp_df['Bdot_out'] + processed[d2df(system, unit, flow, "Bdot")]
+                else:
[…]

### 2017-07-19 Python files/auxiliaryDemand.py

@@ -9,7 +9,7 @@ def auxPowerAnalysis(processed, CONSTANTS, dict_structure):
     processed = propPowerDemand(processed, CONSTANTS)
     # Calculating the contribution from the thrusters
     #processed = thrusters(processed, CONSTANTS)  # The second method works better. But I still keep it
-    processed = thrusters2(processed, CONSTANTS)
+    processed = thrusters(processed, CONSTANTS)
     processed = HVAC(processed, CONSTANTS)
     processed = heatDemand(processed, CONSTANTS, dict_structure)
     processed = HTHR(processed, CONSTANTS)
@@ -64,28 +64,6 @@ def HVAC(processed, CONSTANTS):
 
 
 def thrusters(processed, CONSTANTS):
-    aaa = 0 # This is only defined as a trick for the cases when there are two consecutive time steps with thrusters on
-    for idx in range(len(processed.index)):
-        if aaa == 1:
-            aaa = 0
-            continue
-        idt = processed.index[idx]
-        if idx > len(processed.index) - 2:
-            processed.loc[idt, "Demands:Electricity:Thrusters:Edot"] = 0
-        else:
-            if processed["Demands:Electricity:Total:Edot"][processed.index[idx]] - processed["Demands:Electricity:Total:Edot"][processed.index[idx-1]] > 240:
-                if abs(processed["Demands:Electricity:Total:Edot"][processed.index[idx+1]] - processed["Demands:Electricity:Total:Edot"][processed.index[idx]]) < 50:
-                    processed.loc[processed.index[idx], "Demands:Electricity:Thrusters:Edot"] = processed["Demands:Electricity:Total:Edot"][processed.index[idx]] - (processed["Demands:Electricity:Total:Edot"][processed.index[idx - 1]] + (processed["Demands:Electricity:Total:Edot"][processed.index[idx + 2]] - processed["Demands:Electricity:Total:Edot"][processed.index[idx-1]]) / 3)
[…]

### 2017-07-19 Python files/consistencycheck.py

@@ -20,7 +20,7 @@ def systemCheck(processed, CONSTANTS, dict_structure, dataset_raw):
 def enginesCheck(processed, CONSTANTS):
     # This function checks that all relative values are consistent
     text_file = open(CONSTANTS["filenames"]["consistency_check_report"], "a")
-    print("Started consistency check...")
+    print("Started consistency check for the engines...", end="", flush=True)
     text_file.write("\n *** CONSISTENTCY CHECK FOR ENGINES *** \n")
     for type in ["MainEngines", "AuxEngines"]:
         for name in CONSTANTS["General"]["NAMES"][type]:
@@ -66,7 +66,7 @@ def HTHRcheck(processed, CONSTANTS):
     # This function checks that all relative values are consistent
     tot = len(processed.index)
     text_file = open(CONSTANTS["filenames"]["consistency_check_report"], "a")
-    print("Started consistency check...")
+    print("Started consistency check for the HTHR systems...", end="", flush=True)
     text_file.write("\n *** CONSISTENTCY CHECK FOR THE HTHR SYSTEM *** \n")
     # Checking that the temperature at the end of the circuit is lower than 90 degC
     temp = sum(processed["HTHR:HTHR24:HRWater_in:T"] <= (90 + 273.15)) / tot * 100
@@ -104,27 +104,27 @@ def HTHRcheck(processed, CONSTANTS):
     temp = sum(processed["HTHR:HotWaterHeater:HRWater_in:T"] >= processed["HTHR:HotWaterHeater:HRWater_out:T"]) / tot * 100
     warn = "WARNING " if temp < 95 else ""
     text_file.write(warn + " Hot Water Heater (HR side) temperatures are consistent for " + str(temp) + " % of the datapoints \n")
-
     text_file.close()
+    print("...done!")
 
 def SteamCheck(processed, CONSTANTS, dict_structure):
     # This function checks that all relative values are consistent
     tot = len(processed.index)
     text_file = open(CONSTANTS["filenames"]["consistency_check_report"], "a")
-    print("Started consistency check...")
+    print("Started consistency check for the steam systems...", end="", flush=True)
     text_file.write("\n *** CONSISTENTCY CHECK FOR THE STEAM SYSTEMS *** \n")
     for unit in dict_structure["systems"]["Steam"]["units"]:
[…]

### 2017-07-19 Python files/coolingsystems.py

@@ -55,6 +55,8 @@ def engineCoolingSystemsCalculation(processed, CONSTANTS, engine_type):
             processed[d2df(system,"HRSG","Steam_in","mdot")] = ((processed[d2df(system,"HRSG","Mix_in","T")] - processed[d2df(system,"HRSG","Mix_out","T")]) *
                 processed[d2df(system, "HRSG", "Mix_in", "mdot")] * processed[system+":CP_MIX"] /
                 (CONSTANTS["Steam"]["H_STEAM_VS"] - CONSTANTS["Steam"]["H_STEAM_LS"]))
+            temp = processed[d2df(system,"HRSG","Mix_in","T")] < processed[d2df(system,"HRSG","Mix_out","T")]
+            processed.loc[temp, d2df(system, "HRSG", "Steam_in", "mdot")] = 0
         # Calculation of the mass recirculating from the LT outlet to the HT inlet (not exactly...)
         processed[d2df(system, "HTmerge", "LTWater_in", "mdot")] = processed[d2df(system,"JWC","HTWater_out","mdot")] * (
             processed[d2df(system, "CAC_HT", "HTWater_out", "T")] - processed[d2df(system,"JWC","HTWater_in","T")]) / (
@@ -113,13 +115,6 @@ def centralCoolingSystems(processed, CONSTANTS):
     processed["CoolingSystems:LTHTmerge24:LTWater_in:mdot"] = mdot_HTHR24_tot - processed["CoolingSystems:LTHTmerge24:HTWater_in:mdot"]
 
     return processed
-
-
-
-
-
-def coolingSystemsPost(processed, CONSTANTS, dict_structure):
-    processed["CoolingSystems:LTHTmerge13:HTWater_in:T"]
 
 
 

### 2017-07-19 Python files/input.py

@@ -1,4 +1,8 @@
 import os
+from helpers import d2df
+import numpy as np
+
+
 
 def filenames(project_path):
     #project_path = os.path.realpath('..')
@@ -17,5 +21,57 @@ def filenames(project_path):
     text_file = open(output["consistency_check_report"],"w") # Cleaning the file
     text_file.write("=== STARTING THE REPORT FILE === \n \n")
     text_file.close()
-
     return output
+
+
+
+
+def assumptions(raw, processed, CONSTANTS, hd):
+    # This function includes generic assumed values in the main structure
+    # ALL ENGINES
+    for system in {"ME1", "ME2", "ME3", "ME4", "AE1", "AE2", "AE3", "AE4"}:
+        # The pressure at the turbocharger air inlet and exhaust outlet is equal to the atmospheric pressure
+        processed.loc[:,d2df(system,"Comp","Air_in","p")] = CONSTANTS["General"]["P_ATM"]
+        processed.loc[:,d2df(system, "Turbine", "Mix_out", "p")] = CONSTANTS["General"]["P_ATM"]
+        # Temperature in the engine room, i.e. inlet to the compressor of the TC
+        processed[d2df(system, "Comp", "Air_in", "T")] = raw[hd["ER_AIR_T_"]] + 273.15
+        # Assuming that the pressure in the exhaust gas is 90% of the pressure in the inlet manifold. Somewhat reasonable
+        processed[d2df(system, "Cyl", "EG_out", "p")] = (0.9 * raw[hd[system + "-CAC_AIR_P_OUT"]] + 1.01325) * 100000
+        # Assuming the pressure of the fuel to be around 9 barg, based on measurements from ME4
+        processed.loc[:,d2df(system, "Cyl", "FuelPh_in", "p")] = (9 + 1.01325) * 10e5
+        # Assuming the temperature of the cylinder wall to be 150 degC
+        processed.loc[:,d2df(system, "Cyl", "QdotJW_out", "T")] = 150 + 273.15
+        processed.loc[:,d2df(system, "JWC", "QdotJW_in", "T")] = 150 + 273.15
+        # Assuming a temperature of 100 degC for heat losses from the TC shaft
[…]

### 2017-07-19 Python files/preprocessingAE.py

@@ -1,5 +1,6 @@
 import numpy as np
 import pandas as pd
+import fillerfunctions as ff
 import preprocessingO as ppo
 import coolingsystems as cs
 from helpers import d2df
@@ -11,8 +12,8 @@ def auxEngineProcessing(raw, processed, dict_structure, CONSTANTS, hd):
     processed = ppo.engineStatusCalculation("AuxEngines", raw, processed, CONSTANTS, hd)
     # Calculating the cooling flows, so that they can be later assigned easily
     processed = cs.coolingFlows(processed, CONSTANTS, "AuxEngines")
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-1.1")
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-1.2")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-1.1")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-1.2")
     # Calculating the power, including the generator efficiency
     processed = auxEnginePowerCalculation(processed, CONSTANTS)
     # Calculating engine load, that is used many times later on
@@ -21,12 +22,12 @@ def auxEngineProcessing(raw, processed, dict_structure, CONSTANTS, hd):
     processed = auxEngineFuelFlowCalculation(raw, processed, CONSTANTS)
     # Calculate air and exhaust gas flows in the main engines
     processed = auxEngineAirFlowCalculation(raw, processed, CONSTANTS)
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-2.1")
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-2.2")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-2.1")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-2.2")
     # Calculating cooling flows
     processed = cs.engineCoolingSystemsCalculation(processed, CONSTANTS, "AuxEngines")
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-3.1")
[…]

### 2017-07-19 Python files/preprocessingME.py

@@ -1,7 +1,8 @@
 import numpy as np
 import pandas as pd
-import preprocessingO as ppo
+import fillerfunctions as ff
 import coolingsystems as cs
+import preprocessingO as ppo
 import CoolProp.CoolProp as cp
 from helpers import d2df
 from helpers import polyvalHelperFunction
@@ -13,8 +14,8 @@ def mainEngineProcessing(raw, processed, dict_structure, CONSTANTS, hd):
     processed = readMainEnginesExistingValues(raw, processed, CONSTANTS, hd)
     processed = ppo.engineStatusCalculation("MainEngines", raw, processed, CONSTANTS, hd)
     processed = cs.coolingFlows(processed, CONSTANTS, "MainEngines")
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-1.1")
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-1.2")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-1.1")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-1.2")
     # Calculating the main engines fuel flows
     processed = mainEngineFuelFlowCalculation(raw, processed, CONSTANTS, hd)
     # Calculating the main engines power output
@@ -23,12 +24,12 @@ def mainEngineProcessing(raw, processed, dict_structure, CONSTANTS, hd):
     processed = ppo.engineLoadCalculation("MainEngines", raw, processed, CONSTANTS, hd)
     # Calculating air and exhaust gas flows in the main engines
     processed = mainEngineAirFlowCalculation(raw, processed, dict_structure, CONSTANTS)
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-2.1")
-    processed = ppo.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-2.2")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-2.1")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-2.2")
     # Calculating cooling flows
[…]

### 2017-07-19 Python files/preprocessingO.py

@@ -5,223 +5,6 @@ from helpers import polyvalHelperFunction
 import CoolProp.CoolProp as cp
 
 
-
-def assumptions(raw, processed, CONSTANTS, hd):
-    # This function includes generic assumed values in the main structure
-    # ALL ENGINES
-    for system in {"ME1", "ME2", "ME3", "ME4", "AE1", "AE2", "AE3", "AE4"}:
-        # The pressure at the turbocharger air inlet and exhaust outlet is equal to the atmospheric pressure
-        processed.loc[:,d2df(system,"Comp","Air_in","p")] = CONSTANTS["General"]["P_ATM"]
-        processed.loc[:,d2df(system, "Turbine", "Mix_out", "p")] = CONSTANTS["General"]["P_ATM"]
-        # Temperature in the engine room, i.e. inlet to the compressor of the TC
-        processed[d2df(system, "Comp", "Air_in", "T")] = raw[hd["ER_AIR_T_"]] + 273.15
-        # Assuming that the pressure in the exhaust gas is 90% of the pressure in the inlet manifold. Somewhat reasonable
-        processed[d2df(system, "Cyl", "EG_out", "p")] = (0.9 * raw[hd[system + "-CAC_AIR_P_OUT"]] + 1.01325) * 100000
-        # Assuming the pressure of the fuel to be around 9 barg, based on measurements from ME4
-        processed.loc[:,d2df(system, "Cyl", "FuelPh_in", "p")] = (9 + 1.01325) * 10e5
-        # Assuming the temperature of the cylinder wall to be 150 degC
-        processed.loc[:,d2df(system, "Cyl", "QdotJW_out", "T")] = 150 + 273.15
-        processed.loc[:,d2df(system, "JWC", "QdotJW_in", "T")] = 150 + 273.15
-        # Assuming a temperature of 100 degC for heat losses from the TC shaft
-        processed.loc[:,d2df(system, "TCshaft", "Losses_out", "T")] = 100 + 273.15
-        # We assume the mixgin temperature at the LT inlet of the HT mixer to be constant
-        processed.loc[:,d2df(system, "HTmerge", "LTWater_in", "T")] = CONSTANTS["MainEngines"]["T_COOLING_MIX"]
-        # Assuming the steam pressure and temperature in the HRSG to be constant...
-        hrsg_pressure_assumption = (6 + 1.01325) * 100000
[…]

### 2017-07-19 Python files/unitstructures.py

@@ -65,61 +65,107 @@ def flowStructure():
                                                      "CAC_HT": {}, "CAC_LT": {}, "LOC": {}, "JWC": {}, "Cyl": {},
                                                      "HTsplit": {}, "HTmerge": {}, "TCshaft": {}}
             # Compressor
-            structure["systems"][system]["units"]["Comp"]["flows"] = {"Air_in": {"type": "CPF"}, "Air_out": {"type": "CPF"}, "Power_in": {"type": "Wdot"}} # TC compressor
+            structure["systems"][system]["units"]["Comp"]["flows"] = {
+                "Air_in": {"type": "CPF", "IO": "output"},
+                "Air_out": {"type": "CPF", "IO": "output"},
+                "Power_in": {"type": "Wdot", "IO": "input"}} # TC compressor
             structure["systems"][system]["units"]["Comp"]["equations"] = ["MassBalance"]
             # Bypass Split
-            structure["systems"][system]["units"]["BPsplit"]["flows"] = {"Air_in": {"type": "CPF"}, "Air_out": {"type": "CPF"}, "BP_out": {"type": "CPF"}}  # Bypass valve
+            structure["systems"][system]["units"]["BPsplit"]["flows"] = {
+                "Air_in": {"type": "CPF"},
+                "Air_out": {"type": "CPF"},
+                "BP_out": {"type": "CPF"}}  # Bypass valve
             structure["systems"][system]["units"]["BPsplit"]["equations"] = ["MassBalance", "ConstantPressure", "ConstantTemperature"]
             # Bypass Merge
-            structure["systems"][system]["units"]["BPmerge"]["flows"] = {"EG_in": {"type": "CPF"}, "Mix_out": {"type": "CPF"}, "BP_in": {"type": "CPF"}}
+            structure["systems"][system]["units"]["BPmerge"]["flows"] = {
+                "EG_in": {"type": "CPF"},
+                "Mix_out": {"type": "CPF"},
+                "BP_in": {"type": "CPF"}}
             structure["systems"][system]["units"]["BPmerge"]["equations"] = ["MassBalance", "ConstantPressure"]
             # Bypass Valve
-            structure["systems"][system]["units"]["BPvalve"]["flows"] = {"BP_in": {"type": "CPF"}, "BP_out": {"type": "CPF"}}
+            structure["systems"][system]["units"]["BPvalve"]["flows"] = {
+                "BP_in": {"type": "CPF"},
+                "BP_out": {"type": "CPF"}}
[…]

### 2017-07-19 Manuscript/Manuscript.tex

@@
 
-This section should be written by FA :-)
+The operational data was collected on board from the ships' machine logging and surveillance system.
+The on board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB.
+The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average.
+The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python \cite{mckinney-proc-scipy-2010}.
+A new structured naming of headers and a consistent time frequency over all data points were created, and 245 selected data points were saved in a HDF5 table time series database which is the base for the analysis.
 
-Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse.
+All data-points were checked individually by creating a descriptive statistic and a histogram of each data-point.
+A filter was applied by setting up a pre-defined maximum and minimum value.
+As we do not know, or had any meaningful way of deriving, each individual sensors placement, sensor type or calibration status all data must be filtered and checked accordingly for outliers or bad data.
+
+Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska Högarna for the ambient temperature.
+The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska Högarna lighthouse is along the ship route, in between the Swedish archipelago and Åland.
 The assumption was validated based on  June-December period, for which onboard measurements were available.
 This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
-% We did the calculation, resampling the data from SHMI from a 1(3)h to 15min frequency using linear interpolation.
+The fit the SMHI-data with the rest of the data-set the SMHI data was resampled from 1h for the seawater and 3h for the ambient temperature, to 15min frequency using a linear interpolation.
 

### 2017-10-02 Manuscript/HeatDemandMethodology.tex
Implementation of the new heat balance model

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}

%opening
\title{Estimation of the heat demand}
\author{Francesco Baldi}

\begin{document}

\maketitle

As the heat demand is not measured, it is necessary to determine it based on the available indirect measurements.

The heat demand and generation can be summarized according to the following equations:
\begin{eqnarray}
\dot{Q}_{gen} & = & \dot{Q}_{EGB} + \dot{Q}_{HTHR} + \dot{Q}_{AB} \\
\dot{Q}_{dem} & = & \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + \dot{Q}_{HTH} + \dot{Q}_{MSH} \\
\end{eqnarray}

\section{Heat balance parameter estimation}

As not enough information and measurements are available to determine the various components of the heat balance, in this paper we determined them by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters. The parameter estimation problem is hence written as a minimization problem:
\begin{eqnarray}
	min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
\end{eqnarray}

where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections. A list of the parameters $\textbf{p}$ is shown in Table REF, together with the chosen upper and lower boundaries for the calibration procedure.
\begin{table}
	\centering
	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
		\hline 
		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
		\hline
		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 & 137 \\ 
		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 & 177 \\
		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 & 0.59 \\ 
		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 & 0.98 \\ 
		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 & 0.71 \\ 
		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 & 0.55 \\ 
		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 & 0.27 \\ 
		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 & 345 \\
		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 & 0.72 \\
		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 & 5580 \\ 
		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 & 2920 \\ 
		\hline
	\end{tabular} 
\end{table}

\section{Heat demand}

Based on the ship's heat balance, we calculate the heat demand as the sum of the contributions of the following elements:

Given that no direct measurement of these quantities is available, they had to be estimated based on the following assumptions:
\begin{table}
	\centering
	\begin{tabular}{p{3cm}lp{1.6cm}}
		\hline 
		Heat flow name & Equation \\
		\hline
		HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \frac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
		HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \frac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
		Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
		Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
		Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \frac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
		HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \frac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
		Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \frac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
		HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
		\hline
	\end{tabular}
\end{table}

where all $f_i$ factors are treated as calibration parameters. The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively. The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure REF.

\section{Heat generation}

\subsection{Exhaust gas boilers}

The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty. The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
\begin{equation}
\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
\end{equation}\label{eq:egb}

where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is assumed to be a function of the exhaust gas composition, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF

\subsection{High Temperature Heat Recovery}

It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines. However, measurements of this contribution and of other variables that could lead to its straight-forward identification, thus requiring certain assumptions.

\begin{eqnarray}
\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \\
 & = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
\end{eqnarray}

where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem. The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.

\subsection{Auxiliary boilers}

The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems. The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit. This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value. The calculation can be represented as follows

	\begin{algorithmic}[1]
		\State{Calculate the heat balance with no contribution from oil fired boilers}
		\While{$\int{(\dot{Q}_{HTHR}(t) + \dot{Q}_{EGB}(t) - \dot{Q}_{dem}(t)) dt} < - Q_{ab,max}$}
			\State{Find $ t^* | \int_{t_0}^{t^*}{(\dot{Q}_{HTHR}(t) + \dot{Q}_{EGB}(t) - \dot{Q}_{dem}(t))dt} = 0 $}
			\State{$\dot{Q}_{ab}(t^*:t^*+\frac{Q_{ab,max}}{\dot{Q}_{ab,des}}) = \dot{Q}_{ab,des}$}
		\EndWhile
	\end{algorithmic}

where the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers.

\section{Uncertainty quantification}

Given the lack of input values for the estimation of the heat demand, the provided estimate based on the procedure described above should be integrated with an estimation of the uncertainty.

In this study, we base the estimation of the uncertainty on the production side, as it is the one that has the largest amount of information available. In these regards, the uncertainty can be reduced to the contribution of three elements: $U(\dot{Q}_{EGB})$, $U(\dot{Q}_{HTHR})$, $U(\dot{Q}_{AB})$.

The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$. In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT]. The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions as described in Appendix REF, and can be estimated of being up to 10\%. The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value. These assumptions lead to the following estimation of the uncertainty:
\begin{equation}
\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
\end{equation}

That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements (result obtained for $\pm 20$ K uncertainty. The uncertainty is reduced to 20\% if the measurement uncertainty on the temperature is reduced to  $\pm 20$ K)

The uncertainty of the heat generated by the ABs can be reduced to the contribution of two elements: the uncertainty on $\dot{m}_{fuel,AB}$ and that on $\eta_{AB}$. The former will be at least as large as the calibration error (35\%) and will be considered equal to 50\% to be conservative and accounting also for errors in the aggregated boiler fuel measurements. In addition, the uncertainty on the efficiency can be considered to be around 10\% based on the discrepancy between the considered sources and on the expected variability of the efficiency with load. Similarly to the previous case, the combination of these efficiencies lead to a total uncertainty of 51\%, where the main contribution comes from the uncertainty of the model output compared to the actual fuel consumption.

The estimation of the uncertainty of $\dot{Q}_{HTHR}$ can be based on \ref{eqn:HTHR2} having assigned an acceptable variation of 20\% to the effectiveness of the heat exchanger, 20\% on the estimation of the reference mass flow, $\pm 5$ K uncertainty on water temperature measurements and considering negligible the uncertainty on $c_{p,w}$, leading to a 145\% uncertainty on this measurement.

\end{document}

### 2017-10-02 Main.py

@@ -25,7 +25,7 @@
 
 
 do_processed_data_preparation = "no"
-do_data_processing = "no"
+do_data_processing = "yes"
 
 
 

### 2017-10-02 Python files/Constants.py

@@ -167,6 +167,22 @@ def auxiliaryEngines(CONSTANTS):
 
 def otherUnits(CONSTANTS):
     output = {}  # Initializing the output dictionary
+    # Providing the optimal optimization vector
+    param = [1.36929841e+02, 1.76569058e+02, 5.90926596e-01, 7.06974933e-01, 9.98138178e-01, 5.47476618e-01,
+            2.69942265e-01, 3.45398252e+02, 7.18364651e-01, 3 * 2.06480953e+03, 2.92188518e+03]
+    Qdot_HTHR_constant = param[0]
+    Qdot_steam_constant = param[1]
+    HVACreheaterCoefficient = param[2]
+    HotWaterHeaterCoefficient = param[3]
+    HVACpreheaterCoefficient = param[4]
+    GalleyCoefficient = param[5]
+    OtherCoefficient = param[6]
+    HTHRinletTemperature = param[7]
+    HTHRhexEffectiveness = param[8]
+    BoilerStorageSize = param[9]
+    BoilerReferencePower = param[10]
+
+    # Boiler
     output["BOILER"] = {}  # Initializing the boiler sub-dictionary
     output["BOILER"]["ETA_DES"] = 0.8
     output["BOILER"]["ETA_REGR_X"] = [6.79E-02, 1.20E-01, 1.62E-01, 2.12E-01, 2.86E-01, 3.52E-01, 4.03E-01, 4.41E-01, 4.90E-01, 5.40E-01, 5.89E-01, 6.54E-01, 7.16E-01, 7.67E-01, 8.31E-01, 8.94E-01, 9.47E-01, 9.89E-01, 1.04E+00, 1.09E+00, 1.14E+00, 1.20E+00]
@@ -177,18 +193,33 @@ def otherUnits(CONSTANTS):
                                   ) / (1 - 100/21 * output["BOILER"]["OXYGEN_EG"]) * (32 + 100 / 21 * 28)
     for idx in range(len(output["BOILER"]["ETA_REGR_Y"])):
         output["BOILER"]["ETA_REGR_Y"][idx] = output["BOILER"]["ETA_REGR_Y"][idx] / max(output["BOILER"]["ETA_REGR_Y"])
[…]

### 2017-10-02 Python files/auxiliaryDemand.py

@@ -1,6 +1,7 @@
 import numpy as np
 import pandas as pd
 from helpers import d2df
+import datetime as dt
 
 
 def auxPowerAnalysis(processed, CONSTANTS, dict_structure):
@@ -103,24 +104,52 @@ def heatDemand(processed, CONSTANTS, dict_structure):
     #   midsummer and the 15th of August
     # - During the rest of the season, we take for each day the average number of passengers during that day for the season
 
-    # Since the number of passenergers is only known for a certain period, we take a factor 0.5 for the hot water demand
-    processed["Demands:Heat:HotWaterHeater:Edot"] = processed["Passengers_calc"] / 1800 * CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HOT_WATER_HEATER"] * pd.Series((round(len(processed)/96)*np.ndarray.tolist(np.repeat(CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HWH_HOURLY"],4)))[:len(processed)], index=db_index)
-    temp = CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HVAC_PREHEATER"] * (CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["T_AIR_REF_MAX"] - processed["T_air"]) / (CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["T_AIR_REF_MAX"] - CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["T_AIR_REF_MIN"])
-    temp[temp < 0.1 * CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HVAC_PREHEATER"]] = 0
-    processed["Demands:Heat:HVACpreheater:Edot"] = temp # The pre-heater is not working during summer
-    processed["Demands:Heat:HVACreheater:Edot"] = CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HVAC_REHEATER"] * processed["Demands:Electricity:HVAC:Edot"] / CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HVAC_POWER_DES"] * processed["Passengers_calc"] / 1800
-    processed["Demands:Heat:TankHeating:Edot"] = 0.5 * CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["TANK_HEATING"]
[…]

### 2017-10-02 Python files/input.py

@@ -65,7 +65,7 @@ def assumptions(raw, processed, CONSTANTS, hd):
     # HTHR system
     processed.loc[:,"HTHR:SteamHeater:HRWater_out:T"] = 90 + 273.15 # From the heat balance, the temperature needs to rise at 90 degrees
     # processed.loc[:,"HTHR:SteamHeater:HRWater_out:mdot"] = 298 / 3600 * CONSTANTS["General"]["RHO_W"] # the original value is in m3/h
-    processed.loc[:,"HTHR:HTHR24:HRWater_in:T"] = 70 + 273.15 # Assumption on the Temperature at the HTHR-24 inlet (HT side)
+    processed.loc[:,"HTHR:HTHR24:HRWater_in:T"] = CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["HTHR_INLET_TEMPERATURE"] # Assumption on the Temperature at the HTHR-24 inlet (HT side)
     processed.loc[:, "HTHR:HVACpreheater:Qdot_out:T"] = (50 - 23) / np.log((50+273.15)/(23+273.15))
     processed.loc[:, "HTHR:HVACreheater:Qdot_out:T"] = (80 - 60) / np.log((80 + 273.15) / (60 + 273.15))
     processed.loc[:, "HTHR:HotWaterHeater:Qdot_out:T"] = 70 + 273.15

### 2017-10-02 Python files/psoTuning2.py

@@ -35,45 +35,64 @@ def preparation():
     # Prepare the input
     constant_input = (data, CONSTANTS, "optimization")
     # Set the boundaries
-    lb = [0.0, 0.5, 0.5, 0.5, 0.0, 0.0, 343, 1e4, 2000]
-    ub = [1.2, 1.5, 1.2, 1.2, 1.0, 1.0, 353, 1e5, 8000]
-    # param[0] = weighting factor of the HVAC re-heater     (HVAC,RH)
-    # param[1] = weighting factor of the hot water heater   (HWH)
-    # param[2] = weighting factor of the HVAC pre-heater    (HVAC,PH)
-    # param[3] = weighting factor of the Galley             (G)
-    # param[4] = 0-lev weighting factor of the Other users        (TH, OT, HFOT)
-    # param[5] = 1-lev weighting factor of the Other users        (TH, OT, HFOT)
-    # param[6] = Fixed temperature of the HRHT
-    # param[7] = Boiler drum steam storage capacity [MJ, energy]
-    # param[8] = Boiler heat rate when in use [kW]
+    lb = [0.0,  0.0,  0.0, 0.5, 0.5, 0.5, 0.2, 343, 0.5, 1e2, 2000]
+    ub = [1000, 1000, 1.0, 1.0, 1.0, 1.0, 1.0, 353, 0.9, 1e5, 8000]
+    # param[0] = Constant demand to the HTHR systems
+    # param[1] = Constant steam demand
+    # param[2] = a1 factor of the HVAC re-heater     (HVAC,RH)
+    # param[3] = a1 factor of the hot water heater   (HWH)
+    # param[4] = a1 factor of the HVAC pre-heater    (HVAC,PH)
+    # param[5] = a1 factor of the Galley             (G)
+    # param[6] = a1 factor of the Other users        (TH, OT, HFOT)
+    # param[7] = Fixed temperature of the HRHT
+    # param[8] = Effectiveness of the HTHR heat exchangers
+    # param[9] = Boiler drum steam storage capacity [MJ, energy]
+    # param[10] = Boiler heat rate when in use [kW]
+
     # Launch the optimization
[…]

### 2017-10-02 Python files/psoTuning3.py

@@ -5,6 +5,7 @@ import constants as kk
 from pyswarm import pso
 import os
 import input
+import datetime as dt
 
 
 def preparation():
@@ -34,12 +35,23 @@ def preparation():
     # Prepare the input
     constant_input = (data, CONSTANTS, "optimization")
     # Set the boundaries
-    lb = [0.0, 343, 1e4, 2000]
-    ub = [1.0, 353, 1e5, 8000]
+    lb = [0.0, 0.5, 0.5, 0.5, 0.0, 0.0, 50/3.6, 1e4, 2000]
+    ub = [1.2, 1.5, 1.2, 1.2, 1.0, 1.0, 298/3.6, 1e5, 8000] # Design value is 298 m3/h
+    # param[0] = weighting factor of the HVAC re-heater     (HVAC,RH)
+    # param[1] = weighting factor of the hot water heater   (HWH)
+    # param[2] = weighting factor of the HVAC pre-heater    (HVAC,PH)
+    # param[3] = weighting factor of the Galley             (G)
+    # param[4] = 0-lev weighting factor of the Other users        (TH, OT, HFOT)
+    # param[5] = 1-lev weighting factor of the Other users        (TH, OT, HFOT)
+    # param[6] = Fixed temperature of the HRHT
+    # param[7] = Boiler drum steam storage capacity [MJ, energy]
+    # param[8] = Boiler heat rate when in use [kW]
     # Launch the optimization
-    xopt, fopt = pso(fitnessFunction, lb, ub, args=constant_input, debug=True)
-    # xopt = [6.71988240e-01, 5.60706250e-01, 2.97130271e+02, 2.49548290e+02, 8.10322609e-01, 5.39518203e-01, 3.51315316e+02, 5.42492594e+03]
-    # xopt = [6.06116231e-01, 9.19705648e-01, 2.96474980e+02, 2.52603434e+02, 9.85583957e-01, 5.23062343e-01, 3.44799481e+02, 1.00268573e+04, 7.25138168e+03]
+    # xopt, fopt = pso(fitnessFunction, lb, ub, args=constant_input, debug=True)
[…]

### 2017-10-10 Manuscript/HeatDemandMethodology.tex
(almost) final changes to the code
Most of the bugs have been eliminated

@@
 \usepackage[noend]{algpseudocode}
+\usepackage{tabls}
 
@@
 	\end{tabular}
+	\caption{Parameters optimized in the parameter estimation for the heat balance}
+	\label{tab:ParameterEstimation} 
 \end{table}
@@
 
-Based on the ship's heat balance, we calculate the heat demand as the sum of the contributions of the following elements:
-
-Given that no direct measurement of these quantities is available, they had to be estimated based on the following assumptions:
+We calculated the heat demand as the sum of the contributions of the elements listed in the ship's heat balance documentation.
+As no direct measurement of these quantities was available in the dataset, they had to be estimated based on the following assumptions:
 \begin{table}
 	\centering
-	\begin{tabular}{p{3cm}lp{1.6cm}}
+	{\tablinesep=2ex\tabcolsep=10pt
+	\begin{tabular}{p{2.8cm}l}
 		\hline 
@@
 		\hline
-		HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \frac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
-		HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \frac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
+		HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
+		HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
 		Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
 		Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
-		Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \frac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
-		HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \frac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
-		Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \frac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
+		Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
+		HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
+		Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
 		HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
 		\hline
-	\end{tabular}
+	\end{tabular}}
+	\caption{Summary of the heat demand contributions and their calculation}
+	\label{tab:HeatDemand}
 \end{table}
 
-where all $f_i$ factors are treated as calibration parameters.
+where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
 The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
@@
 
-where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is assumed to be a function of the exhaust gas composition, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF
+where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF
 
@@
 It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
-However, measurements of this contribution and of other variables that could lead to its straight-forward identification, thus requiring certain assumptions.
+However, measurements of this contribution and of other variables that could lead to its straight-forward identification are missing.
+In this work, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},
 
 \begin{eqnarray}
-\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \\
+\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
  & = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
@@
 
-where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem.
+where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
 The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
@@
 
-where the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers.
+where the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
 
-\section{Uncertainty quantification}
+\section{Parameter estimation and uncertainty quantification}
+
+The results of the parameter estimation are represented graphically and quantitatively in Figure REF and Table REF.
 

### 2017-10-10 Manuscript/Manuscript.tex
(almost) final changes to the code
Most of the bugs have been eliminated

--- 2017-07-19 Manuscript/Manuscript.tex
@@
 \usepackage{amsmath}
+%% The package tabls helps in having nicer tables with higher spaces
+\usepackage{tabls}
+%% The package multirow allows merging together multiple cells vertifcally
+\usepackage{multirow}
+%% 
+\usepackage{subcaption}
+
 \def\bibsection{\section*{References}}
@@
 
-According to the third IMO GHG Study, in 2012 CO$_2$ emissions from shipping amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions \cite{Smith2014}.
+CO$_2$ emissions from shipping in 2012 amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions \cite{Smith2014}.
 Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios \cite{Smith2014}.
 As an example, global transport demand has increased by 3.4\% in 2014, compared to a global GBP growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}.
+The OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3].
+The fact that shipping needs to even further reduce its CO$_2$ emissions in the near future is essential for being able to achieve the goals of maintaining the climate below a 2-degree level in 2050 [4].
 
-International Energy Agency data show that the OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3].
-The fact that shipping needs to even further reduce its CO$_2$ emissions in the near future is essential for being able to achieve the goals of maintaining the climate below a 2-degree level in 2050 [4].
-Finally, in the Baltic Sea an emission control area is enforced by the International Maritime Organisation since January 2015 which stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels.
+In addition to considerations related to GHG emissions, an emission control area (ECA) is enforced in the Baltic Sea  by the International Maritime Organisation since January 2015.
+This ECA stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels.
 More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}
@@
 	\item Analyse the demand of a cruise ship in terms of propulsion power, electric power and heat, based on operational measurements.
-	\item Analyse the current efficiency of the system, and potential ways to improve it, by means of applying exergy analysis
-	\item Provide a more detailed analysis of the heat demand of a cruise ship, as a basis for further studies on how to optimize the whole efficiency of the propulsion plant.
+The lack of existing information with reference to the heat demand is considered a particular element of novelty when compared to existing literature in the subject.
+	\item Analyse the current efficiency of the system, and potential ways to improve it, by means of applying exergy analysis.
+	\item Provide reference operational conditions for further use in the field of energy systems optimization.
 \end{itemize}
@@
 
-Here an overview of the method is provided: 
-\begin{itemize}
-	\item Data gathering
-	\item Data filtering
-	\item Data pre-processing
-	\item Heat demand estimation
-	\item Statistical analysis
-	\item Energy and exergy analysis
-\end{itemize}
+In this paper, we present the application of energy and exergy analysis (described in section \ref{sec:met:energyExergy}) to a cruise ship.
+This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; the specific assumptions and methods used for processing the available information into  energy and exergy flows are specified in Section \ref{sec:met:processing}.
+In these regards, the estimation of the heat demand is treated in a separate section (Sec.
+\ref{sec:met:heat}) as it constitutes a particularly challenging task and it is one of the central aspects of this article.
 
-In this paper, we present the application of energy and exergy analysis (described in sections \ref{set:met:energy} and \ref{sec:met:exergy} to a cruise ship.
-This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; these are used to generate the required information for calculating energy and exergy flows (Section \ref{sec:met:processing}.
-In these regards, the estimation of the heat demand is treated in a separate section (Sec.
-\ref{sec:met:heat} as it constituted a particularly challenging task and it is one of the central aspects of this article.
+\subsection{Energy and exergy analysis} \label{sec:met:energyExergy}
 
-\subsection{Energy analysis}
+\subsubsection{Energy analysis}
 
@@
 
-\subsection{Exergy analysis}
+\subsubsection{Exergy analysis}
 
@@
 
-\subsection{System performance}
+\subsubsection{System performance}
 
@@
 The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller.
-The main engines are four Wärtsilä 4-stroke Diesel engines (ME) rated 5850 kW each All engines are equipped with selective catalytic reactors (SCR) for NOX emissions abatement.
+The main engines are four Wärtsilä 4-stroke Diesel engines (ME) rated 5850 kW each.
+All engines are equipped with selective catalytic reactors (SCR) for NOX emissions abatement.
 Propulsion power is needed whenever the ship is sailing; however, it should be noted that the ship rarely sails at full speed, and most of the time it only needs one or two engines operated simultaneously.
@@
 
-Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska Högarna for the ambient temperature.
-The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska Högarna lighthouse is along the ship route, in between the Swedish archipelago and Åland.
+Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska H\"{o}garna for the ambient temperature.
+The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska H\"{o}garna lighthouse is along the ship route, in between the Swedish archipelago and Åland.
 The assumption was validated based on  June-December period, for which onboard measurements were available.
@@
 
-\subsection{Data processing: Ship energy system modeling} \label{sec:met:modeling}
+\subsection{Data processing: Ship energy system modeling} \label{sec:met:processing}
 
@@
 
-\subsubsection*{Main engines}
+\subsubsection{Diesel engines}
 
-Make a figure representing the main engines, with specified the variables for which we have measurements of.
+The setup of the engines and their connections to the other ship systems are summarized in Figure REF and REF for the main and auxiliary engines, respectively.
+The defining net energy flows are listed in Table REF.
+It should be noted that, in the case of the main engines, the engine power output was not measured and needed to be estimated.
+In this work, we calculated the engine power output based on measurements of engine fuel rack position (used as a proxy of the mass of fuel injected per cycle) and of the engine speed:
+\begin{eqnarray}
+\dot{m}_{fuel} & = & \dot{m}_{fuel,des} \left(a_0 + a_1 \frac{frp}{frp_{des}}\right) \frac{\omega}{\omega_{des}} \\
+\eta_{ME} & = & a_0 + a_1 \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} + a_2 \left( \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} \right)^2 \\
+\dot{W}_{ME} & = & \dot{m}_{fuel} \eta_{ME} LHV
+\end{eqnarray}
 
-List the main assumptions
+The validity of this assumption can be seen by observing Figure REF, representing the calculated total power from the main engines as a function of the ship speed
 
-\subsubsection*{Auxiliary engines}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/DieselEngines}
+	\caption{Schematic representation of the Diesel engines and their connections to the cooling systems}
+	\label{fig:DieselEngines}
+\end{figure}
 
-Make a figure representing the main engines, with specified the variables for which we have measurements of.
+In the case of the auxiliary engines, instead, direct measurements of the power generated by each engine were available.
 
-List the main assumptions
+For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine.
+The model is explained in detail in Appendix REF.
 
-\subsubsection*{Auxiliary boilers}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pme_vs_vship_Neng}
+	\caption{Scatter plot, Propulsion power versus ship speed}
+	\label{fig:Pme_vs_vship_Neng}
+\end{figure}
 
-\subsubsection*{Cooling systems}
+ALSO INCLUDE THE POWER VS TC SPEED RELATIONSHIP
 
-\subsection{Estimation of the heat demand} \label{sec:met:heat}
+\begin{table}
+	\small
+	\centering
+	\begin{tabular}{llccrrrrr}
+		\hline
+		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
+		\hline
+		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
+		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
+		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
+		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
+		&   $\beta_{comp}$		& - 	& M & & & & &  \\
+		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
+		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
+		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
+		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
+		&   $\beta_{exp}$		& - 	& C & & & & &  \\
+		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
+		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
+		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
+		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
+		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
+		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
+		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
+		&	$T_{eg,out}$  		& $^o$C & M  & & & & &  \\ 
+		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
+		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
+		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
+		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
+		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
+		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
+		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
+		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
+		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
+		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
+		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
+		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
+		\multicolumn{2}{l}{LOC} & & & & & & &  \\
+		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
+		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
+		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
+		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
+		\multicolumn{2}{l}{JWC} & & & & & & &  \\
+		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
+		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
+		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
+		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
+		\hline
+		\end{tabular}
+	\caption{Main engines, measured and calculated temperatures and flows at different engine loads}
+\end{table}
 
-\subsubsection*{Production-side}
+INCLUDE VALIDATION AGAINST PROJECT GUIDE DATA
 
-In a top-down model, the demand of an energy system is modeled based on the production side of the energy balance.
-According to the top-down approach, the total heat consumption is calculated based on the sum of the heat provided by the three main heat sources available on board:
-\begin{itemize}
-	\item Exhaust gas boilers
-	\item Auxiliary oil-fired boilers
-	\item Heat recovery on the HT cooling systems
-\end{itemize}
+\begin{table}
+	\small
+	\centering
+	\begin{tabular}{llccrrrrr}
+		\hline
+		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
+		\hline
+		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
+		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
+		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
+		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
+		&   $\beta_{comp}$		& - 	& M & & & & &  \\
+		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
+		&	$T_{air,in}$   		& $^o$C & M  & & & & &  \\ 
+		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
+		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
+		&   $\beta_{exp}$		& - 	& C & & & & &  \\
+		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
+		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
+		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
+		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
+		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
+		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
+		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
+		&	$T_{eg,out}$  		& $^o$C & C  & & & & &  \\ 
+		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
+		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
+		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
+		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
+		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
+		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
+		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
+		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
+		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
+		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
+		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
+		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
+		\multicolumn{2}{l}{LOC} & & & & & & &  \\
+		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
+		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
+		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
+		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
+		\multicolumn{2}{l}{JWC} & & & & & & &  \\
+		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
+		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
+		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
+		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
+		\hline
+	\end{tabular}
+	\caption{Auxiliary engines, measured and calculated temperatures and flows at different engine loads}
+\end{table}
 
-For the case of the EGBs, the heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
-\begin{equation}
-\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
-\end{equation}\label{eq:egb}
+INCLUDE VALIDATION AGAINST PROJECT GUIDE DATA
 
-Where measurements of the exhaust gas temperature before and after the EGBs (respectively $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$) are available from the DLS; the mass flow of exhaust gas ($\dot{m}_{eg}$) is calculated as described in appendices \ref{app:me} and \ref{app:ae}; and the specific heat at constant pressure of the exhaust gas ($c_{p,eg}$) is assumed to be constant and equal to 1.08 $\frac{kJ}{kgK}$.
+\begin{table}
+	\small
+	\centering
+	\begin{tabular}{lcrrrrr}
+		\hline
+		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
+		\hline
+		Power output & Mech &  &  &  &  &  \\
+		Exhaust gas & Heat &  &  &  &  &  \\
+		CAC-LT & Heat &  &  &  &  &  \\
+		CAC-HT & Heat &  &  &  &  &  \\
+		JWC & Heat &  &  &  &  &  \\
+		LOC & Heat &  &  &  &  &  \\
+		\hline
+	\end{tabular}
+	\caption{Main engines, energy flows at different engine loads}
+\end{table}
 
-For the case of the ABs, the heat transferred to the steam ($\dot{Q}_{AB}$) is calculated according to equation \ref{eq:ab}:
-\begin{equation}
-\dot{Q}_{AB} = \dot{m}_{fuel,AB} \eta_{AB}
-\end{equation} \label{eq:ab}
+\begin{table}
+	\small
+	\centering
+	\begin{tabular}{lcrrrrr}
+		\hline
+		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
+		\hline
+		Power output & Mech &  &  &  &  &  \\
+		Exhaust gas & Heat &  &  &  &  &  \\
+		CAC-LT & Heat &  &  &  &  &  \\
+		CAC-HT & Heat &  &  &  &  &  \\
+		JWC & Heat &  &  &  &  &  \\
+		LOC & Heat &  &  &  &  &  \\
+		\hline
+	\end{tabular}
+	\caption{Auxiliary engines, energy flows at different engine loads}
+\end{table}
 
-Where the fuel consumption of the ABs ($\dot{m}_{fuel,AB}$) is available on a monthly basis.
-No information was available concerning the first-law efficiency of the boilers ($ \eta_{AB}$).
-\cite{Cohen1962} estimated it to vary between 0.83 and 0.89 depending on the load of the boiler, while according to a more recent experimental campaign presented in \cite{Mrzljak2017} the efficiency varies between 0.7 and 0.79.
-In the absence of more specific available data, a value of 0.8 was selected for $ \eta_{AB}$
+\subsubsection{Cooling systems}
 
-The contribution from the heat recovery on the high temperature cooling systems of the engines ($\dot{Q}_{HRHT}$) represents the main uncertainty related to the top-down estimations.
-This contribution cannot be measured, neither directly or indirectly, and hence needs to be estimated.
+The ship's cooling systems are designed similarly to most ships and divided in two cooling systems of two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C).
+The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board.
+A representation of the ship cooling systems is provided in Figure XXX.
 
-For this reason, two calculated values for the top-down heat demand are provided:
-\begin{itemize}
-	\item A \textbf{high boundary} case, where it is assumed that almost all of the heat available from the HT cooling systems (an utilization factor of 0.9 is assumed to account for losses) is used.
-This corresponds to the assumption that the use of the waste heat from the HT cooling systems is prioritized over the use of waste heat from the exhaust gas of the engines.
-	\begin{equation}
-	\dot{Q}_{HRHT} = f_{HRHT} \dot{Q}_{HT,tot} 
-	\end{equation}
-	\item a \textbf{low boundary} case, where it is assumed that the waste heat from the cooling systems is only used when the ship is in port (again assuming a utilization factor of 0.9). 
-	\item a reference case, (NOTE: We should discuss what the assumption could be here. The one we had in the ECOS paper? Or a value halfway between the highest and lowest estimate?)
-\end{itemize}
+All values not directly measured in the cooling systems were calculated based on mass and energy balances.
 
-In both cases, the available heat from the HT cooling systems is estimated as explained in Section \ref{eq:met:modeling}.
-
-\subsubsection*{Consumers-side}
-Following, the assumptions that we have made related to the bottom-up estimations of ship heat demand:
-\begin{itemize}
-	\item In the HVAC unit, the "\textbf{Preheater}" is only used when heating (i.e., in winter) while the "\textbf{Reheater}" is only used when cooling (i.e., in summer)
-	\item \textbf{Spacial heating} demand during \textbf{winter} is linearly dependent on the external air temperature, from a minimum of 0 for an external temperature of 20$^o$C to a maximum of 3500 kW (design power of the preheater)  for an outer air temperature of -20$^o$C.
-	\item \textbf{Spacial heating} demand during \textbf{summer} is related to the need for dehumidification, and harder to estimate based on available measurements.
-As a first assumption, it is estimated to be equal to the cooling demand (estimated as detailed in Sec.
-ref{sec:met:electric})
-	\item Every passenger contributes with a total of 150 W of free heat during the day (8-23, ref.
-to "walking, seated" activity from \cite{Wang2000}) and 100 W during night (23-8, ref.
-to "seated at rest" activity from \cite{Wang2000}).
-For the crew, as it is assumed that they perform a more intense activity than the passengers, the daily contribution is considered equal to 160 (ref.
-to "moderate work" activity from \cite{Wang2000})
-	\item The heat required for \textbf{tank heating} is linearly dependent on the external sea water temperature. The maximum heat is required for an external seawater temperature of 0$^o$C
-	\item The heat required from the \textbf{galley} depends on two parameters:
-	\begin{itemize}
-		\item The \textbf{number of passengers} (linear dependence)
-		\item the \textbf{hour of the day}. See Figure \ref{fig:hwhvstime}
-	\end{itemize}
-	\item The heat required from the \textbf{hot water} depends on two parameters:
-	\begin{itemize}
-		\item The \textbf{number of passengers} (linear dependence)
-		\item the \textbf{hour of the day}. The hour-dependent factor is assumed based on the estimations for a land-based hotel reported in \cite{Chung2015}.
-	\end{itemize}   
-\end{itemize}
-
-In Table \ref{tab:heatConsumers} we summarize all heat consumers on board the case study vessel.
-For each entry, we provide the design heat demand and the variables we assumed the actual heat demand to depend on:
-\begin{table}
-	\begin{tabular}{lrccccc}
-		\hline 
-		Name & $\dot{Q}_{des}$ [kW] & \multicolumn{5}{c}{Depends on...} \\ 
-		\hline 
-		 & & $T_{air}$ & t & $N_{pax}$ & $f_{TL}$ & $\lambda$ \\ 
-		\hline 
-		Hot water heater & 1200 &  & x & x &  &  \\ 
-		HVAC Preheater & 3500 & x & x & x &  &  \\ 
-		HVAC Reheater & 1780 & x & x & x &  &  \\ 
-		Tank heating & 208 &  &  &  & x &  \\ 
-		Other tanks & 138 &  &  &  & x &  \\ 
-		HFO Tank heating & 271 &  &  &  & x &  \\ 
-		Machinery space heaters & 281 & x &  &  &  &  \\ 
-		Bilge water separator & 26 &  &  &  &  &  \\ 
-		Hot water calorifier & 366 &  & x &  &  &  \\ 
-		Fuel oil heater & 103 &  &  &  &  & x \\ 
-		HFO Separator & 37 &  &  &  &  &  \\ 
-		Galley & 602 &  & x & x &  &  \\
-		\hline 
-	\end{tabular} \label{tab:heatConsumers}
-\end{table}
+Figure cooling systems
 
@@
 
-\subsubsection*{HVAC systems}
+\subsubsection{HVAC systems}
 
@@
 
-\subsubsection*{Thrusters}
+\subsubsection{Thrusters}
 
@@
 In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.
-
-The observation of Figure XX shows how the power demand from thrusters can be identified as "spikes" in the total electric power demand.
 In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
@@
 
+\subsection{Estimation of the heat demand} \label{sec:met:heat}
+
+As the heat demand is not measured, it is necessary to determine it based on the available indirect measurements.
+
+The heat demand and generation can be summarized according to the following equations:
+\begin{eqnarray}
+\dot{Q}_{gen} & = & \dot{Q}_{EGB} + \dot{Q}_{HTHR} + \dot{Q}_{AB} \\
+\dot{Q}_{dem} & = & \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \\
+\end{eqnarray}
+
+The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.999\linewidth]{Figures/Heating_cooling_systems}
+	\caption{General view of the ship heating and cooling systems}
+	\label{fig:ship:general}
+\end{figure}
+
+\subsubsection{Heat balance parameter estimation}
+
+As not enough information and measurements are available to determine the various components of the heat balance, in this paper we determined them by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters.
+The parameter estimation problem is hence written as a minimization problem:
+\begin{eqnarray}
+min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
+\end{eqnarray}
+
+where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections.
+A list of the parameters $\textbf{p}$ is shown in Table REF, together with the chosen upper and lower boundaries for the calibration procedure.
+\begin{table}
+	\centering
+	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
+		\hline 
+		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
+		\hline
+		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 &  \\ 
+		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 &  \\
+		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 &  \\ 
+		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 &  \\ 
+		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 &  \\ 
+		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 &  \\ 
+		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 &  \\ 
+		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 &  \\
+		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 &  \\
+		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 &  \\ 
+		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 &  \\ 
+		\hline
+	\end{tabular}
+	\caption{Parameters optimized in the parameter estimation for the heat balance}
+	\label{tab:ParameterEstimation} 
+\end{table}
+
+\subsubsection{Heat demand}
+
+We calculated the heat demand as the sum of the contributions of the elements listed in the ship's heat balance documentation.
+As no direct measurement of these quantities was available in the dataset, they had to be estimated based on the following assumptions:
+\begin{table}
+	\centering
+	{\tablinesep=2ex\tabcolsep=10pt
+		\begin{tabular}{p{2.8cm}l}
+			\hline 
+			Heat flow name & Equation \\
+			\hline
+			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
+			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
+			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
+			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
+			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
+			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
+			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
+			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
+			\hline
+	\end{tabular}}
+	\caption{Summary of the heat demand contributions and their calculation}
+	\label{tab:HeatDemand}
+\end{table}
+
+where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
+The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
+The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure REF.
+
+\subsubsection{Heat generation}
+
+The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty.
+The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
+\begin{equation}
+\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
+\end{equation}\label{eq:egb}
+
+where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF
+
+It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
+However, measurements of this contribution and of other variables that could lead to its straight-forward identification are missing.
+In this work, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},
+
+\begin{eqnarray}
+\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
+& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
+\end{eqnarray}
+
+where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
+The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
+
+The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems.
+The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit.
+This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value.
+In this calculation, the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
+
+\subsubsection{Parameter estimation and uncertainty quantification}
+
+The results of the parameter estimation are represented graphically and quantitatively in Figure REF and Table REF.
+
+Given the lack of input values for the estimation of the heat demand, the provided estimate based on the procedure described above should be integrated with an estimation of the uncertainty.
+
+In this study, we base the estimation of the uncertainty on the production side, as it is the one that has the largest amount of information available.
+In these regards, the uncertainty can be reduced to the contribution of three elements: $U(\dot{Q}_{EGB})$, $U(\dot{Q}_{HTHR})$, $U(\dot{Q}_{AB})$.
+
+The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$.
+In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT].
+The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions as described in Appendix REF, and can be estimated of being up to 10\%.
+The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value.
+These assumptions lead to the following estimation of the uncertainty:
+\begin{equation}
+\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
+\end{equation}
+
+That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements (result obtained for $\pm 20$ K uncertainty.
+The uncertainty is reduced to 20\% if the measurement uncertainty on the temperature is reduced to  $\pm 20$ K)
+
+The uncertainty of the heat generated by the ABs can be reduced to the contribution of two elements: the uncertainty on $\dot{m}_{fuel,AB}$ and that on $\eta_{AB}$.
+The former will be at least as large as the calibration error (35\%) and will be considered equal to 50\% to be conservative and accounting also for errors in the aggregated boiler fuel measurements.
+In addition, the uncertainty on the efficiency can be considered to be around 10\% based on the discrepancy between the considered sources and on the expected variability of the efficiency with load.
+Similarly to the previous case, the combination of these efficiencies lead to a total uncertainty of 51\%, where the main contribution comes from the uncertainty of the model output compared to the actual fuel consumption.
+
+The estimation of the uncertainty of $\dot{Q}_{HTHR}$ can be based on \ref{eqn:HTHR2} having assigned an acceptable variation of 20\% to the effectiveness of the heat exchanger, 20\% on the estimation of the reference mass flow, $\pm 5$ K uncertainty on water temperature measurements and considering negligible the uncertainty on $c_{p,w}$, leading to a 145\% uncertainty on this measurement.
+
+\subsection{Determination of typical operational days}
+
 \section{Results (and discussion)}
@@
 
-Figure \ref{fig:Tsea_vs_time} represents the distribution and the evolution over the year of the ambient air and sea water temperature.
+Figures \ref{fig:TseaTairTIME} and \ref{fig:TseaTairDIST} represent the time evolution and the distribution over the year of the ambient air and sea water temperature.
 The relatively low temperatures that are experienced by the case study ship during its operations suggest that heating demand can be expected to be higher than cooling demand.
@@
 
-The corresponding distributions are presented in Figure \ref{fig:Tsea_hist}
-
-\begin{figure}
+\begin{figure}[htbp]
 	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Tsea_vs_time}
-	\caption{Yearly evolution of sea water and ambient air temperatures. Landsort, 2014}
-	\label{fig:Tsea_vs_time}
-\end{figure}
-
-\begin{figure}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centering
+		\includegraphics[width=0.999\linewidth]{Figures/Tsea_vs_time}
+		\caption{Time series}
+		\label{fig:TseaTairTIME}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.45\textwidth}
 		\centering
 		\includegraphics[width=0.9\linewidth]{Figures/Tsea_hist}
-	\caption{Yearly distribution of sea water and ambient air temperatures. Landsort, 2014}
-	\label{fig:Tsea_hist}
+		\caption{Distribution}
+		\label{fig:TseaTairDIST}
+	\end{subfigure}
+	\caption{Statistical representation of measured air and sea water temperatures based on yearly data. Landsort, 2014}
+	\label{fig:TseaTair}
 \end{figure}
@@
 
-NOTE: It would be interesting to find a way to categorize all periods in each phase (sailing in the archipelago - sailing at open sea) and trying to analyze statistically their duration / mean power / etc.
-
-\begin{figure}
+\begin{figure}[htbp]
 	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/vship_hist}
+	\includegraphics[width=0.8\linewidth]{Figures/vship_hist}
 	\caption{Yearly distribution of ship speed}
@@
 
-\begin{figure}
+figure --> Operational mode distribution
+\begin{figure}[htbp]
 	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pme_vs_vship}
-	\caption{Scatter plot, Propulsion power versus ship speed}
-	\label{fig:Pme_vs_vship}
+	\includegraphics[width=0.8\linewidth]{Figures/Hist_mainEngines}
+	\caption{Yearly load distribution, main engines}
+	\label{fig:loadME_hist}
 \end{figure}
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/Hist_auxEngines}
+	\caption{Yearly load distribution, auxiliary engines}
+	\label{fig:loadAE_hist}
+\end{figure}
+
+ANYTHING ELSE THAT CAN BE INTERESTING TO SHOW IN THE EXPLORATORY DATA ANALYSIS?
 
@@
 
-The results are presented in three main forms:
-\begin{itemize}
-	\item Time series (comparing some relevant trips with a daily time scale)
-	\item Distribution plot (or box plot...box plots are easier to compare, distribution plots are easier to interpret)
-	\item Sankey diagrams
-\end{itemize}
+Here we can see the typical example (note: the day was not selected as "typical", but simply randomly) of the energy demand during a winter and a summer day (Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}).
+It can be seen clearly how, as expected, the heating demand in winter is much higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduce need for compartment heating.
+On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors.
 
-The results should be presented both for the full database and for each operational mode (port-sailing-drifting)
-
-Here we can see the typical example (note: the day was not selected as "typical", but simply randomly) of the energy demand during a winter and a summer day (Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}).
-
-\begin{figure}
+\begin{figure}[htbp]
 	\centering
@@
 
-\begin{figure}
+\begin{figure}[htbp]
 	\centering
@@
 
-\subsection{Heat demand} \label{sec:res:heat}
-
-In Figure \ref{fig:Heat_vs_time} the heat generated by different hot utilities, as estimated by the model, is shown.
-It should be noticed that, while the heat generated by the HRSGs represents a rather accurate estimation, all other data series (auxiliary boilers, HTHR) are the consequence of rather strong assumptions.
-
-The figure also shows the operation of validation of the assumptions by comparison of the measured values for the boilers fuel consumption (here translated into an energy-per-day basis to ease the comparison).
-It should be noted that, although the absolute value is still quite different compared to the calculated one (mean error is 255\%, minimum 3\%), there is evidence that the order of magnitude is similar, and that some trends are followed.
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyDemand}
+	\caption{Energy demand, aggregated data for one year of operations}
+	\label{fig:Pie_EnergyDemand}
+\end{figure}
 
@@
 	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/Heat_vs_time}
-	\caption{Yearly evolution of heat generated by different sources. "Measured" and "Calculated" relate to the auxiliary oil-fired boilers}
-	\label{fig:Heat_vs_time}
+	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyGeneration}
+	\caption{Energy generation, aggregated data for one year of operations}
+	\label{fig:Pie_EnergyGeneration}
 \end{figure}
 
-Estimated heat demand
-Maximum heat demand 
-Minimum heat demand
+From Figure \ref{fig:Pie_EnergyDemand} it can be noted that nearly 50\% of the total energy demand is related to ship propulsion, while the remaining portion is approximately equally split between electric energy and heat demand.
+
+Within the electric energy demand, the HVAC systems only represent a minor contribution (XX\%), as it is only present for a few months during summer, and at rather low power demand.
+This is not surprising, since it can be observed that (see Figure \ref{fig:TseaTairDIST}) the air is always below 27$^o$C, and rarely above 17$^o$C.
+Also the thrusters, although they represent a rather high instantaneous demand, they are only used for a very short time and, hence, their total contribution to the ship energy demand is limited to XX\%.
+As a consequence, most of the electric energy demand should be allocated to other consumers, and efforts should be directed towards providing additional measurements of how different users contribute to the total electric power demand.
+
+\subsection{Heat demand and generation} \label{sec:res:heat}
+
+FIGURE --> Pie, heat generated by source
+FIGURE --> filled lines, heat generated by source.
+
+Looking at Figure \ref{fig:HeatGen_hist} it is possible to appreciate the amount of heat that is recovered from on board available waste heat (cooling water and exhaust gas).
+It can be observed that the amount of heat recovered from the HT cooling systems is much lower than expected, with an average of around 350 kW.
+This value is quite unreasonable compared to the values found in the technical specifications from the yard, that go from 700 kW (when only one auxiliary engine is operated) to 4800 kW (when all main engines and two auxiliary engines are operated).
+\\
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/HeatGen_hist}
+	\caption{Distribution of thermal power generated from on board waste heat recovery}
+	\label{fig:Hist_WHR}
+\end{figure}
+
+Regarding the heat demand onboard, it can be seen that the HVAC re-heater represents the largest contributor (XX\%) followed by the hot water heater (XX\%).
+The large contribution of the HVAC reheater can be explained by the former being important for long times and in relation to the low ambient temperatures encountered by the case study vessel during operation.
+The relatively large contribution of the hot water heater can instead be related to the fact that these systems are rather important consumers and are necessary during the whole ship operations (i.e.
+also during the summer months).
 
@@
 
-\section{Discussion}
-\label{sec:discussion}
+The results of the exergy analysis of the ship energy systems are shown in Figure REF and Tables REF to REF.
+
+Table(s) --> Exergy efficiencies
+- General at system level
+- One per system (MEs, AEs, CoolingSystems, SteamSystems)
+
+Figure --> Grassman diagram
+
+\subsection{Typical operational days}  \label{sec:res:typicalDays}
 
@@
 
+\section{Full tables of energy and exergy flows}
+
+\section{Full equation list}
+
+\begin{table}[htbp]
+	\centering
+	\begin{tabular}{p{4cm}l}
+		\hline
+		\multirow{2}{*}{Before compressor}		& $ T_{comp,in} = $ Measured \\
+												& $ p_{comp,in} = $ Measured \\
+		\multirow{3}{*}{After compressor}		& $ T_{comp,out} = T_{comp,in} + \frac{T_{comp,out,is} - T_{comp,in} }{\eta_{comp,is}}  $  \\
+												& $ p_{comp,out} = $ Measured \\
+												& $ \eta_{comp,is} = P_2(\frac{p_{comp,out}}{p_{comp,in}}) $ \\
+		\multirow{3}{*}{After CAC-HT stage}		& $ p_{CAC-HT,out} = p_{comp,out} $ \\
+												& $ T_{CAC-HT,out} = T_{comp,out} - \frac{\dot{Q}_{CAC,HT}}{\dot{m}_{air,cyl}} $ \\
+												& $ \dot{Q}_{CAC,HT} = P_2(\lambda) $ \\
+		\multirow{3}{*}{After CAC-LT stage}		& $ p_{CAC-LT,out} = p_{comp,out} $ \\
+												& $ T_{CAC-LT,out} = T_{CAC-HT,out} - \frac{\dot{Q}_{CAC,LT}}{\dot{m}_{air,cyl}} $ \\
+												& $ \dot{Q}_{CAC,LT} = P_2(\lambda) $ \\
+		\multirow{3}{*}{After CAC-LT stage}
+
+	\end{tabular}
+\end{table}
+
 \section{Estimation of air and exhaust gas flows in the main engines}
 
-\section{Estimation of air and exhaust gas flows in the auxiliary engines}
+\subsection{Mass balances}
+
+Mass balance on the split after the compressor, with one flow going to the engine inlet valves and the other flow going to the bypass valve
+\begin{equation}
+\dot{m}_{air,comp} = \dot{m}_{air,cyl} + \dot{m}_{air,bypass} \\
+\end{equation}
+
+Mass balance on the mixer before the turbine, with one flow coming from the engine exhaust valves and the other flow coming from the bypass valve
+\begin{equation}
+\dot{m}_{eg,turb} = \dot{m}_{eg,cyl} + \dot{m}_{air,bypass} \\
+\end{equation}
+
+Note that the exhaust gas mass flow from the cylinders is fully defined as follows:
+\begin{equation}
+\dot{m}_{eg,cyl} = \dot{m}_{air,cyl} + \dot{m}_{fuel,cyl} \\
+\end{equation}
+
+\subsection{Energy balances}
+
+Energy balance on the turbocharger: the power generated by the turbine must be equal to the power absorbed by the compressor
+\begin{equation}
+\dot{m}_{air,comp} \Delta h_{comp} = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
+\end{equation}
+
+Energy balance on the mixer between the air flow after the bypass valve and the exhaust gas flow after the exhaust valve
+\begin{equation}
+\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0)
+\end{equation}
+
+The system of four equations is to be solved in the four following unknowns:
+\begin{itemize}
+	\item $ \dot{m}_{eg,turb} $
+	\item $ \dot{m}_{air,bypass} $
+	\item $ \dot{m}_{air,comp} $
+	\item $ T_{turb,in} $ (In the case of the Main engines)
+	\item $ T_{cyl,out} $ (In the case of the Auxiliary engines)
+\end{itemize}
+
+Include other assumptions:
+- TC efficiency
+- Compressor efficiency
+- Turbine efficiency
+
+\subsection{Main Engines}
+
+To calculate what we need, we must make the system explicit on the variables we need to calculate.
+This process is different for the main and auxiliary engines, mostly because the temperature measurement of the exhaust gas before the turbine is positioned differently.
+In the case of the main engines, the temperature at the \textbf{cylinder outlet} is measured and, hence, known.
+
+First, we make all equations explicit in $ \dot{m}_{air,bypass} $ thus eliminating both 
+$ \dot{m}_{eg,turb} $ and $ \dot{m}_{air,comp} $
+
+$$
+(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech} 
+$$
+$$
+\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_0)
+$$
+
+At this point, we use the first equation to simplify the system and make it explicit in the $ \dot{m}_{air,bypass} $ variable alone:
+
+\begin{eqnarray*}
+	(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}  & = & (\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} \\
+	T_{turb,in} - T_{turb,out} & = & \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} \\
+	T_{turb,in} & = & T_{turb,out} + \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}}  \\
+\end{eqnarray*}
+
+We can now substitute this expression in the second equation:
+
+\begin{multline}
+\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\ 
+(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,out} + \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} - T_0)
+\end{multline}
+
+We first execute the multiplication in the term on the right of the equal:
+\begin{multline}
+\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\
+\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_0) + \\
+\dot{m}_{air,bypass} c_{p,eg} (T_{turb,out} - T_0) + \\ 
+\frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}}(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) c_{p,eg}
+\end{multline}
+
+Now, we can group together the elements in the variable of interest:
+\begin{multline}
+\dot{m}_{air,bypass} \left[ c_{p,air} (T_{comp,out} - T_0) - c_{p,eg} (T_{turb,out} - T_0) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
+\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_0) - \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
+\end{multline}
+
+This can be further simplified into:
+\begin{multline}
+\dot{m}_{air,bypass} \left[ c_{p,air} (T_{comp,out} - T_0) - c_{p,eg} (T_{turb,out} - T_0) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
+\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
+\end{multline}
+
+Finally, we can assume the specific heat at constant pressure of air and exhaust gas to be sufficiently similar for considering them equal in the term at the numerator (we ignored changes in the properties of the mixture in other points as well):
+\begin{multline}
+\dot{m}_{air,bypass} \left[ c_{p,eg} (T_{comp,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
+\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
+\end{multline}
+
+From which we can derive the final form:
+\begin{equation}
+\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}}{c_{p,eg} (T_{comp,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}}
+\end{equation}
+
+Which can be rewritten, in order to have both numerator and denominator positive, as follows:
+\begin{equation}
+\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{cyl,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}}{c_{p,eg} (T_{turb,out} - T_{comp,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}}
+\end{equation}
+
+\subsection{Auxiliary Engines}
+
+In the case of the auxiliary engines, what we know instead is the temperature of the exhaust gas \textbf{before the turbine}, hence after the merging between the exhaust gas and the bypass.
+
+In this case, however, things appear to be simple: the energy balance on the turbocharger can be calculated as there is only one variable unknown: the total mass flow in the compressor, and hence the bypass flow:
+$$
+(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
+$$
+$$
+\dot{m}_{air,cyl} \Delta h_{comp} + \dot{m}_{air,bypass} \Delta h_{comp} = \dot{m}_{eg,cyl} c_{p,eg} (T_{turb,in} - T_{turb,out}) + \dot{m}_{air,bypass} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
+$$
+$$
+\dot{m}_{air,bypass} (\Delta h_{comp} - c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}) = \dot{m}_{eg,cyl} c_{p,eg} (T_{turb,in} - T_{turb,out}) - \dot{m}_{air,cyl} \Delta h_{comp}
+$$
+\begin{equation}
+\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech} - \dot{m}_{air,cyl} \Delta h_{comp}}{\Delta h_{comp} - c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}}
+\end{equation}
+
+Now that we have the value for the bypass flow, it is easy to calculate backwards the temperature of the exhaust gas after the cylinder outlet valves, before the merging with the bypass flow:
+
+$$
+\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0)
+$$
+$$
+\dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0) - \dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0)
+$$
+\begin{equation}
+T_{cyl,out} = T_0 + \frac{\dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0) - \dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0)}{\dot{m}_{eg,cyl} c_{p,eg}}
+\end{equation}
+
+\section{Full list of available measurements}
 

### 2017-10-10 Main.py

@@ -25,7 +25,7 @@
 
 
 do_processed_data_preparation = "no"
-do_data_processing = "yes"
+do_data_processing = "no"
 
 
 
@@ -113,31 +113,27 @@ elif do_data_processing == "yes":
     # Calculating the central cooling systems
     processed = cs.centralCoolingSystems(processed, CONSTANTS)
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-3")
-    # Assigning defined values to all flows for engines off
+    processed = cs.seaWaterCoolers(processed, CONSTANTS, dict_structure)
+    # Calculating energy and exergy properties
     processed = ea.energyAnalysisLauncher(processed, dict_structure, CONSTANTS)
     # Re-doing the calculation of the connected points
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-4")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Demands-2")
-
+    # Assigning the operational mode
+    processed = ppo.operationalModeCalculator(processed, dataset_raw, CONSTANTS, header_names)
     # Saving the processed data
     processed.to_hdf(CONSTANTS["filenames"]["dataset_output"], "processed", format='fixed', mode='w')
-
-
-######################################
-## RESULTS CHECK                	##
-######################################
-cc.systemCheck(processed, CONSTANTS, dict_structure, dataset_raw)
-######################################
-## EXPORT EFFICIENCIES                	##
-######################################
+    # Result check log
+    cc.systemCheck(processed, CONSTANTS, dict_structure, dataset_raw)
+    # Efficiency results export
 ex.exportEfficiecies(processed, CONSTANTS, dict_structure)
 
 #%%
 ######################################
-## PLOTTING	##
+## POSTPROCESSING	##
 ######################################
 
-plot.predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure, ["TimeSeries:Heat_vs_time"])
+# plot.predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure,["Pie:OperationalMode" ])
 # plot.plotMain("prompt", dict_structure, processed)
 
 

### 2017-10-10 Python files/Energyanalysis.py

@@ -9,19 +9,22 @@ import os
 
 
 def energyAnalysisLauncher(processed, dict_structure, CONSTANTS):
-    processed = propertyCalculator(processed, dict_structure, CONSTANTS)  # Calculates the energy and exergy flows
+    processed = propertyCalculator(processed, dict_structure, CONSTANTS, dict_structure["systems"])  # Calculates the energy and exergy flows
     processed = efficiencyCalculator(processed, dict_structure, CONSTANTS)
     return processed
 
 
 
 
-def propertyCalculator(processed, dict_structure, CONSTANTS):
+def propertyCalculator(processed, dict_structure, CONSTANTS, system_list):
     print("Started calculating flow properties...", end="", flush=True)
     df_index = processed.index
-    for system in dict_structure["systems"]:
+    for system in system_list:
         for unit in dict_structure["systems"][system]["units"]:
             for flow in dict_structure["systems"][system]["units"][unit]["flows"]:
+                if system+":"+unit+":"+flow == "CoolingSystems:SWC13:SeaWater_out":
+                    aaa = 0
+                #  The calculation of the properties is done if and only if the exergy flow is completely empty
                 if processed[d2df(system,unit,flow,"Bdot")].isnull().sum() == len(processed[d2df(system,unit,flow,"Bdot")]):
                     if dict_structure["systems"][system]["units"][unit]["flows"][flow]["type"] == "CPF":
                         # Only doing the calculations if the values for p and T are not NaN
@@ -87,6 +90,11 @@ def propertyCalculator(processed, dict_structure, CONSTANTS):
                     elif dict_structure["systems"][system]["units"][unit]["flows"][flow]["type"] == "IPF":
                         if processed[d2df(system,unit,flow,"T")].isnull().sum() != len(processed[d2df(system,unit,flow,"T")]):
                             if "Water" in flow:
+                                if flow == "SeaWater_in":
+                                    dh = pd.Series(index = df_index)
+                                    ds = pd.Series(index = df_index)
+                                    dh[:] = 0
+                                    ds[:] = 0
                                 dh = CONSTANTS["General"]["CP_WATER"] * (processed[d2df(system, unit, flow, "T")] - processed["T_0"])
[…]

### 2017-10-10 Python files/auxiliaryDemand.py

@@ -107,7 +107,7 @@ def heatDemand(processed, CONSTANTS, dict_structure):
     # HVAC Reheater
     processed["Demands:Heat:HVACreheater:Edot"] = CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["DESIGN"]["HVAC_REHEATER"] * (
         CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["LIN"]["HVAC_REHEATER"] * processed["Demands:Electricity:HVAC:Edot"] / processed["Demands:Electricity:HVAC:Edot"].max())
-    processed["Demands:Heat:HVACreheater:Edot"][processed["Demands:Electricity:HVAC:Edot"] <= 0] = 0
+    processed.loc[processed["Demands:Electricity:HVAC:Edot"] <= 0,"Demands:Heat:HVACreheater:Edot"] = 0
     # Hot water heater
     processed["Demands:Heat:HotWaterHeater:Edot"] = (CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["DESIGN"]["HOT_WATER_HEATER"] * (
         CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["LIN"]["HOT_WATER_HEATER"] * processed["Passengers_calc"] / 1800 * (
@@ -116,7 +116,7 @@ def heatDemand(processed, CONSTANTS, dict_structure):
         CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["DESIGN"]["OTHER_HTHR"])
     # HVAC Preheater
     temp = CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["DESIGN"]["HVAC_PREHEATER"] * (
-        CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["LIN"]["HVAC_PREHEATER"] * (297 - processed["T_air"]) / (297 - 248)) - 0.100 * processed["Passengers_calc"]
+        CONSTANTS["OtherUnits"]["HEAT_DEMAND"]["LIN"]["HVAC_PREHEATER"] * (297 - processed["T_air"]) / (297 - 248))  - 0.100 * processed["Passengers_calc"]
     temp[processed["Demands:Electricity:HVAC:Edot"] > 20] = 0  # The preheater is never used when the AC compressors are on
     processed["Demands:Heat:HVACpreheater:Edot"] = temp  # The pre-heater is not working during summer
     # Galley
@@ -151,6 +151,7 @@ def heatDemand(processed, CONSTANTS, dict_structure):
[…]

### 2017-10-10 Python files/coolingsystems.py

@@ -2,6 +2,7 @@ import numpy as np
 import pandas as pd
 from helpers import d2df
 from helpers import polyvalHelperFunction
+import fillerfunctions as ff
 
 
 def engineCoolingSystemsCalculation(processed, CONSTANTS, engine_type):
@@ -9,14 +10,11 @@ def engineCoolingSystemsCalculation(processed, CONSTANTS, engine_type):
     # This function calculates the different flows related to the cooling systems of the main engines.
     for system in CONSTANTS["General"]["NAMES"][engine_type]:
         # Calculating the total energy flow going to the cooling systems, based on the energy balance on the engine
-        energy_2_cooling = (processed[d2df(system,"Cyl","FuelCh_in","Edot")] -
-            processed[d2df(system,"Cyl","Power_out","Edot")] +
-            CONSTANTS["General"]["HFO"]["CP"] * processed[d2df(system,"Cyl","FuelPh_in","mdot")] *
-                (processed[d2df(system,"Cyl","FuelPh_in","T")] - processed["T_0"]) -
-            CONSTANTS["General"]["CP_EG"] * processed[d2df(system,"Turbine","Mix_in","mdot")] *
-                (processed[d2df(system,"Turbine","Mix_out","T")] - processed["T_0"]) +
-            CONSTANTS["General"]["CP_AIR"] * processed[d2df(system,"Comp","Air_out","mdot")] *
-                (processed[d2df(system,"Comp","Air_in","T")] - processed["T_0"]))
+        energy_2_cooling = (processed[d2df(system,"Cyl","FuelCh_in","Edot")] +
+            CONSTANTS["General"]["HFO"]["CP"] * processed[d2df(system,"Cyl","FuelPh_in","mdot")] * (processed[d2df(system,"Cyl","FuelPh_in","T")] - processed["T_0"]) -
+            processed[d2df(system, "Cyl", "Power_out", "Edot")] -
+            processed[d2df(system,"Turbine","Mix_out","Edot")] +
+            processed[d2df(system,"Comp","Air_in","Edot")])
         # Calculating the energy going to the charge air cooler, based on the estimated temperatures on the air line
[…]

### 2017-10-10 Python files/fillerfunctions.py

@@ -72,18 +72,16 @@ def unitOffCheck(processed, dict_structure, CONSTANTS, call_ID):
 
 
 def connectionAssignment(processed, dict_structure, CONSTANTS, system, unit, call_ID):
-    text_file = open(CONSTANTS["filenames"]["consistency_check_report"], "a")
+    text_file = open(CONSTANTS["filenames"]["consistency_check_report"], "a")  # Opens the log file
     counter = 0
     for flow in dict_structure["systems"][system]["units"][unit]["flows"]:
-        if system+":"+unit+":"+flow == "ME1:HTmerge:LTWater_in":
-            a = 0
         if "Connections" in dict_structure["systems"][system]["units"][unit]["flows"][flow]:
             for connected_flow in dict_structure["systems"][system]["units"][unit]["flows"][flow]["Connections"]:
                 for property in CONSTANTS["General"]["PROPERTY_LIST"][dict_structure["systems"][system]["units"][unit]["flows"][flow]["type"]]:
-                    ID = d2df(system,unit,flow,property)
-                    ID_c = connected_flow + ":" + property
-                    if processed[ID].isnull().sum() != len(processed[ID]):
-                        if (processed[ID_c].isnull().sum() == len(processed[ID_c])) or (sum(processed[ID_c]) == 0):
+                    ID = d2df(system,unit,flow,property)  # Writing the property full ID of the flow we are checking
+                    ID_c = connected_flow + ":" + property  # Writing the property full ID of the flow connected to the flow we are checking
+                    if processed[ID].isnull().sum() != len(processed[ID]):  # If the flow is not empty
+                        if (processed[ID_c].isnull().sum() == len(processed[ID_c])) or (sum(processed[ID_c]) == 0):  # If the connected flow IS empty
                             processed[ID_c] = processed[ID]
                             counter = counter + 1
                         elif (processed[ID] - processed[ID_c]).sum() > processed[ID].max():
@@ -99,7 +97,7 @@ def massFill(processed, dict_structure, CONSTANTS, system, unit, streams, call_I
     # This function applies the mass balance over one component:
     # - If some of the flows have not been assigned yet, they are calculated based on the mass balance
     counter = 0
-    if system+":"+unit == "ME1:HTmerge":
+    if system+":"+unit == "CoolingSystems:LTcollector13":
         aaa = 0
     for stream in streams:
         # If there is only one flow, something is tricky...

### 2017-10-10 Python files/plotting.py

@@ -1,5 +1,6 @@
 import matplotlib.pyplot as plt
 from helpers import d2df
+import pandas as pd
 
 def plotMain(type, dict_structure, processed, *args):
     # This function plots the contents of the analysed files.
@@ -214,7 +215,7 @@ def plottingFunction(plot_info, processed, data_type):
             else:
                 fig = plt.figure()
                 for plot in figure["variables"]:
-                    x = data[plot]
+                    x = processed[plot]
                     if figure["plot_mode"] == "hist":
                         num_bins = 50
                         # the histogram of the data
@@ -233,6 +234,9 @@ def plottingFunction(plot_info, processed, data_type):
 def predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure, filenames):
     for filename in filenames:
         plt.figure()
+
+        ### TIME SERIES ###
+
         if filename == "TimeSeries:Heat_vs_time":
             # Contribution from the HRSGs
             temp = (processed["ME2:HRSG:Steam_in:mdot"] * CONSTANTS["Steam"]["DH_STEAM"] + processed["ME3:HRSG:Steam_in:mdot"] * CONSTANTS["Steam"]["DH_STEAM"] + processed["AE1:HRSG:Steam_in:mdot"] * CONSTANTS["Steam"]["DH_STEAM"] + processed["AE2:HRSG:Steam_in:mdot"] * CONSTANTS["Steam"]["DH_STEAM"] + processed["AE3:HRSG:Steam_in:mdot"] * CONSTANTS["Steam"]["DH_STEAM"] + processed["AE4:HRSG:Steam_in:mdot"] * CONSTANTS["Steam"]["DH_STEAM"])
@@ -258,26 +262,7 @@ def predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure, filenames
             plt.plot(boilers_calculated, 'r:', label="Boilers (C)")
             plt.plot(total, 'g-', label='Total heating demand')
             plt.legend()
-        if filename == "Pie:TotalEnergy":
-            quantities = [processed["Demands:Mechanical:Total:Edot"].sum() , processed["Demands:Electricity:Total:Edot"].sum() , processed["Demands:Heat:Total:Edot"].sum()]
-            labels = ["Mechanical Power" , "Electric Power" , "Thermal Power"]
-            explode = (0.05 , 0.05 , 0.05)
[…]

### 2017-10-10 Python files/preprocessingAE.py

@@ -3,6 +3,7 @@ import pandas as pd
 import fillerfunctions as ff
 import preprocessingO as ppo
 import coolingsystems as cs
+from energyanalysis import propertyCalculator
 from helpers import d2df
 
 def auxEngineProcessing(raw, processed, dict_structure, CONSTANTS, hd):
@@ -24,6 +25,8 @@ def auxEngineProcessing(raw, processed, dict_structure, CONSTANTS, hd):
     processed = auxEngineAirFlowCalculation(raw, processed, CONSTANTS)
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-2.1")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-2.2")
+    processed = propertyCalculator(processed, dict_structure, CONSTANTS, ["AE1", "AE2", "AE3", "AE4"])
+    processed = auxEngineAirFlowPostCalculation(processed, CONSTANTS)
     # Calculating cooling flows
     processed = cs.engineCoolingSystemsCalculation(processed, CONSTANTS, "AuxEngines")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "AuxEngines", "AE-3.1")
@@ -143,26 +146,24 @@ def auxEngineAirFlowCalculation(raw, processed, CONSTANTS):
             processed[d2df(system,"BPmerge","BP_in","mdot")] * CONSTANTS["General"]["CP_AIR"]) * processed[d2df(system,"Turbine","Mix_in","T")] -
             processed[d2df(system,"BPmerge","BP_in","mdot")] * CONSTANTS["General"]["CP_AIR"] * processed[d2df(system, "Comp", "Air_out", "T")]) / (
             processed[d2df(system,"Cyl","EG_out","mdot")] * CONSTANTS["General"]["CP_EG"])
-        processed[system + ":CP_MIX"] = (processed[d2df(system, "BPmerge", "BP_in", "mdot")] * CONSTANTS["General"]["CP_AIR"] +
-            processed[d2df(system, "Cyl", "EG_out", "mdot")] * CONSTANTS["General"]["CP_EG"]) / processed[d2df(system, "BPmerge", "Mix_out", "mdot")]
+        # Assiging the mass flow values for the HRSGs otherwise it makes a mess
+        processed[d2df(system, "HRSG", "Mix_in", "mdot")] = processed[d2df(system, "BPmerge", "Mix_out", "mdot")]
[…]

### 2017-10-10 Python files/preprocessingME.py

@@ -4,6 +4,7 @@ import fillerfunctions as ff
 import coolingsystems as cs
 import preprocessingO as ppo
 import CoolProp.CoolProp as cp
+from energyanalysis import propertyCalculator
 from helpers import d2df
 from helpers import polyvalHelperFunction
 
@@ -26,6 +27,8 @@ def mainEngineProcessing(raw, processed, dict_structure, CONSTANTS, hd):
     processed = mainEngineAirFlowCalculation(raw, processed, dict_structure, CONSTANTS)
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-2.1")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-2.2")
+    processed = propertyCalculator(processed, dict_structure, CONSTANTS, ["ME1", "ME2", "ME3", "ME4"])
+    processed = mainEngineAirFlowPostCalculation(processed, dict_structure, CONSTANTS)
     # Calculating cooling flows
     processed = cs.engineCoolingSystemsCalculation(processed, CONSTANTS, "MainEngines")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-3.1")
@@ -171,26 +174,24 @@ def mainEngineAirFlowCalculation(raw, processed, dict_structure, CONSTANTS):
         processed[d2df(system, "BPsplit", "Air_in", "mdot")] = processed[d2df(system,"BPsplit","BP_out","mdot")] + processed[d2df(system,"Cyl","Air_in","mdot")]
         # The flow through the turbine is equal to the sum of the bypass flow and the exhaust coming from the cylinders
         processed[d2df(system,"BPmerge","Mix_out","mdot")] = processed[d2df(system,"BPsplit","BP_out","mdot")] + processed[d2df(system,"Cyl","EG_out","mdot")]
-        processed[system+":CP_MIX"] = (processed[d2df(system,"BPsplit","BP_out","mdot")] * CONSTANTS["General"]["CP_AIR"] +
-            processed[d2df(system, "Cyl", "EG_out", "mdot")] * CONSTANTS["General"]["CP_EG"]) / processed[d2df(system,"BPmerge","Mix_out","mdot")]
+        # Assiging the mass flow values for the HRSGs otherwise it makes a mess
+        if system in {"ME2", "ME3"}:
[…]

### 2017-10-10 Python files/preprocessingO.py

@@ -5,6 +5,17 @@ from helpers import polyvalHelperFunction
 import CoolProp.CoolProp as cp
 
 
+def operationalModeCalculator(processed, raw, CONSTANTS, hd):
+    # Initializing the data series
+    temp = pd.Series(index=processed.index)
+    temp[raw[hd["SHIP_SPEED_KNOT_"]] > 15] = "High Speed Sailing"
+    temp[(raw[hd["SHIP_SPEED_KNOT_"]] < 15) * (raw[hd["SHIP_SPEED_KNOT_"]] > 4)] = "Low Speed Sailing"
+    temp[((processed["Demands:Electricity:Thrusters:Edot"] > 0) + ((raw[hd["SHIP_SPEED_KNOT_"]] < 4) * (raw[hd["SHIP_SPEED_KNOT_"]] > 2)))] = "Maneuvering"
+    temp[temp.isnull()] = "Port/Stay"
+    processed["operationalMode"] = pd.Series(temp.values)
+    return processed
+
+
 def engineStatusCalculation(type, raw, processed, CONSTANTS, hd):
     for system in CONSTANTS["General"]["NAMES"][type]:
         processed[system + ":" + "on"] = raw[hd[system+"-TC__RPM_"]] > 5000

### 2017-10-10 Python files/psoTuning2.py

@@ -143,6 +143,7 @@ def fitnessFunction(param, *args):
 
     Qdot_ab = pd.Series(index=db_index)
     Qdot_ab[:] = 0
+
     Qdot_deficit = -Qdot_steam + Qdot_hrsg
     Qcumul_deficit = Qdot_deficit.cumsum()
     total_deficit = Qcumul_deficit[db_index[-1]]

### 2017-10-10 Python files/psoTuning3.py

@@ -19,15 +19,21 @@ def preparation():
     dataset_raw = pd.read_hdf(filenames["dataset_raw"], 'table')
     data = pd.DataFrame(index=processed_temp.index)
     indexes_2_save_processed = {"Passengers_calc", "Demands:Electricity:HVAC:Edot", "T_air",
-                      "AE1:Cyl:FuelPh_in:mdot", "AE2:Cyl:FuelPh_in:mdot", "AE3:Cyl:FuelPh_in:mdot", "AE4:Cyl:FuelPh_in:mdot",
-                      "ME1:Cyl:FuelPh_in:mdot", "ME2:Cyl:FuelPh_in:mdot", "ME3:Cyl:FuelPh_in:mdot","ME4:Cyl:FuelPh_in:mdot",
-                      "ME2:HRSG:Steam_in:mdot", "ME3:HRSG:Steam_in:mdot",
-                      "AE1:HRSG:Steam_in:mdot", "AE2:HRSG:Steam_in:mdot", "AE3:HRSG:Steam_in:mdot", "AE4:HRSG:Steam_in:mdot",
-                      "CoolingSystems:HTcollector24:HTWater_out:mdot", "CoolingSystems:HTcollector13:HTWater_out:mdot",
-                      "CoolingSystems:HTcollector24:HTWater_out:T", "CoolingSystems:HTcollector13:HTWater_out:T",
-                      "ME1:on", "ME2:on", "ME3:on", "ME4:on", "AE1:on", "AE2:on", "AE3:on", "AE4:on",
-                      }
-    indexes_2_save_raw = {"SBO 1 OUTLET PRESSUR:7313:bar:Average:900", "SBO 2 OUTLET PRESSUR:7333:bar:Average:900", "Boiler_Port", "Boiler_starbord"}
+                                "AE1:Cyl:FuelPh_in:mdot", "AE2:Cyl:FuelPh_in:mdot", "AE3:Cyl:FuelPh_in:mdot",
+                                "AE4:Cyl:FuelPh_in:mdot",
+                                "ME1:Cyl:FuelPh_in:mdot", "ME2:Cyl:FuelPh_in:mdot", "ME3:Cyl:FuelPh_in:mdot",
+                                "ME4:Cyl:FuelPh_in:mdot",
+                                "ME2:HRSG:Steam_in:mdot", "ME3:HRSG:Steam_in:mdot",
[…]

### 2017-10-10 Python files/unitstructures.py

@@ -194,6 +194,7 @@ def flowStructure():
                 "LTWater_AE3_in": {"type": "IPF"},
                 "LTWater_ME1_in": {"type": "IPF"},
                 "LTWater_ME3_in": {"type": "IPF"},
+                "LTWater_AB1_in": {"type": "IPF"},
                 "LTWater_out": {"type": "IPF"}, "HTWater_out": {"type": "IPF"}}
             structure["systems"][system]["units"]["LTcollector13"]["equations"] = ["MassBalance"]
             structure["systems"][system]["units"]["LTcollector24"]["flows"] = {
@@ -209,7 +210,8 @@ def flowStructure():
                 "LTWater_AE3_out": {"type": "IPF"},
                 "LTWater_AE1_out": {"type": "IPF"},
                 "LTWater_ME1_out": {"type": "IPF"},
-                "LTWater_ME3_out": {"type": "IPF"}}
+                "LTWater_ME3_out": {"type": "IPF"},
+                "LTWater_AB1_out": {"type": "IPF"}}
             structure["systems"][system]["units"]["LTdistribution13"]["equations"] = ["MassBalance"]
             structure["systems"][system]["units"]["LTdistribution24"]["flows"] = {
                 "LTWater_in": {"type": "IPF"},
@@ -311,7 +313,7 @@ def flowStructure():
 
 
         elif system == "Steam":
-            structure["systems"][system]["units"] = {"Boiler1": {}, "Boiler2": {}, "SteamCollector": {}, "SteamDistribution": {},
+            structure["systems"][system]["units"] = {"Boiler1": {}, "Boiler2": {}, "HotWell": {}, "SteamDistribution": {},
                                                      "TankHeating": {}, "OtherTanks": {}, "HFOtankHeating": {}, "MachinerySpaceHeaters": {},
                                                      "HFOheater": {}, "Galley": {}}
             # Auxiliary boiler
@@ -347,10 +349,11 @@ def flowStructure():
                 "Steam_HH_out": {"type": "SF", "state": "SV"},
                 "Steam_G_out": {"type": "SF", "state": "SV"},
                 "Steam_SH_out": {"type": "SF", "state": "SV"},
-                "Steam_OT_out": {"type": "SF", "state": "SV"}}
[…]

### 2017-10-13 Manuscript/Manuscript.tex
Other small fixes (we are almost there!)

@@
 
-The setup of the engines and their connections to the other ship systems are summarized in Figure REF and REF for the main and auxiliary engines, respectively.
-The defining net energy flows are listed in Table REF.
+The setup of the engines and their connections to the other ship systems are summarized in Figure \ref{fig:DieselEngines} for the main and auxiliary engines, respectively.
+The most relevant operative values are provided in Tables \ref{tab:ME_values} and \ref{tab:AE_values} and the main energy flows are listed in Tables \ref{tab:ME_flows} and \ref{tab:AE_flows}.
 It should be noted that, in the case of the main engines, the engine power output was not measured and needed to be estimated.
@@
 
-The validity of this assumption can be seen by observing Figure REF, representing the calculated total power from the main engines as a function of the ship speed
+The validity of this assumption can be seen by observing Figures \ref{fig:Pme_vs_vship_Neng} and \ref{fig:Pme_vs_TCspeed} where the calculated  power from the main engines is represented against the ship speed and the turbocharger speed, respectively.
 
@@
 
-ALSO INCLUDE THE POWER VS TC SPEED RELATIONSHIP
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Scatter_Pme_vs_TCspeed}
+	\caption{Scatter plot, Main engine power versus turbocharger speed (ME1)}
+	\label{fig:Pme_vs_TCspeed}
+\end{figure}
 
@@
 	\centering
-	\begin{tabular}{llccrrrrr}
-		\hline
-		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
-		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
-		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
-		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{comp}$		& - 	& M & & & & &  \\
-		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
-		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
-		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{exp}$		& - 	& C & & & & &  \\
-		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
-		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
-		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
-		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
-		&	$T_{eg,out}$  		& $^o$C & M  & & & & &  \\ 
-		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
-		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
-		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
-		\multicolumn{2}{l}{LOC} & & & & & & &  \\
-		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
-		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		\multicolumn{2}{l}{JWC} & & & & & & &  \\
-		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
+	\begin{tabular}{llccrrrr}
+		\hline
+		\multicolumn{2}{l}{Variable} & Unit & M/C & 30\% & 50\% & 70\% & 90\% \\
+		\hline
+		\multicolumn{2}{l}{Compressor} & & & & & & \\ 
+		& 	$T_{air,in}$  		& $^o$C & M & \multicolumn{4}{c}{302} \\ 
+		&	$T_{air,out}$  		& $^o$C & C  & 356.0 & 409.4 & 434.1 & 461.7 \\ 
+		&	$\dot{m}_{air}$  	& $kg/s$ & C  & 4.24 & 7.62 & 8.51 & 10.28 \\
+		&   $\beta_{comp}$		& - 	& M & & & &  \\
+		\multicolumn{2}{l}{Turbine} & & & & & &  \\
+		&	$T_{eg,in}$   		& $^o$C & C  & 698.8 & 702.2 & 763.3 & 786.0 \\ 
+		&	$T_{eg,out}$  		& $^o$C & M  & 639.3 & 592.9 & 633.2 & 631.0 \\ 
+		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & 4.35 & 7.79 & 8.73 & 10.57 \\
+		&   $\beta_{exp}$		& - 	& C & & & &  \\
+		\multicolumn{2}{l}{Bypass valve} & & & & & &  \\
+		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & 0.98 & 2.05 & 1.02 & 0.39 \\  
+		\multicolumn{2}{l}{Cylinders} & & & & & &   \\
+		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & 3.25 & 5.57 & 7.49 & 9.89 \\
+		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & 0.11 & 0.17 & 0.22 & 0.28 \\ 
+		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & 3.37 & 5.74 & 7.71 & 10.17 \\
+		&	$T_{air,in}$  		& $^o$C & M  & 323.7 & 323.2 & 323.1 & 323.5 \\ 
+		&	$T_{eg,out}$  		& $^o$C & M  & 698.8 & 702.2 & 763.3 & 786.0 \\ 
+		\multicolumn{2}{l}{CAC-LT stage} & & & & & &  \\
+		&	$T_{air,in}$  		& $^o$C & C  & 355.2 & 350.1 & 360.0 & 368.7 \\ 
+		&	$T_{air,out}$ 		& $^o$C & M  & 323.7 & 323.2 & 323.1 & 323.5 \\
+		&	$T_{w,in}$   		& $^o$C & M  & 309.3 & 309.6 & 309.5 & 309.6 \\
+		&	$T_{w,out}$   		& $^o$C & C  & 310.4 & 311.0 & 311.8 & 313.0 \\  
+		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & C& 23.36 & 26.09 & 29.1 & 31.53 \\
+		\multicolumn{2}{l}{CAC-HT stage} & & & & & & \\
+		&	$T_{air,in}$   		& $^o$C & C  & 356.0 & 409.4 & 434.1 & 461.7 \\
+		&	$T_{air,out}$   	& $^o$C & C  & 355.2 & 350.1 & 360.0 & 368.7 \\ 
+		&	$T_{w,in}$    		& $^o$C & C  & 360.8 & 360.2 & 359.6 & 358.7 \\ 
+		&	$T_{w,out}$   		& $^o$C & C  & 360.8 & 362.6 & 363.6 & 365.4 \\
+		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & 33.21 & 33.18 & 33.33 & 33.33 \\
+		\multicolumn{2}{l}{LOC} & & & & & &  \\
+		&	$T_{lo,in}$   		& $^o$C & C  &346.6 & 348.9 & 349.8 & 350.5 \\ 
+		&	$T_{lo,out}$    	& $^o$C & M  & 335.1 & 336.8 & 338.1 & 338.3 \\ 
+		&	$T_{w,in}$    		& $^o$C & C  & 310.4 & 311.0 & 311.8 & 313.0 \\
+		&	$T_{w,out}$   		& $^o$C & C  & 317.4 & 317.6 & 317.5 & 318.5 \\ 
+		\multicolumn{2}{l}{JWC} & & & & & &  \\
+		&	$T_{wall}$    		& $^o$C & C  & \multicolumn{4}{c}{423}  \\ 
+		&	$T_{w,in}$    		& $^o$C & M  & 358.1 & 358.4 & 357.6 & 356.4 \\ 
+		&	$T_{w,out}$	   		& $^o$C & C  & 360.8 & 360.2 & 359.6 & 358.7 \\ 
 		\hline
@@
 	\caption{Main engines, measured and calculated temperatures and flows at different engine loads}
+	\label{tab:ME_values}
 \end{table}
@@
 	\centering
-	\begin{tabular}{llccrrrrr}
-		\hline
-		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
-		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
-		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
-		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{comp}$		& - 	& M & & & & &  \\
-		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
-		&	$T_{air,in}$   		& $^o$C & M  & & & & &  \\ 
-		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
-		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{exp}$		& - 	& C & & & & &  \\
-		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
-		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
-		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
-		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
-		&	$T_{eg,out}$  		& $^o$C & C  & & & & &  \\ 
-		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
-		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
-		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
-		\multicolumn{2}{l}{LOC} & & & & & & &  \\
-		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
-		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		\multicolumn{2}{l}{JWC} & & & & & & &  \\
-		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
+	\begin{tabular}{llccrrrr}
+		\hline
+		\multicolumn{2}{l}{Variable} & Unit & M/C & 30\% & 50\% & 70\% & 90\% \\
+		\hline
+		\multicolumn{2}{l}{Compressor} & & & & & & \\ 
+		& 	$T_{air,in}$  		& $^o$C & M & 284.2 & 285.3 & 284.9 & 294.6 \\ 
+		&	$T_{air,out}$  		& $^o$C & C  & 323.2 & 357.9 & 391.6 & 430.9 \\
+		&	$\dot{m}_{air}$  	& $kg/s$ & C  & 1.96 & 2.65 & 3.5 & 4.27 \\
+		&   $\beta_{comp}$		& - 	& M & & & &  \\
+		\multicolumn{2}{l}{Turbine} & & & & & &  \\
+		&	$T_{eg,in}$   		& $^o$C & M  & 729.5 & 762.1 & 770.3 & 798.5 \\
+		&	$T_{eg,out}$  		& $^o$C & M  & 652.8 & 667.8 & 654.2 & 659.4 \\
+		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & 2.02 & 2.74 & 3.61 & 4.41 \\
+		&   $\beta_{exp}$		& - 	& C & & & &  \\
+		\multicolumn{2}{l}{Cylinders} & & & & & &   \\
+		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & 1.96 & 2.65 & 3.5 & 4.27 \\ 
+		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & 0.05 & 0.08 & 0.11 & 0.14 \\
+		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & 2.02 & 2.74 & 3.61 & 4.41 \\
+		&	$T_{air,in}$  		& $^o$C & M  & 320.8 & 322.3 & 327.6 & 329.1 \\
+		&	$T_{eg,out}$  		& $^o$C & C  & 729.5 & 762.1 & 770.3 & 798.5 \\
+		\multicolumn{2}{l}{CAC-LT stage} & & & & & &  \\
+		&	$T_{air,in}$  		& $^o$C & C  & 323.2 & 356.3 & 366.7 & 372.0 \\ 
+		&	$T_{air,out}$ 		& $^o$C & M  & 320.8 & 322.3 & 327.6 & 329.1 \\ 
+		&	$T_{w,in}$   		& $^o$C & M  & 317.1 & 316.3 & 317.5 & 314.1 \\
+		&	$T_{w,out}$   		& $^o$C & C  & 317.2 & 318.2 & 320.5 & 318.3 \\
+		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & C & 11.16 & 11.16 & 10.98 & 10.78 \\ 
+		\multicolumn{2}{l}{CAC-HT stage} & & & & & & \\
+		&	$T_{air,in}$   		& $^o$C & C  & 323.2 & 357.9 & 391.6 & 430.9 \\ 
+		&	$T_{air,out}$   	& $^o$C & C  & 323.2 & 356.3 & 366.7 & 372.0 \\ 
+		&	$T_{w,in}$    		& $^o$C & C  & 362.5 & 364.2 & 368.0 & 368.2 \\
+		&	$T_{w,out}$   		& $^o$C & C  & 362.5 & 364.3 & 369.3 & 371.9 \\
+		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & 16.66 & 16.67 & 16.55 & 16.67 \\
+		\multicolumn{2}{l}{LOC} & & & & & &  \\
+		&	$T_{lo,in}$   		& $^o$C & C  & 348.2 & 350.3 & 354.0 & 357.3 \\
+		&	$T_{lo,out}$    	& $^o$C & M  & 336.6 & 337.4 & 339.7 & 340.4 \\
+		&	$T_{w,in}$    		& $^o$C & C & 317.2 & 318.2 & 320.5 & 318.3 \\ 
+		&	$T_{w,out}$   		& $^o$C & C  & 325.9 & 327.8 & 331.4 & 331.3 \\
+		\multicolumn{2}{l}{JWC} & & & & & & \\
+		&	$T_{wall}$    		& $^o$C & C  & \multicolumn{4}{c}{423}  \\
+		&	$T_{w,in}$    		& $^o$C & M  & 360.0 & 359.7 & 362.3 & 361.5 \\
+		&	$T_{w,out}$	   		& $^o$C & C  & 362.5 & 364.2 & 368.0 & 368.2 \\  
 		\hline
@@
 	\caption{Auxiliary engines, measured and calculated temperatures and flows at different engine loads}
+	\label{tab:AE_values}
 \end{table}
@@
 	\centering
-	\begin{tabular}{lcrrrrr}
-		\hline
-		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		Power output & Mech &  &  &  &  &  \\
-		Exhaust gas & Heat &  &  &  &  &  \\
-		CAC-LT & Heat &  &  &  &  &  \\
-		CAC-HT & Heat &  &  &  &  &  \\
-		JWC & Heat &  &  &  &  &  \\
-		LOC & Heat &  &  &  &  &  \\
+	\begin{tabular}{lcrrrr}
+		\hline
+		Energy flow & type & 30\% & 50\% & 70\% & 90\% \\
+		\hline
+		Power output & Mech &  1767 & 2946 & 4118 & 5293 \\	
+		Exhaust gas (after turbine) & Heat & 1672 & 2566 & 3227 & 3938 \\
+		CAC-LT & Heat & 104 & 152 & 279 & 454 \\
+		CAC-HT & Heat & 2 & 337  & 569 & 843  \\
+		JWC & Heat & 381 & 249 & 278 & 318 \\
+		LOC & Heat & 682 & 721 & 699 & 725  \\
 		\hline
@@
 	\caption{Main engines, energy flows at different engine loads}
+	\label{tab:ME_flows}
 \end{table}
@@
 	\centering
-	\begin{tabular}{lcrrrrr}
-		\hline
-		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		Power output & Mech &  &  &  &  &  \\
-		Exhaust gas & Heat &  &  &  &  &  \\
-		CAC-LT & Heat &  &  &  &  &  \\
-		CAC-HT & Heat &  &  &  &  &  \\
-		JWC & Heat &  &  &  &  &  \\
-		LOC & Heat &  &  &  &  &  \\
+	\begin{tabular}{lcrrrr}
+		\hline
+		Energy flow & type & 30\% & 50\% & 70\% & 90\% \\
+		\hline
+		Power output & Mech & 828 & 1381 & 1932 & 2482 \\
+
+		Exhaust gas (after turbine) & Heat & 808 & 1144 & 1456 & 1753 \\
+		CAC-LT & Heat & 5 & 91 & 139 & 216 \\
+		CAC-HT & Heat & 0 & 5 & 89 & 257  \\
+		JWC & Heat & 176 & 314 & 396 & 467 \\ 
+		LOC & Heat & 405 & 448 & 499 & 587  \\
 		\hline
@@
 	\caption{Auxiliary engines, energy flows at different engine loads}
+	\label{tab:AE_flows}
 \end{table}
@@
 
-\subsection{Heat demand and generation} \label{sec:res:heat}
-
 FIGURE --> Pie, heat generated by source
@@
 The results of the exergy analysis of the ship energy systems are shown in Figure REF and Tables REF to REF. 
-
-Table(s) --> Exergy efficiencies
-- General at system level
-- One per system (MEs, AEs, CoolingSystems, SteamSystems)
-
-Figure --> Grassman diagram
+\begin{itemize}
+	\item Table(s) --> Exergy efficiencies
+	\begin{itemize}
+		\item General at system level
+		\item One per system (MEs, AEs, CoolingSystems, SteamSystems) 
+	\end{itemize}
+	\item Figure --> Grassman diagram
+\end{itemize}
+
+It can be observed that there are large irreversibilities concentrated in some specific parts of the system that show potential for improvement.
+This is particularly the case of the LT/HT merging valves, where exergy is destroyed in the mixing process.
+Another substantial source of exergy destruction is the steam heater in the HTHR system, where heat at medium pressure (6 bar) is used to heat up water at 90$^o$C.
+The analysis of exergy destruction also points out the HVAC reheater as a potential source of improvement, based on the use of relatively high-temperature water for heating up air to around 30$^o$C and on the large energy demand of the component.
+Other important exergy destruction sources can be located in the engines, as expected, but the improvement of engine processes and technology goes beyond the scope of this work.
+
+From the point of view of the exergy losses, it appears that the largest exergy losses are located in the exhaust gas.
+It should be noted that, even though a large part of this exergy cannot be recovered given the limitations on the exhaust gas outlet temperature to avoid the condensation of sulphuric acid, there is still a significant potential to be harvested if all the available exergy in the exhaust gas was recovered.
+This potential is further increased by the exergy destroyed in the process of dumping steam when it is produced in excess during summer.
+The exergy lost in the sea water coolers is much lower, highlighting the fact that the potential for improving the efficiency of the system is located in other parts of the cooling systems.
 
 \subsection{Typical operational days}  \label{sec:res:typicalDays}
+
+\subsection{Potential for system improvement}
+
+Based on the results of the energy and exergy analysis, it is possible to identify a number of possible measures that could potentially improve the efficiency of the system.
+
+Given the fact that many engines are operated at low load (both main and auxiliary engines), the system would benefit from both an electrification of the system and from the installation of batteries.
+The system electrification was explored by the authors \cite{Baldi2016b} and showed a relevant potential from both an environmental and an economic point of view.
+The installation of batteries was also considered in a previous publication by the authors \cite{Baldi2017} and proved the potential for yearly savings of 1-2\%.
+In this second case, however, it should be noted that the economic performance was not ensured.
+
+The availability of waste heat that is not already recovered by the system suggests that the system could benefit from the installation of a Heat-to-power WHR.
+This possibility was tested by the authors both in the case that no additional retrofit is performed \cite{Ahlgren2016,Mondejar2017}, with estimated savings of 22\% of the auxiliary power demand for a reference voyge, and in the case of a full integration with the rest of the system \cite{Baldi2017}, with estimated savings of up to 6\% of the total fuel consumption for a year of ship operations.
+
+It should be noted that in the second case the advantages in terms of system performance were obtained through a system-wide optimization of the system, where an improved utilization of the low-temperature waste heat available allowed "freeing" part of the high-temperature heat from the exhaust gas to the heat-to-power WHR system, hence improving the overall exergy efficiency of the system.
+This includes, for instance, the concept of using heat from the LT cooling systems (currently unused) for the HVAC preheater, hence reducing exergy destruction in the HVAC preheater itself and in the steam heater.
+This concept could be extended to other low-temperature heat consumers (such as the machinery space heating), but it would clearly constitute a larger benefit if applied to the HVAC preheater, which is the largest heat demand on board.
+
+OTHER?
 

### 2017-10-13 Main.py

@@ -25,7 +25,7 @@
 
 
 do_processed_data_preparation = "no"
-do_data_processing = "no"
+do_data_processing = "yes"
 
 
 
@@ -60,6 +60,7 @@ import auxiliaryDemand as aux
 import coolingsystems as cs
 from helpers import d2df
 import export as ex
+import postProcessing as post
 
 #%%
 ######################################
@@ -107,7 +108,7 @@ elif do_data_processing == "yes":
     processed = ppa.auxEngineProcessing(dataset_raw, processed, dict_structure, CONSTANTS, header_names)
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-1")
     # Calculating the auxiliary power demands: heating and electric power
-    processed = aux.auxPowerAnalysis(processed, CONSTANTS, dict_structure)
+    processed = aux.auxPowerAnalysis(processed, CONSTANTS, dict_structure, dataset_raw, header_names)
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-2")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Demands", "Demands-1")
     # Calculating the central cooling systems
@@ -119,8 +120,6 @@ elif do_data_processing == "yes":
     # Re-doing the calculation of the connected points
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-4")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Demands-2")
-    # Assigning the operational mode
-    processed = ppo.operationalModeCalculator(processed, dataset_raw, CONSTANTS, header_names)
     # Saving the processed data
     processed.to_hdf(CONSTANTS["filenames"]["dataset_output"], "processed", format='fixed', mode='w')
     # Result check log

### 2017-10-13 Python files/Constants.py

@@ -80,25 +80,17 @@ def mainEngines(CONSTANTS):
     output["FRP_2_MFR"]["POLY"] = {"ME1": [-159.612, 24.23254], "ME2": [-159.612, 28.282788], "ME3": [-159.612, 28.282788], "ME4": [-159.612, 28.282788]}
     output["FRP_2_MFR"]["FRP_MIN"] = {"ME1": 18.0, "ME2": 16.6, "ME3": 17.5, "ME4": 16.4}
     output["FRP_2_MFR"]["FRP_MAX"] = {"ME1": 51, "ME2": 47, "ME3": 47, "ME4": 46}
-    output["POLY_FUEL_LOAD_2_BSFC_ISO"] = np.polyfit(np.array([336.3, 587.8, 836.6, 953.1, 1141])/1141, [216.9, 187.1, 178.5, 179.2, 181.4], 2)   # Fits a 2nd degree polynomial relating relative fuel rack position to fuel flow in kg/s
-# output["POLY_RPM_2_POWER"] = polyfit([315 397 454 474 500 516], [1463 2925 4388 4973 5890 6435], 3)  
+    # output["POLY_FUEL_LOAD_2_BSFC_ISO"] = np.polyfit(np.array([336.3, 587.8, 836.6, 953.1, 1141])/1141, [216.9, 187.1, 178.5, 179.2, 181.4], 2)   # Fits a 2nd degree polynomial relating relative fuel rack position to fuel flow in kg/s
+    output["POLY_FUEL_LOAD_2_BSFC_ISO"] = np.polyfit(np.array([336.3, 587.8, 836.6, 953.1, 1141]) / 1141,
+                                                     [205, 187.1, 178.5, 179.2, 181.4], 2)  # Fits a 2nd degree polynomial relating relative fuel rack position to fuel flow in kg/s
     output["POLY_RPM_2_ISO_BSFC"] = np.polyfit([315.0, 397.0, 454.0, 474.0, 500.0, 516.0], [np.mean([216.1, 207.6, 225.5, 209.9]), 188.2, 179.7, 181.6, 185, 191.1], 2)
-# output["POLY_PCA_2_LOAD"] = [0.25/0.24, 0, 0.2577, 0.1438, 0.5, 0]  
[…]

### 2017-10-13 Python files/auxiliaryDemand.py

@@ -2,19 +2,23 @@ import numpy as np
 import pandas as pd
 from helpers import d2df
 import datetime as dt
+import preprocessingO as ppo
 
 
-def auxPowerAnalysis(processed, CONSTANTS, dict_structure):
+def auxPowerAnalysis(processed, CONSTANTS, dict_structure, dataset_raw, header_names):
     # Calculating the total auxiliary power demand
     processed = auxPowerTotalDemand(processed)
     processed = propPowerDemand(processed, CONSTANTS)
     # Calculating the contribution from the thrusters
     #processed = thrusters(processed, CONSTANTS)  # The second method works better. But I still keep it
     processed = thrusters(processed, CONSTANTS)
+    # Assigning the operational mode
+    processed = ppo.operationalModeCalculator(processed, dataset_raw, CONSTANTS, header_names)
+
     processed = HVAC(processed, CONSTANTS)
     processed = heatDemand(processed, CONSTANTS, dict_structure)
     processed = HTHR(processed, CONSTANTS)
-    processed = steamSystems(processed, CONSTANTS)
+    processed = steamSystems(processed, CONSTANTS, dataset_raw, header_names)
     return processed
 
 
@@ -252,7 +256,7 @@ def HTHR(processed, CONSTANTS):
 
 
 
-def steamSystems(processed, CONSTANTS):
+def steamSystems(processed, CONSTANTS, dataset_raw, header_names):
     db_index = processed.index
     # This function assigns all calculations to the steam systems
     Qdot_ab = pd.Series(index=db_index)
@@ -299,7 +303,14 @@ def steamSystems(processed, CONSTANTS):
     Qdot_steam = pd.Series(data=Qdot_steam, index=db_index)
 
     # Adding the boiler flow
-    processed["Steam:Boiler1:Steam_in:mdot"] = Qdot_ab / CONSTANTS["Steam"]["DH_STEAM"]
+    processed["Steam:Boiler1:Steam_out:mdot"] = (Qdot_steam + Qdot_dumped) / CONSTANTS["Steam"]["DH_STEAM"]
[…]

### 2017-10-13 Python files/consistencycheck.py

@@ -63,7 +63,8 @@ def enginesCheck(processed, CONSTANTS):
                 processed[d2df(name,"Cyl","Power_out","Edot")] +
                 processed[d2df(name,"Turbine","Mix_out","Edot")] +
                 (processed[d2df(name,"LOC","LTWater_out","Edot")] - processed[d2df(name,"CAC_LT","LTWater_in","Edot")]) +
-                (processed[d2df(name, "CAC_HT", "HTWater_out", "Edot")] - processed[d2df(name, "JWC", "HTWater_in", "Edot")]))
+                (processed[d2df(name, "CAC_HT", "HTWater_out", "Edot")] - processed[d2df(name, "JWC", "HTWater_in", "Edot")]) +
+                processed[d2df(name, "Cyl", "QdotRad_out", "Edot")])
             energyInput = (
                 processed[d2df(name, "Cyl", "FuelCh_in", "Edot")] +
                 processed[d2df(name, "Cyl", "FuelPh_in", "Edot")] +

### 2017-10-13 Python files/export.py

@@ -5,7 +5,7 @@ def exportEfficiecies(processed, CONSTANTS, dict_structure):
     text_file.write("Variable name,EnergyEfficiency,ExergyEfficiency,EfficiencyDefect,IrreversibilityShare\n")
     for system in ["ME1", "ME2", "ME3", "ME4", "AE1", "AE2", "AE3", "AE4", "CoolingSystems", "HTHR", "Steam"]:
         text_file.write("\n")
-        text_file.write(system+"\n")
+        text_file.write(system+", , , ," + str(processed[system+":"+"delta"].mean()) + "\n")
         for unit in dict_structure["systems"][system]["units"]:
             text_file.write(unit)
             for efficiency in ["eta", "eps", "lambda", "delta"]:

### 2017-10-13 Python files/postProcessing.py

@@ -1,0 +1,50 @@
+from helpers import d2df
+import pandas as pd
+import matplotlib.pylab as plt
+
+def predefinedTables(processed, name, CONSTANTS, dict_structure, system, load_vector):
+    file = open(CONSTANTS["filenames"]["latex_table"], "w")
+    if name == "MainEnginesModel":
+        for unit in dict_structure["systems"]["ME1"]["units"]:
+            for flow in dict_structure["systems"]["ME1"]["units"][unit]["flows"]:
+                if dict_structure["systems"]["ME1"]["units"][unit]["flows"][flow]["type"] in {"IPF", "CPF"}:
+                    for property in {"T", "mdot"}:
+                        value = []
+                        for load in [0.3, 0.5, 0.7, 0.9]:
+                            if property == "T":
+                                rounding = 1
+                            else:
+                                rounding = 2
+                            value.append(round(processed[d2df("ME1", unit, flow, property)][
+                                                   (processed["ME1:load"] < load + 0.005) & (
+                                                   processed["ME1:load"] > load - 0.005)].mean(), rounding))
+                        print(property + "_" + unit + "_" + flow + " = " + str(value[0]) + " & " + str(
+                            value[1]) + " & " + str(value[2]) + " & " + str(value[3]) + " \\\\ ")
+
+    if name == "MainEnginesEnergyFlows":
+        df = pd.DataFrame(index=processed.index)
+        df["Wdot_mech"] = processed[d2df(system, "Cyl", "Power_out", "Edot")]
+        df["Qdot_eg"] = processed[d2df(system, "Turbine", "Mix_out", "Edot")]
+        df["Qdot_jwc"] = processed[d2df(system, "Cyl", "QdotJW_out", "Edot")]
+        df["Qdot_loc"] = processed[d2df(system, "LOC", "LTWater_out", "Edot")] - processed[d2df(system, "LOC", "LTWater_in", "Edot")]
+        df["Qdot_caclt"] = processed[d2df(system, "CAC_LT", "LTWater_out", "Edot")] - processed[d2df(system, "CAC_LT", "LTWater_in", "Edot")]
+        df["Qdot_cacht"] = processed[d2df(system, "CAC_HT", "HTWater_out", "Edot")] - processed[d2df(system, "CAC_HT", "HTWater_in", "Edot")]
+        df["Qdot_ht"] = df["Qdot_cacht"] + df["Qdot_jwc"]
+        df["Qdot_lt"] = df["Qdot_caclt"] + df["Qdot_loc"]
[…]

### 2017-10-13 Python files/preprocessingME.py

@@ -13,7 +13,7 @@ def mainEngineProcessing(raw, processed, dict_structure, CONSTANTS, hd):
     # This script summarizes all the functions that calculate the required data for the Main Engines different flows
     # Reading existing values
     processed = readMainEnginesExistingValues(raw, processed, CONSTANTS, hd)
-    processed = ppo.engineStatusCalculation("MainEngines", raw, processed, CONSTANTS, hd)
+    processed = ppo.engineStatusCalculation("MainEngines", raw, processed, CONSTANTS, hd, dict_structure)
     processed = cs.coolingFlows(processed, CONSTANTS, "MainEngines")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-1.1")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "MainEngines", "ME-1.2")
@@ -144,7 +144,7 @@ def mainEngineAirFlowCalculation(raw, processed, dict_structure, CONSTANTS):
         processed[d2df(system, "Comp", "Air_out", "T")] = T_Comp_out
         # Calculating the air inflow aspired by the cylinder: calculated as inlet air density times the maximum volume,
         # times the engine speed
-        processed[d2df(system,"Cyl","Air_in","mdot")] = CONSTANTS["MainEngines"]["V_MAX"] * (
+        processed[d2df(system,"Cyl","Air_in","mdot")] = CONSTANTS["MainEngines"]["AIR_FLOW_MULT"] * CONSTANTS["MainEngines"]["V_MAX"] * (
             processed[d2df(system,"Comp","Air_out","p")]) / (
             CONSTANTS["General"]["R_AIR"] * processed[d2df(system,"Cyl","Air_in","T")]) * (
             processed[d2df(system,"Cyl","Power_out","omega")] / 60 / 2 * CONSTANTS["General"]["ETA_VOL"]) * (

### 2017-10-13 Python files/tables.py

@@ -1,3 +1,0 @@
-def predefinedTables(processed, name):
-    open()
-    if name == "MainEnginesModel":

### 2017-10-13 Python files/unitstructures.py

@@ -129,11 +129,12 @@ def flowStructure():
             # Engine cylinders
             structure["systems"][system]["units"]["Cyl"]["flows"] = {
                 "Air_in": {"type": "CPF", "IO": "input"},
-                "FuelPh_in": {"type": "CPF", "IO": "input"},
+                "FuelPh_in": {"type": "IPF", "IO": "input"},
                 "FuelCh_in": {"type": "CEF", "IO": "input"},
                 "EG_out": {"type":"CPF"},
                 "Power_out": {"type": "Wdot", "IO": "output"},
                 "QdotJW_out": {"type":"Qdot"},
+                "QdotRad_out": {"type":"Qdot"},
                 "LubOil_in": {"type": "IPF"},
                 "LubOil_out": {"type": "IPF"}}
             structure["systems"][system]["units"]["Cyl"]["equations"] = ["MassBalance"]
@@ -313,7 +314,7 @@ def flowStructure():
 
 
         elif system == "Steam":
-            structure["systems"][system]["units"] = {"Boiler1": {}, "Boiler2": {}, "HotWell": {}, "SteamDistribution": {},
+            structure["systems"][system]["units"] = {"Boiler1": {}, "HotWell": {}, "SteamDistribution": {},
                                                      "TankHeating": {}, "OtherTanks": {}, "HFOtankHeating": {}, "MachinerySpaceHeaters": {},
                                                      "HFOheater": {}, "Galley": {}}
             # Auxiliary boiler
@@ -322,27 +323,20 @@ def flowStructure():
                 "EG_out": {"type": "CPF"},
                 "FuelCh_in": {"type": "CEF", "IO": "input"},
                 "FuelPh_in": {"type": "IPF", "IO": "input"},
-                "Steam_in": {"type": "SF", "state": "SL", "IO": "output"},
-                "Steam_out": {"type": "SF", "state": "SV", "IO": "output"}}
+                "Steam_HotWell_in": {"type": "SF", "state": "SL", "IO": "output"},
+                "Steam_out": {"type": "SF", "state": "SV", "IO": "output"},
+                "Steam_HRSG_ME2_in": {"type": "SF", "state": "SV"},
+                "Steam_HRSG_ME3_in": {"type": "SF", "state": "SV", "IO": "input"},
[…]

### 2017-10-13 Manuscript/Manuscript.tex
Manuscript and results update

@@
 
+\subsection{Operational mode}
+
+The ship operates in different conditions, and it can be useful to provide a separate analysis depending on the operational mode.
+The four operational modes identified for the selected case study, together with the conditions applied to the dataset to uniquely identify each time step to a separate operational mode, are given as follows:
+\begin{description}
+	\item[High speed sailing] Ship speed above 15 kn
+	\item[Low speed sailing] Ship speed between 4 and 15 kn
+	\item[Maneuvering] Ship speed between 2 and 4 knots OR Thrusters power larger than 0
+	\item[Port stays / ship drifting] All other points
+\end{description}
+
 \subsection{Determination of typical operational days}
@@
 
-figure --> Operational mode distribution
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/Pie_operationalMode}
+	\caption{Operational time by mode}
+	\label{fig:opMode_pie}
+\end{figure}
+
 \begin{figure}[htbp]
@@
 	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_W}
-	\caption{Energy demand during a day of operations, Winter}
+	\caption{Energy demand during a day of operations, Winter (31 Jan)}
 	\label{fig:Demand_vs_time_W}
@@
 	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_S}
-	\caption{Energy demand during a day of operations, Summer}
+	\caption{Energy demand during a day of operations, Summer (31 Jul)}
 	\label{fig:Demand_vs_time_S}
@@
 
-FIGURE --> Pie, heat generated by source
-FIGURE --> filled lines, heat generated by source.
-
-Looking at Figure \ref{fig:HeatGen_hist} it is possible to appreciate the amount of heat that is recovered from on board available waste heat (cooling water and exhaust gas).
-It can be observed that the amount of heat recovered from the HT cooling systems is much lower than expected, with an average of around 350 kW.
-This value is quite unreasonable compared to the values found in the technical specifications from the yard, that go from 700 kW (when only one auxiliary engine is operated) to 4800 kW (when all main engines and two auxiliary engines are operated).
-\\
+Looking at Figure \ref{fig:Pie_HeatGeneration} it is possible to appreciate the amount of heat that is recovered from on board available waste heat (cooling water and exhaust gas).
+It can be observed that  of heat recovered from the HT cooling systems represents approximately one quarter of the total heat supply.
+Figure \ref{fig:TimeSeries_HeatGeneration} shows how the contribution of the HRSG and the HTHR is substantially constant throughout the year, while the auxiliary boilers are used to cover the peaks and the winter increased baseload.
+
 \begin{figure}
 	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/HeatGen_hist}
-	\caption{Distribution of thermal power generated from on board waste heat recovery}
-	\label{fig:Hist_WHR}
+	\includegraphics[width=0.99\linewidth]{Figures/Pie_HeatGeneration}
+	\caption{Share of thermal power generated from on board heat systems}
+	\label{fig:Pie_HeatGeneration}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/TimeSeries_HeatGeneration}
+	\caption{Time series representation of the thermal power generated from on board heat systems}
+	\label{fig:TimeSeries_HeatGeneration}
 \end{figure}
@@
 The exergy lost in the sea water coolers is much lower, highlighting the fact that the potential for improving the efficiency of the system is located in other parts of the cooling systems.
+
+The efficiency of the recovery systems is also shown in Figure \ref{fig:Bar_efficiencyWHR}, where we show the fraction of the total energy (exergy) lost by the cooling systems and the exhaust gas of the ship's engines.
+It can be noted that, even when only looking at the energy that can be recovered based on the existing systems (i.e.
+not including the LT cooling systems, for instance), the recovery efficiency is located at around 25-35\% depending on whether the energy or the exergy efficiency is considered.
+Worth noting is also the fact that the efficiency on the exhaust gas side is lowered by the fact that two of the main engines are currently not equipped with HRSGs.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/Bar_WHRefficiency}
+	\caption{Energy and exergy efficiency of the WHR systems}
+	\label{fig:Bar_efficiencyWHR}
+\end{figure}
 

### 2017-10-13 Python files/Energyanalysis.py

@@ -138,6 +138,8 @@ def efficiencyCalculator(processed, dict_structure, CONSTANTS):
             temp_flow_list = {"Edot_in", "Edot_in_useful", "Edot_out", "Edot_out_useful", "Bdot_in", "Bdot_in_useful", "Bdot_out", "Bdot_out_useful"}
             temp_df = pd.DataFrame(0, columns=temp_flow_list, index=df_index)
             for flow in dict_structure["systems"][system]["units"][unit]["flows"]:
+                dict_structure["systems"][system]["units"][unit]["flows"][flow]["E"] = processed[d2df(system, unit, flow, "Edot")].sum() * 60 * 15 * 1e-6
+                dict_structure["systems"][system]["units"][unit]["flows"][flow]["B"] = processed[d2df(system, unit, flow, "Bdot")].sum() * 60 * 15 * 1e-6
                 if flow[-2:] == "in":
                     temp_df['Edot_in'] = temp_df['Edot_in'] + processed[d2df(system, unit, flow, "Edot")]
                     temp_df['Bdot_in'] = temp_df['Bdot_in'] + processed[d2df(system, unit, flow, "Bdot")]
@@ -168,17 +170,50 @@ def efficiencyCalculator(processed, dict_structure, CONSTANTS):
                             text_file.write("Flow {}:{}:{} is not recognised as either input or output \n".format(system, unit,flow))
                     else:
                         text_file.write("Flow {}:{}:{} is not recognised as either USEFEUL input or output \n".format(system, unit, flow))
+            # Saving the total energy flows
+            dict_structure["systems"][system]["units"][unit]["E_in"] = temp_df["Edot_in"].sum() * 60 * 15 * 1e-6
+            dict_structure["systems"][system]["units"][unit]["E_out"] = temp_df["Edot_out"].sum() * 60 * 15 * 1e-6
+            dict_structure["systems"][system]["units"][unit]["E_in_useful"] = temp_df["Edot_in_useful"].sum() * 60 * 15 * 1e-6
+            dict_structure["systems"][system]["units"][unit]["E_out_useful"] = temp_df["Edot_out_useful"].sum() * 60 * 15 * 1e-6
+            # Saving total exergy flows
+            dict_structure["systems"][system]["units"][unit]["B_in"] = temp_df["Bdot_in"].sum() * 60 * 15 * 1e-6
+            dict_structure["systems"][system]["units"][unit]["B_out"] = temp_df["Bdot_out"].sum() * 60 * 15 * 1e-6
[…]

### 2017-10-13 Python files/export.py

@@ -10,7 +10,7 @@ def exportEfficiecies(processed, CONSTANTS, dict_structure):
             text_file.write(unit)
             for efficiency in ["eta", "eps", "lambda", "delta"]:
                 if system + ":" + unit + ":" + efficiency in processed.columns:
-                    text_file.write("," + str(processed[system+":"+unit+":"+ efficiency].mean()))
+                    text_file.write("," + str(dict_structure["systems"][system]["units"][unit][efficiency]))
                 else:
                     text_file.write(",")
             text_file.write("\n")

### 2017-10-13 Python files/input.py

@@ -40,7 +40,7 @@ def assumptions(raw, processed, CONSTANTS, hd):
         # Temperature in the engine room, i.e. inlet to the compressor of the TC
         processed[d2df(system, "Comp", "Air_in", "T")] = raw[hd["ER_AIR_T_"]] + 273.15
         # Assuming that the pressure in the exhaust gas is 90% of the pressure in the inlet manifold. Somewhat reasonable
-        processed[d2df(system, "Cyl", "EG_out", "p")] = (0.9 * raw[hd[system + "-CAC_AIR_P_OUT"]] + 1.01325) * 100000
+        processed[d2df(system, "Cyl", "EG_out", "p")] = (0.98 * raw[hd[system + "-CAC_AIR_P_OUT"]] + 1.01325) * 100000
         # Assuming the pressure of the fuel to be around 9 barg, based on measurements from ME4
         processed.loc[:,d2df(system, "Cyl", "FuelPh_in", "p")] = (9 + 1.01325) * 10e5
         # Assuming the temperature of the cylinder wall to be 150 degC

### 2017-10-13 Python files/plotting.py

@@ -1,6 +1,8 @@
 import matplotlib.pyplot as plt
 from helpers import d2df
 import pandas as pd
+import numpy as np
+from matplotlib.patches import Rectangle
 
 def plotMain(type, dict_structure, processed, *args):
     # This function plots the contents of the analysed files.
@@ -233,7 +235,7 @@ def plottingFunction(plot_info, processed, data_type):
 
 def predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure, filenames):
     for filename in filenames:
-        plt.figure()
+        fig, ax = plt.subplots()
 
         ### TIME SERIES ###
 
@@ -255,23 +257,40 @@ def predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure, filenames
             # Total demand
             total = processed["Demands:Heat:Total:Edot"].resample("D").sum() * 60 * 15
             # Actual plotting
-            plt.plot(hrsg, 'k-', label="HRSG")
-            plt.plot(hthr, 'b-', label="HTHR")
+            ax.plot(hrsg, 'k-', label="HRSG")
+            ax.plot(hthr, 'b-', label="HTHR")
             # plt.plot(hthr_max, 'b:', label="HTHR max")
-            plt.plot(boilers_measured, 'r-', label="Boilers (M)")
-            plt.plot(boilers_calculated, 'r:', label="Boilers (C)")
-            plt.plot(total, 'g-', label='Total heating demand')
+            ax.plot(boilers_measured, 'r-', label="Boilers (M)")
+            ax.plot(boilers_calculated, 'r:', label="Boilers (C)")
+            ax.plot(total, 'g-', label='Total heating demand')
+            plt.legend()
+
+        if filename == "TimeSeries:TypicalWinterDay":
+            ax.plot(processed["Demands:Mechanical:Total:Edot"]["2014-01-31"], "g-", label = "Propulsion power")
+            ax.plot(processed["Demands:Electricity:Total:Edot"]["2014-01-31"], "b--", label="Electric power")
+            ax.plot(processed["Demands:Heat:Total:Edot"]["2014-01-31"], "r:", label="Heat")
+            fig.autofmt_xdate()
+            plt.xlabel("Time [MM-DD HH]")
+            plt.ylabel("Power [kW]")
+            plt.legend()
+
[…]

### 2017-10-13 Python files/preprocessingO.py

@@ -12,7 +12,7 @@ def operationalModeCalculator(processed, raw, CONSTANTS, hd):
     temp[(raw[hd["SHIP_SPEED_KNOT_"]] < 15) * (raw[hd["SHIP_SPEED_KNOT_"]] > 4)] = "Low Speed Sailing"
     temp[((processed["Demands:Electricity:Thrusters:Edot"] > 0) + ((raw[hd["SHIP_SPEED_KNOT_"]] < 4) * (raw[hd["SHIP_SPEED_KNOT_"]] > 2)))] = "Maneuvering"
     temp[temp.isnull()] = "Port/Stay"
-    processed["operationalMode"] = pd.Series(temp.values)
+    processed["operationalMode"] = temp
     return processed
 
 

### 2017-11-26 Manuscript/Other/EngineBalance.tex
Added output for Sankey and Grassmann diagrams

--- 2017-10-10 Manuscript/Manuscript.tex
@@
-%% 
-%% Copyright 2007, 2008, 2009 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
+\documentclass[]{article}
+\usepackage{amsmath}
+\usepackage{verbatim}
 
-%% Template article for Elsevier's document class `elsarticle'
-%% with numbered style bibliographic references
-%% SP 2008/03/01
-
-\documentclass[preprint,12pt]{elsarticle}
-
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn]{elsarticle}
-%% \documentclass[final,3p,times]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn]{elsarticle}
-%% \documentclass[final,5p,times]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
-%% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-\usepackage{amsthm}
-%% Tthe amsmath package provides the "align" environment for nicer equation formatting
-\usepackage{amsmath}
-%% The package tabls helps in having nicer tables with higher spaces
-\usepackage{tabls}
-%% The package multirow allows merging together multiple cells vertifcally
-\usepackage{multirow}
-%% 
-\usepackage{subcaption}
-
-\def\bibsection{\section*{References}}
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-%% \usepackage{lineno}
-
-\journal{Energy}
+%opening
+\title{Demonstration of the solution of the system of mass and energy balances for MS Birka Main Engines}
+\author{Francesco Baldi}
 
@@
 
-\begin{frontmatter}
-
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \address{Address\fnref{label3}}
-%% \fntext[label3]{}
-
-\title{Energy and exergy analysis of a cruise ship}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \address[label1]{}
-%% \address[label2]{}
-
-\author[EPFL]{Francesco Baldi}
-\author[Linnaeus]{Fredrik Ahlgren}
-\author[DTU]{Tuong-Van Nguyen}
-\author[CTU]{Cecilia Gabrielii}
-\author[CTU]{Karin Andersson}
-
-\address[EPFL]{Industrial Process Energy Systems Engineering (IPESE), \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, 1950, Sion, Switzerland}
-\address[Linnaeus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
-\address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
-\address[CTU]{Department of Mechanics and Maritime Sciences, Chalmers University of technology, Gothenburg, Sweden}
+\maketitle
 
 \begin{abstract}
-%% Text of abstract
 
@@
 
-\begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-
-%% PACS codes here, in the form: \PACS code \sep code
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-
-low carbon shipping \sep energy analysis \sep exergy analysis \sep energy efficiency
-
-\end{keyword}
-
-\end{frontmatter}
-
-%% \linenumbers
-
-%% main text
-
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
-%% \appendix
-
-%% \section{}
-%% \label{}
-
-%% If you have bibdatabase file and want bibtex to generate the
-%% bibitems, please use
-%%
-%%  \bibliographystyle{elsarticle-num} 
-%%  \bibliography{<your bibdatabase>}
-
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
-
-\section{Introduction} \label{sec:introduction}
-
-\subsection{Background}
-
-CO$_2$ emissions from shipping in 2012 amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions \cite{Smith2014}.
-Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios \cite{Smith2014}.
-As an example, global transport demand has increased by 3.4\% in 2014, compared to a global GBP growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}.
-The OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3].
-The fact that shipping needs to even further reduce its CO$_2$ emissions in the near future is essential for being able to achieve the goals of maintaining the climate below a 2-degree level in 2050 [4].
-
-In addition to considerations related to GHG emissions, an emission control area (ECA) is enforced in the Baltic Sea  by the International Maritime Organisation since January 2015.
-This ECA stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels.
-More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}
-
-Altogether, these conditions present a challenge to the shipping companies who attempt to reduce their fuel consumption, environmental impact, and operative costs.
-A wide range of fuel saving solutions for shipping are available and partially implemented in the existing fleet, both from the design and operational perspective; several specific studies have been conducted on these technologies, and a more detailed treatise would be out of the scope of this work.
-In this context, it has been acknowledged that the world fleet is heterogeneous, and measures need to be evaluated on a ship-to-ship basis \cite{Bouman2017}.
-In this process, a deeper understanding of energy use on board of the specific ship is vital.
-
-\subsection{Previous work}
-
-The idea of improving the understanding of the behavior of the energy system of a ship is not new.
-Most of the work published around this subject relates to the use  of mathematical models of the ship systems.
-
-Most authors focused on the propulsion part of the problem, as this is often the most relevant energy demand on board.
-Shi et al.
-proposed a modeling approach for predicting ship fuel consumption of a cargo/passenger ferry \cite{Shi2009}; Theotokoats and Tzelepis applied a similar procedure to the case of a Handymax product carrier \cite{Theotokatos2015}, similarly to Tillig et al., who also added the dynamic element to their predictive model \cite{Tillig2016}.
-
-The work referenced above allowed improving the understanding of how many operational (such as the ship's speed) and environmental (such as wave height and wind speed) parameters influence the ship's energy performance.
-These studies, however, are not based on actual measurements of the ship's operations.
-Also, they focus entirely on the ship's power demand for propulsion.
-This is a very reasonable practice for most ship types, given that propulsion represent the largest part of the total energy demand, but does not allow improving the knowledge of the remaining part of the system.
-
-% Data-driven analysis
-Other authors filled the gap by both including electric energy demand, and by basing their work on measurements from actual ship operations.
-The work presented by Thomas et al.
-\cite{Thomas2010} and Basurko et al.
-\cite{Basurko2013} shows the application of energy auditing methods to fishing vessels.
-This approach represents a step forward towards improving the understanding of how energy is used on board during actual ship operations.
-
-More generally, several authors have highlighted the importance of a detailed knowledge of the ship's operational profile in order to appropriately assess and optimize possible alternatives for improving ship energy efficiency \cite{Banks2013}.
-Coraddu et al., started from a statistical analysis of measured ship operations and used the aggregated data, together with a computational model of the vessel, to provide a better prediction of the actual ship's operational efficiency \cite{Coraddu2014}.
-The importance of considering the operational profile was also showed in the case of the optimization of engine-propeller interaction \cite{Baldi2015c}, in the process of retrofitting existing systems \cite{Baldi2015b,Choi2013}, and in ship design \cite{Ghassemi2017,Solem2015}.
-
-% Heat availability
-Heat demand is rarely a subject of concern on board ships, with some notable exceptions.
-This is due to a combination of generally low demand and high availability from the waste heat of the engines.
-In a previous publications by the authors, for instance, it is shown that in the case of a product tanker, although the heat demand was estimated to account for roughly 20\% of the total energy demand of the ship on a yearly basis, it only contributes to 4.1\% of the total fuel consumption (contribution of the auxiliary boilers), while the rest of the demand is fulfilled using waste heat \cite{Baldi2015a}.
-
-%% WHR and heat availability
-It should be noted that, however, much research effort has been devoted, especially in recent times, to the improvement of the efficiency of ship energy systems by recovering the waste heat available from the engines.
-With reference to different types of technologies, case studies, and designs,the several authors showed the existance of a quite significant potential for energy saving when WHR systems are employed, ranging from around 1\% for single-pressure steam cycles applied to two-stroke engines \cite{Theotokatos2012} to more complex systems based on ORCs (up to 10\%, \cite{Hountalas2012}) or including the cooling systems as a source of waste heat (over10\%, \cite{Dimopoulos2012}).
-The case of the installation of an ORC on board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2015}, showing a similar potential.
-
-The potential uses for waste heat on board are not limited to improving the efficiency of the power plant.
-Waste heat is commonly used for fulfilling on board heat demand for spatial heating and freshwater generation \cite{Molland2011,Baldi2015a,Mccarthy1990};  Balaji et al.
-proposed the use of waste heat for ballast-water systems \cite{Balaji2012}; Salmi et al.
-suggested its use for adsorption refrigeration systems \cite{Salmi2017}.
-A detailed review of potential uses for waste heat from marine engines is presented by Shu et al \cite{Shu2013}.
-
-% Cruise ships
-Some ship types constitute notable examples.
-On cruise ships the heat demand is significantly higher compared to standard cargo ships.
-Referring to winter conditions, Marty et al.
-estimated the instantaneous heat demand of a selected cruise ship to reach roughly 23 MW, compared to an estimated peak of 49 MW for propulsion and electric auxiliaries combined \cite{Marty2012}.
-
-This shows how the heat from the engines should be considered as a potential resource, rather than a waste, and that there is potential for improvement based on the optimal use of these heat sources.
-The work presented by the authors in \cite{Baldi2016} represent a step in this direction; however, there is the need for an increased detail in the estimation of the heat demand.
-
-% Exergy analysis
-Exergy analysis provides a more accurate estimation of the potential for energy recovery on board.
-The application of exergy analysis to the case of ship energy systems is, however, still limited.
-Dimopoulos et al.
-showed how the process optimizing the WHR system of a container ship can be more efficient if exergy efficiency, rather than energy efficiency, is used as the target of the optimization \cite{Dimopoulos2012}; Baldi et al.
-also analyzed the exergy flows on board of a product tanker, showing that this allows having a more accurate understanding of what parts of the system show potential for improvement \cite{Baldi2015a}; similar results were obtained by Marty et al., who focused on the power plant of a cruise ship \cite{Marty2016}.
-Koroglu et al.
-made a step futher also including advanced exergy analysis in their study \cite{Koroglu2017}.
-
-\subsection{Aim}
-
-Given the existing limitations of the current available information in scientific literature highlighted in the previous section, the objective of this work is the following:
-\begin{itemize}
-	\item Analyse the demand of a cruise ship in terms of propulsion power, electric power and heat, based on operational measurements.
-The lack of existing information with reference to the heat demand is considered a particular element of novelty when compared to existing literature in the subject.
-	\item Analyse the current efficiency of the system, and potential ways to improve it, by means of applying exergy analysis.
-	\item Provide reference operational conditions for further use in the field of energy systems optimization.
-\end{itemize}
-
-\section{Method} \label{sec:method}
-
-In this paper, we present the application of energy and exergy analysis (described in section \ref{sec:met:energyExergy}) to a cruise ship.
-This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; the specific assumptions and methods used for processing the available information into  energy and exergy flows are specified in Section \ref{sec:met:processing}.
-In these regards, the estimation of the heat demand is treated in a separate section (Sec.
-\ref{sec:met:heat}) as it constitutes a particularly challenging task and it is one of the central aspects of this article.
-
-\subsection{Energy and exergy analysis} \label{sec:met:energyExergy}
-
-\subsubsection{Energy analysis}
-
-\emph{Energy} can be stored, transformed from one form to another (e.g.
-heat to power) and transferred between systems, but can neither be created nor destroyed (conservation law)~\textbf{Bejan et al.}.
-The system under study is the \emph{ship energy system} and is thus taken as control volume.
-The energy balance can be expressed as:
-
-\begin{align}
-	\sum_{\mathrm{in}} \dot{H}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{H}_{\mathrm{out}} \\
-	\dot{H}_{\mathrm{fuel}} + \dot{H}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}+\dot{W}_{\mathrm{el}}+\dot{Q}_{\mathrm{heating}}
-\end{align}
-
-The left-hand side term represents, on a time rate basis, the energy associated with the fuel consumed in the boilers and engines ($\dot{H}_{\mathrm{fuel}}$) and the air used in the combustion processes ($\dot{H}_{\mathrm{air}}$).
-The right-hand side term denotes the power ($\dot{W}_{\mathrm{el}}$) and heat ($\dot{Q}_{\mathrm{heating}}$) required on-site (e.g.
-propulsion and fuel heating), and the heat discharged into the environment ($\sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}$) with, for instance, the exhaust gases.
-
-The energy flow associated with a material stream is calculated as the sum of the physical and chemical enthalpies, and kinetic and potential energies are neglected.
-The physical energy is taken as the relative enthalpy, as underlined in~\textbf{Kotas et al.}, and the chemical energy is taken as the lower/higher heating value.
-The environmental conditions taken for the present analysis are the ambient pressure and seawater temperature.
-
-\subsubsection{Exergy analysis}
-
-\emph{Exergy} may be defined as the `maximum theoretical useful work (shaft work or electrical work) as the system is brought into complete thermodynamic equilibrium with the thermodynamic environment while the system interacts with it only'~\citep{Moran1989}.
-Unlike energy, exergy is not conserved but some is destroyed because of the irreversible phenomena taking place in real processes (e.g.
-chemical reactions like combustion).
-The exergy balance for the system under study can be expressed as:
-
-\begin{align}
-	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \\
-	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d} 
-\end{align}
-
-The left-hand side term represents, on a time rate basis, the exergy associated with the fuel and air.
-The right-hand side term denotes the exergy of the waste streams, the exergy transfers with heat and power, and the exergy destroyed in the ship system.
-Kinetic and potential exergies are neglected, and the exergy of a given material stream is derived as the sum of the physical and chemical exergies.
-The chemical exergy is calculated based on the reference environment of \textbf{Szargut (1989)}, and its value is approximatively equal to the higher heating value for hydrocarbon fuels.
-The exergy destruction can be calculated from the Gouy-Stodola theorem~\textbf:{Kotas (1995)}.
-
-The exergy balance may alternatively be formulated as:
-\begin{equation}
-\dot{E}_{\mathrm{p}} = \dot{E}_{\mathrm{f}} - \dot{E}_{\mathrm{d}} - \dot{E}_{\mathrm{l}} 
-\end{equation}
-
-where $\dot{E}_{\mathrm{p}}$ is called the exergy product, and corresponds to the desired output of the system, in exergy terms (for example, the power produced in an engine).
-$\dot{E}_{\mathrm{f}}$ denotes the exergy fuel, and represents the resources spent to drive the studied process (for instance, the fuel used in a combustion process).
-The last term $\dot{E}_{\mathrm{l}}$ corresponds to the losses of a system, such as the heat discharged into the environment with cooling water.
-
-\subsubsection{System performance}
-
-The following indicators are used to evaluate the system performance:
-\begin{itemize}
-	\item the exergy efficiency $\varepsilon$, defined as the ratio between the exergy product and fuel of a given component or system
-	\begin{equation} \varepsilon = \frac{\dot{E}_{\mathrm{p}}}{\dot{E}_{\mathrm{f}}} \end{equation}
-	\item the efficiency defect $\lambda$, presented in the work of Kotas, defined as the fraction of the total exergy input destroyed in the successive irreversible processes
-	\begin{equation} \lambda = \frac{\dot{E}_{\mathrm{d}}}{\dot{E}_{\mathrm{in,tot}}} \end{equation}
-	\item  the irreversibility share $\delta$, suggested in the work of Tsatsaronis, defined as the ratio between the exergy destroyed in the $i$-th component in relation to the exergy destroyed in the entire system
-	\begin{equation} \delta_i = \frac{\dot{E}_{\mathrm{d},i}}{\dot{E}_{\mathrm{d,tot}}} \end{equation}
-
-\end{itemize}
-
-\subsection{Case study vessel} \label{sec:met:case}
-
-The energy and exergy analysis are here applied to a specific  cruise ship operating daily cruises in the Baltic Sea between Stockholm and the island of \.{A}land.
-The ship is 176.9 m long and has a beam of 28.6 m, and has a design speed of 21 knots.
-The ship was built in Aker Finnyards, Raumo Finland in 2004.
-
-The ship has a capacity of 1800 passengers and has several restaurants, night clubs and bars, as well as saunas and pools.
-This means that the energy system regarding the heat and electricity demand is more complex than a regular cargo vessel in the same size.
-Typical ship operations, although they can vary slightly between different days, are represented in Figure \ref{fig:typicalShipOperations}.
-It should be noted that the ship stops and drifts in open sea during night hours before mooring at its destination in the morning, if allowed by weather conditions.
-
-The ship systems are summarized in Figure \ref{fig:shipSystems}.
-The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller.
-The main engines are four Wärtsilä 4-stroke Diesel engines (ME) rated 5850 kW each.
-All engines are equipped with selective catalytic reactors (SCR) for NOX emissions abatement.
-Propulsion power is needed whenever the ship is sailing; however, it should be noted that the ship rarely sails at full speed, and most of the time it only needs one or two engines operated simultaneously.
-
-Auxiliary power is provided by four auxiliary engines (AE) rated 2760 kW each.
-Auxiliary power is needed on board for a number of alternative functions, from pumps in the engine room to lights, restaurants, ventilation and entertainment for the passengers.
-
-Auxiliary heat needs are fulfilled by the exhaust gas steam generators (HRSG) located on all four AEs and on two of the four MEs or by oil-fired auxiliary boilers (mainly when in port, or during winter), by the heat recovery on the HT cooling water systems (HRHT), and by the auxiliary, oil fired boilers (AB).
-The heat is needed for passenger and crew accommodation, as well as for the heating of the highly viscous heavy fuel oil used for engines and boilers.
-This last part, however, is drastically reduced since the 1st of January 2015, as new regulations entering into force require the use of low-sulphur fuels, which require a much more limited heating.
-
-\subsection{Data gathering and pre-processing} \label{sec:met:gathering}
-
-The operational data was collected on board from the ships' machine logging and surveillance system.
-The on board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB.
-The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average.
-The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python \cite{mckinney-proc-scipy-2010}.
-A new structured naming of headers and a consistent time frequency over all data points were created, and 245 selected data points were saved in a HDF5 table time series database which is the base for the analysis.
-
-All data-points were checked individually by creating a descriptive statistic and a histogram of each data-point.
-A filter was applied by setting up a pre-defined maximum and minimum value.
-As we do not know, or had any meaningful way of deriving, each individual sensors placement, sensor type or calibration status all data must be filtered and checked accordingly for outliers or bad data.
-
-Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska H\"{o}garna for the ambient temperature.
-The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska H\"{o}garna lighthouse is along the ship route, in between the Swedish archipelago and Åland.
-The assumption was validated based on  June-December period, for which onboard measurements were available.
-This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
-The fit the SMHI-data with the rest of the data-set the SMHI data was resampled from 1h for the seawater and 3h for the ambient temperature, to 15min frequency using a linear interpolation.
-
-\subsection{Data processing: Ship energy system modeling} \label{sec:met:processing}
-
-Not all of the variables required to perform a full energy and exergy analysis of the system are available from measurements.
-In some cases, they are measurable, but not measured (e.g.
-some temperatures, mass flows, etc.).
-In other cases, they are simply impossible, or impractical, to measure (e.g.
-specific enthalpy, specific entropy).
-For this reason, part of the ship systems needed to be modeled in order to derive the unknown variables.
-
-\subsubsection{Diesel engines}
-
-The setup of the engines and their connections to the other ship systems are summarized in Figure REF and REF for the main and auxiliary engines, respectively.
-The defining net energy flows are listed in Table REF.
-It should be noted that, in the case of the main engines, the engine power output was not measured and needed to be estimated.
-In this work, we calculated the engine power output based on measurements of engine fuel rack position (used as a proxy of the mass of fuel injected per cycle) and of the engine speed:
-\begin{eqnarray}
-\dot{m}_{fuel} & = & \dot{m}_{fuel,des} \left(a_0 + a_1 \frac{frp}{frp_{des}}\right) \frac{\omega}{\omega_{des}} \\
-\eta_{ME} & = & a_0 + a_1 \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} + a_2 \left( \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} \right)^2 \\
-\dot{W}_{ME} & = & \dot{m}_{fuel} \eta_{ME} LHV
-\end{eqnarray}
-
-The validity of this assumption can be seen by observing Figure REF, representing the calculated total power from the main engines as a function of the ship speed
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/DieselEngines}
-	\caption{Schematic representation of the Diesel engines and their connections to the cooling systems}
-	\label{fig:DieselEngines}
-\end{figure}
-
-In the case of the auxiliary engines, instead, direct measurements of the power generated by each engine were available.
-
-For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine.
-The model is explained in detail in Appendix REF.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pme_vs_vship_Neng}
-	\caption{Scatter plot, Propulsion power versus ship speed}
-	\label{fig:Pme_vs_vship_Neng}
-\end{figure}
-
-ALSO INCLUDE THE POWER VS TC SPEED RELATIONSHIP
-
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{llccrrrrr}
-		\hline
-		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
-		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
-		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
-		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{comp}$		& - 	& M & & & & &  \\
-		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
-		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
-		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{exp}$		& - 	& C & & & & &  \\
-		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
-		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
-		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
-		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
-		&	$T_{eg,out}$  		& $^o$C & M  & & & & &  \\ 
-		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
-		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
-		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
-		\multicolumn{2}{l}{LOC} & & & & & & &  \\
-		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
-		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		\multicolumn{2}{l}{JWC} & & & & & & &  \\
-		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
-		\hline
-		\end{tabular}
-	\caption{Main engines, measured and calculated temperatures and flows at different engine loads}
-\end{table}
-
-INCLUDE VALIDATION AGAINST PROJECT GUIDE DATA
-
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{llccrrrrr}
-		\hline
-		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
-		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
-		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
-		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{comp}$		& - 	& M & & & & &  \\
-		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
-		&	$T_{air,in}$   		& $^o$C & M  & & & & &  \\ 
-		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
-		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{exp}$		& - 	& C & & & & &  \\
-		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
-		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
-		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
-		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
-		&	$T_{eg,out}$  		& $^o$C & C  & & & & &  \\ 
-		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
-		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
-		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
-		\multicolumn{2}{l}{LOC} & & & & & & &  \\
-		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
-		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		\multicolumn{2}{l}{JWC} & & & & & & &  \\
-		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
-		\hline
-	\end{tabular}
-	\caption{Auxiliary engines, measured and calculated temperatures and flows at different engine loads}
-\end{table}
-
-INCLUDE VALIDATION AGAINST PROJECT GUIDE DATA
-
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{lcrrrrr}
-		\hline
-		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		Power output & Mech &  &  &  &  &  \\
-		Exhaust gas & Heat &  &  &  &  &  \\
-		CAC-LT & Heat &  &  &  &  &  \\
-		CAC-HT & Heat &  &  &  &  &  \\
-		JWC & Heat &  &  &  &  &  \\
-		LOC & Heat &  &  &  &  &  \\
-		\hline
-	\end{tabular}
-	\caption{Main engines, energy flows at different engine loads}
-\end{table}
-
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{lcrrrrr}
-		\hline
-		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		Power output & Mech &  &  &  &  &  \\
-		Exhaust gas & Heat &  &  &  &  &  \\
-		CAC-LT & Heat &  &  &  &  &  \\
-		CAC-HT & Heat &  &  &  &  &  \\
-		JWC & Heat &  &  &  &  &  \\
-		LOC & Heat &  &  &  &  &  \\
-		\hline
-	\end{tabular}
-	\caption{Auxiliary engines, energy flows at different engine loads}
-\end{table}
-
-\subsubsection{Cooling systems}
-
-The ship's cooling systems are designed similarly to most ships and divided in two cooling systems of two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C).
-The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board.
-A representation of the ship cooling systems is provided in Figure XXX.
-
-All values not directly measured in the cooling systems were calculated based on mass and energy balances.
-
-Figure cooling systems
-
-\subsection{Estimation of the electric power demand} \label{sec:met:electric}
-
-The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy
-\begin{equation}
-P_{el,tot} = \sum_{1}^{4} AE_i
-\end{equation}
-
-The determination of how different consumers contribute to the total electric power demand is less trivial.
-The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc.
-None of the individual contributions is directly measured, hence estimations need to be based on indirect measurements.
-
-\subsubsection{HVAC systems}
-
-For what concerns the air conditioning unit demand (HVAC), the observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a clear period, corresponding to the summer months, during which there is a significant increase of the demand.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pel_vs_time}
-	\caption[Total electric power demand versus time (Daily average)]{Total electric power demand versus time (Daily average)}
-	\label{fig:pelvstime}
-\end{figure}
-
-Since the total demand seems to be rather constant otherwise, it can be concluded that the demand of the HVAC compressors related to the need for cooling is concentrated during the summer months.
-Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a random summer and winter day, shows that the daily variation is comparable.
-This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/PelRel_vs_time}
-	\caption{Instantaneous power demand over daily average, winter versus summer day}
-	\label{fig:pelrelvstime}
-\end{figure}
-
-Hence, the HVAC electric power demand was estimated as follows:
-\begin{equation}
-P_{el,HVAC} =
-\begin{cases}
-P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
-0, & \text{otherwise}
-\end{cases}
-\end{equation}
-where $P_{el,ref}(t)$ is calculated as:
-\begin{equation}
-P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
-\end{equation}
-
-\subsubsection{Thrusters}
-
-When entering port areas, the ship needs to use thrusters to maneuver and berth.
-In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.
-In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
-From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).
-
-By comparing the reference profile to the instantaneous one, the points where the former is more than 10\% higher than the latter are identified as "thruster-on" points and treated consequently.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Thruster_selection}
-	\caption{Example of the procedure of selection of thruster power demand}
-	\label{fig:thruster_selection}
-\end{figure}
-
-\subsection{Estimation of the heat demand} \label{sec:met:heat}
-
-As the heat demand is not measured, it is necessary to determine it based on the available indirect measurements.
-
-The heat demand and generation can be summarized according to the following equations:
-\begin{eqnarray}
-\dot{Q}_{gen} & = & \dot{Q}_{EGB} + \dot{Q}_{HTHR} + \dot{Q}_{AB} \\
-\dot{Q}_{dem} & = & \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \\
-\end{eqnarray}
-
-The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.999\linewidth]{Figures/Heating_cooling_systems}
-	\caption{General view of the ship heating and cooling systems}
-	\label{fig:ship:general}
-\end{figure}
-
-\subsubsection{Heat balance parameter estimation}
-
-As not enough information and measurements are available to determine the various components of the heat balance, in this paper we determined them by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters.
-The parameter estimation problem is hence written as a minimization problem:
-\begin{eqnarray}
-min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
-\end{eqnarray}
-
-where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections.
-A list of the parameters $\textbf{p}$ is shown in Table REF, together with the chosen upper and lower boundaries for the calibration procedure.
-\begin{table}
-	\centering
-	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
-		\hline 
-		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
-		\hline
-		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 &  \\ 
-		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 &  \\
-		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 &  \\ 
-		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 &  \\ 
-		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 &  \\
-		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 &  \\
-		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 &  \\ 
-		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 &  \\ 
-		\hline
-	\end{tabular}
-	\caption{Parameters optimized in the parameter estimation for the heat balance}
-	\label{tab:ParameterEstimation} 
-\end{table}
-
-\subsubsection{Heat demand}
-
-We calculated the heat demand as the sum of the contributions of the elements listed in the ship's heat balance documentation.
-As no direct measurement of these quantities was available in the dataset, they had to be estimated based on the following assumptions:
-\begin{table}
-	\centering
-	{\tablinesep=2ex\tabcolsep=10pt
-		\begin{tabular}{p{2.8cm}l}
-			\hline 
-			Heat flow name & Equation \\
-			\hline
-			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
-			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
-			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
-			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
-			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
-			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
-			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
-			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
-			\hline
-	\end{tabular}}
-	\caption{Summary of the heat demand contributions and their calculation}
-	\label{tab:HeatDemand}
-\end{table}
-
-where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
-The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
-The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure REF.
-
-\subsubsection{Heat generation}
-
-The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty.
-The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
-\begin{equation}
-\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
-\end{equation}\label{eq:egb}
-
-where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF
-
-It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
-However, measurements of this contribution and of other variables that could lead to its straight-forward identification are missing.
-In this work, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},
-
-\begin{eqnarray}
-\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
-& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
-\end{eqnarray}
-
-where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
-The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
-
-The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems.
-The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit.
-This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value.
-In this calculation, the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
-
-\subsubsection{Parameter estimation and uncertainty quantification}
-
-The results of the parameter estimation are represented graphically and quantitatively in Figure REF and Table REF.
-
-Given the lack of input values for the estimation of the heat demand, the provided estimate based on the procedure described above should be integrated with an estimation of the uncertainty.
-
-In this study, we base the estimation of the uncertainty on the production side, as it is the one that has the largest amount of information available.
-In these regards, the uncertainty can be reduced to the contribution of three elements: $U(\dot{Q}_{EGB})$, $U(\dot{Q}_{HTHR})$, $U(\dot{Q}_{AB})$.
-
-The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$.
-In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT].
-The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions as described in Appendix REF, and can be estimated of being up to 10\%.
-The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value.
-These assumptions lead to the following estimation of the uncertainty:
-\begin{equation}
-\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
-\end{equation}
-
-That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements (result obtained for $\pm 20$ K uncertainty.
-The uncertainty is reduced to 20\% if the measurement uncertainty on the temperature is reduced to  $\pm 20$ K)
-
-The uncertainty of the heat generated by the ABs can be reduced to the contribution of two elements: the uncertainty on $\dot{m}_{fuel,AB}$ and that on $\eta_{AB}$.
-The former will be at least as large as the calibration error (35\%) and will be considered equal to 50\% to be conservative and accounting also for errors in the aggregated boiler fuel measurements.
-In addition, the uncertainty on the efficiency can be considered to be around 10\% based on the discrepancy between the considered sources and on the expected variability of the efficiency with load.
-Similarly to the previous case, the combination of these efficiencies lead to a total uncertainty of 51\%, where the main contribution comes from the uncertainty of the model output compared to the actual fuel consumption.
-
-The estimation of the uncertainty of $\dot{Q}_{HTHR}$ can be based on \ref{eqn:HTHR2} having assigned an acceptable variation of 20\% to the effectiveness of the heat exchanger, 20\% on the estimation of the reference mass flow, $\pm 5$ K uncertainty on water temperature measurements and considering negligible the uncertainty on $c_{p,w}$, leading to a 145\% uncertainty on this measurement.
-
-\subsection{Determination of typical operational days}
-
-\section{Results (and discussion)}
-
-\subsection{Exploratory analysis} \label{sec:res:exploratory}
-
-In this section, we observe and analyze direct measurements from ship systems, with no pre-processing, that are expected to have an influence on the energy analysis.
-
-Figures \ref{fig:TseaTairTIME} and \ref{fig:TseaTairDIST} represent the time evolution and the distribution over the year of the ambient air and sea water temperature.
-The relatively low temperatures that are experienced by the case study ship during its operations suggest that heating demand can be expected to be higher than cooling demand.
-On the other hand, the fact that, in summer, air temperatures reach up to 26 degrees Celsius justify the existence of an HVAC unit that can also operate in cooling mode.
-
-\begin{figure}[htbp]
-	\centering
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centering
-		\includegraphics[width=0.999\linewidth]{Figures/Tsea_vs_time}
-		\caption{Time series}
-		\label{fig:TseaTairTIME}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{Figures/Tsea_hist}
-		\caption{Distribution}
-		\label{fig:TseaTairDIST}
-	\end{subfigure}
-	\caption{Statistical representation of measured air and sea water temperatures based on yearly data. Landsort, 2014}
-	\label{fig:TseaTair}
-\end{figure}
-
-Figure \ref{fig:vship_hist} represents the distribution of the ship speed, both in terms of statistical distribution and in terms of a reference 24h period.
-The ship operates almost for the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
-This can also be seen from Figure \ref{fig:Pme_vs_vship}, where the main cluster of operations at middle-low speed can be clearly identified, as well as the smaller area of operations at high speed.
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/vship_hist}
-	\caption{Yearly distribution of ship speed}
-	\label{fig:vship_hist}
-\end{figure}
-
-figure --> Operational mode distribution
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/Hist_mainEngines}
-	\caption{Yearly load distribution, main engines}
-	\label{fig:loadME_hist}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/Hist_auxEngines}
-	\caption{Yearly load distribution, auxiliary engines}
-	\label{fig:loadAE_hist}
-\end{figure}
-
-ANYTHING ELSE THAT CAN BE INTERESTING TO SHOW IN THE EXPLORATORY DATA ANALYSIS?
-
-\subsection{Energy analysis} \label{sec:res:energy}
-
-Here we can see the typical example (note: the day was not selected as "typical", but simply randomly) of the energy demand during a winter and a summer day (Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}).
-It can be seen clearly how, as expected, the heating demand in winter is much higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduce need for compartment heating.
-On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors.
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_W}
-	\caption{Energy demand during a day of operations, Winter}
-	\label{fig:Demand_vs_time_W}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_S}
-	\caption{Energy demand during a day of operations, Summer}
-	\label{fig:Demand_vs_time_S}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyDemand}
-	\caption{Energy demand, aggregated data for one year of operations}
-	\label{fig:Pie_EnergyDemand}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyGeneration}
-	\caption{Energy generation, aggregated data for one year of operations}
-	\label{fig:Pie_EnergyGeneration}
-\end{figure}
-
-From Figure \ref{fig:Pie_EnergyDemand} it can be noted that nearly 50\% of the total energy demand is related to ship propulsion, while the remaining portion is approximately equally split between electric energy and heat demand.
-
-Within the electric energy demand, the HVAC systems only represent a minor contribution (XX\%), as it is only present for a few months during summer, and at rather low power demand.
-This is not surprising, since it can be observed that (see Figure \ref{fig:TseaTairDIST}) the air is always below 27$^o$C, and rarely above 17$^o$C.
-Also the thrusters, although they represent a rather high instantaneous demand, they are only used for a very short time and, hence, their total contribution to the ship energy demand is limited to XX\%.
-As a consequence, most of the electric energy demand should be allocated to other consumers, and efforts should be directed towards providing additional measurements of how different users contribute to the total electric power demand.
-
-\subsection{Heat demand and generation} \label{sec:res:heat}
-
-FIGURE --> Pie, heat generated by source
-FIGURE --> filled lines, heat generated by source.
-
-Looking at Figure \ref{fig:HeatGen_hist} it is possible to appreciate the amount of heat that is recovered from on board available waste heat (cooling water and exhaust gas).
-It can be observed that the amount of heat recovered from the HT cooling systems is much lower than expected, with an average of around 350 kW.
-This value is quite unreasonable compared to the values found in the technical specifications from the yard, that go from 700 kW (when only one auxiliary engine is operated) to 4800 kW (when all main engines and two auxiliary engines are operated).
-\\
-\begin{figure}
-	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/HeatGen_hist}
-	\caption{Distribution of thermal power generated from on board waste heat recovery}
-	\label{fig:Hist_WHR}
-\end{figure}
-
-Regarding the heat demand onboard, it can be seen that the HVAC re-heater represents the largest contributor (XX\%) followed by the hot water heater (XX\%).
-The large contribution of the HVAC reheater can be explained by the former being important for long times and in relation to the low ambient temperatures encountered by the case study vessel during operation.
-The relatively large contribution of the hot water heater can instead be related to the fact that these systems are rather important consumers and are necessary during the whole ship operations (i.e.
-also during the summer months).
-
-\subsection{Exergy Analysis} \label{sec:res:exergy}
-
-The results of the exergy analysis of the ship energy systems are shown in Figure REF and Tables REF to REF.
-
-Table(s) --> Exergy efficiencies
-- General at system level
-- One per system (MEs, AEs, CoolingSystems, SteamSystems)
-
-Figure --> Grassman diagram
-
-\subsection{Typical operational days}  \label{sec:res:typicalDays}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-\appendix
-
-\section{Full tables of energy and exergy flows}
-
-\section{Full equation list}
-
-\begin{table}[htbp]
-	\centering
-	\begin{tabular}{p{4cm}l}
-		\hline
-		\multirow{2}{*}{Before compressor}		& $ T_{comp,in} = $ Measured \\
-												& $ p_{comp,in} = $ Measured \\
-		\multirow{3}{*}{After compressor}		& $ T_{comp,out} = T_{comp,in} + \frac{T_{comp,out,is} - T_{comp,in} }{\eta_{comp,is}}  $  \\
-												& $ p_{comp,out} = $ Measured \\
-												& $ \eta_{comp,is} = P_2(\frac{p_{comp,out}}{p_{comp,in}}) $ \\
-		\multirow{3}{*}{After CAC-HT stage}		& $ p_{CAC-HT,out} = p_{comp,out} $ \\
-												& $ T_{CAC-HT,out} = T_{comp,out} - \frac{\dot{Q}_{CAC,HT}}{\dot{m}_{air,cyl}} $ \\
-												& $ \dot{Q}_{CAC,HT} = P_2(\lambda) $ \\
-		\multirow{3}{*}{After CAC-LT stage}		& $ p_{CAC-LT,out} = p_{comp,out} $ \\
-												& $ T_{CAC-LT,out} = T_{CAC-HT,out} - \frac{\dot{Q}_{CAC,LT}}{\dot{m}_{air,cyl}} $ \\
-												& $ \dot{Q}_{CAC,LT} = P_2(\lambda) $ \\
-		\multirow{3}{*}{After CAC-LT stage}
-
-	\end{tabular}
-\end{table}
-
-\section{Estimation of air and exhaust gas flows in the main engines}
+\section{Initial system}
 
@@
 
-Include other assumptions:
-- TC efficiency
-- Compressor efficiency
-- Turbine efficiency
+\section{Calculating the explicit system}
+
+To calculate what we need, we must make the system explicit on the variables we need to calculate.
+This process is different for the main and auxiliary engines, mostly because the temperature measurement of the exhaust gas before the turbine is positioned differently
 
@@
 
-To calculate what we need, we must make the system explicit on the variables we need to calculate.
-This process is different for the main and auxiliary engines, mostly because the temperature measurement of the exhaust gas before the turbine is positioned differently.
 In the case of the main engines, the temperature at the \textbf{cylinder outlet} is measured and, hence, known.
@@
 \end{multline}
+
+We now have to simplify this equation in order to make the term $ \dot{m}_{air,bypass} $ explicit. 
+\begin{comment}
+One first thing we can do is noticing that the fraction:
+$$
+\frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}}
+$$
+Can be rewritten as:
+$$
+\frac{1}{1 + \frac{\dot{m}_{fuel,cyl}}{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}}
+$$
+Since the numerator of the "subfraction" is much smaller than the denominator, we can make the assumption that this fraction behaves as:
+$$
+\frac{1}{1 + \epsilon}
+$$
+For which we know that:
+$$
+\frac{1}{1 + \epsilon} \approx 1 - \epsilon 
+$$
+That corresponds to:
+$$
+1 - \frac{\dot{m}_{fuel,cyl}}{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}
+$$
+\begin{multline}
+\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\ 
+(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} \left[T_{turb,out} + \left( 1 - \frac{\dot{m}_{fuel,cyl}}{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}} \right) \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} - T_0\right]
+\end{multline}
+\end{comment}
 
@@
 
-\section{Full list of available measurements}
-
-\bibliographystyle{unsrt}
-\bibliography{bibliography}
-
-%% \begin{thebibliography}{widestlabel}
-%% \bibitem{label}
-%% Text of bibliographic item
-%%\end{thebibliography}
-
 \end{document}
-\endinput
-%%
-%% End of file `elsarticle-template-num.tex'.

### 2017-11-26 Main.py

@@ -60,6 +60,7 @@ import auxiliaryDemand as aux
 import coolingsystems as cs
 from helpers import d2df
 import export as ex
+import clustering
 import postProcessing as post
 
 #%%
@@ -110,22 +111,23 @@ elif do_data_processing == "yes":
     # Calculating the auxiliary power demands: heating and electric power
     processed = aux.auxPowerAnalysis(processed, CONSTANTS, dict_structure, dataset_raw, header_names)
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-2")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-3")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Demands", "Demands-1")
     # Calculating the central cooling systems
     processed = cs.centralCoolingSystems(processed, CONSTANTS)
-    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-3")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-4")
     processed = cs.seaWaterCoolers(processed, CONSTANTS, dict_structure)
     # Calculating energy and exergy properties
     processed = ea.energyAnalysisLauncher(processed, dict_structure, CONSTANTS)
     # Re-doing the calculation of the connected points
-    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-4")
+    processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Other-5")
     processed = ff.systemFill(processed, dict_structure, CONSTANTS, "Other", "Demands-2")
     # Saving the processed data
     processed.to_hdf(CONSTANTS["filenames"]["dataset_output"], "processed", format='fixed', mode='w')
     # Result check log
     cc.systemCheck(processed, CONSTANTS, dict_structure, dataset_raw)
     # Efficiency results export
-    ex.exportEfficiecies(processed, CONSTANTS, dict_structure)
+
 
 #%%
 ######################################
@@ -133,10 +135,15 @@ elif do_data_processing == "yes":
 ######################################
 
 processed = ppo.operationalModeCalculator(processed, dataset_raw, CONSTANTS, header_names)
+processed = ppo.seasonCalculator(processed)
+exported = ex.exportAggregatedEyergyFlows(processed, CONSTANTS, dict_structure)
[…]

### 2017-11-26 Python files/Energyanalysis.py

@@ -132,7 +132,7 @@ def efficiencyCalculator(processed, dict_structure, CONSTANTS):
     df_index = processed.index
     for system in dict_structure["systems"]:
         for unit in dict_structure["systems"][system]["units"]:
-            if system+":"+unit == "Steam:SteamDistribution":
+            if system+":"+unit == "CoolingSystems:LTdistribution13":
                 aaa = 0
             # Here I create eight (8) series: 4 Edot and 4 Bdot. 2 "Full" and 2 "useful".
             temp_flow_list = {"Edot_in", "Edot_in_useful", "Edot_out", "Edot_out_useful", "Bdot_in", "Bdot_in_useful", "Bdot_out", "Bdot_out_useful"}
@@ -171,7 +171,7 @@ def efficiencyCalculator(processed, dict_structure, CONSTANTS):
                     else:
                         text_file.write("Flow {}:{}:{} is not recognised as either USEFEUL input or output \n".format(system, unit, flow))
             # Saving the total energy flows
-            dict_structure["systems"][system]["units"][unit]["E_in"] = temp_df["Edot_in"].sum() * 60 * 15 * 1e-6
+                        dict_structure["systems"][system]["units"][unit]["E_in"] = temp_df["Edot_in"].sum() * 60 * 15 * 1e-6
             dict_structure["systems"][system]["units"][unit]["E_out"] = temp_df["Edot_out"].sum() * 60 * 15 * 1e-6
             dict_structure["systems"][system]["units"][unit]["E_in_useful"] = temp_df["Edot_in_useful"].sum() * 60 * 15 * 1e-6
             dict_structure["systems"][system]["units"][unit]["E_out_useful"] = temp_df["Edot_out_useful"].sum() * 60 * 15 * 1e-6

### 2017-11-26 Python files/clustering.py

@@ -1,0 +1,116 @@
+from sklearn.cluster import KMeans
+from sklearn.metrics.pairwise import pairwise_distances
+import random
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+
+def clusteringTest(exported, method, nTests, nTestClusters):
+    # nTestClusters is a tuple with the range of clusters to be tested. So, for instance, it is (10,20) if a minimum of 10 to a max of 20 clusters are to be tested
+    results = np.zeros((nTests,nTestClusters[1]-nTestClusters[0]))
+    for idx in range(nTests):
+        temp = punctualClustering(exported, method, nTestClusters)
+        results[idx][:] = temp
+    plt.boxplot(results)
+    plt.show()
+
+
+def punctualClustering(exported, method, nTestClusters):
+    storeResults = []
+    nAttempts = 10
+    X = pd.DataFrame(index=exported.index)
+    X["totalElectricDemand"] = exported["totalElectricDemand"]
+    X["heatDemandLowTemperature"] = exported["heatDemandLowTemperature"]
+    X["heatDemandHighTemperature"] = exported["heatDemandHighTemperature"]
+    # Doing the clustering X times for each number of clusters
+    clusteringEvaluation = [] # Estimation of the goodness-of-fit for each cluster
+    clusterCenters = [] # Storage of the cluster centers
+    clusterDuration = [] # Storage of the duration of the time steps for each cluster
+    clusterNumber = [] # Storing the number of clusters used
+    for nClusters in range(nTestClusters[0], nTestClusters[1]):
+        if method=="kmeans":
+            [labels, centroids] = kmeansClustering(X, nClusters)
+        elif method=="kmedoids":
+            [labels, centroids] = kmedoidsClustering(X, nClusters)
+        else:
+            print("Error in the punctualClustering function: the method given as input should be either kmeans or kmedoids")
+        # Adding the centers to the output
+        clusterCenters.append(centroids)
+        clusterNumber.append(nClusters)
+        distanceMatrix = np.zeros((nClusters,len(X.keys())+1))
+        clusterDurationVector = np.zeros((nClusters))
+        for iCluster in range(nClusters):
+            # First, let's count how many clusters of each type there are
+            clusterNumber.append(iCluster)
[…]

### 2017-11-26 Python files/coolingsystems.py

@@ -68,48 +68,63 @@ def engineCoolingSystemsCalculation(processed, CONSTANTS, engine_type):
 
 
 def centralCoolingSystems(processed, CONSTANTS):
+
     # Doing the calculations for the 1-3 engine room
     ER13set = {"AE1", "AE3", "ME1", "ME3", "AB1"}
     # Calculating the temperature at the LT collector outlet
+    mdot_temp = sum(processed[d2df("CoolingSystems", "LTcollector13", "LTWater_" + idx + "_in", "mdot")] for idx in ER13set)
     processed[d2df("CoolingSystems", "LTcollector13", "LTWater_out", "T")] = (
         (sum(processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "T")] *
             processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "mdot")] for idx in ER13set)) / (
-        sum(processed[d2df("CoolingSystems", "LTcollector13", "LTWater_"+idx+"_in", "mdot")] for idx in ER13set)))
+            mdot_temp))
+    processed.loc[mdot_temp == 0, d2df("CoolingSystems", "LTcollector13", "LTWater_out", "T")] = processed['T_0']
     # The temperature at the other output of the collector is the same
     processed["CoolingSystems:LTcollector13:HTWater_out:T"] = processed["CoolingSystems:LTcollector13:LTWater_out:T"]
-    # Note that the temperature at the LT distribution inlet is calculated based on the average temperature at the LT collector outlets
+    # Note that the temperature at the LT distribution inlet is calculated based on the average temperature at the LT collector outlets.
+    # The idea is that we have different temperatures at different instants, and we need to average them
+    mdot_temp = sum(processed[d2df("CoolingSystems", "LTdistribution13", "LTWater_" + idx + "_out", "mdot")] for idx in ER13set)
     processed[d2df("CoolingSystems", "LTdistribution13", "LTWater_in", "T")] = (
         sum(processed[d2df("CoolingSystems", "LTdistribution13", "LTWater_" + idx + "_out", "T")] *
[…]

### 2017-11-26 Python files/input.py

@@ -18,8 +18,9 @@ def filenames(project_path):
     output["dataset_output"] = project_path + os.sep +'Data_Process' + os.sep + 'database_out.h5'
     output["consistency_check_report"] = project_path + os.sep + 'Data_Process' + os.sep + 'check_report.txt'
     output["aggregated_efficiencies"] = project_path + os.sep + 'Results' + os.sep + 'aggregated_efficiencies.csv'
-    output["aggregated_flows_total"] = project_path + os.sep + 'Results' + os.sep + 'aggregated_flows_total.csv'
-    output["aggregated_flows_opmode"] = project_path + os.sep + 'Results' + os.sep + 'aggregated_flows_opmode.csv'
+    output["aggregated_flows_energy"] = project_path + os.sep + 'Results' + os.sep + 'aggregated_flows_energy.csv'
+    output["aggregated_flows_exergy"] = project_path + os.sep + 'Results' + os.sep + 'aggregated_flows_exergy.csv'
+    output["flows_clustering"] = project_path + os.sep + 'Results' + os.sep + 'flows_clustering.csv'
     output["latex_table"] = project_path + os.sep + 'Manuscript' + os.sep + 'Figures' + os.sep + 'table.tex'
     # opening the report file and inititing it
     text_file = open(output["consistency_check_report"],"w") # Cleaning the file
@@ -63,6 +64,7 @@ def assumptions(raw, processed, CONSTANTS, hd):
     # Others
     processed.loc[:,"T_0"] = raw[hd["water_T_forsmark_smhi-opendata"]] + 273.15
     processed.loc[:,"T_air"] = raw[hd["air_T_sv_hogarna_smhi-opendata"]] + 273.15
+    processed.loc[:,"ShipSpeed"] = raw[hd["SHIP_SPEED_KNOT_"]]
     processed[d2df("CoolingSystems","SWC13","SeaWater_out","T")] = raw[hd["SWC13_SW_T_OUT"]] + 273.15 # CHECK IF IT IS IN OR OUT
     processed[d2df("CoolingSystems","SWC24","SeaWater_out","T")] = raw[hd["SWC24_SW_T_OUT"]] + 273.15 # CHECK IF IT IS IN OR OUT
     # Boilers

### 2017-11-26 Python files/preprocessingO.py

@@ -6,13 +6,23 @@ import CoolProp.CoolProp as cp
 
 
 def operationalModeCalculator(processed, raw, CONSTANTS, hd):
+    # This function assigns to each point an operational mode based on a set of assumptions
     # Initializing the data series
     temp = pd.Series(index=processed.index)
     temp[raw[hd["SHIP_SPEED_KNOT_"]] > 15] = "High Speed Sailing"
     temp[(raw[hd["SHIP_SPEED_KNOT_"]] < 15) * (raw[hd["SHIP_SPEED_KNOT_"]] > 4)] = "Low Speed Sailing"
     temp[((processed["Demands:Electricity:Thrusters:Edot"] > 0) + ((raw[hd["SHIP_SPEED_KNOT_"]] < 4) * (raw[hd["SHIP_SPEED_KNOT_"]] > 2)))] = "Maneuvering"
     temp[temp.isnull()] = "Port/Stay"
-    processed["operationalMode"] = temp
+    processed["OperationalMode"] = temp
+    return processed
+
+def seasonCalculator(processed):
+    # This function assigns to each point a season identifier based on a set of assumptions
+    temp = pd.Series(index=processed.index)
+    temp[processed["T_0"] < 280] = "Winter"
+    temp[processed["T_0"] >= 280] = "Mid-Season"
+    temp[processed["Demands:Electricity:HVAC:Edot"] > 0] = "Summer"
+    processed["Season"] = temp
     return processed
 
 

### 2018-04-01 Manuscript/EngineModellingAppendix.tex
Updates in the manuscript

--- 2017-10-10 Manuscript/Manuscript.tex
@@
-%% 
-%% Copyright 2007, 2008, 2009 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
-
-%% Template article for Elsevier's document class `elsarticle'
-%% with numbered style bibliographic references
-%% SP 2008/03/01
-
-\documentclass[preprint,12pt]{elsarticle}
-
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn]{elsarticle}
-%% \documentclass[final,3p,times]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn]{elsarticle}
-%% \documentclass[final,5p,times]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
-%% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-\usepackage{amsthm}
-%% Tthe amsmath package provides the "align" environment for nicer equation formatting
-\usepackage{amsmath}
-%% The package tabls helps in having nicer tables with higher spaces
-\usepackage{tabls}
-%% The package multirow allows merging together multiple cells vertifcally
-\usepackage{multirow}
-%% 
-\usepackage{subcaption}
-
-\def\bibsection{\section*{References}}
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-%% \usepackage{lineno}
-
-\journal{Energy}
-
-\begin{document}
-
-\begin{frontmatter}
-
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \address{Address\fnref{label3}}
-%% \fntext[label3]{}
-
-\title{Energy and exergy analysis of a cruise ship}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \address[label1]{}
-%% \address[label2]{}
-
-\author[EPFL]{Francesco Baldi}
-\author[Linnaeus]{Fredrik Ahlgren}
-\author[DTU]{Tuong-Van Nguyen}
-\author[CTU]{Cecilia Gabrielii}
-\author[CTU]{Karin Andersson}
-
-\address[EPFL]{Industrial Process Energy Systems Engineering (IPESE), \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, 1950, Sion, Switzerland}
-\address[Linnaeus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
-\address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
-\address[CTU]{Department of Mechanics and Maritime Sciences, Chalmers University of technology, Gothenburg, Sweden}
-
-\begin{abstract}
-%% Text of abstract
-
-\end{abstract}
-
-\begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-
-%% PACS codes here, in the form: \PACS code \sep code
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-
-low carbon shipping \sep energy analysis \sep exergy analysis \sep energy efficiency
-
-\end{keyword}
-
-\end{frontmatter}
-
-%% \linenumbers
-
-%% main text
-
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
-%% \appendix
-
-%% \section{}
-%% \label{}
-
-%% If you have bibdatabase file and want bibtex to generate the
-%% bibitems, please use
-%%
-%%  \bibliographystyle{elsarticle-num} 
-%%  \bibliography{<your bibdatabase>}
-
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
-
-\section{Introduction} \label{sec:introduction}
-
-\subsection{Background}
-
-CO$_2$ emissions from shipping in 2012 amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions \cite{Smith2014}.
-Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios \cite{Smith2014}.
-As an example, global transport demand has increased by 3.4\% in 2014, compared to a global GBP growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}.
-The OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3].
-The fact that shipping needs to even further reduce its CO$_2$ emissions in the near future is essential for being able to achieve the goals of maintaining the climate below a 2-degree level in 2050 [4].
-
-In addition to considerations related to GHG emissions, an emission control area (ECA) is enforced in the Baltic Sea  by the International Maritime Organisation since January 2015.
-This ECA stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels.
-More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}
-
-Altogether, these conditions present a challenge to the shipping companies who attempt to reduce their fuel consumption, environmental impact, and operative costs.
-A wide range of fuel saving solutions for shipping are available and partially implemented in the existing fleet, both from the design and operational perspective; several specific studies have been conducted on these technologies, and a more detailed treatise would be out of the scope of this work.
-In this context, it has been acknowledged that the world fleet is heterogeneous, and measures need to be evaluated on a ship-to-ship basis \cite{Bouman2017}.
-In this process, a deeper understanding of energy use on board of the specific ship is vital.
-
-\subsection{Previous work}
-
-The idea of improving the understanding of the behavior of the energy system of a ship is not new.
-Most of the work published around this subject relates to the use  of mathematical models of the ship systems.
-
-Most authors focused on the propulsion part of the problem, as this is often the most relevant energy demand on board.
-Shi et al.
-proposed a modeling approach for predicting ship fuel consumption of a cargo/passenger ferry \cite{Shi2009}; Theotokoats and Tzelepis applied a similar procedure to the case of a Handymax product carrier \cite{Theotokatos2015}, similarly to Tillig et al., who also added the dynamic element to their predictive model \cite{Tillig2016}.
-
-The work referenced above allowed improving the understanding of how many operational (such as the ship's speed) and environmental (such as wave height and wind speed) parameters influence the ship's energy performance.
-These studies, however, are not based on actual measurements of the ship's operations.
-Also, they focus entirely on the ship's power demand for propulsion.
-This is a very reasonable practice for most ship types, given that propulsion represent the largest part of the total energy demand, but does not allow improving the knowledge of the remaining part of the system.
-
-% Data-driven analysis
-Other authors filled the gap by both including electric energy demand, and by basing their work on measurements from actual ship operations.
-The work presented by Thomas et al.
-\cite{Thomas2010} and Basurko et al.
-\cite{Basurko2013} shows the application of energy auditing methods to fishing vessels.
-This approach represents a step forward towards improving the understanding of how energy is used on board during actual ship operations.
-
-More generally, several authors have highlighted the importance of a detailed knowledge of the ship's operational profile in order to appropriately assess and optimize possible alternatives for improving ship energy efficiency \cite{Banks2013}.
-Coraddu et al., started from a statistical analysis of measured ship operations and used the aggregated data, together with a computational model of the vessel, to provide a better prediction of the actual ship's operational efficiency \cite{Coraddu2014}.
-The importance of considering the operational profile was also showed in the case of the optimization of engine-propeller interaction \cite{Baldi2015c}, in the process of retrofitting existing systems \cite{Baldi2015b,Choi2013}, and in ship design \cite{Ghassemi2017,Solem2015}.
-
-% Heat availability
-Heat demand is rarely a subject of concern on board ships, with some notable exceptions.
-This is due to a combination of generally low demand and high availability from the waste heat of the engines.
-In a previous publications by the authors, for instance, it is shown that in the case of a product tanker, although the heat demand was estimated to account for roughly 20\% of the total energy demand of the ship on a yearly basis, it only contributes to 4.1\% of the total fuel consumption (contribution of the auxiliary boilers), while the rest of the demand is fulfilled using waste heat \cite{Baldi2015a}.
-
-%% WHR and heat availability
-It should be noted that, however, much research effort has been devoted, especially in recent times, to the improvement of the efficiency of ship energy systems by recovering the waste heat available from the engines.
-With reference to different types of technologies, case studies, and designs,the several authors showed the existance of a quite significant potential for energy saving when WHR systems are employed, ranging from around 1\% for single-pressure steam cycles applied to two-stroke engines \cite{Theotokatos2012} to more complex systems based on ORCs (up to 10\%, \cite{Hountalas2012}) or including the cooling systems as a source of waste heat (over10\%, \cite{Dimopoulos2012}).
-The case of the installation of an ORC on board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2015}, showing a similar potential.
-
-The potential uses for waste heat on board are not limited to improving the efficiency of the power plant.
-Waste heat is commonly used for fulfilling on board heat demand for spatial heating and freshwater generation \cite{Molland2011,Baldi2015a,Mccarthy1990};  Balaji et al.
-proposed the use of waste heat for ballast-water systems \cite{Balaji2012}; Salmi et al.
-suggested its use for adsorption refrigeration systems \cite{Salmi2017}.
-A detailed review of potential uses for waste heat from marine engines is presented by Shu et al \cite{Shu2013}.
-
-% Cruise ships
-Some ship types constitute notable examples.
-On cruise ships the heat demand is significantly higher compared to standard cargo ships.
-Referring to winter conditions, Marty et al.
-estimated the instantaneous heat demand of a selected cruise ship to reach roughly 23 MW, compared to an estimated peak of 49 MW for propulsion and electric auxiliaries combined \cite{Marty2012}.
-
-This shows how the heat from the engines should be considered as a potential resource, rather than a waste, and that there is potential for improvement based on the optimal use of these heat sources.
-The work presented by the authors in \cite{Baldi2016} represent a step in this direction; however, there is the need for an increased detail in the estimation of the heat demand.
-
-% Exergy analysis
-Exergy analysis provides a more accurate estimation of the potential for energy recovery on board.
-The application of exergy analysis to the case of ship energy systems is, however, still limited.
-Dimopoulos et al.
-showed how the process optimizing the WHR system of a container ship can be more efficient if exergy efficiency, rather than energy efficiency, is used as the target of the optimization \cite{Dimopoulos2012}; Baldi et al.
-also analyzed the exergy flows on board of a product tanker, showing that this allows having a more accurate understanding of what parts of the system show potential for improvement \cite{Baldi2015a}; similar results were obtained by Marty et al., who focused on the power plant of a cruise ship \cite{Marty2016}.
-Koroglu et al.
-made a step futher also including advanced exergy analysis in their study \cite{Koroglu2017}.
-
-\subsection{Aim}
-
-Given the existing limitations of the current available information in scientific literature highlighted in the previous section, the objective of this work is the following:
-\begin{itemize}
-	\item Analyse the demand of a cruise ship in terms of propulsion power, electric power and heat, based on operational measurements.
-The lack of existing information with reference to the heat demand is considered a particular element of novelty when compared to existing literature in the subject.
-	\item Analyse the current efficiency of the system, and potential ways to improve it, by means of applying exergy analysis.
-	\item Provide reference operational conditions for further use in the field of energy systems optimization.
-\end{itemize}
-
-\section{Method} \label{sec:method}
-
-In this paper, we present the application of energy and exergy analysis (described in section \ref{sec:met:energyExergy}) to a cruise ship.
-This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; the specific assumptions and methods used for processing the available information into  energy and exergy flows are specified in Section \ref{sec:met:processing}.
-In these regards, the estimation of the heat demand is treated in a separate section (Sec.
-\ref{sec:met:heat}) as it constitutes a particularly challenging task and it is one of the central aspects of this article.
-
-\subsection{Energy and exergy analysis} \label{sec:met:energyExergy}
-
-\subsubsection{Energy analysis}
-
-\emph{Energy} can be stored, transformed from one form to another (e.g.
-heat to power) and transferred between systems, but can neither be created nor destroyed (conservation law)~\textbf{Bejan et al.}.
-The system under study is the \emph{ship energy system} and is thus taken as control volume.
-The energy balance can be expressed as:
-
-\begin{align}
-	\sum_{\mathrm{in}} \dot{H}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{H}_{\mathrm{out}} \\
-	\dot{H}_{\mathrm{fuel}} + \dot{H}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}+\dot{W}_{\mathrm{el}}+\dot{Q}_{\mathrm{heating}}
-\end{align}
-
-The left-hand side term represents, on a time rate basis, the energy associated with the fuel consumed in the boilers and engines ($\dot{H}_{\mathrm{fuel}}$) and the air used in the combustion processes ($\dot{H}_{\mathrm{air}}$).
-The right-hand side term denotes the power ($\dot{W}_{\mathrm{el}}$) and heat ($\dot{Q}_{\mathrm{heating}}$) required on-site (e.g.
-propulsion and fuel heating), and the heat discharged into the environment ($\sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}$) with, for instance, the exhaust gases.
-
-The energy flow associated with a material stream is calculated as the sum of the physical and chemical enthalpies, and kinetic and potential energies are neglected.
-The physical energy is taken as the relative enthalpy, as underlined in~\textbf{Kotas et al.}, and the chemical energy is taken as the lower/higher heating value.
-The environmental conditions taken for the present analysis are the ambient pressure and seawater temperature.
-
-\subsubsection{Exergy analysis}
-
-\emph{Exergy} may be defined as the `maximum theoretical useful work (shaft work or electrical work) as the system is brought into complete thermodynamic equilibrium with the thermodynamic environment while the system interacts with it only'~\citep{Moran1989}.
-Unlike energy, exergy is not conserved but some is destroyed because of the irreversible phenomena taking place in real processes (e.g.
-chemical reactions like combustion).
-The exergy balance for the system under study can be expressed as:
-
-\begin{align}
-	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \\
-	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d} 
-\end{align}
-
-The left-hand side term represents, on a time rate basis, the exergy associated with the fuel and air.
-The right-hand side term denotes the exergy of the waste streams, the exergy transfers with heat and power, and the exergy destroyed in the ship system.
-Kinetic and potential exergies are neglected, and the exergy of a given material stream is derived as the sum of the physical and chemical exergies.
-The chemical exergy is calculated based on the reference environment of \textbf{Szargut (1989)}, and its value is approximatively equal to the higher heating value for hydrocarbon fuels.
-The exergy destruction can be calculated from the Gouy-Stodola theorem~\textbf:{Kotas (1995)}.
-
-The exergy balance may alternatively be formulated as:
-\begin{equation}
-\dot{E}_{\mathrm{p}} = \dot{E}_{\mathrm{f}} - \dot{E}_{\mathrm{d}} - \dot{E}_{\mathrm{l}} 
-\end{equation}
-
-where $\dot{E}_{\mathrm{p}}$ is called the exergy product, and corresponds to the desired output of the system, in exergy terms (for example, the power produced in an engine).
-$\dot{E}_{\mathrm{f}}$ denotes the exergy fuel, and represents the resources spent to drive the studied process (for instance, the fuel used in a combustion process).
-The last term $\dot{E}_{\mathrm{l}}$ corresponds to the losses of a system, such as the heat discharged into the environment with cooling water.
-
-\subsubsection{System performance}
-
-The following indicators are used to evaluate the system performance:
-\begin{itemize}
-	\item the exergy efficiency $\varepsilon$, defined as the ratio between the exergy product and fuel of a given component or system
-	\begin{equation} \varepsilon = \frac{\dot{E}_{\mathrm{p}}}{\dot{E}_{\mathrm{f}}} \end{equation}
-	\item the efficiency defect $\lambda$, presented in the work of Kotas, defined as the fraction of the total exergy input destroyed in the successive irreversible processes
-	\begin{equation} \lambda = \frac{\dot{E}_{\mathrm{d}}}{\dot{E}_{\mathrm{in,tot}}} \end{equation}
-	\item  the irreversibility share $\delta$, suggested in the work of Tsatsaronis, defined as the ratio between the exergy destroyed in the $i$-th component in relation to the exergy destroyed in the entire system
-	\begin{equation} \delta_i = \frac{\dot{E}_{\mathrm{d},i}}{\dot{E}_{\mathrm{d,tot}}} \end{equation}
-
-\end{itemize}
-
-\subsection{Case study vessel} \label{sec:met:case}
-
-The energy and exergy analysis are here applied to a specific  cruise ship operating daily cruises in the Baltic Sea between Stockholm and the island of \.{A}land.
-The ship is 176.9 m long and has a beam of 28.6 m, and has a design speed of 21 knots.
-The ship was built in Aker Finnyards, Raumo Finland in 2004.
-
-The ship has a capacity of 1800 passengers and has several restaurants, night clubs and bars, as well as saunas and pools.
-This means that the energy system regarding the heat and electricity demand is more complex than a regular cargo vessel in the same size.
-Typical ship operations, although they can vary slightly between different days, are represented in Figure \ref{fig:typicalShipOperations}.
-It should be noted that the ship stops and drifts in open sea during night hours before mooring at its destination in the morning, if allowed by weather conditions.
-
-The ship systems are summarized in Figure \ref{fig:shipSystems}.
-The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller.
-The main engines are four Wärtsilä 4-stroke Diesel engines (ME) rated 5850 kW each.
-All engines are equipped with selective catalytic reactors (SCR) for NOX emissions abatement.
-Propulsion power is needed whenever the ship is sailing; however, it should be noted that the ship rarely sails at full speed, and most of the time it only needs one or two engines operated simultaneously.
-
-Auxiliary power is provided by four auxiliary engines (AE) rated 2760 kW each.
-Auxiliary power is needed on board for a number of alternative functions, from pumps in the engine room to lights, restaurants, ventilation and entertainment for the passengers.
-
-Auxiliary heat needs are fulfilled by the exhaust gas steam generators (HRSG) located on all four AEs and on two of the four MEs or by oil-fired auxiliary boilers (mainly when in port, or during winter), by the heat recovery on the HT cooling water systems (HRHT), and by the auxiliary, oil fired boilers (AB).
-The heat is needed for passenger and crew accommodation, as well as for the heating of the highly viscous heavy fuel oil used for engines and boilers.
-This last part, however, is drastically reduced since the 1st of January 2015, as new regulations entering into force require the use of low-sulphur fuels, which require a much more limited heating.
-
-\subsection{Data gathering and pre-processing} \label{sec:met:gathering}
-
-The operational data was collected on board from the ships' machine logging and surveillance system.
-The on board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB.
-The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average.
-The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python \cite{mckinney-proc-scipy-2010}.
-A new structured naming of headers and a consistent time frequency over all data points were created, and 245 selected data points were saved in a HDF5 table time series database which is the base for the analysis.
-
-All data-points were checked individually by creating a descriptive statistic and a histogram of each data-point.
-A filter was applied by setting up a pre-defined maximum and minimum value.
-As we do not know, or had any meaningful way of deriving, each individual sensors placement, sensor type or calibration status all data must be filtered and checked accordingly for outliers or bad data.
-
-Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska H\"{o}garna for the ambient temperature.
-The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska H\"{o}garna lighthouse is along the ship route, in between the Swedish archipelago and Åland.
-The assumption was validated based on  June-December period, for which onboard measurements were available.
-This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
-The fit the SMHI-data with the rest of the data-set the SMHI data was resampled from 1h for the seawater and 3h for the ambient temperature, to 15min frequency using a linear interpolation.
-
-\subsection{Data processing: Ship energy system modeling} \label{sec:met:processing}
-
-Not all of the variables required to perform a full energy and exergy analysis of the system are available from measurements.
-In some cases, they are measurable, but not measured (e.g.
-some temperatures, mass flows, etc.).
-In other cases, they are simply impossible, or impractical, to measure (e.g.
-specific enthalpy, specific entropy).
-For this reason, part of the ship systems needed to be modeled in order to derive the unknown variables.
-
-\subsubsection{Diesel engines}
-
-The setup of the engines and their connections to the other ship systems are summarized in Figure REF and REF for the main and auxiliary engines, respectively.
-The defining net energy flows are listed in Table REF.
-It should be noted that, in the case of the main engines, the engine power output was not measured and needed to be estimated.
-In this work, we calculated the engine power output based on measurements of engine fuel rack position (used as a proxy of the mass of fuel injected per cycle) and of the engine speed:
-\begin{eqnarray}
-\dot{m}_{fuel} & = & \dot{m}_{fuel,des} \left(a_0 + a_1 \frac{frp}{frp_{des}}\right) \frac{\omega}{\omega_{des}} \\
-\eta_{ME} & = & a_0 + a_1 \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} + a_2 \left( \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} \right)^2 \\
-\dot{W}_{ME} & = & \dot{m}_{fuel} \eta_{ME} LHV
-\end{eqnarray}
-
-The validity of this assumption can be seen by observing Figure REF, representing the calculated total power from the main engines as a function of the ship speed
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/DieselEngines}
-	\caption{Schematic representation of the Diesel engines and their connections to the cooling systems}
-	\label{fig:DieselEngines}
-\end{figure}
-
-In the case of the auxiliary engines, instead, direct measurements of the power generated by each engine were available.
-
-For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine.
-The model is explained in detail in Appendix REF.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pme_vs_vship_Neng}
-	\caption{Scatter plot, Propulsion power versus ship speed}
-	\label{fig:Pme_vs_vship_Neng}
-\end{figure}
-
-ALSO INCLUDE THE POWER VS TC SPEED RELATIONSHIP
-
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{llccrrrrr}
-		\hline
-		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
-		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
-		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
-		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{comp}$		& - 	& M & & & & &  \\
-		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
-		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
-		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{exp}$		& - 	& C & & & & &  \\
-		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
-		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
-		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
-		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
-		&	$T_{eg,out}$  		& $^o$C & M  & & & & &  \\ 
-		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
-		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
-		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
-		\multicolumn{2}{l}{LOC} & & & & & & &  \\
-		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
-		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		\multicolumn{2}{l}{JWC} & & & & & & &  \\
-		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
-		\hline
-		\end{tabular}
-	\caption{Main engines, measured and calculated temperatures and flows at different engine loads}
-\end{table}
-
-INCLUDE VALIDATION AGAINST PROJECT GUIDE DATA
-
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{llccrrrrr}
-		\hline
-		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
-		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
-		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
-		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{comp}$		& - 	& M & & & & &  \\
-		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
-		&	$T_{air,in}$   		& $^o$C & M  & & & & &  \\ 
-		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
-		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
-		&   $\beta_{exp}$		& - 	& C & & & & &  \\
-		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
-		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
-		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
-		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
-		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
-		&	$T_{eg,out}$  		& $^o$C & C  & & & & &  \\ 
-		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
-		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
-		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
-		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
-		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
-		\multicolumn{2}{l}{LOC} & & & & & & &  \\
-		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
-		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
-		\multicolumn{2}{l}{JWC} & & & & & & &  \\
-		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
-		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
-		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
-		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
-		\hline
-	\end{tabular}
-	\caption{Auxiliary engines, measured and calculated temperatures and flows at different engine loads}
-\end{table}
-
-INCLUDE VALIDATION AGAINST PROJECT GUIDE DATA
-
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{lcrrrrr}
-		\hline
-		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		Power output & Mech &  &  &  &  &  \\
-		Exhaust gas & Heat &  &  &  &  &  \\
-		CAC-LT & Heat &  &  &  &  &  \\
-		CAC-HT & Heat &  &  &  &  &  \\
-		JWC & Heat &  &  &  &  &  \\
-		LOC & Heat &  &  &  &  &  \\
-		\hline
-	\end{tabular}
-	\caption{Main engines, energy flows at different engine loads}
-\end{table}
-
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{lcrrrrr}
-		\hline
-		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
-		\hline
-		Power output & Mech &  &  &  &  &  \\
-		Exhaust gas & Heat &  &  &  &  &  \\
-		CAC-LT & Heat &  &  &  &  &  \\
-		CAC-HT & Heat &  &  &  &  &  \\
-		JWC & Heat &  &  &  &  &  \\
-		LOC & Heat &  &  &  &  &  \\
-		\hline
-	\end{tabular}
-	\caption{Auxiliary engines, energy flows at different engine loads}
-\end{table}
-
-\subsubsection{Cooling systems}
-
-The ship's cooling systems are designed similarly to most ships and divided in two cooling systems of two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C).
-The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board.
-A representation of the ship cooling systems is provided in Figure XXX.
-
-All values not directly measured in the cooling systems were calculated based on mass and energy balances.
-
-Figure cooling systems
-
-\subsection{Estimation of the electric power demand} \label{sec:met:electric}
-
-The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy
-\begin{equation}
-P_{el,tot} = \sum_{1}^{4} AE_i
-\end{equation}
-
-The determination of how different consumers contribute to the total electric power demand is less trivial.
-The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc.
-None of the individual contributions is directly measured, hence estimations need to be based on indirect measurements.
-
-\subsubsection{HVAC systems}
-
-For what concerns the air conditioning unit demand (HVAC), the observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a clear period, corresponding to the summer months, during which there is a significant increase of the demand.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pel_vs_time}
-	\caption[Total electric power demand versus time (Daily average)]{Total electric power demand versus time (Daily average)}
-	\label{fig:pelvstime}
-\end{figure}
-
-Since the total demand seems to be rather constant otherwise, it can be concluded that the demand of the HVAC compressors related to the need for cooling is concentrated during the summer months.
-Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a random summer and winter day, shows that the daily variation is comparable.
-This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/PelRel_vs_time}
-	\caption{Instantaneous power demand over daily average, winter versus summer day}
-	\label{fig:pelrelvstime}
-\end{figure}
-
-Hence, the HVAC electric power demand was estimated as follows:
-\begin{equation}
-P_{el,HVAC} =
-\begin{cases}
-P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
-0, & \text{otherwise}
-\end{cases}
-\end{equation}
-where $P_{el,ref}(t)$ is calculated as:
-\begin{equation}
-P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
-\end{equation}
-
-\subsubsection{Thrusters}
-
-When entering port areas, the ship needs to use thrusters to maneuver and berth.
-In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.
-In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
-From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).
-
-By comparing the reference profile to the instantaneous one, the points where the former is more than 10\% higher than the latter are identified as "thruster-on" points and treated consequently.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Thruster_selection}
-	\caption{Example of the procedure of selection of thruster power demand}
-	\label{fig:thruster_selection}
-\end{figure}
-
-\subsection{Estimation of the heat demand} \label{sec:met:heat}
-
-As the heat demand is not measured, it is necessary to determine it based on the available indirect measurements.
-
-The heat demand and generation can be summarized according to the following equations:
-\begin{eqnarray}
-\dot{Q}_{gen} & = & \dot{Q}_{EGB} + \dot{Q}_{HTHR} + \dot{Q}_{AB} \\
-\dot{Q}_{dem} & = & \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \\
-\end{eqnarray}
-
-The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.999\linewidth]{Figures/Heating_cooling_systems}
-	\caption{General view of the ship heating and cooling systems}
-	\label{fig:ship:general}
-\end{figure}
-
-\subsubsection{Heat balance parameter estimation}
-
-As not enough information and measurements are available to determine the various components of the heat balance, in this paper we determined them by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters.
-The parameter estimation problem is hence written as a minimization problem:
-\begin{eqnarray}
-min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
-\end{eqnarray}
-
-where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections.
-A list of the parameters $\textbf{p}$ is shown in Table REF, together with the chosen upper and lower boundaries for the calibration procedure.
-\begin{table}
-	\centering
-	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
-		\hline 
-		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
-		\hline
-		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 &  \\ 
-		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 &  \\
-		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 &  \\ 
-		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 &  \\ 
-		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 &  \\
-		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 &  \\
-		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 &  \\ 
-		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 &  \\ 
-		\hline
-	\end{tabular}
-	\caption{Parameters optimized in the parameter estimation for the heat balance}
-	\label{tab:ParameterEstimation} 
-\end{table}
-
-\subsubsection{Heat demand}
-
-We calculated the heat demand as the sum of the contributions of the elements listed in the ship's heat balance documentation.
-As no direct measurement of these quantities was available in the dataset, they had to be estimated based on the following assumptions:
-\begin{table}
-	\centering
-	{\tablinesep=2ex\tabcolsep=10pt
-		\begin{tabular}{p{2.8cm}l}
-			\hline 
-			Heat flow name & Equation \\
-			\hline
-			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
-			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
-			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
-			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
-			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
-			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
-			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
-			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
-			\hline
-	\end{tabular}}
-	\caption{Summary of the heat demand contributions and their calculation}
-	\label{tab:HeatDemand}
-\end{table}
-
-where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
-The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
-The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure REF.
-
-\subsubsection{Heat generation}
-
-The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty.
-The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
-\begin{equation}
-\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
-\end{equation}\label{eq:egb}
-
-where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF
-
-It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
-However, measurements of this contribution and of other variables that could lead to its straight-forward identification are missing.
-In this work, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},
-
-\begin{eqnarray}
-\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
-& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
-\end{eqnarray}
-
-where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
-The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
-
-The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems.
-The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit.
-This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value.
-In this calculation, the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
-
-\subsubsection{Parameter estimation and uncertainty quantification}
-
-The results of the parameter estimation are represented graphically and quantitatively in Figure REF and Table REF.
-
-Given the lack of input values for the estimation of the heat demand, the provided estimate based on the procedure described above should be integrated with an estimation of the uncertainty.
-
-In this study, we base the estimation of the uncertainty on the production side, as it is the one that has the largest amount of information available.
-In these regards, the uncertainty can be reduced to the contribution of three elements: $U(\dot{Q}_{EGB})$, $U(\dot{Q}_{HTHR})$, $U(\dot{Q}_{AB})$.
-
-The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$.
-In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT].
-The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions as described in Appendix REF, and can be estimated of being up to 10\%.
-The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value.
-These assumptions lead to the following estimation of the uncertainty:
-\begin{equation}
-\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
-\end{equation}
-
-That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements (result obtained for $\pm 20$ K uncertainty.
-The uncertainty is reduced to 20\% if the measurement uncertainty on the temperature is reduced to  $\pm 20$ K)
-
-The uncertainty of the heat generated by the ABs can be reduced to the contribution of two elements: the uncertainty on $\dot{m}_{fuel,AB}$ and that on $\eta_{AB}$.
-The former will be at least as large as the calibration error (35\%) and will be considered equal to 50\% to be conservative and accounting also for errors in the aggregated boiler fuel measurements.
-In addition, the uncertainty on the efficiency can be considered to be around 10\% based on the discrepancy between the considered sources and on the expected variability of the efficiency with load.
-Similarly to the previous case, the combination of these efficiencies lead to a total uncertainty of 51\%, where the main contribution comes from the uncertainty of the model output compared to the actual fuel consumption.
-
-The estimation of the uncertainty of $\dot{Q}_{HTHR}$ can be based on \ref{eqn:HTHR2} having assigned an acceptable variation of 20\% to the effectiveness of the heat exchanger, 20\% on the estimation of the reference mass flow, $\pm 5$ K uncertainty on water temperature measurements and considering negligible the uncertainty on $c_{p,w}$, leading to a 145\% uncertainty on this measurement.
-
-\subsection{Determination of typical operational days}
-
-\section{Results (and discussion)}
-
-\subsection{Exploratory analysis} \label{sec:res:exploratory}
-
-In this section, we observe and analyze direct measurements from ship systems, with no pre-processing, that are expected to have an influence on the energy analysis.
-
-Figures \ref{fig:TseaTairTIME} and \ref{fig:TseaTairDIST} represent the time evolution and the distribution over the year of the ambient air and sea water temperature.
-The relatively low temperatures that are experienced by the case study ship during its operations suggest that heating demand can be expected to be higher than cooling demand.
-On the other hand, the fact that, in summer, air temperatures reach up to 26 degrees Celsius justify the existence of an HVAC unit that can also operate in cooling mode.
-
-\begin{figure}[htbp]
-	\centering
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centering
-		\includegraphics[width=0.999\linewidth]{Figures/Tsea_vs_time}
-		\caption{Time series}
-		\label{fig:TseaTairTIME}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{Figures/Tsea_hist}
-		\caption{Distribution}
-		\label{fig:TseaTairDIST}
-	\end{subfigure}
-	\caption{Statistical representation of measured air and sea water temperatures based on yearly data. Landsort, 2014}
-	\label{fig:TseaTair}
-\end{figure}
-
-Figure \ref{fig:vship_hist} represents the distribution of the ship speed, both in terms of statistical distribution and in terms of a reference 24h period.
-The ship operates almost for the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
-This can also be seen from Figure \ref{fig:Pme_vs_vship}, where the main cluster of operations at middle-low speed can be clearly identified, as well as the smaller area of operations at high speed.
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/vship_hist}
-	\caption{Yearly distribution of ship speed}
-	\label{fig:vship_hist}
-\end{figure}
-
-figure --> Operational mode distribution
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/Hist_mainEngines}
-	\caption{Yearly load distribution, main engines}
-	\label{fig:loadME_hist}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/Hist_auxEngines}
-	\caption{Yearly load distribution, auxiliary engines}
-	\label{fig:loadAE_hist}
-\end{figure}
-
-ANYTHING ELSE THAT CAN BE INTERESTING TO SHOW IN THE EXPLORATORY DATA ANALYSIS?
-
-\subsection{Energy analysis} \label{sec:res:energy}
-
-Here we can see the typical example (note: the day was not selected as "typical", but simply randomly) of the energy demand during a winter and a summer day (Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}).
-It can be seen clearly how, as expected, the heating demand in winter is much higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduce need for compartment heating.
-On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors.
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_W}
-	\caption{Energy demand during a day of operations, Winter}
-	\label{fig:Demand_vs_time_W}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_S}
-	\caption{Energy demand during a day of operations, Summer}
-	\label{fig:Demand_vs_time_S}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyDemand}
-	\caption{Energy demand, aggregated data for one year of operations}
-	\label{fig:Pie_EnergyDemand}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyGeneration}
-	\caption{Energy generation, aggregated data for one year of operations}
-	\label{fig:Pie_EnergyGeneration}
-\end{figure}
-
-From Figure \ref{fig:Pie_EnergyDemand} it can be noted that nearly 50\% of the total energy demand is related to ship propulsion, while the remaining portion is approximately equally split between electric energy and heat demand.
-
-Within the electric energy demand, the HVAC systems only represent a minor contribution (XX\%), as it is only present for a few months during summer, and at rather low power demand.
-This is not surprising, since it can be observed that (see Figure \ref{fig:TseaTairDIST}) the air is always below 27$^o$C, and rarely above 17$^o$C.
-Also the thrusters, although they represent a rather high instantaneous demand, they are only used for a very short time and, hence, their total contribution to the ship energy demand is limited to XX\%.
-As a consequence, most of the electric energy demand should be allocated to other consumers, and efforts should be directed towards providing additional measurements of how different users contribute to the total electric power demand.
-
-\subsection{Heat demand and generation} \label{sec:res:heat}
-
-FIGURE --> Pie, heat generated by source
-FIGURE --> filled lines, heat generated by source.
-
-Looking at Figure \ref{fig:HeatGen_hist} it is possible to appreciate the amount of heat that is recovered from on board available waste heat (cooling water and exhaust gas).
-It can be observed that the amount of heat recovered from the HT cooling systems is much lower than expected, with an average of around 350 kW.
-This value is quite unreasonable compared to the values found in the technical specifications from the yard, that go from 700 kW (when only one auxiliary engine is operated) to 4800 kW (when all main engines and two auxiliary engines are operated).
-\\
-\begin{figure}
-	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/HeatGen_hist}
-	\caption{Distribution of thermal power generated from on board waste heat recovery}
-	\label{fig:Hist_WHR}
-\end{figure}
-
-Regarding the heat demand onboard, it can be seen that the HVAC re-heater represents the largest contributor (XX\%) followed by the hot water heater (XX\%).
-The large contribution of the HVAC reheater can be explained by the former being important for long times and in relation to the low ambient temperatures encountered by the case study vessel during operation.
-The relatively large contribution of the hot water heater can instead be related to the fact that these systems are rather important consumers and are necessary during the whole ship operations (i.e.
-also during the summer months).
-
-\subsection{Exergy Analysis} \label{sec:res:exergy}
-
-The results of the exergy analysis of the ship energy systems are shown in Figure REF and Tables REF to REF.
-
-Table(s) --> Exergy efficiencies
-- General at system level
-- One per system (MEs, AEs, CoolingSystems, SteamSystems)
-
-Figure --> Grassman diagram
-
-\subsection{Typical operational days}  \label{sec:res:typicalDays}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-\appendix
-
-\section{Full tables of energy and exergy flows}
-
-\section{Full equation list}
-
-\begin{table}[htbp]
-	\centering
-	\begin{tabular}{p{4cm}l}
-		\hline
-		\multirow{2}{*}{Before compressor}		& $ T_{comp,in} = $ Measured \\
-												& $ p_{comp,in} = $ Measured \\
-		\multirow{3}{*}{After compressor}		& $ T_{comp,out} = T_{comp,in} + \frac{T_{comp,out,is} - T_{comp,in} }{\eta_{comp,is}}  $  \\
-												& $ p_{comp,out} = $ Measured \\
-												& $ \eta_{comp,is} = P_2(\frac{p_{comp,out}}{p_{comp,in}}) $ \\
-		\multirow{3}{*}{After CAC-HT stage}		& $ p_{CAC-HT,out} = p_{comp,out} $ \\
-												& $ T_{CAC-HT,out} = T_{comp,out} - \frac{\dot{Q}_{CAC,HT}}{\dot{m}_{air,cyl}} $ \\
-												& $ \dot{Q}_{CAC,HT} = P_2(\lambda) $ \\
-		\multirow{3}{*}{After CAC-LT stage}		& $ p_{CAC-LT,out} = p_{comp,out} $ \\
-												& $ T_{CAC-LT,out} = T_{CAC-HT,out} - \frac{\dot{Q}_{CAC,LT}}{\dot{m}_{air,cyl}} $ \\
-												& $ \dot{Q}_{CAC,LT} = P_2(\lambda) $ \\
-		\multirow{3}{*}{After CAC-LT stage}
-
-	\end{tabular}
-\end{table}
-
 \section{Estimation of air and exhaust gas flows in the main engines}
@@
 \end{equation}
-
-\section{Full list of available measurements}
-
-\bibliographystyle{unsrt}
-\bibliography{bibliography}
-
-%% \begin{thebibliography}{widestlabel}
-%% \bibitem{label}
-%% Text of bibliographic item
-%%\end{thebibliography}
-
-\end{document}
-\endinput
-%%
-%% End of file `elsarticle-template-num.tex'.

### 2018-04-01 Manuscript/Manuscript.tex
Updates in the manuscript

@@
 
+In this context, the cruise industry is growing at an even greater pace.
+Cruise ship passengers have increased from 17.8 million in 2009 to 24.7 million in 2015 \cite{CruiseOutlook2017}, and this growth is expected to continue in the coming years \cite{CruiseOutlook2017}.
+Cruise travels, with an estimated average of 160 $kg_{CO_2}$ per passenger and per day, are among the most carbon intensive in the whole tourism industry.
+The contribution of the cruise industry to global CO$_2$ emissions was estimated to 19.3 Mtons annually in 2010 \cite{Eijgelaar2010}.
+
+The cruise industry have also received much attention for its environmental impact other than in terms of GHG emissions.
+Cruise ships were shown to have a remarkable negative impact on both air \cite{Poplawski2011} and water \cite{Butt2007} quality, and were also pointed as sources of environmental damage from a wider perspective \cite{Brida2009}.
+The fact that cruise ships tend to operate extensively in highly populated and environmental sensitive areas makes the situation all the more worth of attention.
+
 Altogether, these conditions present a challenge to the shipping companies who attempt to reduce their fuel consumption, environmental impact, and operative costs.
@@
 It should be noted that, however, much research effort has been devoted, especially in recent times, to the improvement of the efficiency of ship energy systems by recovering the waste heat available from the engines.
-With reference to different types of technologies, case studies, and designs,the several authors showed the existance of a quite significant potential for energy saving when WHR systems are employed, ranging from around 1\% for single-pressure steam cycles applied to two-stroke engines \cite{Theotokatos2012} to more complex systems based on ORCs (up to 10\%, \cite{Hountalas2012}) or including the cooling systems as a source of waste heat (over10\%, \cite{Dimopoulos2012}).
+With reference to different types of technologies, case studies, and designs,the several authors showed the existance of a quite significant potential for energy saving when WHR systems are employed, ranging from around 1\% for single-pressure steam cycles applied to two-stroke engines \cite{Theotokatos2012} to more complex systems based on ORCs (up to 10\%, \cite{Hountalas2012}) or including the cooling systems as a source of waste heat (over 10\%, \cite{Dimopoulos2012}).
 The case of the installation of an ORC on board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2015}, showing a similar potential.
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.95\linewidth]{Figures/ShipOperationalProfile}
+	\caption{Reference operational profile (ship speed vs time) for the case study}
+	\label{fig:typicalShipOperations}
+\end{figure}
+
 The ship systems are summarized in Figure \ref{fig:shipSystems}.
 The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller.
-The main engines are four Wärtsilä 4-stroke Diesel engines (ME) rated 5850 kW each.
-All engines are equipped with selective catalytic reactors (SCR) for NOX emissions abatement.
+The main engines are four W\"{a}rtsil\"{a} 4-stroke Diesel engines (ME) rated 5850 kW each.
+All engines are equipped with selective catalytic reactors (SCR) for NO$_X$ emissions abatement.
 Propulsion power is needed whenever the ship is sailing; however, it should be noted that the ship rarely sails at full speed, and most of the time it only needs one or two engines operated simultaneously.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.95\linewidth]{Figures/ShipLayout}
+	\caption{Layout of the case study energy systems}
+	\label{fig:shipSystems}
+\end{figure}
 
@@
 For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine.
-The model is explained in detail in Appendix REF.
+In particular, the following equations are used:
+\begin{itemize}
+	\item Mass balance on the split after the compressor, with one flow going to the engine inlet valves and the other flow going to the bypass valve. 
+	\begin{equation}
+	\dot{m}_{air,comp} = \dot{m}_{air,cyl} + \dot{m}_{air,bypass} \\
+	\end{equation}
+	\item Mass balance on the mixer before the turbine, with one flow coming from the engine exhaust valves and the other flow coming from the bypass valve
+	\begin{equation}
+	\dot{m}_{eg,turb} = \dot{m}_{eg,cyl} + \dot{m}_{air,bypass} \\
+	\end{equation}
+	\item Mass balance in the engine cylinder:
+	\begin{equation}
+	\dot{m}_{eg,cyl} = \dot{m}_{air,cyl} + \dot{m}_{fuel,cyl} \\
+	\end{equation}
+	\item Energy balance on the turbocharger: the power generated by the turbine must be equal to the power absorbed by the compressor
+	\begin{equation}
+	\dot{m}_{air,comp} \Delta h_{comp} = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
+	\end{equation}
+	\item Energy balance on the mixer between the air flow after the bypass valve and the exhaust gas flow after the exhaust valve
+	\begin{equation}
+	\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0)
+	\end{equation}
+\end{itemize}
+
+The energy balance over the whole engine is presented in Equation \ref{eqn:engineEnergyBalance}.
+\begin{equation}
+\dot{m}_{fuel} (LHV_{fuel} + c_{p,fuel} (T_{fuel,in} - T_0)) + \dot{m}_{air} c_{p,air} (T_{air,comp,in} - T_0) = \dot{W}_{mech} + \dot{m}_{eg} c_{p,eg} (T_{eg,TC,out} - T_0) + \dot{Q}_{cooling} 
+\end{equation} \label{eqn:engineEnergyBalance}
+With: 
+\begin{equation}
+\dot{Q}_{cooling} = \dot{Q}_{CAC,HT} + \dot{Q}_{CAC,LT} + \dot{Q}_{JWC} + \dot{Q}_{LOC}
+\end{equation} \label{eqn:engineCoolingBalance}
+The value of $\dot{Q}_{cooling}$ is determined based on the balance in Equation \ref{eqn:engineEnergyBalance}.
+The share between the four contributions in Equation \ref{eqn:engineCoolingBalance} is calculated based on the data for the heat loss to the HT and LT systems presented in the engines project guides as functions of the engine load.
+The values are interpolated based on a 2$^nd$ degree polynomial and scaled to respect the energy balance.
+In absence of more accurate data, it is then assumed that the contributions of $\dot{Q}_{JWC}$ and $\dot{Q}_{LOC}$ are equal for all loads.
+CHANGE VALUES IN THE TABLE
 
@@
 
-INCLUDE VALIDATION AGAINST PROJECT GUIDE DATA
-
 \begin{table}
@@
 The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board.
-A representation of the ship cooling systems is provided in Figure XXX.
+A representation of the ship cooling systems is provided in Figure \ref{fig:generalsystemssimplified}.
 
@@
 
-Figure cooling systems
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/GeneralSystemsSimplified}
+	\caption{Layout of the ship heating and cooling system}
+	\label{fig:generalsystemssimplified}
+\end{figure}
 
@@
 Figures \ref{fig:TseaTairTIME} and \ref{fig:TseaTairDIST} represent the time evolution and the distribution over the year of the ambient air and sea water temperature.
-The relatively low temperatures that are experienced by the case study ship during its operations suggest that heating demand can be expected to be higher than cooling demand.
+The relatively low temperatures that are experienced by the case study ship during its operations rises expectations for a high space heating demand and to a relatively low for summer HVAC electric power demand.
 On the other hand, the fact that, in summer, air temperatures reach up to 26 degrees Celsius justify the existence of an HVAC unit that can also operate in cooling mode.
@@
 
-Figure \ref{fig:vship_hist} represents the distribution of the ship speed, both in terms of statistical distribution and in terms of a reference 24h period.
+Figure \ref{fig:vship_hist} represents the distribution of the ship speed.
 The ship operates almost for the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
-This can also be seen from Figure \ref{fig:Pme_vs_vship}, where the main cluster of operations at middle-low speed can be clearly identified, as well as the smaller area of operations at high speed.
+The relatively large amount of time the ship spends in port can also be seen in Figure \ref{fig:opMode_pie}.
+This suggests that, in spite of the high installed power for propulsion, the energy demand for this category is expected to be relatively low.
+
+Figures \ref{fig:loadME_hist} and \ref{fig:loadME_hist} show the frequency of the load at which each of the engines (main and auxiliary engines, respectively) is operated.
 
@@
 
-\section{Estimation of air and exhaust gas flows in the main engines}
-
-\subsection{Mass balances}
-
-Mass balance on the split after the compressor, with one flow going to the engine inlet valves and the other flow going to the bypass valve
-\begin{equation}
-\dot{m}_{air,comp} = \dot{m}_{air,cyl} + \dot{m}_{air,bypass} \\
-\end{equation}
-
-Mass balance on the mixer before the turbine, with one flow coming from the engine exhaust valves and the other flow coming from the bypass valve
-\begin{equation}
-\dot{m}_{eg,turb} = \dot{m}_{eg,cyl} + \dot{m}_{air,bypass} \\
-\end{equation}
-
-Note that the exhaust gas mass flow from the cylinders is fully defined as follows:
-\begin{equation}
-\dot{m}_{eg,cyl} = \dot{m}_{air,cyl} + \dot{m}_{fuel,cyl} \\
-\end{equation}
-
-\subsection{Energy balances}
-
-Energy balance on the turbocharger: the power generated by the turbine must be equal to the power absorbed by the compressor
-\begin{equation}
-\dot{m}_{air,comp} \Delta h_{comp} = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
-\end{equation}
-
-Energy balance on the mixer between the air flow after the bypass valve and the exhaust gas flow after the exhaust valve
-\begin{equation}
-\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0)
-\end{equation}
-
-The system of four equations is to be solved in the four following unknowns:
-\begin{itemize}
-	\item $ \dot{m}_{eg,turb} $
-	\item $ \dot{m}_{air,bypass} $
-	\item $ \dot{m}_{air,comp} $
-	\item $ T_{turb,in} $ (In the case of the Main engines)
-	\item $ T_{cyl,out} $ (In the case of the Auxiliary engines)
-\end{itemize}
-
-Include other assumptions:
-- TC efficiency
-- Compressor efficiency
-- Turbine efficiency
-
-\subsection{Main Engines}
-
-To calculate what we need, we must make the system explicit on the variables we need to calculate.
-This process is different for the main and auxiliary engines, mostly because the temperature measurement of the exhaust gas before the turbine is positioned differently.
-In the case of the main engines, the temperature at the \textbf{cylinder outlet} is measured and, hence, known.
-
-First, we make all equations explicit in $ \dot{m}_{air,bypass} $ thus eliminating both 
-$ \dot{m}_{eg,turb} $ and $ \dot{m}_{air,comp} $
-
-$$
-(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech} 
-$$
-$$
-\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_0)
-$$
-
-At this point, we use the first equation to simplify the system and make it explicit in the $ \dot{m}_{air,bypass} $ variable alone:
-
-\begin{eqnarray*}
-	(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}  & = & (\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} \\
-	T_{turb,in} - T_{turb,out} & = & \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} \\
-	T_{turb,in} & = & T_{turb,out} + \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}}  \\
-\end{eqnarray*}
-
-We can now substitute this expression in the second equation:
-
-\begin{multline}
-\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\ 
-(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,out} + \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} - T_0)
-\end{multline}
-
-We first execute the multiplication in the term on the right of the equal:
-\begin{multline}
-\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\
-\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_0) + \\
-\dot{m}_{air,bypass} c_{p,eg} (T_{turb,out} - T_0) + \\ 
-\frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}}(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) c_{p,eg}
-\end{multline}
-
-Now, we can group together the elements in the variable of interest:
-\begin{multline}
-\dot{m}_{air,bypass} \left[ c_{p,air} (T_{comp,out} - T_0) - c_{p,eg} (T_{turb,out} - T_0) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
-\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_0) - \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
-\end{multline}
-
-This can be further simplified into:
-\begin{multline}
-\dot{m}_{air,bypass} \left[ c_{p,air} (T_{comp,out} - T_0) - c_{p,eg} (T_{turb,out} - T_0) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
-\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
-\end{multline}
-
-Finally, we can assume the specific heat at constant pressure of air and exhaust gas to be sufficiently similar for considering them equal in the term at the numerator (we ignored changes in the properties of the mixture in other points as well):
-\begin{multline}
-\dot{m}_{air,bypass} \left[ c_{p,eg} (T_{comp,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
-\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
-\end{multline}
-
-From which we can derive the final form:
-\begin{equation}
-\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}}{c_{p,eg} (T_{comp,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}}
-\end{equation}
-
-Which can be rewritten, in order to have both numerator and denominator positive, as follows:
-\begin{equation}
-\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{cyl,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}}{c_{p,eg} (T_{turb,out} - T_{comp,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}}
-\end{equation}
-
-\subsection{Auxiliary Engines}
-
-In the case of the auxiliary engines, what we know instead is the temperature of the exhaust gas \textbf{before the turbine}, hence after the merging between the exhaust gas and the bypass.
-
-In this case, however, things appear to be simple: the energy balance on the turbocharger can be calculated as there is only one variable unknown: the total mass flow in the compressor, and hence the bypass flow:
-$$
-(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
-$$
-$$
-\dot{m}_{air,cyl} \Delta h_{comp} + \dot{m}_{air,bypass} \Delta h_{comp} = \dot{m}_{eg,cyl} c_{p,eg} (T_{turb,in} - T_{turb,out}) + \dot{m}_{air,bypass} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
-$$
-$$
-\dot{m}_{air,bypass} (\Delta h_{comp} - c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}) = \dot{m}_{eg,cyl} c_{p,eg} (T_{turb,in} - T_{turb,out}) - \dot{m}_{air,cyl} \Delta h_{comp}
-$$
-\begin{equation}
-\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech} - \dot{m}_{air,cyl} \Delta h_{comp}}{\Delta h_{comp} - c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}}
-\end{equation}
-
-Now that we have the value for the bypass flow, it is easy to calculate backwards the temperature of the exhaust gas after the cylinder outlet valves, before the merging with the bypass flow:
-
-$$
-\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0)
-$$
-$$
-\dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0) - \dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0)
-$$
-\begin{equation}
-T_{cyl,out} = T_0 + \frac{\dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0) - \dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0)}{\dot{m}_{eg,cyl} c_{p,eg}}
-\end{equation}
-
 \section{Full list of available measurements}

### 2018-04-01 Python files/plotting.py

@@ -441,13 +441,19 @@ def predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure, filenames
             temp1 = []
             for engine in {"AE1", "AE2", "AE3", "AE4"}:
                 temp1.append(processed[engine+":Cyl:Power_out:Edot"][processed[engine+":on"]]/CONSTANTS["AuxEngines"]["MCR"])
-            ax.hist(tuple(temp1), normed=1, alpha=0.8, label=["AE1", "AE2", "AE3", "AE4"])
+            ax.hist(tuple(temp1), normed=False, alpha=0.8, label=["AE1", "AE2", "AE3", "AE4"])
             plt.legend()
+            plt.xlabel("Engine load")
+            plt.ylabel("Number of observations")
 
         if filename == "Hist:MainEngines":
+            temp1 = []
             for engine in {"ME1", "ME2", "ME3", "ME4"}:
-                ax.hist(processed[engine+":Cyl:Power_out:Edot"][processed[engine+":on"]]/CONSTANTS["MainEngines"]["MCR"], normed=1, alpha=0.5, label=engine)
+                temp1.append(processed[engine + ":Cyl:Power_out:Edot"][processed[engine + ":on"]] / CONSTANTS["MainEngines"]["MCR"])
+            ax.hist(tuple(temp1), normed=False, alpha=0.8, label=["ME1", "ME2", "ME3", "ME4"])
             plt.legend()
+            plt.xlabel("Engine load")
+            plt.ylabel("Number of observations")
 
 
         ### SCATTER PLOTS ###

### 2018-05-08 Manuscript/Manuscript.tex
Update

@@
 \usepackage{multirow}
-%% 
+%% The package subcaption is for helping when using subfigures and their captions
 \usepackage{subcaption}
+%% This package is used for having nicer tables
+\usepackage{booktabs}
 
@@
 For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine.
-In particular, the following equations are used:
-\begin{itemize}
-	\item Mass balance on the split after the compressor, with one flow going to the engine inlet valves and the other flow going to the bypass valve. 
-	\begin{equation}
-	\dot{m}_{air,comp} = \dot{m}_{air,cyl} + \dot{m}_{air,bypass} \\
-	\end{equation}
-	\item Mass balance on the mixer before the turbine, with one flow coming from the engine exhaust valves and the other flow coming from the bypass valve
-	\begin{equation}
-	\dot{m}_{eg,turb} = \dot{m}_{eg,cyl} + \dot{m}_{air,bypass} \\
-	\end{equation}
-	\item Mass balance in the engine cylinder:
-	\begin{equation}
-	\dot{m}_{eg,cyl} = \dot{m}_{air,cyl} + \dot{m}_{fuel,cyl} \\
-	\end{equation}
-	\item Energy balance on the turbocharger: the power generated by the turbine must be equal to the power absorbed by the compressor
-	\begin{equation}
-	\dot{m}_{air,comp} \Delta h_{comp} = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
-	\end{equation}
-	\item Energy balance on the mixer between the air flow after the bypass valve and the exhaust gas flow after the exhaust valve
-	\begin{equation}
-	\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0)
-	\end{equation}
-\end{itemize}
+\begin{align}
+	\dot{m}_{air,comp} & =  \dot{m}_{air,cyl} + \dot{m}_{air,bp} \label{eq:mb:comp} \\ 
+	\dot{m}_{eg,turb} & = \dot{m}_{eg,cyl} + \dot{m}_{air,bp} \label{eq:mb:turb} \\
+	\dot{m}_{eg,cyl} & = \dot{m}_{air,cyl} + \dot{m}_{fuel,cyl} \label{eq:mb:cyl} \\
+	\dot{m}_{air,comp} \Delta h_{comp} & = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech} \label{eq:eb:turbocharger} \\
+	\dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0) & = \dot{m}_{air,bp} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) \label{eq:eb:bpmerge}
+\end{align}
+Where equations \ref{eq:mb:comp} to \ref{eq:mb:cyl} represent the mass balances of the bypass split, bypass merge, and cylinder respectively, while equations \ref{eq:eb:turbocharger} and \ref{eq:eb:bpmerge} represent the energy balances of the turbocharger and of the bypass merge, respectively.
 
@@
 \begin{equation}
-\dot{m}_{fuel} (LHV_{fuel} + c_{p,fuel} (T_{fuel,in} - T_0)) + \dot{m}_{air} c_{p,air} (T_{air,comp,in} - T_0) = \dot{W}_{mech} + \dot{m}_{eg} c_{p,eg} (T_{eg,TC,out} - T_0) + \dot{Q}_{cooling} 
-\end{equation} \label{eqn:engineEnergyBalance}
+\dot{Q}_{fuel} + \dot{Q}_{air,in} = \dot{W}_{mech} + \dot{Q}_{eg} + \dot{Q}_{cooling}  \label{eqn:engineEnergyBalance} \\
+\end{equation}
 With: 
-\begin{equation}
-\dot{Q}_{cooling} = \dot{Q}_{CAC,HT} + \dot{Q}_{CAC,LT} + \dot{Q}_{JWC} + \dot{Q}_{LOC}
-\end{equation} \label{eqn:engineCoolingBalance}
+\begin{align}
+\dot{Q}_{fuel} & =\dot{m}_{fuel} (LHV_{fuel} + c_{p,fuel} (T_{fuel,in} - T_0))\\
+\dot{Q}_{air,in} & = \dot{m}_{air} c_{p,air} (T_{air,comp,in} - T_0) \\ 
+\dot{Q}_{eg} & = \dot{m}_{eg} c_{p,eg} (T_{eg,TC,out} - T_0) \\
+\dot{Q}_{cooling} & = \dot{Q}_{CAC,HT} + \dot{Q}_{CAC,LT} + \dot{Q}_{JWC} + \dot{Q}_{LOC} \label{eq:eb:cooling}
+\end{align}
+
 The value of $\dot{Q}_{cooling}$ is determined based on the balance in Equation \ref{eqn:engineEnergyBalance}.
-The share between the four contributions in Equation \ref{eqn:engineCoolingBalance} is calculated based on the data for the heat loss to the HT and LT systems presented in the engines project guides as functions of the engine load.
+The share between the four contributions in Equation \ref{eq:eb:cooling} is calculated based on the data for the heat loss to the HT and LT systems presented in the engines project guides as functions of the engine load.
 The values are interpolated based on a 2$^nd$ degree polynomial and scaled to respect the energy balance.
@@
 CHANGE VALUES IN THE TABLE
+
+Values for temperatures, mass flows and energy flows for different engine loads are provided in the Appendix \ref{sec:appendix:engines}, for both main and auxiliary engines.
+
+\subsubsection{Cooling systems}
+
+The ship's cooling systems are designed similarly to most ships and divided in two cooling systems of two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C).
+The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board.
+A representation of the ship cooling systems is provided in Figure \ref{fig:generalsystemssimplified}.
+
+All values not directly measured in the cooling systems were calculated based on mass and energy balances.
+The mass flow was determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/GeneralSystemsSimplified}
+	\caption{Layout of the ship heating and cooling system}
+	\label{fig:generalsystemssimplified}
+\end{figure}
+
+\subsection{Electric power demand} \label{sec:met:electric}
+
+The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy
+\begin{equation}
+P_{el,tot} = \sum_{1}^{4} AE_i
+\end{equation}
+
+The determination of how different consumers contribute to the total electric power demand is less trivial.
+The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc.
+None of the individual contributions is directly measured, hence estimations need to be based on indirect measurements.
+
+\subsubsection{HVAC systems}
+
+For what concerns the air conditioning unit demand (HVAC), the observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a clear period, corresponding to the summer months, during which there is a significant increase of the demand.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pel_vs_time}
+	\caption[Total electric power demand versus time (Daily average)]{Total electric power demand versus time (Daily average)}
+	\label{fig:pelvstime}
+\end{figure}
+
+Since the total demand seems to be rather constant otherwise, it can be concluded that the demand of the HVAC compressors related to the need for cooling is concentrated during the summer months.
+Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a random summer and winter day, shows that the daily variation is comparable.
+This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/PelRel_vs_time}
+	\caption{Instantaneous power demand over daily average, winter versus summer day}
+	\label{fig:pelrelvstime}
+\end{figure}
+
+Hence, the HVAC electric power demand was estimated as follows:
+\begin{equation}
+P_{el,HVAC} =
+\begin{cases}
+P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
+0, & \text{otherwise}
+\end{cases}
+\end{equation}
+where $P_{el,ref}(t)$ is calculated as:
+\begin{equation}
+P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
+\end{equation}
+
+\subsubsection{Thrusters}
+
+When entering port areas, the ship needs to use thrusters to maneuver and berth.
+In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.
+In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
+From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).
+
+By comparing the reference profile to the instantaneous one, the points where the former is more than 10\% higher than the latter are identified as "thruster-on" points and treated consequently.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Thruster_selection}
+	\caption{Example of the procedure of selection of thruster power demand}
+	\label{fig:thruster_selection}
+\end{figure}
+
+\subsection{Heat demand} \label{sec:met:heat}
+
+As the heat demand is not measured, it is necessary to determine it based on the available indirect measurements.
+
+The heat demand and generation can be summarized according to the following equations:
+\begin{eqnarray}
+\dot{Q}_{gen} & = & \dot{Q}_{EGB} + \dot{Q}_{HTHR} + \dot{Q}_{AB} \\
+\dot{Q}_{dem} & = & \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \\
+\end{eqnarray}
+
+The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.999\linewidth]{Figures/Heating_cooling_systems}
+	\caption{General view of the ship heating and cooling systems}
+	\label{fig:ship:general}
+\end{figure}
+
+\subsubsection{Heat balance parameter estimation}
+
+As not enough information and measurements are available to determine the various components of the heat balance, in this paper we determined them by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters.
+The parameter estimation problem is hence written as a minimization problem:
+\begin{eqnarray}
+min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
+\end{eqnarray}
+
+where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections.
+A list of the parameters $\textbf{p}$ is shown in Table REF, together with the chosen upper and lower boundaries for the calibration procedure.
+\begin{table}
+	\centering
+	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
+		\hline 
+		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
+		\hline
+		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 &  \\ 
+		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 &  \\
+		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 &  \\ 
+		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 &  \\ 
+		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 &  \\ 
+		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 &  \\ 
+		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 &  \\ 
+		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 &  \\
+		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 &  \\
+		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 &  \\ 
+		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 &  \\ 
+		\hline
+	\end{tabular}
+	\caption{Parameters optimized in the parameter estimation for the heat balance}
+	\label{tab:ParameterEstimation} 
+\end{table}
+
+\subsubsection{Heat demand}
+
+We calculated the heat demand as the sum of the contributions of the elements listed in the ship's heat balance documentation.
+As no direct measurement of these quantities was available in the dataset, they had to be estimated based on the following assumptions:
+\begin{table}
+	\centering
+	{\tablinesep=2ex\tabcolsep=10pt
+		\begin{tabular}{p{2.8cm}l}
+			\hline 
+			Heat flow name & Equation \\
+			\hline
+			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
+			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
+			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
+			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
+			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
+			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
+			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
+			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
+			\hline
+	\end{tabular}}
+	\caption{Summary of the heat demand contributions and their calculation}
+	\label{tab:HeatDemand}
+\end{table}
+
+where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
+The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
+The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure REF.
+
+\subsubsection{Heat generation}
+
+The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty.
+The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
+\begin{equation}
+\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
+\end{equation}\label{eq:egb}
+
+where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF
+
+It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
+However, measurements of this contribution and of other variables that could lead to its straight-forward identification are missing.
+In this work, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},
+
+\begin{eqnarray}
+\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
+& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
+\end{eqnarray}
+
+where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
+The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
+
+The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems.
+The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit.
+This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value.
+In this calculation, the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
+
+\subsubsection{Parameter estimation and uncertainty quantification}
+
+The results of the parameter estimation are represented graphically and quantitatively in Figure REF and Table REF.
+
+Given the lack of input values for the estimation of the heat demand, the provided estimate based on the procedure described above should be integrated with an estimation of the uncertainty.
+
+In this study, we base the estimation of the uncertainty on the production side, as it is the one that has the largest amount of information available.
+In these regards, the uncertainty can be reduced to the contribution of three elements: $U(\dot{Q}_{EGB})$, $U(\dot{Q}_{HTHR})$, $U(\dot{Q}_{AB})$.
+
+The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$.
+In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT].
+The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions as described in Appendix REF, and can be estimated of being up to 10\%.
+The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value.
+These assumptions lead to the following estimation of the uncertainty:
+\begin{equation}
+\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
+\end{equation}
+
+That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements (result obtained for $\pm 20$ K uncertainty.
+The uncertainty is reduced to 20\% if the measurement uncertainty on the temperature is reduced to  $\pm 20$ K)
+
+The uncertainty of the heat generated by the ABs can be reduced to the contribution of two elements: the uncertainty on $\dot{m}_{fuel,AB}$ and that on $\eta_{AB}$.
+The former will be at least as large as the calibration error (35\%) and will be considered equal to 50\% to be conservative and accounting also for errors in the aggregated boiler fuel measurements.
+In addition, the uncertainty on the efficiency can be considered to be around 10\% based on the discrepancy between the considered sources and on the expected variability of the efficiency with load.
+Similarly to the previous case, the combination of these efficiencies lead to a total uncertainty of 51\%, where the main contribution comes from the uncertainty of the model output compared to the actual fuel consumption.
+
+The estimation of the uncertainty of $\dot{Q}_{HTHR}$ can be based on \ref{eqn:HTHR2} having assigned an acceptable variation of 20\% to the effectiveness of the heat exchanger, 20\% on the estimation of the reference mass flow, $\pm 5$ K uncertainty on water temperature measurements and considering negligible the uncertainty on $c_{p,w}$, leading to a 145\% uncertainty on this measurement.
+
+\subsection{Operational mode}
+
+The ship operates in different conditions, and it can be useful to provide a separate analysis depending on the operational mode.
+The four operational modes identified for the selected case study, together with the conditions applied to the dataset to uniquely identify each time step to a separate operational mode, are given as follows:
+\begin{description}
+	\item[High speed sailing] Ship speed above 15 kn
+	\item[Low speed sailing] Ship speed between 4 and 15 kn
+	\item[Maneuvering] Ship speed between 2 and 4 knots OR Thrusters power larger than 0
+	\item[Port stays / ship drifting] All other points
+\end{description}
+
+\subsection{Determination of typical operational conditions}
+
+\section{Results (and discussion)}
+
+\subsection{Exploratory analysis} \label{sec:res:exploratory}
+
+In this section, we observe and analyze direct measurements from ship systems, with no pre-processing, that are expected to have an influence on the energy analysis.
+
+Figures \ref{fig:TseaTairTIME} and \ref{fig:TseaTairDIST} represent the time evolution and the distribution over the year of the ambient air and sea water temperature.
+The relatively low temperatures that are experienced by the case study ship during its operations rise expectations for a high space heating demand and to a relatively low for summer HVAC electric power demand.
+On the other hand, the fact that, in summer, air temperatures reach up to 26 degrees Celsius justify the existence of an HVAC unit that can also operate in cooling mode.
+
+\begin{figure}[htbp]
+	\centering
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centering
+		\includegraphics[width=0.999\linewidth]{Figures/Tsea_vs_time}
+		\caption{Time series}
+		\label{fig:TseaTairTIME}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centering
+		\includegraphics[width=0.9\linewidth]{Figures/Tsea_hist}
+		\caption{Distribution}
+		\label{fig:TseaTairDIST}
+	\end{subfigure}
+	\caption{Statistical representation of measured air and sea water temperatures based on yearly data. Landsort, 2014}
+	\label{fig:TseaTair}
+\end{figure}
+
+Figure \ref{fig:vship_hist} represents the distribution of the ship speed.
+The ship operates almost for the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
+In Figure \ref{fig:opMode_pie} it can be also observed that the ship spends a relatively high amount of time in port.
+These remarks suggest that, in spite of the high installed power of the main engines, the energy demand for propulsion is expected to be relatively low.
+
+Figures \ref{fig:loadME_hist} and \ref{fig:loadME_hist} show the frequency of the load at which each of the engines (main and auxiliary engines, respectively) is operated.
+In the case of the main engines, it can be observed that they are operated mostly at loads between 30\% and 60\%, despite the optimal load for the engines being located around 85\%.
+This is a consequence of both the low speed at which the ship is generally operated, and of the fact that the two shaft lines are operated independently, and hence do not allow the use of only one of the four engines at high load.
+
+In the case of the auxiliary engines, a larger difference between how the engines are operated can be observed.
+AE-3 is generally more often operated, as it is the engine that is run in port with clean fuels.
+On the other hand, it seems that AE-4 is operated rarely, which might be due to maintenance issues.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/vship_hist}
+	\caption{Yearly distribution of ship speed}
+	\label{fig:vship_hist}
+\end{figure}
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/Pie_operationalMode}
+	\caption{Operational time by mode}
+	\label{fig:opMode_pie}
+\end{figure}
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/Hist_mainEngines}
+	\caption{Yearly load distribution, main engines}
+	\label{fig:loadME_hist}
+\end{figure}
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/Hist_auxEngines}
+	\caption{Yearly load distribution, auxiliary engines}
+	\label{fig:loadAE_hist}
+\end{figure}
+
+ANYTHING ELSE THAT CAN BE INTERESTING TO SHOW IN THE EXPLORATORY DATA ANALYSIS?
+
+\subsection{Energy analysis} \label{sec:res:energy}
+
+The ship energy demand is first subdivided among three main consumer categories: propulsion, electric power, and heat.
+The evolution of these demands over time during a typical voyage is shown in Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}, representing a winter and a summer day respectively.
+It can be seen clearly how, as expected, the heating demand in winter is higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduced need for compartment heating.
+On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors.
+Looking at the yearly cumulated demand, nearly 50\% of the total energy demand is related to ship propulsion (see Figure \ref{fig:Pie_EnergyDemand}), while the remaining portion is approximately equally split between electric energy and heat demand.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_W}
+	\caption{Energy demand during a day of operations, Winter (31 Jan)}
+	\label{fig:Demand_vs_time_W}
+\end{figure}
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_S}
+	\caption{Energy demand during a day of operations, Summer (31 Jul)}
+	\label{fig:Demand_vs_time_S}
+\end{figure}
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyDemand}
+	\caption{Energy demand, aggregated data for one year of operations}
+	\label{fig:Pie_EnergyDemand}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyGeneration}
+	\caption{Energy generation, aggregated data for one year of operations}
+	\label{fig:Pie_EnergyGeneration}
+\end{figure}
+
+Within the electric energy demand, the HVAC systems only represent a minor contribution (3\%), as it is only present for a few months during summer.
+This is not surprising, since it can be observed that (see Figure \ref{fig:TseaTairDIST}) the air is always below 27$^o$C, and rarely above 17$^o$C.
+Also the thrusters, although their instantaneous power demand is high, are only used for a very short time each day and, hence, their total contribution to the ship electrical energy demand is limited to 1.5\%.
+
+Heat recovery has a large impact on the overall heat demand (see Fig.
+\ref{fig:Pie_HeatGeneration}).
+The the exhaust gas boilers and the heat recovery on the HT-water fulfill almost 75\% of the yearly demand for heating, leaving only the remaining 25\% to be provided by oil-fired auxiliary boilers.
+The contribution of the HRSG and the HTHR is substantially constant throughout the year (Fig.
+\ref{fig:TimeSeries_HeatGeneration}), while the auxiliary boilers are used to cover the peaks and the winter increased baseload.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/Pie_HeatGeneration}
+	\caption{Share of thermal power generated from on board heat systems}
+	\label{fig:Pie_HeatGeneration}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/TimeSeries_HeatGeneration}
+	\caption{Time series representation of the thermal power generated from on board heat systems}
+	\label{fig:TimeSeries_HeatGeneration}
+\end{figure}
+
+Looking at heating from the demand side, it can be seen that the HVAC pre-heater represents the largest contributor (56\%) to the annual heating demand, followed by the hot water heater (18\%) and by the galley (11\%).In multi-season central HVAC systems the pre-heater is generally used for the main heating contribution during winter, while the re-heater is only used for peak demand, or for adjusting the delivery air temperature during summer, thus explaining the large contribution of the former.
+The relatively large share of the hot water heater can instead be related to the fact that these systems are used during all seasons.
+
+The full Sankey diagram representation of the energy flows on board is provided in Figure \ref{fig:Sankey}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/Sankey_diagram_energy_ship_v9}
+	\caption{Sankey diagram}
+	\label{fig:Sankey}
+\end{figure}
+
+\subsection{Exergy Analysis} \label{sec:res:exergy}
+
+The results of the exergy analysis of the ship energy systems are shown in Figure REF and Tables REF to REF. 
+\begin{itemize}
+	\item Table(s) --> Exergy efficiencies
+	\begin{itemize}
+		\item General at system level
+		\item One per system (MEs, AEs, CoolingSystems, SteamSystems) 
+	\end{itemize}
+	\item Figure --> Grassman diagram
+\end{itemize}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/Grassmann_diagram_v5}
+	\caption{Grassmann diagram}
+	\label{fig:Grassmann}
+\end{figure}
+
+It can be observed that there are large irreversibilities concentrated in some specific parts of the system that show potential for improvement.
+This is particularly the case of the LT/HT merging valves, where exergy is destroyed in the mixing process.
+Another substantial source of exergy destruction is the steam heater in the HTHR system, where heat at medium pressure (6 bar) is used to heat up water at 90$^o$C.
+The analysis of exergy destruction also points out the HVAC reheater as a potential source of improvement, based on the use of relatively high-temperature water for heating up air to around 30$^o$C and on the large energy demand of the component.
+Other important exergy destruction sources can be located in the engines, as expected, but the improvement of engine processes and technology goes beyond the scope of this work.
+\begin{table}
+	\centering
+	\begin{tabular}{cc}
+		\hline 
+		Component name & Contribution to exergy losses \\ 
+		\hline 
+		LT-HT merge & 13.4\% \\ 
+		\hline 
+		Sea water cooler & 10.6\% \\ 
+		\hline 
+		Steam heater & 6.0\% \\ 
+		\hline 
+		HVAC preheater & 6.0\% \\ 
+		\hline 
+	\end{tabular} 
+	\caption{Major contributors to ship exergy losses. Losses related to combustion are not included in the calculation}
+	\label{tab:exergyDestruction}
+\end{table}
+
+From the point of view of the exergy losses, it appears that the largest exergy losses are located in the exhaust gas.
+It should be noted that, even though a large part of this exergy cannot be recovered given the limitations on the exhaust gas outlet temperature to avoid the condensation of sulphuric acid, there is still a significant potential to be harvested if all the available exergy in the exhaust gas was recovered.
+This potential is further increased by the exergy destroyed in the process of dumping steam when it is produced in excess during summer.
+The exergy lost in the sea water coolers as low-temperature sea water discharged to the environment is lower, highlighting the fact that the potential for improving the efficiency of the system is located in other parts of the cooling systems.
+
+The efficiency of the recovery systems is also shown in Figure \ref{fig:Bar_efficiencyWHR}, where we show the fraction of the total energy (exergy) lost by the cooling systems and the exhaust gas of the ship's engines.
+It can be noted that, even when only looking at the energy that can be recovered based on the existing systems (i.e.
+not including the LT cooling systems, for instance), the recovery efficiency is located at around 25-35\% depending on whether the energy or the exergy efficiency is considered.
+Worth noting is also the fact that the efficiency on the exhaust gas side is lowered by the fact that two of the main engines are currently not equipped with HRSGs.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/Bar_WHRefficiency}
+	\caption{Energy and exergy efficiency of the WHR systems}
+	\label{fig:Bar_efficiencyWHR}
+\end{figure}
+
+\subsection{Typical operational days}  \label{sec:res:typicalDays}
+
+\subsection{Potential for system improvement}
+
+Based on the results of the energy and exergy analysis, it is possible to identify a number of possible measures that could potentially improve the efficiency of the system.
+
+Given the fact that many engines are operated at low load (both main and auxiliary engines), the system would benefit from both an electrification of the system and from the installation of batteries.
+The system electrification was explored by the authors \cite{Baldi2016b} and showed a relevant potential from both an environmental and an economic point of view.
+The installation of batteries was also considered in a previous publication by the authors \cite{Baldi2017} and proved the potential for yearly savings of 1-2\%.
+In this second case, however, it should be noted that the economic performance was not ensured.
+
+The availability of waste heat that is not already recovered by the system suggests that the system could benefit from the installation of a Heat-to-power WHR.
+This possibility was tested by the authors both in the case that no additional retrofit is performed \cite{Ahlgren2016,Mondejar2017}, with estimated savings of 22\% of the auxiliary power demand for a reference voyge, and in the case of a full integration with the rest of the system \cite{Baldi2017}, with estimated savings of up to 6\% of the total fuel consumption for a year of ship operations.
+
+It should be noted that in the second case the advantages in terms of system performance were obtained through a system-wide optimization of the system, where an improved utilization of the low-temperature waste heat available allowed "freeing" part of the high-temperature heat from the exhaust gas to the heat-to-power WHR system, hence improving the overall exergy efficiency of the system.
+This includes, for instance, the concept of using heat from the LT cooling systems (currently unused) for the HVAC preheater, hence reducing exergy destruction in the HVAC preheater itself and in the steam heater.
+This concept could be extended to other low-temperature heat consumers (such as the machinery space heating), but it would clearly constitute a larger benefit if applied to the HVAC preheater, which is the largest heat demand on board.
+
+Add reduction of non-HVAC el demand. Say absorption chillers are useless
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+\appendix
+
+\section{Engine modelling, figures and tables}
 
@@
 		Exhaust gas (after turbine) & Heat & 1672 & 2566 & 3227 & 3938 \\
-		CAC-LT & Heat & 104 & 152 & 279 & 454 \\
-		CAC-HT & Heat & 2 & 337  & 569 & 843  \\
-		JWC & Heat & 381 & 249 & 278 & 318 \\
-		LOC & Heat & 682 & 721 & 699 & 725  \\
+		CAC-LT & Heat & 80 & 197 & 353 & 613 \\
+		CAC-HT & Heat & 0 & 256  & 470 & 763  \\
+		JWC & Heat & 501 & 507 & 549 & 620 \\
+		LOC & Heat & 501 & 507 & 549 & 620 \\
 		\hline
@@
 		Exhaust gas (after turbine) & Heat & 808 & 1144 & 1456 & 1753 \\
-		CAC-LT & Heat & 5 & 91 & 139 & 216 \\
-		CAC-HT & Heat & 0 & 5 & 89 & 257  \\
-		JWC & Heat & 176 & 314 & 396 & 467 \\ 
-		LOC & Heat & 405 & 448 & 499 & 587  \\
+		CAC-LT & Heat & 5 & 92 & 139 & 185 \\
+		CAC-HT & Heat & 0 & 4 & 87 & 256  \\
+		JWC & Heat & 278 & 366 & 436 & 510 \\ 
+		LOC & Heat & 278 & 366 & 436 & 510 \\ 
 		\hline
@@
 
-\subsubsection{Cooling systems}
-
-The ship's cooling systems are designed similarly to most ships and divided in two cooling systems of two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C).
-The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board.
-A representation of the ship cooling systems is provided in Figure \ref{fig:generalsystemssimplified}.
-
-All values not directly measured in the cooling systems were calculated based on mass and energy balances.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/GeneralSystemsSimplified}
-	\caption{Layout of the ship heating and cooling system}
-	\label{fig:generalsystemssimplified}
-\end{figure}
-
-\subsection{Estimation of the electric power demand} \label{sec:met:electric}
-
-The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy
-\begin{equation}
-P_{el,tot} = \sum_{1}^{4} AE_i
-\end{equation}
-
-The determination of how different consumers contribute to the total electric power demand is less trivial.
-The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc.
-None of the individual contributions is directly measured, hence estimations need to be based on indirect measurements.
-
-\subsubsection{HVAC systems}
-
-For what concerns the air conditioning unit demand (HVAC), the observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a clear period, corresponding to the summer months, during which there is a significant increase of the demand.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pel_vs_time}
-	\caption[Total electric power demand versus time (Daily average)]{Total electric power demand versus time (Daily average)}
-	\label{fig:pelvstime}
-\end{figure}
-
-Since the total demand seems to be rather constant otherwise, it can be concluded that the demand of the HVAC compressors related to the need for cooling is concentrated during the summer months.
-Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a random summer and winter day, shows that the daily variation is comparable.
-This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/PelRel_vs_time}
-	\caption{Instantaneous power demand over daily average, winter versus summer day}
-	\label{fig:pelrelvstime}
-\end{figure}
-
-Hence, the HVAC electric power demand was estimated as follows:
-\begin{equation}
-P_{el,HVAC} =
-\begin{cases}
-P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
-0, & \text{otherwise}
-\end{cases}
-\end{equation}
-where $P_{el,ref}(t)$ is calculated as:
-\begin{equation}
-P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
-\end{equation}
-
-\subsubsection{Thrusters}
-
-When entering port areas, the ship needs to use thrusters to maneuver and berth.
-In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.
-In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
-From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).
-
-By comparing the reference profile to the instantaneous one, the points where the former is more than 10\% higher than the latter are identified as "thruster-on" points and treated consequently.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Thruster_selection}
-	\caption{Example of the procedure of selection of thruster power demand}
-	\label{fig:thruster_selection}
-\end{figure}
-
-\subsection{Estimation of the heat demand} \label{sec:met:heat}
-
-As the heat demand is not measured, it is necessary to determine it based on the available indirect measurements.
-
-The heat demand and generation can be summarized according to the following equations:
-\begin{eqnarray}
-\dot{Q}_{gen} & = & \dot{Q}_{EGB} + \dot{Q}_{HTHR} + \dot{Q}_{AB} \\
-\dot{Q}_{dem} & = & \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \\
-\end{eqnarray}
-
-The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.999\linewidth]{Figures/Heating_cooling_systems}
-	\caption{General view of the ship heating and cooling systems}
-	\label{fig:ship:general}
-\end{figure}
-
-\subsubsection{Heat balance parameter estimation}
-
-As not enough information and measurements are available to determine the various components of the heat balance, in this paper we determined them by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters.
-The parameter estimation problem is hence written as a minimization problem:
-\begin{eqnarray}
-min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
-\end{eqnarray}
-
-where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections.
-A list of the parameters $\textbf{p}$ is shown in Table REF, together with the chosen upper and lower boundaries for the calibration procedure.
+\section{Full tables of energy and exergy flows}
+
+\subsection{Exergy flows}
+
 \begin{table}
 	\centering
-	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
-		\hline 
-		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
-		\hline
-		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 &  \\ 
-		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 &  \\
-		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 &  \\ 
-		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 &  \\ 
-		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 &  \\
-		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 &  \\
-		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 &  \\ 
-		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 &  \\ 
-		\hline
+	\begin{tabular}{clclcrrrrrrrr}
+\toprule
+\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & Port/Stay & Low Speed Sailing & Maneuvering & High Speed Sailing & Winter & Mid-Season & Summer\\
+\midrule
+E & Env. & ME & Comp. & Ph & 0.16 & 0.000 & 0.141 & 0.013 & 0.010 & 0.102 & 0.048 & 0.014\\
+ME & Comp. & ME & CAC-HT & Ph & 13.64 & 0.004 & 9.932 & 1.148 & 2.555 & 4.976 & 5.626 & 3.037\\
+ME & CAC-HT & ME & CAC-LT & Ph & 12.20 & 0.004 & 9.008 & 1.023 & 2.164 & 4.476 & 5.079 & 2.643\\
+ME & CAC-LT & ME & Cyl. & Ph & 11.04 & 0.004 & 8.144 & 0.932 & 1.960 & 4.019 & 4.619 & 2.401\\
+ME & Turb. & ME & Comp. & W & 20.16 & 0.006 & 15.287 & 1.659 & 3.207 & 7.539 & 8.523 & 4.098\\
+ME & Cyl. & ME & Turb. & Ph & 55.91 & 0.037 & 43.951 & 4.839 & 7.085 & 22.177 & 23.408 & 10.327\\
+ME & Turb. & E & HRSG & Ph & 15.45 & 0.014 & 12.621 & 1.353 & 1.458 & 6.605 & 6.320 & 2.520\\
+ME & Turb. & E & Losses & Q & 0.81 & 0.001 & 0.653 & 0.069 & 0.087 & 0.341 & 0.338 & 0.131\\
+ME & Turb. & E & Env. & Ph & 14.04 & 0.014 & 11.256 & 1.285 & 1.489 & 5.694 & 5.941 & 2.409\\
+ME & Comp. & ME & BP valve & Ph & 17.44 & 0.004 & 13.179 & 1.425 & 2.831 & 6.557 & 7.323 & 3.559\\
+ME & BP valve & ME & Turb. & Ph & 3.80 & 0.001 & 3.246 & 0.277 & 0.275 & 1.581 & 1.697 & 0.522\\
+ME & CAC-HT & CS & HT Cool. & Q & 1.16 & 0.000 & 0.924 & 0.125 & 0.392 & 0.501 & 0.547 & 0.394\\
+ME & CAC-LT & CS & LT Cool. & Q & 0.59 & 0.000 & 0.864 & 0.091 & 0.203 & 0.456 & 0.460 & 0.242\\
+ME & Cyl. & ME & LOC & Q & 4.18 & 0.011 & 3.442 & 0.443 & 0.286 & 1.860 & 1.758 & 0.565\\
+ME & Cyl. & ME & JWC & Q & 8.13 & 0.022 & 6.660 & 0.879 & 0.565 & 3.424 & 3.514 & 1.188\\
+ME & JWC & CS & HT Cool. & Q & 5.30 & 0.022 & 6.660 & 0.879 & 0.565 & 3.424 & 3.514 & 1.188\\
+ME & LOC & CS & LT Cool. & Q & 2.45 & 0.011 & 3.442 & 0.443 & 0.286 & 1.860 & 1.758 & 0.565\\
+ME & Cyl. & E & Losses & Q & 0.63 & 0.002 & 0.525 & 0.067 & 0.040 & 0.275 & 0.271 & 0.088\\
+ME & HRSG & S & Boiler & Q & 4.63 & 0.008 & 5.622 & 0.628 & 0.498 & 3.015 & 2.677 & 1.064\\
+ME & HRSG & E & Env. & Ph & 8.69 & 0.006 & 6.999 & 0.725 & 0.961 & 3.590 & 3.643 & 1.456\\
+ME & Cyl. & P & Prop. shaft & W & 107.39 & 0.141 & 85.158 & 10.141 & 11.947 & 42.238 & 45.916 & 19.233\\
+ME & Fuel tank & ME & Cyl. & Ch & 266.02 & 0.382 & 212.237 & 25.450 & 27.949 & 105.146 & 114.197 & 46.675\\
+ME & Fuel tank & ME & Cyl. & Q & 0.15 & 0.000 & 0.125 & 0.014 & 0.014 & 0.068 & 0.063 & 0.022\\
+\midrule
+E & Env. & AE & Comp. & Ph & 0.01 & 0.002 & 0.003 & 0.001 & 0.000 & 0.003 & 0.001 & 0.001\\
+AE & Comp. & AE & CAC-HT & Ph & 7.91 & 2.960 & 4.008 & 0.659 & 0.288 & 3.392 & 3.091 & 1.432\\
+AE & CAC-HT & AE & CAC-LT & Ph & 7.65 & 2.835 & 3.906 & 0.644 & 0.269 & 3.302 & 3.017 & 1.335\\
+AE & CAC-LT & AE & Cyl. & Ph & 7.09 & 2.589 & 3.650 & 0.606 & 0.248 & 3.066 & 2.802 & 1.227\\
+AE & Turb. & AE & Comp. & W & 9.54 & 3.504 & 4.853 & 0.841 & 0.338 & 4.087 & 3.762 & 1.687\\
+AE & Cyl. & AE & Turb. & Ph & 36.26 & 11.776 & 19.815 & 3.592 & 1.079 & 15.705 & 14.961 & 5.597\\
+AE & Turb. & E & HRSG & Ph & 21.04 & 6.378 & 11.936 & 2.157 & 0.571 & 9.166 & 8.866 & 3.011\\
+AE & Turb. & E & Losses & Q & 0.46 & 0.132 & 0.259 & 0.060 & 0.012 & 0.209 & 0.194 & 0.060\\
+AE & Comp. & AE & BP valve & Ph & 7.91 & 2.960 & 4.008 & 0.659 & 0.288 & 3.392 & 3.091 & 1.432\\
+AE & BP valve & AE & Turb. & Ph & 0.00 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
+AE & CAC-HT & CS & HT Cool. & Q & 0.24 & 0.125 & 0.102 & 0.015 & 0.018 & 0.090 & 0.074 & 0.097\\
+AE & CAC-LT & CS & LT Cool. & Q & 0.36 & 0.245 & 0.256 & 0.037 & 0.021 & 0.236 & 0.215 & 0.108\\
+AE & Cyl. & AE & LOC & Q & 3.14 & 0.962 & 1.757 & 0.344 & 0.075 & 1.476 & 1.275 & 0.388\\
+AE & Cyl. & AE & JWC & Q & 5.84 & 1.772 & 3.270 & 0.656 & 0.144 & 2.608 & 2.451 & 0.783\\
+AE & JWC & CS & HT Cool. & Q & 3.90 & 1.772 & 3.270 & 0.656 & 0.144 & 2.608 & 2.451 & 0.783\\
+AE & LOC & CS & LT Cool. & Q & 2.18 & 0.962 & 1.757 & 0.344 & 0.075 & 1.476 & 1.275 & 0.388\\
+AE & Cyl. & E & Losses & Q & 0.37 & 0.096 & 0.217 & 0.047 & 0.008 & 0.166 & 0.159 & 0.044\\
+AE & HRSG & S & Boiler & Q & 6.32 & 2.869 & 5.504 & 1.011 & 0.237 & 4.309 & 3.901 & 1.412\\
+AE & HRSG & E & Env. & Ph & 11.42 & 3.509 & 6.432 & 1.145 & 0.334 & 4.857 & 4.965 & 1.599\\
+AE & Cyl. & EL & SB & W & 62.02 & 20.781 & 34.328 & 6.249 & 1.851 & 26.971 & 26.344 & 9.894\\
+AE & Cyl. & E & AG losses & Q & 0.85 & 0.228 & 0.499 & 0.105 & 0.020 & 0.384 & 0.363 & 0.104\\
+AE & Fuel tank & AE & Cyl. & Ch & 172.09 & 54.452 & 94.876 & 17.916 & 4.849 & 73.199 & 72.744 & 26.150\\
+AE & Fuel tank & AE & Cyl. & Q & 0.08 & 0.019 & 0.054 & 0.009 & 0.003 & 0.041 & 0.033 & 0.010\\
+\midrule
+CS & HT Cool. & CS & LT Cool. & Q & 7.36 & 1.46 & 8.63 & 1.36 & 0.97 & 5.10 & 5.25 & 2.08\\
+CS & HT Cool. & S & HRHT & Q & 3.24 & 0.46 & 2.33 & 0.31 & 0.14 & 1.53 & 1.34 & 0.38\\
+S & Boiler & S & HRHT & Q & 10.87 & 4.94 & 5.09 & 0.75 & 0.09 & 6.51 & 3.82 & 0.53\\
+S & Aux Boiler & S & Boiler & Ph & 5.71 & 4.66 & 0.90 & 0.17 & -0.03 & 3.77 & 1.85 & 0.10\\
+S & Aux Boiler & E & Env. & Ph & 1.51 & 1.04 & 0.38 & 0.08 & 0.00 & 0.99 & 0.50 & 0.03\\
+S & Fuel tank & S & Aux Boiler & Ch & 20.12 & 14.37 & 4.73 & 1.02 & 0.01 & 12.99 & 6.76 & 0.37\\
+S & HRHT & D & HVAC-PH & Q & 7.43 & 1.22 & 1.89 & 0.25 & 0.05 & 2.29 & 1.11 & 0.00\\
+S & HRHT & D & HVAC-RH & Q & 0.33 & 0.09 & 0.13 & 0.03 & 0.03 & 0.00 & 0.00 & 0.28\\
+S & HRHT & D & HWH  & Q & 2.27 & 0.64 & 1.09 & 0.17 & 0.05 & 0.84 & 0.81 & 0.31\\
+S & Boiler & D & MSH & Q & 0.68 & 0.06 & 0.10 & 0.01 & 0.00 & 0.09 & 0.07 & 0.02\\
+S & Boiler & D & HFO TH & Q & 0.66 & 0.13 & 0.20 & 0.03 & 0.01 & 0.16 & 0.16 & 0.04\\
+S & Boiler & D & HFO H & Q & 0.44 & 0.04 & 0.21 & 0.03 & 0.02 & 0.13 & 0.12 & 0.04\\
+S & Boiler & D & TH & Q & 0.51 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
+S & Boiler & D & G & Q & 2.45 & 0.44 & 0.98 & 0.14 & 0.04 & 0.71 & 0.69 & 0.19\\
+S & Boiler & D & OT & Q & 0.34 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
+EL & SB & EL & Thrusters & W & 0.94 & 0.00 & 0.00 & 0.94 & 0.00 & 0.36 & 0.43 & 0.15\\
+EL & SB & EL & HVAC & W & 1.79 & 0.59 & 0.85 & 0.18 & 0.18 & 0.00 & 0.00 & 1.79\\
+EL & SB & EL & Other el. Dem. & W & 59.29 & 19.99 & 32.66 & 4.99 & 1.64 & 26.06 & 25.45 & 7.78\\
+P & Prop. shaft & P & Prop. & W & 104.19 & 0.14 & 82.62 & 9.84 & 11.59 & 40.98 & 44.55 & 18.66\\
+P & Prop. shaft & E & Losses & Q & 3.20 & 0.00 & 2.54 & 0.30 & 0.36 & 1.26 & 1.37 & 0.57\\
+S & Boiler & E & Dumped steam & Q & 0.81 & 0.00 & 0.43 & 0.13 & 0.25 & 0.05 & 0.20 & 0.55\\
+\midrule
+& HVAC-PH & D & Dem. & Q & 3.41 & 1.22 & 1.89 & 0.25 & 0.05 & 2.29 & 1.11 & 0.00\\
+& HVAC-RH & D & Dem. & Q & 0.28 & 0.09 & 0.13 & 0.03 & 0.03 & 0.00 & 0.00 & 0.28\\
+& HWH  & D & Dem. & Q & 1.96 & 0.64 & 1.09 & 0.17 & 0.05 & 0.84 & 0.81 & 0.31\\
+& MSH & D & Dem. & Q & 0.18 & 0.06 & 0.10 & 0.01 & 0.00 & 0.09 & 0.07 & 0.02\\
+& HFO TH & D & Dem. & Q & 0.37 & 0.13 & 0.20 & 0.03 & 0.01 & 0.16 & 0.16 & 0.04\\
+& HFO H & D & Dem. & Q & 0.29 & 0.04 & 0.21 & 0.03 & 0.02 & 0.13 & 0.12 & 0.04\\
+& TH & D & Dem. & Q & 0.23 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
+& G & D & Dem. & Q & 1.59 & 0.44 & 0.98 & 0.14 & 0.04 & 0.71 & 0.69 & 0.19\\
+& OT & D & Dem. & Q & 0.15 & 0.05 & 0.08 & 0.01 & 0.00 & 0.07 & 0.07 & 0.02\\
+\bottomrule
 	\end{tabular}
-	\caption{Parameters optimized in the parameter estimation for the heat balance}
-	\label{tab:ParameterEstimation} 
 \end{table}
-
-\subsubsection{Heat demand}
-
-We calculated the heat demand as the sum of the contributions of the elements listed in the ship's heat balance documentation.
-As no direct measurement of these quantities was available in the dataset, they had to be estimated based on the following assumptions:
-\begin{table}
-	\centering
-	{\tablinesep=2ex\tabcolsep=10pt
-		\begin{tabular}{p{2.8cm}l}
-			\hline 
-			Heat flow name & Equation \\
-			\hline
-			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
-			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
-			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
-			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
-			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
-			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
-			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
-			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
-			\hline
-	\end{tabular}}
-	\caption{Summary of the heat demand contributions and their calculation}
-	\label{tab:HeatDemand}
-\end{table}
-
-where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
-The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
-The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure REF.
-
-\subsubsection{Heat generation}
-
-The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty.
-The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
-\begin{equation}
-\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
-\end{equation}\label{eq:egb}
-
-where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF
-
-It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
-However, measurements of this contribution and of other variables that could lead to its straight-forward identification are missing.
-In this work, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},
-
-\begin{eqnarray}
-\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
-& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
-\end{eqnarray}
-
-where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
-The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
-
-The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems.
-The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit.
-This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value.
-In this calculation, the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
-
-\subsubsection{Parameter estimation and uncertainty quantification}
-
-The results of the parameter estimation are represented graphically and quantitatively in Figure REF and Table REF.
-
-Given the lack of input values for the estimation of the heat demand, the provided estimate based on the procedure described above should be integrated with an estimation of the uncertainty.
-
-In this study, we base the estimation of the uncertainty on the production side, as it is the one that has the largest amount of information available.
-In these regards, the uncertainty can be reduced to the contribution of three elements: $U(\dot{Q}_{EGB})$, $U(\dot{Q}_{HTHR})$, $U(\dot{Q}_{AB})$.
-
-The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$.
-In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT].
-The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions as described in Appendix REF, and can be estimated of being up to 10\%.
-The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value.
-These assumptions lead to the following estimation of the uncertainty:
-\begin{equation}
-\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
-\end{equation}
-
-That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements (result obtained for $\pm 20$ K uncertainty.
-The uncertainty is reduced to 20\% if the measurement uncertainty on the temperature is reduced to  $\pm 20$ K)
-
-The uncertainty of the heat generated by the ABs can be reduced to the contribution of two elements: the uncertainty on $\dot{m}_{fuel,AB}$ and that on $\eta_{AB}$.
-The former will be at least as large as the calibration error (35\%) and will be considered equal to 50\% to be conservative and accounting also for errors in the aggregated boiler fuel measurements.
-In addition, the uncertainty on the efficiency can be considered to be around 10\% based on the discrepancy between the considered sources and on the expected variability of the efficiency with load.
-Similarly to the previous case, the combination of these efficiencies lead to a total uncertainty of 51\%, where the main contribution comes from the uncertainty of the model output compared to the actual fuel consumption.
-
-The estimation of the uncertainty of $\dot{Q}_{HTHR}$ can be based on \ref{eqn:HTHR2} having assigned an acceptable variation of 20\% to the effectiveness of the heat exchanger, 20\% on the estimation of the reference mass flow, $\pm 5$ K uncertainty on water temperature measurements and considering negligible the uncertainty on $c_{p,w}$, leading to a 145\% uncertainty on this measurement.
-
-\subsection{Operational mode}
-
-The ship operates in different conditions, and it can be useful to provide a separate analysis depending on the operational mode.
-The four operational modes identified for the selected case study, together with the conditions applied to the dataset to uniquely identify each time step to a separate operational mode, are given as follows:
-\begin{description}
-	\item[High speed sailing] Ship speed above 15 kn
-	\item[Low speed sailing] Ship speed between 4 and 15 kn
-	\item[Maneuvering] Ship speed between 2 and 4 knots OR Thrusters power larger than 0
-	\item[Port stays / ship drifting] All other points
-\end{description}
-
-\subsection{Determination of typical operational days}
-
-\section{Results (and discussion)}
-
-\subsection{Exploratory analysis} \label{sec:res:exploratory}
-
-In this section, we observe and analyze direct measurements from ship systems, with no pre-processing, that are expected to have an influence on the energy analysis.
-
-Figures \ref{fig:TseaTairTIME} and \ref{fig:TseaTairDIST} represent the time evolution and the distribution over the year of the ambient air and sea water temperature.
-The relatively low temperatures that are experienced by the case study ship during its operations rises expectations for a high space heating demand and to a relatively low for summer HVAC electric power demand.
-On the other hand, the fact that, in summer, air temperatures reach up to 26 degrees Celsius justify the existence of an HVAC unit that can also operate in cooling mode.
-
-\begin{figure}[htbp]
-	\centering
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centering
-		\includegraphics[width=0.999\linewidth]{Figures/Tsea_vs_time}
-		\caption{Time series}
-		\label{fig:TseaTairTIME}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{Figures/Tsea_hist}
-		\caption{Distribution}
-		\label{fig:TseaTairDIST}
-	\end{subfigure}
-	\caption{Statistical representation of measured air and sea water temperatures based on yearly data. Landsort, 2014}
-	\label{fig:TseaTair}
-\end{figure}
-
-Figure \ref{fig:vship_hist} represents the distribution of the ship speed.
-The ship operates almost for the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
-The relatively large amount of time the ship spends in port can also be seen in Figure \ref{fig:opMode_pie}.
-This suggests that, in spite of the high installed power for propulsion, the energy demand for this category is expected to be relatively low.
-
-Figures \ref{fig:loadME_hist} and \ref{fig:loadME_hist} show the frequency of the load at which each of the engines (main and auxiliary engines, respectively) is operated.
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/vship_hist}
-	\caption{Yearly distribution of ship speed}
-	\label{fig:vship_hist}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/Pie_operationalMode}
-	\caption{Operational time by mode}
-	\label{fig:opMode_pie}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/Hist_mainEngines}
-	\caption{Yearly load distribution, main engines}
-	\label{fig:loadME_hist}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/Hist_auxEngines}
-	\caption{Yearly load distribution, auxiliary engines}
-	\label{fig:loadAE_hist}
-\end{figure}
-
-ANYTHING ELSE THAT CAN BE INTERESTING TO SHOW IN THE EXPLORATORY DATA ANALYSIS?
-
-\subsection{Energy analysis} \label{sec:res:energy}
-
-Here we can see the typical example (note: the day was not selected as "typical", but simply randomly) of the energy demand during a winter and a summer day (Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}).
-It can be seen clearly how, as expected, the heating demand in winter is much higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduce need for compartment heating.
-On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors.
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_W}
-	\caption{Energy demand during a day of operations, Winter (31 Jan)}
-	\label{fig:Demand_vs_time_W}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Demand_vs_time_S}
-	\caption{Energy demand during a day of operations, Summer (31 Jul)}
-	\label{fig:Demand_vs_time_S}
-\end{figure}
-
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyDemand}
-	\caption{Energy demand, aggregated data for one year of operations}
-	\label{fig:Pie_EnergyDemand}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyGeneration}
-	\caption{Energy generation, aggregated data for one year of operations}
-	\label{fig:Pie_EnergyGeneration}
-\end{figure}
-
-From Figure \ref{fig:Pie_EnergyDemand} it can be noted that nearly 50\% of the total energy demand is related to ship propulsion, while the remaining portion is approximately equally split between electric energy and heat demand.
-
-Within the electric energy demand, the HVAC systems only represent a minor contribution (XX\%), as it is only present for a few months during summer, and at rather low power demand.
-This is not surprising, since it can be observed that (see Figure \ref{fig:TseaTairDIST}) the air is always below 27$^o$C, and rarely above 17$^o$C.
-Also the thrusters, although they represent a rather high instantaneous demand, they are only used for a very short time and, hence, their total contribution to the ship energy demand is limited to XX\%.
-As a consequence, most of the electric energy demand should be allocated to other consumers, and efforts should be directed towards providing additional measurements of how different users contribute to the total electric power demand.
-
-Looking at Figure \ref{fig:Pie_HeatGeneration} it is possible to appreciate the amount of heat that is recovered from on board available waste heat (cooling water and exhaust gas).
-It can be observed that  of heat recovered from the HT cooling systems represents approximately one quarter of the total heat supply.
-Figure \ref{fig:TimeSeries_HeatGeneration} shows how the contribution of the HRSG and the HTHR is substantially constant throughout the year, while the auxiliary boilers are used to cover the peaks and the winter increased baseload.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/Pie_HeatGeneration}
-	\caption{Share of thermal power generated from on board heat systems}
-	\label{fig:Pie_HeatGeneration}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/TimeSeries_HeatGeneration}
-	\caption{Time series representation of the thermal power generated from on board heat systems}
-	\label{fig:TimeSeries_HeatGeneration}
-\end{figure}
-
-Regarding the heat demand onboard, it can be seen that the HVAC re-heater represents the largest contributor (XX\%) followed by the hot water heater (XX\%).
-The large contribution of the HVAC reheater can be explained by the former being important for long times and in relation to the low ambient temperatures encountered by the case study vessel during operation.
-The relatively large contribution of the hot water heater can instead be related to the fact that these systems are rather important consumers and are necessary during the whole ship operations (i.e.
-also during the summer months).
-
-\subsection{Exergy Analysis} \label{sec:res:exergy}
-
-The results of the exergy analysis of the ship energy systems are shown in Figure REF and Tables REF to REF. 
-\begin{itemize}
-	\item Table(s) --> Exergy efficiencies
-	\begin{itemize}
-		\item General at system level
-		\item One per system (MEs, AEs, CoolingSystems, SteamSystems) 
-	\end{itemize}
-	\item Figure --> Grassman diagram
-\end{itemize}
-
-It can be observed that there are large irreversibilities concentrated in some specific parts of the system that show potential for improvement.
-This is particularly the case of the LT/HT merging valves, where exergy is destroyed in the mixing process.
-Another substantial source of exergy destruction is the steam heater in the HTHR system, where heat at medium pressure (6 bar) is used to heat up water at 90$^o$C.
-The analysis of exergy destruction also points out the HVAC reheater as a potential source of improvement, based on the use of relatively high-temperature water for heating up air to around 30$^o$C and on the large energy demand of the component.
-Other important exergy destruction sources can be located in the engines, as expected, but the improvement of engine processes and technology goes beyond the scope of this work.
-
-From the point of view of the exergy losses, it appears that the largest exergy losses are located in the exhaust gas.
-It should be noted that, even though a large part of this exergy cannot be recovered given the limitations on the exhaust gas outlet temperature to avoid the condensation of sulphuric acid, there is still a significant potential to be harvested if all the available exergy in the exhaust gas was recovered.
-This potential is further increased by the exergy destroyed in the process of dumping steam when it is produced in excess during summer.
-The exergy lost in the sea water coolers is much lower, highlighting the fact that the potential for improving the efficiency of the system is located in other parts of the cooling systems.
-
-The efficiency of the recovery systems is also shown in Figure \ref{fig:Bar_efficiencyWHR}, where we show the fraction of the total energy (exergy) lost by the cooling systems and the exhaust gas of the ship's engines.
-It can be noted that, even when only looking at the energy that can be recovered based on the existing systems (i.e.
-not including the LT cooling systems, for instance), the recovery efficiency is located at around 25-35\% depending on whether the energy or the exergy efficiency is considered.
-Worth noting is also the fact that the efficiency on the exhaust gas side is lowered by the fact that two of the main engines are currently not equipped with HRSGs.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/Bar_WHRefficiency}
-	\caption{Energy and exergy efficiency of the WHR systems}
-	\label{fig:Bar_efficiencyWHR}
-\end{figure}
-
-\subsection{Typical operational days}  \label{sec:res:typicalDays}
-
-\subsection{Potential for system improvement}
-
-Based on the results of the energy and exergy analysis, it is possible to identify a number of possible measures that could potentially improve the efficiency of the system.
-
-Given the fact that many engines are operated at low load (both main and auxiliary engines), the system would benefit from both an electrification of the system and from the installation of batteries.
-The system electrification was explored by the authors \cite{Baldi2016b} and showed a relevant potential from both an environmental and an economic point of view.
-The installation of batteries was also considered in a previous publication by the authors \cite{Baldi2017} and proved the potential for yearly savings of 1-2\%.
-In this second case, however, it should be noted that the economic performance was not ensured.
-
-The availability of waste heat that is not already recovered by the system suggests that the system could benefit from the installation of a Heat-to-power WHR.
-This possibility was tested by the authors both in the case that no additional retrofit is performed \cite{Ahlgren2016,Mondejar2017}, with estimated savings of 22\% of the auxiliary power demand for a reference voyge, and in the case of a full integration with the rest of the system \cite{Baldi2017}, with estimated savings of up to 6\% of the total fuel consumption for a year of ship operations.
-
-It should be noted that in the second case the advantages in terms of system performance were obtained through a system-wide optimization of the system, where an improved utilization of the low-temperature waste heat available allowed "freeing" part of the high-temperature heat from the exhaust gas to the heat-to-power WHR system, hence improving the overall exergy efficiency of the system.
-This includes, for instance, the concept of using heat from the LT cooling systems (currently unused) for the HVAC preheater, hence reducing exergy destruction in the HVAC preheater itself and in the steam heater.
-This concept could be extended to other low-temperature heat consumers (such as the machinery space heating), but it would clearly constitute a larger benefit if applied to the HVAC preheater, which is the largest heat demand on board.
-
-OTHER?
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-\appendix
-
-\section{Full tables of energy and exergy flows}
 

### 2018-05-08 Python files/fillerfunctions.py

@@ -19,7 +19,7 @@ def systemFill(processed, dict_structure, CONSTANTS, system_type, call_ID):
     counter_p_tot = 0
     counter_T_tot = 0
     counter_c_tot = 0
-    processed = unitOffCheck(processed, dict_structure, CONSTANTS, call_ID)
+    processed = unitOffCheck(processed, dict_structure, CONSTANTS, system_set, call_ID)
     for system in system_set:
         for unit in dict_structure["systems"][system]["units"]:
             # (counter_c, processed) = connectionAssignment(processed, dict_structure, CONSTANTS, system, unit, call_ID)
@@ -37,7 +37,7 @@ def systemFill(processed, dict_structure, CONSTANTS, system_type, call_ID):
                 else:
                     print("Equation not recognized")
                         # Trivial assignment: doing this if the flow is connected with other flows
-    processed = unitOffCheck(processed, dict_structure, CONSTANTS, call_ID)
+    processed = unitOffCheck(processed, dict_structure, CONSTANTS, system_set, call_ID)
     for system in system_set:
         for unit in dict_structure["systems"][system]["units"]:
             (counter_c,processed) = connectionAssignment(processed, dict_structure, CONSTANTS, system, unit, call_ID)
@@ -47,9 +47,9 @@ def systemFill(processed, dict_structure, CONSTANTS, system_type, call_ID):
 
 
 
-def unitOffCheck(processed, dict_structure, CONSTANTS, call_ID):
+def unitOffCheck(processed, dict_structure, CONSTANTS, system_set, call_ID):
     # Assigns to 0 all values related to components when they are off
-    for system in dict_structure["systems"]:
+    for system in system_set:
         if processed[system + ":" + "on"].isnull().sum() == 0:
             for unit in dict_structure["systems"][system]["units"]:
                 for flow in dict_structure["systems"][system]["units"][unit]["flows"]:
@@ -85,7 +85,12 @@ def connectionAssignment(processed, dict_structure, CONSTANTS, system, unit, cal
                             processed[ID_c] = processed[ID]
                             counter = counter + 1
                         elif (processed[ID] - processed[ID_c]).sum() > processed[ID].max():
-                            text_file.write("ERROR. FUN: ppo.connectionAssignment. Something is wrong. Flows {} and {} should be the same, they are not. \n".format(ID, ID_c))
+                            error = abs(processed[ID] - processed[ID_c])
+                            relativeError = error / processed[ID].mean()
+                            averageError = relativeError.mean()
[…]

### 2018-05-10 Manuscript/Manuscript.tex
Updates to the manuscript

@@
 \usepackage{booktabs}
+% Package for improving table handling
+\usepackage{array}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 
-\subsection{Data processing: Ship energy system modeling} \label{sec:met:processing}
+\subsection{Ship energy system modeling} \label{sec:met:processing}
 
@@
 In absence of more accurate data, it is then assumed that the contributions of $\dot{Q}_{JWC}$ and $\dot{Q}_{LOC}$ are equal for all loads.
-CHANGE VALUES IN THE TABLE
-
-Values for temperatures, mass flows and energy flows for different engine loads are provided in the Appendix \ref{sec:appendix:engines}, for both main and auxiliary engines.
+
+Values for temperatures, mass flows and energy flows for different engine loads are provided in the \ref{sec:appendix:engines}, for both main and auxiliary engines.
 
@@
 All values not directly measured in the cooling systems were calculated based on mass and energy balances.
-The mass flow was determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps.
+Mass flows were determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps as provided in the engine project guides.
 
@@
 \begin{equation}
-P_{el,tot} = \sum_{1}^{4} AE_i
+P_{el,tot} = \sum_{1}^{4} P_{AE_i}
 \end{equation}
 
-The determination of how different consumers contribute to the total electric power demand is less trivial.
 The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc.
-None of the individual contributions is directly measured, hence estimations need to be based on indirect measurements.
+However, none of the individual contributions is directly measured, hence estimations of the energy demand of individual consumers need to be based on indirect measurements.
+In this paper we identified methods for the estimation of the contribution of two consumers: the HVAC systems and the bow thrusters.
 
@@
 
-For what concerns the air conditioning unit demand (HVAC), the observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a clear period, corresponding to the summer months, during which there is a significant increase of the demand.
+The observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a  period, corresponding to the summer months, when the electric power demand is significantly higher then during the remaining months.
+Based on this observation, in this work we assumed that the demand of the HVAC compressors related to the need for space cooling is concentrated during the summer months.
+Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a typical summer and winter day, shows that the daily variation is comparable.
+This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.
 
@@
 \end{figure}
-
-Since the total demand seems to be rather constant otherwise, it can be concluded that the demand of the HVAC compressors related to the need for cooling is concentrated during the summer months.
-Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a random summer and winter day, shows that the daily variation is comparable.
-This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.
 
@@
 In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.
-In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
-From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).
+In order to isolate the demand from the thrusters, we extracted from ship operations a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
+From this daily profile (made of 96 points), we manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).
 
@@
 
-As the heat demand is not measured, it is necessary to determine it based on the available indirect measurements.
-
-The heat demand and generation can be summarized according to the following equations:
-\begin{eqnarray}
-\dot{Q}_{gen} & = & \dot{Q}_{EGB} + \dot{Q}_{HTHR} + \dot{Q}_{AB} \\
-\dot{Q}_{dem} & = & \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \\
-\end{eqnarray}
-
-The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}
+The case study ship is not provided with any means to directly measure the heat demand (e.g.
+steam flow meters), and it was hence necessary to provide estimations based on other measurements and assumptions.
+The heat demand is subdivided among the contributions of the HVAC preheater and reheater ($\dot{Q}_{HVAC,PH}$ and $\dot{Q}_{HVAC,RH}$), the hot water heater ($\dot{Q}_{HWH}$), the HFO heaters ($\dot{Q}_{HH}$) the tank heating systems for HFO and other liquids ($\dot{Q}_{HTH}$ and $\dot{Q}_{TH}$), the machinery space heating ($\dot{Q}_{MSH}$) and other tanks ($\dot{Q}_{OT}$).
+The heat generation is instead composed of the contribution of the exhaust gas boilers ($\dot{Q}_{EGB,ME,i}$ and $\dot{Q}_{EGB,AE,i}$), the heat recovery from the high temperature cooling systems ($\dot{Q}_{HTHR,j,i}$) and the boilers ($\dot{Q}_{AB}$).
+ The heat demand and generation are summarized in equations \ref{eq:Qgen} and \ref{eq:Qdem}.
+\begin{align}
+\dot{Q}_{gen} & = \sum_{i=2,3} \dot{Q}_{EGB,ME,i} +  \sum_{i=1:4} \dot{Q}_{EGB,AE,i} + \sum_{j=ME,AE, i=1:4}\dot{Q}_{HTHR,j,i} + \dot{Q}_{AB} \label{eq:Qgen} \\
+\dot{Q}_{dem} & = \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{HH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \label{eq:Qdem} \\
+\end{align}
+
+The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}.
 
@@
 
-As not enough information and measurements are available to determine the various components of the heat balance, in this paper we determined them by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters.
+The individual contributions of different consumers were determined by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters.
 The parameter estimation problem is hence written as a minimization problem:
@@
 where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections.
-A list of the parameters $\textbf{p}$ is shown in Table REF, together with the chosen upper and lower boundaries for the calibration procedure.
+A list of the parameters $\textbf{p}$ is shown in Table \ref{tab:ParameterEstimation}, together with the chosen upper and lower boundaries for the calibration procedure.
 \begin{table}
@@
 
-We calculated the heat demand as the sum of the contributions of the elements listed in the ship's heat balance documentation.
-As no direct measurement of these quantities was available in the dataset, they had to be estimated based on the following assumptions:
+The heat demand was calculated as the sum of the contributions of the elements listed in the ship's heat balance documentation.
+The assumptions used in the estimation are summarized in Table \ref{tab:HeatDemand}:
 \begin{table}
@@
 The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
-The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure REF.
+The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure \ref{fig:galleyhotwater}.
+
+\begin{figure}[tbp!]
+	\centering
+	\includegraphics[width=0.7\linewidth]{Figures/galleyHotWater.pdf}
+	\caption{Hourly relative demand of hot water heating and galley}
+	\label{fig:galleyhotwater}
+\end{figure}
 
@@
 
-where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF
+where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance.
 
 It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
-However, measurements of this contribution and of other variables that could lead to its straight-forward identification are missing.
-In this work, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},
+Based on the available measurements, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},
 
@@
 
-The results of the parameter estimation are represented graphically and quantitatively in Figure REF and Table REF.
-
-Given the lack of input values for the estimation of the heat demand, the provided estimate based on the procedure described above should be integrated with an estimation of the uncertainty.
-
 In this study, we base the estimation of the uncertainty on the production side, as it is the one that has the largest amount of information available.
@@
 In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT].
-The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions as described in Appendix REF, and can be estimated of being up to 10\%.
+The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions and can be estimated of being up to 10\%.
 The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value.
@@
 
-That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements (result obtained for $\pm 20$ K uncertainty.
-The uncertainty is reduced to 20\% if the measurement uncertainty on the temperature is reduced to  $\pm 20$ K)
+That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements.
 
@@
 Figure \ref{fig:vship_hist} represents the distribution of the ship speed.
-The ship operates almost for the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
-In Figure \ref{fig:opMode_pie} it can be also observed that the ship spends a relatively high amount of time in port.
+The ship operates for almost  the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
+In Figure \ref{fig:opMode_pie} it can be also observed that the ship spends a relatively large amount of time in port.
 These remarks suggest that, in spite of the high installed power of the main engines, the energy demand for propulsion is expected to be relatively low.
 
-Figures \ref{fig:loadME_hist} and \ref{fig:loadME_hist} show the frequency of the load at which each of the engines (main and auxiliary engines, respectively) is operated.
+Figures \ref{fig:loadME_hist} and \ref{fig:loadAE_hist} show the frequency of the load at which each of the engines (main and auxiliary engines, respectively) is operated.
 In the case of the main engines, it can be observed that they are operated mostly at loads between 30\% and 60\%, despite the optimal load for the engines being located around 85\%.
@@
 AE-3 is generally more often operated, as it is the engine that is run in port with clean fuels.
-On the other hand, it seems that AE-4 is operated rarely, which might be due to maintenance issues.
+On the other hand, it seems that AE-4 is operated rarely, which might be due to maintenance issues during the selected period of analysis.
 
@@
 
-ANYTHING ELSE THAT CAN BE INTERESTING TO SHOW IN THE EXPLORATORY DATA ANALYSIS?
-
 \subsection{Energy analysis} \label{sec:res:energy}
 
-The ship energy demand is first subdivided among three main consumer categories: propulsion, electric power, and heat.
-The evolution of these demands over time during a typical voyage is shown in Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}, representing a winter and a summer day respectively.
-It can be seen clearly how, as expected, the heating demand in winter is higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduced need for compartment heating.
+The ship energy demand is first subdivided among  The evolution of the demands of propulsion, electric power, and heat over time during a typical voyage is shown in Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}, representing a winter and a summer day respectively.
+As expected, the heating demand in winter is higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduced need for compartment heating.
 On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors.
@@
 \ref{fig:Pie_HeatGeneration}).
-The the exhaust gas boilers and the heat recovery on the HT-water fulfill almost 75\% of the yearly demand for heating, leaving only the remaining 25\% to be provided by oil-fired auxiliary boilers.
+The exhaust gas boilers and the heat recovery on the HT-water fulfill almost 75\% of the yearly demand for heating, leaving only the remaining 25\% to be provided by oil-fired auxiliary boilers.
 The contribution of the HRSG and the HTHR is substantially constant throughout the year (Fig.
@@
 
-Looking at heating from the demand side, it can be seen that the HVAC pre-heater represents the largest contributor (56\%) to the annual heating demand, followed by the hot water heater (18\%) and by the galley (11\%).In multi-season central HVAC systems the pre-heater is generally used for the main heating contribution during winter, while the re-heater is only used for peak demand, or for adjusting the delivery air temperature during summer, thus explaining the large contribution of the former.
+On the heating demand side, the HVAC pre-heater represents the largest contributor (56\%) to the annual heating demand, followed by the hot water heater (18\%) and by the galley (11\%).
+In multi-season central HVAC systems the pre-heater is generally used for the main heating contribution during winter, while the re-heater is only used for peak demand, or for adjusting the delivery air temperature during summer, thus explaining the large contribution of the former.
 The relatively large share of the hot water heater can instead be related to the fact that these systems are used during all seasons.
@@
 
-The results of the exergy analysis of the ship energy systems are shown in Figure REF and Tables REF to REF. 
+The results of the exergy analysis of the ship energy systems are shown in Figure \ref{fig:Grassmann} and in Tables REF to REF in the Appendix. 
 \begin{itemize}
@@
 It can be observed that there are large irreversibilities concentrated in some specific parts of the system that show potential for improvement.
-This is particularly the case of the LT/HT merging valves, where exergy is destroyed in the mixing process.
-Another substantial source of exergy destruction is the steam heater in the HTHR system, where heat at medium pressure (6 bar) is used to heat up water at 90$^o$C.
-The analysis of exergy destruction also points out the HVAC reheater as a potential source of improvement, based on the use of relatively high-temperature water for heating up air to around 30$^o$C and on the large energy demand of the component.
-Other important exergy destruction sources can be located in the engines, as expected, but the improvement of engine processes and technology goes beyond the scope of this work.
+The individual largest contribution to exergy destruction in the ship systems is associated to combustion, both in the engines and in the boilers.
+The loss of energy resulting from the conversion of chemical energy to heat contributes to more than 76\% of the total exergy destruction on board.
+Once this contribution is isolated, the largest sources of exergy destruction are located in the engine turbochargers (28.9\%), in the exhaust gas boilers (10.7\%), in the steam heater of the HTHR system (10.5\%), the HVAC preheater (10.4\%) and the sea water cooler (8.5\%).
+In all the cases above, excluding that of the turbochargers, the exergy destruction is a consequence of a combination of large exergy flows and of the mismatch between hot and cold streams of the heat exchangers.
+In the exhaust boilers, exhaust gas are cooled in average from more than 300$^o$C to 200$^o$C, while the heat is used to generate steam at relatively low pressure (6 bar); in the steam heater, the steam is used to heat up water at 90$^o$C; in the HVAC preheater, the relatively high-temperature water from the HTHR systems is used for heating up air to around 30$^o$C.
+
 \begin{table}
@@
 		\hline 
-		LT-HT merge & 13.4\% \\ 
-		\hline 
-		Sea water cooler & 10.6\% \\ 
-		\hline 
-		Steam heater & 6.0\% \\ 
-		\hline 
-		HVAC preheater & 6.0\% \\ 
+		Turbochargers & 28.9\% \\ 
+		HRSGs & 10.7\% \\
+		Steam heater & 10.5\% \\ 
+		HVAC preheater & 10.4\% \\
+		Sea water cooler & 8.5\% \\ 
+		Electric generators & 6.1\% \\ 
 		\hline 
@@
 
-Add reduction of non-HVAC el demand. Say absorption chillers are useless
+The analysis also suggests that the electric power demand resulting from the use of the HVAC on board is relatively low, a consequence of the fact that the case study ship mostly operates in cold climates.
+This suggests that demand-related efforts should be focused on other consumers (e.g.
+lighting) rather than on the HVAC compressor.
+Also, the analysis suggests that the use of related systems for WHR, such as absorption chillers, would not be particularly beneficial in the specific case here studied.
 
@@
 
-\section{Engine modelling, figures and tables}
+\section{Engine modelling, figures and tables} \label{sec:appendix:engines}
 
@@
 
-\section{Full tables of energy and exergy flows}
-
-\subsection{Exergy flows}
-
-\begin{table}
-	\centering
+\clearpage
+
+\section{Full tables of energy and exergy flows} \label{sec:appendix:energy}
+
+\subsection{Energy flows}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
 		\begin{tabular}{clclcrrrrrrrr}
 			\toprule
-\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & Port/Stay & Low Speed Sailing & Maneuvering & High Speed Sailing & Winter & Mid-Season & Summer\\
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+			\midrule
+E & Env. & ME & Comp. & Ph & 4.57 & 0.005 & 3.819 & 0.385 & 0.359 & 2.327 & 1.671 & 0.571\\
+ME & Comp. & ME & CAC-HT & Ph & 19.27 & 0.010 & 14.412 & 1.645 & 3.208 & 7.494 & 7.822 & 3.958\\
+ME & CAC-HT & ME & CAC-LT & Ph & 13.98 & 0.010 & 10.900 & 1.190 & 1.885 & 5.717 & 5.728 & 2.541\\
+ME & CAC-LT & ME & Cyl. & Ph & 7.35 & 0.009 & 5.951 & 0.655 & 0.737 & 3.340 & 2.969 & 1.043\\
+ME & Turb. & ME & Comp. & W & 20.16 & 0.006 & 15.287 & 1.659 & 3.207 & 7.539 & 8.523 & 4.098\\
+ME & Cyl. & ME & Turb. & Ph & 103.85 & 0.093 & 82.544 & 9.285 & 11.924 & 41.276 & 43.973 & 18.597\\
+ME & Turb. & E & HRSG & Ph & 44.77 & 0.044 & 36.434 & 3.974 & 4.318 & 18.383 & 18.840 & 7.546\\
+ME & Turb. & E & Losses & Q & 3.32 & 0.003 & 2.652 & 0.289 & 0.378 & 1.307 & 1.425 & 0.590\\
+ME & Comp. & ME & BP valve & Ph & 24.73 & 0.011 & 19.106 & 2.045 & 3.567 & 9.866 & 10.194 & 4.669\\
+ME & Bypass valve & ME & Turb. & Ph & 5.45 & 0.002 & 4.694 & 0.399 & 0.359 & 2.372 & 2.372 & 0.711\\
+ME & CAC-HT & CS & HT Cool. & Q & 5.29 & 0.000 & 3.512 & 0.455 & 1.323 & 1.778 & 2.094 & 1.418\\
+ME & CAC-LT & CS & LT Cool. & Q & 6.63 & 0.001 & 4.949 & 0.535 & 1.147 & 2.377 & 2.758 & 1.497\\
+ME & Cyl. & ME & LOC & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
+ME & Cyl. & ME & JWC & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
+ME & JWC & CS & HT Cool. & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
+ME & LOC & CS & LT Cool. & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
+ME & Cyl. & E & Losses & Q & 2.59 & 0.007 & 2.133 & 0.280 & 0.172 & 1.054 & 1.143 & 0.396\\
+ME & HRSG & S & Boiler & Q & 13.41 & 0.018 & 11.130 & 1.277 & 0.987 & 5.810 & 5.439 & 2.163\\
+ME & HRSG & E & Env. & Ph & 31.36 & 0.025 & 25.304 & 2.697 & 3.331 & 12.573 & 13.402 & 5.383\\
+ME & Cyl. & P & Prop. sh. & W & 107.39 & 0.141 & 85.158 & 10.141 & 11.947 & 42.238 & 45.916 & 19.233\\
+ME & Fuel tank & ME & Cyl. & Chl & 250.62 & 0.360 & 199.949 & 23.977 & 26.331 & 99.059 & 107.585 & 43.973\\
+ME & Fuel tank & ME & Cyl. & Q & 1.09 & 0.002 & 0.875 & 0.102 & 0.109 & 0.455 & 0.458 & 0.174\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of energy flows for the main engines}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+			\midrule
+E & Env. & AE & Comp. & Ph & 0.42 & 0.13 & 0.23 & 0.05 & 0.01 & 0.25 & 0.09 & 0.08\\
+AE & Comp. & AE & CAC-HT & Ph & 9.96 & 3.64 & 5.08 & 0.89 & 0.35 & 4.34 & 3.85 & 1.76\\
+AE & CAC-HT & AE & CAC-LT & Ph & 8.91 & 3.13 & 4.68 & 0.83 & 0.27 & 3.99 & 3.54 & 1.38\\
+AE & CAC-LT & AE & Cyl. & Ph & 5.62 & 1.76 & 3.13 & 0.59 & 0.14 & 2.75 & 2.20 & 0.67\\
+AE & Turb. & AE & Comp. & W & 9.54 & 3.50 & 4.85 & 0.84 & 0.34 & 4.09 & 3.76 & 1.69\\
+AE & Cyl. & AE & Turb. & Ph & 70.69 & 21.90 & 39.33 & 7.46 & 2.00 & 30.33 & 29.78 & 10.57\\
+AE & Turb. & E & HRSG & Ph & 56.90 & 17.21 & 32.13 & 6.00 & 1.57 & 24.42 & 24.18 & 8.30\\
+AE & Turb. & E & Losses & Q & 1.89 & 0.54 & 1.05 & 0.25 & 0.05 & 0.80 & 0.82 & 0.27\\
+AE & Comp. & AE & BP valve & Ph & 9.96 & 3.64 & 5.08 & 0.89 & 0.35 & 4.34 & 3.85 & 1.76\\
+AE & Bypass valve & AE & Turb. & Ph & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
+AE & CAC-HT & CS & HT Cool. & Q & 1.04 & 0.51 & 0.41 & 0.06 & 0.07 & 0.34 & 0.31 & 0.39\\
+AE & CAC-LT & CS & LT Cool. & Q & 3.29 & 1.37 & 1.55 & 0.24 & 0.13 & 1.24 & 1.34 & 0.70\\
+AE & Cyl. & AE & LOC & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
+AE & Cyl. & AE & JWC & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
+AE & JWC & CS & HT Cool. & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
+AE & LOC & CS & LT Cool. & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
+AE & Cyl. & E & Losses & Q & 1.50 & 0.39 & 0.88 & 0.19 & 0.04 & 0.64 & 0.67 & 0.20\\
+AE & HRSG & S & Boiler & Q & 18.30 & 5.43 & 10.43 & 1.99 & 0.45 & 8.08 & 7.49 & 2.73\\
+AE & HRSG & E & Env. & Ph & 38.60 & 11.78 & 21.69 & 4.01 & 1.11 & 16.34 & 16.68 & 5.58\\
+AE & Cyl. & EL & Switchboard & W & 62.02 & 20.78 & 34.33 & 6.25 & 1.85 & 26.97 & 26.34 & 9.89\\
+AE & Cyl. & E & AG losses & Q & 3.47 & 0.93 & 2.03 & 0.43 & 0.09 & 1.47 & 1.53 & 0.47\\
+AE & Fuel tank & AE & Cyl. & Ch & 162.13 & 51.30 & 89.38 & 16.88 & 4.57 & 68.96 & 68.53 & 24.64\\
+AE & Fuel tank & AE & Cyl. & Q & 0.62 & 0.15 & 0.38 & 0.06 & 0.02 & 0.28 & 0.25 & 0.08\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of energy flows for the auxiliary engines}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+			\midrule
+CS & HT Cool. & CS & LT Cool. & Q & 31.84 & 3.48 & 21.94 & 3.57 & 2.85 & 12.29 & 13.62 & 5.93\\
+CS & HT Cool. & S & HRHT & Q & 16.33 & 2.39 & 11.60 & 1.59 & 0.75 & 7.15 & 7.01 & 2.17\\
+S & Boiler & S & HRHT & Q & 31.06 & 14.15 & 14.50 & 2.15 & 0.26 & 18.16 & 11.25 & 1.66\\
+S & Aux Boiler & S & Boiler & Ph & 16.22 & 13.56 & 2.32 & 0.43 & 0.09 & 10.50 & 5.42 & 0.30\\
+S & Aux Boiler & E & Env. & Ph & 4.05 & 2.83 & 1.00 & 0.22 & 0.00 & 2.62 & 1.35 & 0.07\\
+S & Fuel Tank & S & Aux Boiler & Ch. & 20.32 & 14.19 & 5.04 & 1.08 & 0.01 & 13.13 & 6.81 & 0.37\\
+S & HRHT & D & Preheater & Q & 34.51 & 12.33 & 19.12 & 2.56 & 0.50 & 21.02 & 13.49 & 0.00\\
+S & HRHT & D & Reheater & Q & 1.83 & 0.60 & 0.86 & 0.19 & 0.18 & 0.00 & 0.00 & 1.83\\
+S & HRHT & D & HWH & Q & 11.05 & 3.60 & 6.12 & 1.00 & 0.33 & 4.29 & 4.76 & 2.00\\
+S & Boiler & D & MSH & Q & 1.99 & 0.69 & 1.07 & 0.17 & 0.06 & 0.81 & 0.90 & 0.28\\
+S & Boiler & D & HFO TH & Q & 1.92 & 0.67 & 1.03 & 0.16 & 0.05 & 0.79 & 0.86 & 0.27 \\ 
+S & Boiler & D & HFO H & Q & 1.29 & 0.16 & 0.90 & 0.13 & 0.10 & 0.52 & 0.55 & 0.21\\
+S & Boiler & D & TH & Q & 1.49 & 0.52 & 0.80 & 0.13 & 0.04 & 0.61 & 0.67 & 0.21\\
+S & Boiler & D & G & Q & 7.11 & 1.96 & 4.33 & 0.64 & 0.17 & 2.97 & 3.19 & 0.95\\
+S & Boiler & D & OT & Q & 1.49 & 0.52 & 0.80 & 0.13 & 0.04 & 0.61 & 0.67 & 0.21\\
+EL & Switchboard & EL & Thrusters & W & 0.94 & 0.00 & 0.00 & 0.94 & 0.00 & 0.36 & 0.43 & 0.15\\
+EL & Switchboard & EL & HVAC & W & 1.79 & 0.59 & 0.85 & 0.18 & 0.18 & 0.00 & 0.00 & 1.79\\
+EL & Switchboard & EL & Other el. dem. & W & 59.29 & 19.99 & 32.66 & 4.99 & 1.64 & 26.06 & 25.45 & 7.78\\
+P & Prop. sh. shaft & P & Prop. sh. & W & 104.19 & 0.14 & 82.62 & 9.84 & 11.59 & 40.98 & 44.55 & 18.66\\
+P & Prop. sh. shaft & E & Losses & Q & 3.20 & 0.00 & 2.54 & 0.30 & 0.36 & 1.26 & 1.37 & 0.57\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of energy flows for the boiler/steam s}
+\end{table}
+
+\clearpage
+
+\subsection{Exergy flows}   \label{sec:appendix:exergy}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
 \midrule
@@
 ME & HRSG & E & Env. & Ph & 8.69 & 0.006 & 6.999 & 0.725 & 0.961 & 3.590 & 3.643 & 1.456\\
-ME & Cyl. & P & Prop. shaft & W & 107.39 & 0.141 & 85.158 & 10.141 & 11.947 & 42.238 & 45.916 & 19.233\\
+ME & Cyl. & P & Prop. sh. & W & 107.39 & 0.141 & 85.158 & 10.141 & 11.947 & 42.238 & 45.916 & 19.233\\
 ME & Fuel tank & ME & Cyl. & Ch & 266.02 & 0.382 & 212.237 & 25.450 & 27.949 & 105.146 & 114.197 & 46.675\\
 ME & Fuel tank & ME & Cyl. & Q & 0.15 & 0.000 & 0.125 & 0.014 & 0.014 & 0.068 & 0.063 & 0.022\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of exergy flows for the main engines}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
 			\midrule
@@
 AE & Fuel tank & AE & Cyl. & Q & 0.08 & 0.019 & 0.054 & 0.009 & 0.003 & 0.041 & 0.033 & 0.010\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of exergy flows for the auxiliary engines}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
 			\midrule
@@
 S & Boiler & E & Dumped steam & Q & 0.81 & 0.00 & 0.43 & 0.13 & 0.25 & 0.05 & 0.20 & 0.55\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of exergy flows for the boiler and steam systems}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
 			\midrule
-& HVAC-PH & D & Dem. & Q & 3.41 & 1.22 & 1.89 & 0.25 & 0.05 & 2.29 & 1.11 & 0.00\\
-& HVAC-RH & D & Dem. & Q & 0.28 & 0.09 & 0.13 & 0.03 & 0.03 & 0.00 & 0.00 & 0.28\\
-& HWH  & D & Dem. & Q & 1.96 & 0.64 & 1.09 & 0.17 & 0.05 & 0.84 & 0.81 & 0.31\\
-& MSH & D & Dem. & Q & 0.18 & 0.06 & 0.10 & 0.01 & 0.00 & 0.09 & 0.07 & 0.02\\
-& HFO TH & D & Dem. & Q & 0.37 & 0.13 & 0.20 & 0.03 & 0.01 & 0.16 & 0.16 & 0.04\\
-& HFO H & D & Dem. & Q & 0.29 & 0.04 & 0.21 & 0.03 & 0.02 & 0.13 & 0.12 & 0.04\\
-& TH & D & Dem. & Q & 0.23 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
-& G & D & Dem. & Q & 1.59 & 0.44 & 0.98 & 0.14 & 0.04 & 0.71 & 0.69 & 0.19\\
-& OT & D & Dem. & Q & 0.15 & 0.05 & 0.08 & 0.01 & 0.00 & 0.07 & 0.07 & 0.02\\
+D& HVAC-PH & D & Dem. & Q & 3.41 & 1.22 & 1.89 & 0.25 & 0.05 & 2.29 & 1.11 & 0.00\\
+D& HVAC-RH & D & Dem. & Q & 0.28 & 0.09 & 0.13 & 0.03 & 0.03 & 0.00 & 0.00 & 0.28\\
+D& HWH  & D & Dem. & Q & 1.96 & 0.64 & 1.09 & 0.17 & 0.05 & 0.84 & 0.81 & 0.31\\
+D& MSH & D & Dem. & Q & 0.18 & 0.06 & 0.10 & 0.01 & 0.00 & 0.09 & 0.07 & 0.02\\
+D& HFO TH & D & Dem. & Q & 0.37 & 0.13 & 0.20 & 0.03 & 0.01 & 0.16 & 0.16 & 0.04\\
+D& HFO H & D & Dem. & Q & 0.29 & 0.04 & 0.21 & 0.03 & 0.02 & 0.13 & 0.12 & 0.04\\
+D& TH & D & Dem. & Q & 0.23 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
+D& G & D & Dem. & Q & 1.59 & 0.44 & 0.98 & 0.14 & 0.04 & 0.71 & 0.69 & 0.19\\
+D& OT & D & Dem. & Q & 0.15 & 0.05 & 0.08 & 0.01 & 0.00 & 0.07 & 0.07 & 0.02\\
 \bottomrule
-	\end{tabular}
+\end{tabular}}
+\caption{Summary of exergy flows for the heating demand}
 \end{table}
 
-\section{Full equation list}
-
-\begin{table}[htbp]
-	\centering
-	\begin{tabular}{p{4cm}l}
-		\hline
-		\multirow{2}{*}{Before compressor}		& $ T_{comp,in} = $ Measured \\
-												& $ p_{comp,in} = $ Measured \\
-		\multirow{3}{*}{After compressor}		& $ T_{comp,out} = T_{comp,in} + \frac{T_{comp,out,is} - T_{comp,in} }{\eta_{comp,is}}  $  \\
-												& $ p_{comp,out} = $ Measured \\
-												& $ \eta_{comp,is} = P_2(\frac{p_{comp,out}}{p_{comp,in}}) $ \\
-		\multirow{3}{*}{After CAC-HT stage}		& $ p_{CAC-HT,out} = p_{comp,out} $ \\
-												& $ T_{CAC-HT,out} = T_{comp,out} - \frac{\dot{Q}_{CAC,HT}}{\dot{m}_{air,cyl}} $ \\
-												& $ \dot{Q}_{CAC,HT} = P_2(\lambda) $ \\
-		\multirow{3}{*}{After CAC-LT stage}		& $ p_{CAC-LT,out} = p_{comp,out} $ \\
-												& $ T_{CAC-LT,out} = T_{CAC-HT,out} - \frac{\dot{Q}_{CAC,LT}}{\dot{m}_{air,cyl}} $ \\
-												& $ \dot{Q}_{CAC,LT} = P_2(\lambda) $ \\
-		\multirow{3}{*}{After CAC-LT stage}
-
-	\end{tabular}
-\end{table}
-
-\section{Full list of available measurements}
+\clearpage
+
+\section{Full list of available measurements}   \label{sec:appendix:measurements}
+
+\clearpage
 

### 2018-05-10 Python files/coolingsystems.py

@@ -111,18 +111,18 @@ def centralCoolingSystems(processed, CONSTANTS):
 
     # We can now calculate the flows into the LTHT merge for the ER 13
     mdot_HTHR13_tot = processed["CoolingSystems:LTHTmerge13:HTWater_ME1_out:mdot"] + processed["CoolingSystems:LTHTmerge13:HTWater_ME3_out:mdot"] + processed["CoolingSystems:LTHTmerge13:HTWater_AE1_out:mdot"] + processed["CoolingSystems:LTHTmerge13:HTWater_AE3_out:mdot"]
-    processed["CoolingSystems:LTHTmerge13:HTWater_in:mdot"] = mdot_HTHR13_tot * (
-        CONSTANTS["MainEngines"]["T_COOLING_MIX"] - processed["CoolingSystems:LTcollector13:HTWater_out:T"]) / (
-        processed["CoolingSystems:LTHTmerge13:HTWater_in:T"] - processed["CoolingSystems:LTcollector13:HTWater_out:T"])
+    #processed["CoolingSystems:LTHTmerge13:HTWater_in:mdot"] = mdot_HTHR13_tot * (
+    #    CONSTANTS["MainEngines"]["T_COOLING_MIX"] - processed["CoolingSystems:LTcollector13:HTWater_out:T"]) / (
+    #    processed["CoolingSystems:LTHTmerge13:HTWater_in:T"] - processed["CoolingSystems:LTcollector13:HTWater_out:T"])
     # If the previously calculated mass flow is NaN, we set it equal to 0
     processed.loc[np.isnan(processed["CoolingSystems:LTHTmerge13:HTWater_in:mdot"]), "CoolingSystems:LTHTmerge13:HTWater_in:mdot"] = 0
     processed["CoolingSystems:LTHTmerge13:LTWater_in:mdot"] = mdot_HTHR13_tot - processed["CoolingSystems:LTHTmerge13:HTWater_in:mdot"]
 
     # We can now calculate the flows into the LTHT merge for the ER 24
     mdot_HTHR24_tot = processed["CoolingSystems:LTHTmerge24:HTWater_ME2_out:mdot"] + processed["CoolingSystems:LTHTmerge24:HTWater_ME4_out:mdot"] + processed["CoolingSystems:LTHTmerge24:HTWater_AE2_out:mdot"] + processed["CoolingSystems:LTHTmerge24:HTWater_AE4_out:mdot"]
[…]

### 2018-05-10 Manuscript/Manuscript.tex
hjl

@@
 
-The results of the exergy analysis of the ship energy systems are shown in Figure \ref{fig:Grassmann} and in Tables REF to REF in the Appendix. 
-\begin{itemize}
-	\item Table(s) --> Exergy efficiencies
-	\begin{itemize}
-		\item General at system level
-		\item One per system (MEs, AEs, CoolingSystems, SteamSystems) 
-	\end{itemize}
-	\item Figure --> Grassman diagram
-\end{itemize}
+The results of the exergy analysis of the ship energy systems are shown in Figure \ref{fig:Grassmann} and in Tables \ref{tab:flows:ex:ME} to \ref{tab:flows:ex:D} in the Appendix.
 
@@
 It can be observed that there are large irreversibilities concentrated in some specific parts of the system that show potential for improvement.
-The individual largest contribution to exergy destruction in the ship systems is associated to combustion, both in the engines and in the boilers.
+As expected, the individual largest contribution to exergy destruction in the ship systems is associated to combustion, both in the engines and in the boilers.
 The loss of energy resulting from the conversion of chemical energy to heat contributes to more than 76\% of the total exergy destruction on board.
-Once this contribution is isolated, the largest sources of exergy destruction are located in the engine turbochargers (28.9\%), in the exhaust gas boilers (10.7\%), in the steam heater of the HTHR system (10.5\%), the HVAC preheater (10.4\%) and the sea water cooler (8.5\%).
-In all the cases above, excluding that of the turbochargers, the exergy destruction is a consequence of a combination of large exergy flows and of the mismatch between hot and cold streams of the heat exchangers.
+Focusing on the remaining parts of the system, the largest sources of exergy destruction are located in the engine turbochargers (28.9\%), in the exhaust gas boilers (10.7\%), in the steam heater of the HTHR system (10.5\%), the HVAC preheater (10.4\%) and the sea water cooler (8.5\%).
+In all the cases above, excluding the turbochargers, the exergy destruction is a consequence of a combination of large exergy flows and of the mismatch between hot and cold stream temperatures in the heat exchangers.
 In the exhaust boilers, exhaust gas are cooled in average from more than 300$^o$C to 200$^o$C, while the heat is used to generate steam at relatively low pressure (6 bar); in the steam heater, the steam is used to heat up water at 90$^o$C; in the HVAC preheater, the relatively high-temperature water from the HTHR systems is used for heating up air to around 30$^o$C.
@@
 
-The efficiency of the recovery systems is also shown in Figure \ref{fig:Bar_efficiencyWHR}, where we show the fraction of the total energy (exergy) lost by the cooling systems and the exhaust gas of the ship's engines.
+The efficiency of the recovery systems is also shown in Figure \ref{fig:Bar_efficiencyWHR}, where the fraction of the total energy (exergy) lost by the cooling systems and the exhaust gas of the ship's engines is represented.
 It can be noted that, even when only looking at the energy that can be recovered based on the existing systems (i.e.
@@
 \caption{Summary of energy flows for the main engines}
+\label{tab:flows:en:ME}
 \end{table}
@@
 \caption{Summary of energy flows for the auxiliary engines}
+\label{tab:flows:en:AE}
 \end{table}
@@
 \end{tabular}}
-\caption{Summary of energy flows for the boiler/steam s}
+\caption{Summary of energy flows for the boiler/steam systems}
+\label{tab:flows:en:S}
 \end{table}
@@
 \caption{Summary of exergy flows for the main engines}
+\label{tab:flows:ex:ME}
 \end{table}
@@
 \caption{Summary of exergy flows for the auxiliary engines}
+\label{tab:flows:ex:AE}
 \end{table}
@@
 \caption{Summary of exergy flows for the boiler and steam systems}
+\label{tab:flows:ex:S}
 \end{table}
@@
 \caption{Summary of exergy flows for the heating demand}
+\label{tab:flows:ex:D}
 \end{table}

### 2018-05-24 Manuscript/Manuscript.tex
Updates to the manuscript with clustering

@@
 \newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+% Package for drawing pictures
+\usepackage{tikz}
 
@@
 With reference to different types of technologies, case studies, and designs,the several authors showed the existance of a quite significant potential for energy saving when WHR systems are employed, ranging from around 1\% for single-pressure steam cycles applied to two-stroke engines \cite{Theotokatos2012} to more complex systems based on ORCs (up to 10\%, \cite{Hountalas2012}) or including the cooling systems as a source of waste heat (over 10\%, \cite{Dimopoulos2012}).
-The case of the installation of an ORC on board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2015}, showing a similar potential.
+The case of the installation of an ORC on board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2017}, showing a similar potential.
 
@@
 
-\emph{Exergy} may be defined as the `maximum theoretical useful work (shaft work or electrical work) as the system is brought into complete thermodynamic equilibrium with the thermodynamic environment while the system interacts with it only'~\citep{Moran1989}.
+\emph{Exergy} may be defined as the `maximum theoretical useful work (shaft work or electrical work) as the system is brought into complete thermodynamic equilibrium with the thermodynamic environment while the system interacts with it only' \cite{Tsatsaronis2007}.
 Unlike energy, exergy is not conserved but some is destroyed because of the irreversible phenomena taking place in real processes (e.g.
@@
 		\hline
-		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 &  \\ 
-		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 &  \\
-		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 &  \\ 
-		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 &  \\ 
-		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 &  \\ 
-		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 &  \\
-		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 &  \\
-		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 &  \\ 
-		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 &  \\ 
+		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 & 20.5 \\ 
+		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 & 0 \\
+		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 & 0.40 \\ 
+		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 & 1.00 \\ 
+		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 & 0.72 \\ 
+		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 & 0.68 \\ 
+		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 & 0.21 \\ 
+		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 & 348 \\
+		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 & 0.5 \\
+		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 & 9.5e6 \\ 
+		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 & 5100 \\ 
 		\hline
@@
 It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
-Based on the available measurements, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},
+Based on the available measurements, we calculated the heat exchanged in the two HTHRs according to equation \ref{eqn:HTHR2},
 
@@
 
-\subsection{Operational mode}
+\subsection{Operational mode} \label{sec:met:operationalMode}
 
@@
 
-\subsection{Determination of typical operational conditions}
-
-\section{Results (and discussion)}
+\subsection{Determination of typical operational conditions} \label{sec:met:typicalDays}
+
+Improving the performance of a ship requires a detailed analysis of the energy demands to optimize the dimensioning and operation of the technologies and processes to install on-board.
+It requires the analysis of the hourly, daily and seasonal variations of the heat/power consumption, operating (e.g.
+ship speed) and external conditions (e.g.
+air temperature).
+However, gathering and processing those data may be time-consuming because of the high number of points to assess.
+It is therefore desirable to reduce this large quantity of information while keeping the relevant details on the relations between the variables of interest.
+One way is to represent these yearly profiles in a limited number of representative (typical) periods, avoiding the repetition of similar data sets.
+Clustering consists of grouping such sets in a single group (cluster) so that the items in the same group are more like each other than to those belonging to other groups.
+Several algorithms are introduced in the literature to partition these datasets.
+The one selected in this work is the Lloyd’s (k-means) algorithm, which is a non-deterministic method computationally faster than conventional hierarchical tools.
+The approach applied by Fazlollahi et al.
+was used to select the optimum number of clusters (Nk).
+It (Nk) should be as low as possible for data handling purposes, while preserving a high accuracy of the retrieved data.
+It builds on the calculation of three criteria:
+\begin{enumerate}
+	\item the average intra-cluster distance to assess the density of each cluster, preferably small;
+	\item the average inter-cluster distance to assess the distance between each, preferably high; 
+	\item the expected square error, which is a statistical measure suggested by Pham et al.. 
+\end{enumerate}
+
+In parallel, the cluster quality was systematically evaluated by calculating the following performance indicators:
+\begin{enumerate}
+	\item the profile deviation (deviation between original and typical period profiles),$\sigma_{profile}$;
+	\item the deviation from the load duration curve of the average values of each period, $\sigma{CDC}$;
+	\item the relative error in load duration curve deviation $ELDC$;
+	\item the maximum duration load curve difference $\Delta_{MLDC}$;
+	\item the number of periods with relative errors $\Delta_{prod}$ higher than 7\%.
+\end{enumerate}
+
+This set of performance objectives is defined as a set of constraints with an upper limit that should be respected when minimizing the number of clusters, applying the e-constraints algorithm.
+The dataset is further reduced by partitioning each typical day into a set of segments using a similar approach.
+The aim of the present clustering is to identify typical periods that are appropriate for improving the thermodynamic performance of ship energy systems.
+It can be achieved through either reducing the external energy demands (better energy management), e.g.
+with storage systems or internal recovery, or through implementing new technologies that result in higher fuel-to-demand efficiencies (better energy conversion).
+Hence, the attributes that are the primary focus of this study are:
+\begin{enumerate}
+	\item the total power consumption, including the mechanical and electrical loads;
+	\item the total heating demand, further divided into the low- and high-temperature needs, which is related to the external temperature; 
+	\item the total exergy destruction on-board, which quantifies the thermodynamic performance of the ship energy systems.
+\end{enumerate}
+A one-year data with 35,040 time steps (sampling of 15 minutes) was selected and divided into typical periods based on these three attributes. 
+The clustering approach proposed by Fazlollahi et al.
+\cite{Fazlollah2014} provides as output typical days having the same data frequency as the original time series (in this case, 15 minutes).
+From a computational perspective, however, this might not necessarily be the most efficient sampling.
+Operating conditions within a certain cluster (typical day) are often nearly constant for a longer time, suggesting that a variable time resolution would lead to a lower dimensionality of the data while retaining most of the original information.
+To this purpose, in this paper we employ the Adaptive Piecewise Constant Approximation (APCA) approach proposed by Keogh et al. \cite{Keogh2001}
+
+\section{Results and discussion} \label{sec:results}
 
@@
 	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/Sankey_diagram_energy_ship_v9}
+	\includegraphics[width=0.95\linewidth]{Figures/Sankey_diagram_energy_ship_v9}
 	\caption{Sankey diagram}
@@
 	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/Grassmann_diagram_v5}
+	\includegraphics[width=0.95\linewidth]{Figures/Grassmann_diagram_v5}
 	\caption{Grassmann diagram}
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.99\linewidth]{Figures/Bar_WHRefficiency}
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Bar_WHRefficiency}
 	\caption{Energy and exergy efficiency of the WHR systems}
@@
 \subsection{Typical operational days}  \label{sec:res:typicalDays}
+
+A data clustering was applied at first on the total power and heating demands along the operation year, using 1000 starting points generated randomly for cluster initialization.
+As presented in Section \ref{sec:met:typicalDays}, the number of typical days depends on the values of the intra- and inter-cluster distances and on the expected squared error, obeying five constraints on the load duration profiles.
+A higher number of typical days would result in smaller data losses, but the calculation of the ESE indicator shows that this gain of information is negligible.
+The average power demand is about 4700 kW with a standard deviation of 3000 kW and a 95\% percentile of 8300 kW.
+These figures denote a high variability, but the number of optimum clusters is only 2, which shows that the power demand follows repeatable trends over time.
+The first typical day corresponds to more than 32000 hours (low to high demands, up to 12000 kW), and the second one to \~{}3000 hours (peak demands, above 12000 kW).
+On the contrary, the total heating demand presents variations of smaller amplitude, but a higher number of clusters is necessary.
+Five typical days appear sufficient to represent the load duration curve and energy demand profiles.
+The cases with the greatest power consumption and different trends are already considered in these five clusters, and the addition of an extreme period is not necessary.
+The large number of typical days (5) compared to those required (2) when clustering only the power demand (2) illustrates the lack of direct correlation between the power and heating requirements.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.95\linewidth]{Figures/ld_multyday_at1.png}
+	\caption{Power profile and duration curves, represented with the 35,040 datapoints (black curve) and five typical days (other colors).}
+	\label{fig:res:powerProfile}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[width=0.95\linewidth]{Figures/ld_multyday_at2.png}
+	\caption{Power profile and duration curves, represented with the 35,040 datapoints (black curve) and five typical days (other colors).}
+	\label{fig:res:heatProfile}
+\end{figure}
+
+The quality of the clustering was assessed by calculating the performance indicators for both attributes (Table \ref{tab:res:clusteringKPI}).
+The suggested typical periods present low relative errors and are slightly better for characterization of the heat demand profiles.
+\begin{table}
+	\centering
+	\begin{tabular}{crr}
+		\toprule 
+								& $P_{el}+P_{mech}$ & $Q_{tot}$ \\
+		\midrule
+		$\sigma_{profile}$		&  0.65				&  0.10		\\ 
+		$\sigma_{CDC}$ 			&  0.11				&  0.11		\\ 
+		$ELDC$ 					&  0.23				&  0.07		\\ 
+		$\Delta_{MLDC}$			&  0.18				&  0.06		\\ 
+		$\delta_{prod} (7\%)$	&  155.00			&  168.00	\\ 
+		\bottomrule 
+		\end{tabular} 
+		\caption{Performance parameters for the 5-cluster power and heat}
+		\label{tab:res:clusteringKPI}
+\end{table}
+
+The five typical days are further segmented and the closeness of the plots for both demands illustrates the quality of the segmentation.
+In most typical days, the power consumption varies in a 2000-8000 kW range, with an average consumption of about 4500-5000 kW and similar trends.
+The maximum values, of about 19000 kW, are reached at highest ship speeds.
+
+\begin{figure}[htbp!]
+	\centering
+	\includegraphics[width=0.95\linewidth]{Figures/segmentation_power_heat.png}
+	\caption{Power (left) and heat (right) profiles, original (dotted) and segmented (full), for the 5 typical days and duration curves.}
+	\label{fig:res:powerProfileSegmented}
+\end{figure}
+
+The same findings can be deduced from the segmentation of the heat profiles as all present similar tendencies.
+The only significant difference is the range between the minimum and maximum values reached in a single day (e.g.
+2000-2800 kW, 1000-2000 kW, 700-1700 kW, etc.).
+These differences are correlated to changes of the outer temperature and are thus seasonal variations.
+Heat storage from low-demand to high-demand days is not feasible, but may be implemented, if responsive enough, on single days.
+
+A data clustering was applied on the total power demand and exergy destruction, following the same approach as above.
+In this case, the optimum number of clusters resulted to be only 2, which shows the direct relation between the power consumption and exergy destruction, and, consequently, the weak relation with the heating needs.
+This result is as expected, since the engines and subsequent components are responsible for the largest irreversibilities in the system.
+High accuracy of the clustering for even such a small number of typical days is reached, as shown, for instance, with the relative load differences ($\Delta_{MLDC}$) are 0.18 for both power and exergy.
+A further segmentation of these two typical days shows that these two variables follow the same trend in both regular and extreme conditions.
+The data clustering shows that, for the energy system of the case study ship, it is critical to reduce the power demand or to improve the power generation system for enhancing the system performance.
+Such findings are valid for all types of operation days.
+
+\begin{figure}[htbp!]
+	\centering
+	\includegraphics[width=0.95\linewidth]{Figures/segmentation_power_exergy.png}
+	\caption{Exergy destruction profile and duration curve (left), represented with the 35,040 datapoints (black curve) and two typical days (dark and clear blue).
+Segmented power and exergy destruction profiles for the two typical days (right).
+}
+	\label{fig:res:exergyDestructionProfile}
+\end{figure}
 
@@
 The availability of waste heat that is not already recovered by the system suggests that the system could benefit from the installation of a Heat-to-power WHR.
-This possibility was tested by the authors both in the case that no additional retrofit is performed \cite{Ahlgren2016,Mondejar2017}, with estimated savings of 22\% of the auxiliary power demand for a reference voyge, and in the case of a full integration with the rest of the system \cite{Baldi2017}, with estimated savings of up to 6\% of the total fuel consumption for a year of ship operations.
+This possibility was tested by the authors both in the case that no additional retrofit is performed \cite{Ahlgren2016,Mondejar2017}, with estimated savings of 22\% of the auxiliary power demand for a reference voyage, and in the case of a full integration with the rest of the system \cite{Baldi2017}, with estimated savings of up to 6\% of the total fuel consumption for a year of ship operations.
 
@@
 
-\section{Conclusion}
-\label{sec:conclusion}
+\section{Conclusion} \label{sec:conclusion}
 

### 2018-05-24 Manuscript/Manuscript.tex
Updated with the conclusion

@@
 \section{Conclusion} \label{sec:conclusion}
+
+% Summary of the background
+Shipping, similarly to any other industry, is facing challenges in relation to its contribution to climate change.
+In this context, cruise ships are particularly under scrutiny because of their large impact on the environment, and on their direct connection to their final users.
+
+% What we did in this paper
+In this paper, we analyzed the energy system of a cruise ship operating in the Baltic Sea, with the aim of providing a better understanding of how energy is used on board, for what purposes, and with what efficiency.
+Although these numbers can vary substantially depending on the ship type, size and operating region, the results presented in this paper represent a unique contribution to scientific literature in terms of the detail of the analysis and of the extent of the data available.
+
+% Conclusion: Energy demand share
+The results showed that propulsion represents the largest share of the energy demand, as in most ships, but it is less dominant compared to other ship types, as it represents 45\% of the yearly energy demand.
+Auxiliary electric power and heat represent an approximately equal share of the remaining part (27\% and 28\% respectively), with the largest share of the latter being related to space heating.
+This is a consequence of the ship being operated in the North of Europe, and results are expected to vary substantially for a ship operating in warmer climates (Mediterranean or Caribbeab Sea, for instance).
+
+% Conclusion: Main sources of exergy destruction
+The application of exergy analysis allowed the identification of the main causes for losses of energy quality.
+In addition to the combustion processes in the main engines and in the boilers, that together contribute to 77\% of the exergy destruction on board, the main potential for improvement was identified in the turbochargers, the exhaust gas boilers (HRSGs), the steam heater in the hot water heat recovery systems, in the HVAC pre-heater and in the sea water cooler.
+The main basic causes of exergy destruction are hence to be considered the low-load operations of the engines and the mismatch in the operating temperatures of the heat exchangers.
+
+% Conclusion: High system efficiency
+The existing ship systems proved to have a high efficiency, at least in relation to other ships of different types.
+From the results of the energy and exergy analysis it can be concluded that the existing systems already make use of a large share of the waste heat, including that at relatively low temperature that is transferred to the HT cooling systems.
+
+% Conclusion: Potential for improvement 
+Despite its good efficiency, the system shows potential for improvement.
+The fact that both main and auxiliary engines are operated at low load can be compensated with a higher level of electrification of the system, and by the use of batteries.
+The installation of an ORC can be a solution to use the high-temperature heat of the exhaust gas of the Diesel engines, while low-temperature heat from the Diesel engines could be used, instead of wasted to the sea water cooler, for space heating.
 

### 2018-06-05 Manuscript/Figures/dataProcessing.tex
Updates after meeting - Part 2

\documentclass{standalone}

\usepackage{tikz}
\usepackage{pgfplots}

\usetikzlibrary{arrows,shadows,positioning,shapes.geometric}
\definecolor{colone}{RGB}{62,99,118}
\definecolor{coltwo}{RGB}{82,131,156}
\definecolor{colthree}{RGB}{139,176,195}
\definecolor{colfour}{RGB}{255,211,71}

% Add layers. Add a "background" layer
\pgfdeclarelayer{bg}    % declare background layer
\pgfsetlayers{bg,main}  % set the order of the layers (main is the standard layer)

\tikzset{
  frame/.style={
    rectangle, 
    text width=8em, text centered, text = white,
    minimum height=4em,drop shadow,fill=colone,
    rounded corners,
  },
  % INPUT BOXES
  frameInput/.style={
  	trapezium,
  	trapezium left angle=60,
  	trapezium right angle=120,
  	drop shadow,
  	text width=8em,
  	text centered,
  	text = white,
  	fill = colone
  },
% Output BOXES
frameOutput/.style={
	rectangle,
	drop shadow,
	text width=10em,
	text centered,
	text = white,
	fill = coltwo
},
  % THESIS SECTIONS
  frameSections/.style={
  rectangle,
  text width = 16em,
  text = black,
  fill = white,
  font = \sffamily\huge
  },
  % PROCESSES 
  frameProcesses/.style={
  rectangle,
  drop shadow,
  text width = 10em,
  text = white,
  fill = colthree,
  text centered
  },
  % VARIABLE BOXES
  frameVar/.style={
  rectangle,
  text = blue!80,
  text width 8em,
  fill = white,
  text centered,
  rounded corners
  },
  % MAIN DATA FLOWS
  lineData/.style={
  draw=colfour,rounded corners=2mm,-latex
  },
  lineInfo/.style={
	draw=colfour,dashed,rounded corners=2mm,-latex
}    
}

\begin{document}

\begin{tikzpicture}[font=\large\sffamily\bfseries,very thick,node distance = 4cm]

% Input nodes for Ship 1
\node [frameInput] (input_data) at (0,-3) {Measurements \\ from ship \\ operations} ;
\node [frameInput] (input_tech) at (0,-5) {Ship technical \\ documentation} ;

% Entries on the left
\node [frameSections] (section1) at (0,3) {Data collection} ;
\node [frameSections] (section2) at (14,2) {Data processing} ;
\node [frameSections] (section3) at (31,2) {Analysis} ;

% Processes for ship 1
\node [frameProcesses] (process_filtering) at (6,-3) {Data \\ pre-treatment \\ and filtering} ;

\node [frameProcesses] (process_ME) at (12,-1) {Main engines \\ analysis} ;
\node [frameProcesses] (process_AE) at (12,-5) {Aux engines \\ analysis} ;
\node [frameProcesses] (process_CS) at (14.5,-3) {Cooling systems \\ analysis} ;
\node [frameProcesses] (process_HTHR) at (14.5,-9) {HTHR \\  systems} ;

\node [frameProcesses] (process_Q) at (20,-8) {Heat demand \\ estimation} ;
\node [frameProcesses] (process_Paux) at (20,-4.5) {Auxiliary power \\ demand \\ estimation} ;

\node [frameOutput] (output_preliminary) at (28,0) {Preliminary \\ analyis} ;
\node [frameOutput] (output_energy) at (28,-3) {Energy \\ analysis} ;
\node [frameOutput] (output_exergy) at (28,-6) {Exergy \\ analysis} ;

% Lines for ship 1
\path [lineData] (input_data.east) -- (process_filtering.west) ;
\path [lineInfo] (input_tech.east) -| (process_filtering.south) ;

\path [lineData] (process_filtering.east) --  (process_CS.west) ;
\path [lineData] (process_filtering.east) -| (9,-2) |- (process_ME.west) ;
\path [lineData] (process_filtering.east) -| (9,-4) |- (process_AE.west) ;

\path [lineData] (process_CS.east) -| (17,-4) |- (15,-7) -| (12,-8.5) |- (process_HTHR.west) ;

\path [lineData] (process_AE.east) -| (14.5,-5) |- (process_Paux.west) ;
\path [lineData] (process_ME.east) -| (process_CS.north) ;
\path [lineData] (process_AE.east) -| (process_CS.south) ;
\path [lineData] (process_HTHR.east) -| (17,-8.5) |- (process_Q.west) ;

\path [lineData] (process_ME.east) -| (22.5,-2) |- (output_energy.west) ;
\path [lineData] (process_Paux.east) -| (22.5,-4) |- (output_energy.west) ;
\path [lineData] (process_Q.east) -| (22.5,-4) |- (output_energy.west) ;
\path [lineData] (process_CS.east) --  (output_energy.west) ;

\path [lineData] (24,-3) |- (output_exergy.west) ;
\path [lineData] (24,-3) |- (output_preliminary.west) ;

% LINE INFO
\path [lineInfo] (11,-6) -| (process_AE.south) ;
\path [lineInfo] (11,-2) -| (process_ME.south) ;
\path [lineInfo] (13.5,-10) -| (process_HTHR.south) ;
\path [lineInfo] (19,-9) -| (process_Q.south) ;
\path [lineInfo] (19,-6) -| (process_Paux.south) ;

% VARIABLES
\node [frameVar] (aa1) at (25,-2.5) {$\dot{W}, \dot{Q}$} ;
\node [frameVar] (aa1) at (25,-5.5) {$\dot{B}$} ;
% \node [frameVar] (aa1) at (25,-6.5) {$\dot{m}, T$} ;

% Papers
%\node [framePapers] (paper1) at (8,-9) {Paper I} ;
%\node [framePapers] (paper2) at (6,-14) {Paper III} ;
%\node [framePapers] (paper3) at (13.3,-14) {Paper IV} ;
%\node [framePapers] (paper4) at (13.3,-15) {Paper V} ;

\end{tikzpicture}

\end{document}

### 2018-06-05 Manuscript/Manuscript.tex
Author update
Updates after meeting - Part 2

@@
 \author[DTU]{Tuong-Van Nguyen}
-\author[CTU]{Cecilia Gabrielii}
+\author[LTU]{Marcus Thern}
 \author[CTU]{Karin Andersson}
@@
 \address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
+\address[LTU]{Energy Sciences, Lund University, Lund SE-22100, Sweden}
 \address[CTU]{Department of Mechanics and Maritime Sciences, Chalmers University of technology, Gothenburg, Sweden}
@@
 
-The cruise industry have also received much attention for its environmental impact other than in terms of GHG emissions.
+The cruise industry has also received much attention for its environmental impact other than in terms of GHG emissions.
 Cruise ships were shown to have a remarkable negative impact on both air \cite{Poplawski2011} and water \cite{Butt2007} quality, and were also pointed as sources of environmental damage from a wider perspective \cite{Brida2009}.
@@
 
-Altogether, these conditions present a challenge to the shipping companies who attempt to reduce their fuel consumption, environmental impact, and operative costs.
-A wide range of fuel saving solutions for shipping are available and partially implemented in the existing fleet, both from the design and operational perspective; several specific studies have been conducted on these technologies, and a more detailed treatise would be out of the scope of this work.
+Altogether, these conditions present a challenge to the shipping companies that attempt to reduce their fuel consumption, environmental impact, and operative costs.
+A wide range of fuel saving solutions for shipping are available and partially implemented in the existing fleet, both from the design and operational perspectives; several specific studies have been conducted on these technologies, and a more detailed treatise would be out of the scope of this work.
 In this context, it has been acknowledged that the world fleet is heterogeneous, and measures need to be evaluated on a ship-to-ship basis \cite{Bouman2017}.
@@
 Shi et al.
-proposed a modeling approach for predicting ship fuel consumption of a cargo/passenger ferry \cite{Shi2009}; Theotokoats and Tzelepis applied a similar procedure to the case of a Handymax product carrier \cite{Theotokatos2015}, similarly to Tillig et al., who also added the dynamic element to their predictive model \cite{Tillig2016}.
+\cite{Shi2009} proposed a modeling approach for predicting ship fuel consumption of a cargo/passenger ferry; Theotokoats and Tzelepis \cite{Theotokatos2015} applied a similar procedure to the case of a Handymax product carrier, similarly to Tillig et al.
+\cite{Tillig2016}, who also added the dynamic element to their predictive model.
 
@@
 More generally, several authors have highlighted the importance of a detailed knowledge of the ship's operational profile in order to appropriately assess and optimize possible alternatives for improving ship energy efficiency \cite{Banks2013}.
-Coraddu et al., started from a statistical analysis of measured ship operations and used the aggregated data, together with a computational model of the vessel, to provide a better prediction of the actual ship's operational efficiency \cite{Coraddu2014}.
+Coraddu et al.
+\cite{Coraddu2014} started from a statistical analysis of measured ship operations and used the aggregated data, together with a computational model of the vessel, to provide a better prediction of the actual ship's operational efficiency .
 The importance of considering the operational profile was also showed in the case of the optimization of engine-propeller interaction \cite{Baldi2015c}, in the process of retrofitting existing systems \cite{Baldi2015b,Choi2013}, and in ship design \cite{Ghassemi2017,Solem2015}.
@@
 This is due to a combination of generally low demand and high availability from the waste heat of the engines.
-In a previous publications by the authors, for instance, it is shown that in the case of a product tanker, although the heat demand was estimated to account for roughly 20\% of the total energy demand of the ship on a yearly basis, it only contributes to 4.1\% of the total fuel consumption (contribution of the auxiliary boilers), while the rest of the demand is fulfilled using waste heat \cite{Baldi2015a}.
+In a previous publications by the authors \cite{Baldi2015a}, for instance, it is shown that in the case of a product tanker, although the heat demand was estimated to account for roughly 20\% of the total energy demand of the ship on a yearly basis, it only contributes to 4.1\% of the total fuel consumption (contribution of the auxiliary boilers), while the rest of the demand is fulfilled using waste heat.
 
@@
 The potential uses for waste heat on board are not limited to improving the efficiency of the power plant.
-Waste heat is commonly used for fulfilling on board heat demand for spatial heating and freshwater generation \cite{Molland2011,Baldi2015a,Mccarthy1990};  Balaji et al.
-proposed the use of waste heat for ballast-water systems \cite{Balaji2012}; Salmi et al.
-suggested its use for adsorption refrigeration systems \cite{Salmi2017}.
+Waste heat is commonly used for fulfilling on board heat demand for spatial heating and freshwater generation \cite{Baldi2015a,Molland2011,Mccarthy1990};  Balaji et al.
+\cite{Balaji2012} proposed the use of waste heat for ballast-water systems ; Salmi et al.
+\cite{Salmi2017} suggested its use for adsorption refrigeration systems.
 A detailed review of potential uses for waste heat from marine engines is presented by Shu et al \cite{Shu2013}.
@@
 Dimopoulos et al.
-showed how the process optimizing the WHR system of a container ship can be more efficient if exergy efficiency, rather than energy efficiency, is used as the target of the optimization \cite{Dimopoulos2012}; Baldi et al.
-also analyzed the exergy flows on board of a product tanker, showing that this allows having a more accurate understanding of what parts of the system show potential for improvement \cite{Baldi2015a}; similar results were obtained by Marty et al., who focused on the power plant of a cruise ship \cite{Marty2016}.
+\cite{Dimopoulos2012} showed how the process optimizing the WHR system of a container ship can be more efficient if exergy efficiency, rather than energy efficiency, is used as the target of the optimization; Baldi et al.
+\cite{Baldi2015a} also analyzed the exergy flows on board of a product tanker, showing that this allows having a more accurate understanding of what parts of the system show potential for improvement ; similar results were obtained by Marty et al.
+\cite{Marty2016}, who focused on the power plant of a cruise ship.
 Koroglu et al.
-made a step futher also including advanced exergy analysis in their study \cite{Koroglu2017}.
+\cite{Koroglu2017} made a step futher also including advanced exergy analysis in their study.
 
@@
 	\item Analyse the current efficiency of the system, and potential ways to improve it, by means of applying exergy analysis.
-	\item Provide reference operational conditions for further use in the field of energy systems optimization.
+	\item Provide reference operational conditions and typical operational days for further use in the field of energy systems optimization.
 \end{itemize}
@@
 
-In this paper, we present the application of energy and exergy analysis (described in section \ref{sec:met:energyExergy}) to a cruise ship.
+In this paper, we present the application of energy and exergy analysis (described in Section \ref{sec:met:energyExergy}) to a cruise ship.
 This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; the specific assumptions and methods used for processing the available information into  energy and exergy flows are specified in Section \ref{sec:met:processing}.
-In these regards, the estimation of the heat demand is treated in a separate section (Sec.
-\ref{sec:met:heat}) as it constitutes a particularly challenging task and it is one of the central aspects of this article.
+In these regards, the estimation of the heat demand is treated in a separate section (Section \ref{sec:met:heat}) as it constitutes a particularly challenging task and is one of the central aspects of this article.
 
@@
 \emph{Energy} can be stored, transformed from one form to another (e.g.
-heat to power) and transferred between systems, but can neither be created nor destroyed (conservation law)~\textbf{Bejan et al.}.
+heat to power) and transferred between systems, but can neither be created nor destroyed (conservation law) \cite{Bejan1996}.
 The system under study is the \emph{ship energy system} and is thus taken as control volume.
@@
 The energy flow associated with a material stream is calculated as the sum of the physical and chemical enthalpies, and kinetic and potential energies are neglected.
-The physical energy is taken as the relative enthalpy, as underlined in~\textbf{Kotas et al.}, and the chemical energy is taken as the lower/higher heating value.
-The environmental conditions taken for the present analysis are the ambient pressure and seawater temperature.
+The physical energy is taken as the relative enthalpy, as underlined in \cite{Kotas2013}, and the chemical energy is taken as the lower/higher heating value.
+The environmental conditions taken for the present analysis are the ambient pressure (1.01 bar) and seawater temperature (measured) .
 
@@
 Kinetic and potential exergies are neglected, and the exergy of a given material stream is derived as the sum of the physical and chemical exergies.
-The chemical exergy is calculated based on the reference environment of \textbf{Szargut (1989)}, and its value is approximatively equal to the higher heating value for hydrocarbon fuels.
-The exergy destruction can be calculated from the Gouy-Stodola theorem~\textbf:{Kotas (1995)}.
+The chemical exergy is calculated based on the reference environment of \cite{Szargut1989}, and its value is approximatively equal to the higher heating value for hydrocarbon fuels.
+The exergy destruction can be calculated from the Gouy-Stodola theorem \cite{Kotas2013}.
 
@@
 	\begin{equation} \varepsilon = \frac{\dot{E}_{\mathrm{p}}}{\dot{E}_{\mathrm{f}}} \end{equation}
-	\item the efficiency defect $\lambda$, presented in the work of Kotas, defined as the fraction of the total exergy input destroyed in the successive irreversible processes
+	\item the efficiency defect $\lambda$, presented in the work of Kotas \cite{Kotas2013}, defined as the fraction of the total exergy input destroyed in the successive irreversible processes
 	\begin{equation} \lambda = \frac{\dot{E}_{\mathrm{d}}}{\dot{E}_{\mathrm{in,tot}}} \end{equation}
-	\item  the irreversibility share $\delta$, suggested in the work of Tsatsaronis, defined as the ratio between the exergy destroyed in the $i$-th component in relation to the exergy destroyed in the entire system
+	\item  the irreversibility share $\delta$, suggested in the work of Tsatsaronis \cite{Tsatsaronis1993}, defined as the ratio between the exergy destroyed in the $i$-th component in relation to the exergy destroyed in the entire system
 	\begin{equation} \delta_i = \frac{\dot{E}_{\mathrm{d},i}}{\dot{E}_{\mathrm{d,tot}}} \end{equation}
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{Figures/ShipLayout}
+	\includegraphics[width=0.95\linewidth]{Figures/SystemGeneral}
 	\caption{Layout of the case study energy systems}
@@
 The operational data was collected on board from the ships' machine logging and surveillance system.
-The on board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB.
-The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average.
-The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python \cite{mckinney-proc-scipy-2010}.
-A new structured naming of headers and a consistent time frequency over all data points were created, and 245 selected data points were saved in a HDF5 table time series database which is the base for the analysis.
-
-All data-points were checked individually by creating a descriptive statistic and a histogram of each data-point.
+The data sampling rate was selected to 15 minutes, automatically calculated as averages over the respective periods by the on board data logging systems, as an acceptable compromise between accuracy and computational time required for the data processing.
+Data consistency was checked individually by creating a descriptive statistic and a histogram of each measurement.
 A filter was applied by setting up a pre-defined maximum and minimum value.
-As we do not know, or had any meaningful way of deriving, each individual sensors placement, sensor type or calibration status all data must be filtered and checked accordingly for outliers or bad data.
-
-Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska H\"{o}garna for the ambient temperature.
-The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska H\"{o}garna lighthouse is along the ship route, in between the Swedish archipelago and Åland.
-The assumption was validated based on  June-December period, for which onboard measurements were available.
-This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
-The fit the SMHI-data with the rest of the data-set the SMHI data was resampled from 1h for the seawater and 3h for the ambient temperature, to 15min frequency using a linear interpolation.
+
+Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska H\"{o}garna for the ambient temperature, both located along the route of the vessel.
 
@@
 
+The main principles applied to the modeling are presented in the following sections and summarized in Figure \ref{fig:method}, while a more thorough description and validation of the different approaches is presented in the Appendix.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.95\linewidth]{Figures/dataProcessing}
+	\caption{Overview of the data processing from inputs to outputs}
+	\label{fig:dataprocessing}
+\end{figure}
+
 \subsubsection{Diesel engines}
 
-The setup of the engines and their connections to the other ship systems are summarized in Figure \ref{fig:DieselEngines} for the main and auxiliary engines, respectively.
-The most relevant operative values are provided in Tables \ref{tab:ME_values} and \ref{tab:AE_values} and the main energy flows are listed in Tables \ref{tab:ME_flows} and \ref{tab:AE_flows}.
-It should be noted that, in the case of the main engines, the engine power output was not measured and needed to be estimated.
-In this work, we calculated the engine power output based on measurements of engine fuel rack position (used as a proxy of the mass of fuel injected per cycle) and of the engine speed:
+The Diesel engines are the only power source of the system, subdivided in four main engines uniquely devoted to propulsion and four auxiliary engines used for on board electric power generation.
+The energy and exergy analysis of the engines is not limited to the boundary of the system, but is brought to the detail of the different components that constitute the engine systems, as shown in Figure \ref{fig:DieselEngines}.
+
+In the case of the main engines, the engine power output was not measured and needed to be estimated.
+In this work, we calculated the engine power output based on measurements of engine fuel rack position (used as a proxy of the mass of fuel injected per cycle) and of the engine speed (see equations \ref{eqn:ME:eq1} to \ref{eqn:ME:eq3}).
+In the case of the auxiliary engines, instead, direct measurements of the power generated by each engine were available.
 \begin{eqnarray}
-\dot{m}_{fuel} & = & \dot{m}_{fuel,des} \left(a_0 + a_1 \frac{frp}{frp_{des}}\right) \frac{\omega}{\omega_{des}} \\
-\eta_{ME} & = & a_0 + a_1 \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} + a_2 \left( \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} \right)^2 \\
-\dot{W}_{ME} & = & \dot{m}_{fuel} \eta_{ME} LHV
+\dot{m}_{fuel} & = & \dot{m}_{fuel,des} \left(a_0 + a_1 \frac{frp}{frp_{des}}\right) \frac{\omega}{\omega_{des}} \label{eqn:ME:eq1} \\
+\eta_{ME} & = & a_0 + a_1 \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} + a_2 \left( \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} \right)^2 \label{eqn:ME:eq2} \\
+\dot{W}_{ME} & = & \dot{m}_{fuel} \eta_{ME} LHV \label{eqn:ME:eq3}
 \end{eqnarray}
-
-The validity of this assumption can be seen by observing Figures \ref{fig:Pme_vs_vship_Neng} and \ref{fig:Pme_vs_TCspeed} where the calculated  power from the main engines is represented against the ship speed and the turbocharger speed, respectively.
 
@@
 
-In the case of the auxiliary engines, instead, direct measurements of the power generated by each engine were available.
-
 For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine.
-\begin{align}
-	\dot{m}_{air,comp} & =  \dot{m}_{air,cyl} + \dot{m}_{air,bp} \label{eq:mb:comp} \\ 
-	\dot{m}_{eg,turb} & = \dot{m}_{eg,cyl} + \dot{m}_{air,bp} \label{eq:mb:turb} \\
-	\dot{m}_{eg,cyl} & = \dot{m}_{air,cyl} + \dot{m}_{fuel,cyl} \label{eq:mb:cyl} \\
-	\dot{m}_{air,comp} \Delta h_{comp} & = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech} \label{eq:eb:turbocharger} \\
-	\dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0) & = \dot{m}_{air,bp} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) \label{eq:eb:bpmerge}
-\end{align}
-Where equations \ref{eq:mb:comp} to \ref{eq:mb:cyl} represent the mass balances of the bypass split, bypass merge, and cylinder respectively, while equations \ref{eq:eb:turbocharger} and \ref{eq:eb:bpmerge} represent the energy balances of the turbocharger and of the bypass merge, respectively.
-
-The energy balance over the whole engine is presented in Equation \ref{eqn:engineEnergyBalance}.
-\begin{equation}
-\dot{Q}_{fuel} + \dot{Q}_{air,in} = \dot{W}_{mech} + \dot{Q}_{eg} + \dot{Q}_{cooling}  \label{eqn:engineEnergyBalance} \\
-\end{equation}
-With: 
-\begin{align}
-\dot{Q}_{fuel} & =\dot{m}_{fuel} (LHV_{fuel} + c_{p,fuel} (T_{fuel,in} - T_0))\\
-\dot{Q}_{air,in} & = \dot{m}_{air} c_{p,air} (T_{air,comp,in} - T_0) \\ 
-\dot{Q}_{eg} & = \dot{m}_{eg} c_{p,eg} (T_{eg,TC,out} - T_0) \\
-\dot{Q}_{cooling} & = \dot{Q}_{CAC,HT} + \dot{Q}_{CAC,LT} + \dot{Q}_{JWC} + \dot{Q}_{LOC} \label{eq:eb:cooling}
-\end{align}
-
-The value of $\dot{Q}_{cooling}$ is determined based on the balance in Equation \ref{eqn:engineEnergyBalance}.
-The share between the four contributions in Equation \ref{eq:eb:cooling} is calculated based on the data for the heat loss to the HT and LT systems presented in the engines project guides as functions of the engine load.
-The values are interpolated based on a 2$^nd$ degree polynomial and scaled to respect the energy balance.
-In absence of more accurate data, it is then assumed that the contributions of $\dot{Q}_{JWC}$ and $\dot{Q}_{LOC}$ are equal for all loads.
-
 Values for temperatures, mass flows and energy flows for different engine loads are provided in the \ref{sec:appendix:engines}, for both main and auxiliary engines.
 
-\subsubsection{Cooling systems}
-
-The ship's cooling systems are designed similarly to most ships and divided in two cooling systems of two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C).
-The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board.
-A representation of the ship cooling systems is provided in Figure \ref{fig:generalsystemssimplified}.
-
-All values not directly measured in the cooling systems were calculated based on mass and energy balances.
-Mass flows were determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps as provided in the engine project guides.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/GeneralSystemsSimplified}
-	\caption{Layout of the ship heating and cooling system}
-	\label{fig:generalsystemssimplified}
-\end{figure}
-
 \subsection{Electric power demand} \label{sec:met:electric}
 
-The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy
+The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy.
 \begin{equation}
@@
 However, none of the individual contributions is directly measured, hence estimations of the energy demand of individual consumers need to be based on indirect measurements.
-In this paper we identified methods for the estimation of the contribution of two consumers: the HVAC systems and the bow thrusters.
-
-\subsubsection{HVAC systems}
-
-The observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a  period, corresponding to the summer months, when the electric power demand is significantly higher then during the remaining months.
-Based on this observation, in this work we assumed that the demand of the HVAC compressors related to the need for space cooling is concentrated during the summer months.
-Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a typical summer and winter day, shows that the daily variation is comparable.
-This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pel_vs_time}
-	\caption[Total electric power demand versus time (Daily average)]{Total electric power demand versus time (Daily average)}
-	\label{fig:pelvstime}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/PelRel_vs_time}
-	\caption{Instantaneous power demand over daily average, winter versus summer day}
-	\label{fig:pelrelvstime}
-\end{figure}
-
-Hence, the HVAC electric power demand was estimated as follows:
-\begin{equation}
-P_{el,HVAC} =
-\begin{cases}
-P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
-0, & \text{otherwise}
-\end{cases}
-\end{equation}
-where $P_{el,ref}(t)$ is calculated as:
-\begin{equation}
-P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
-\end{equation}
-
-\subsubsection{Thrusters}
-
-When entering port areas, the ship needs to use thrusters to maneuver and berth.
-In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.
-In order to isolate the demand from the thrusters, we extracted from ship operations a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
-From this daily profile (made of 96 points), we manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).
-
-By comparing the reference profile to the instantaneous one, the points where the former is more than 10\% higher than the latter are identified as "thruster-on" points and treated consequently.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Thruster_selection}
-	\caption{Example of the procedure of selection of thruster power demand}
-	\label{fig:thruster_selection}
-\end{figure}
-
-\subsection{Heat demand} \label{sec:met:heat}
+In this paper only the contributions of the HVAC and the bow thrusters are identified separately from the remaining parts of the system.
+More details about the methods employed in this process are provided in Appendix REF.
+
+\subsection{Heating and cooling systems} \label{sec:met:heat}
 
@@
 
+The ship's cooling systems are designed similarly to most ships and divided in two cooling systems operated at two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C).
+The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board.
+An overview of the ship heating and cooling systems is provided in Figure \ref{fig:generalsystemssimplified}.
+
+All values not directly measured in the cooling systems were calculated based on mass and energy balances.
+Mass flows were determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps as provided in the engine project guides.
 The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}.
@@
 
-\subsubsection{Heat balance parameter estimation}
-
-The individual contributions of different consumers were determined by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters.
-The parameter estimation problem is hence written as a minimization problem:
-\begin{eqnarray}
-min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
-\end{eqnarray}
-
-where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections.
-A list of the parameters $\textbf{p}$ is shown in Table \ref{tab:ParameterEstimation}, together with the chosen upper and lower boundaries for the calibration procedure.
-\begin{table}
-	\centering
-	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
-		\hline 
-		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
-		\hline
-		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 & 20.5 \\ 
-		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 & 0 \\
-		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 & 0.40 \\ 
-		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 & 1.00 \\ 
-		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 & 0.72 \\ 
-		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 & 0.68 \\ 
-		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 & 0.21 \\ 
-		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 & 348 \\
-		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 & 0.5 \\
-		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 & 9.5e6 \\ 
-		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 & 5100 \\ 
-		\hline
-	\end{tabular}
-	\caption{Parameters optimized in the parameter estimation for the heat balance}
-	\label{tab:ParameterEstimation} 
-\end{table}
-
-\subsubsection{Heat demand}
-
-The heat demand was calculated as the sum of the contributions of the elements listed in the ship's heat balance documentation.
-The assumptions used in the estimation are summarized in Table \ref{tab:HeatDemand}:
-\begin{table}
-	\centering
-	{\tablinesep=2ex\tabcolsep=10pt
-		\begin{tabular}{p{2.8cm}l}
-			\hline 
-			Heat flow name & Equation \\
-			\hline
-			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
-			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
-			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
-			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
-			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
-			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
-			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
-			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
-			\hline
-	\end{tabular}}
-	\caption{Summary of the heat demand contributions and their calculation}
-	\label{tab:HeatDemand}
-\end{table}
-
-where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
-The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
-The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure \ref{fig:galleyhotwater}.
-
-\begin{figure}[tbp!]
-	\centering
-	\includegraphics[width=0.7\linewidth]{Figures/galleyHotWater.pdf}
-	\caption{Hourly relative demand of hot water heating and galley}
-	\label{fig:galleyhotwater}
-\end{figure}
-
-\subsubsection{Heat generation}
-
-The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty.
-The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
-\begin{equation}
-\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
-\end{equation}\label{eq:egb}
-
-where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance.
-
-It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
-Based on the available measurements, we calculated the heat exchanged in the two HTHRs according to equation \ref{eqn:HTHR2},
-
-\begin{eqnarray}
-\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
-& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
-\end{eqnarray}
-
-where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
-The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
-
-The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems.
-The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit.
-This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value.
-In this calculation, the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
-
-\subsubsection{Parameter estimation and uncertainty quantification}
-
-In this study, we base the estimation of the uncertainty on the production side, as it is the one that has the largest amount of information available.
-In these regards, the uncertainty can be reduced to the contribution of three elements: $U(\dot{Q}_{EGB})$, $U(\dot{Q}_{HTHR})$, $U(\dot{Q}_{AB})$.
-
-The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$.
-In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT].
-The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions and can be estimated of being up to 10\%.
-The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value.
-These assumptions lead to the following estimation of the uncertainty:
-\begin{equation}
-\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
-\end{equation}
-
-That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements.
-
-The uncertainty of the heat generated by the ABs can be reduced to the contribution of two elements: the uncertainty on $\dot{m}_{fuel,AB}$ and that on $\eta_{AB}$.
-The former will be at least as large as the calibration error (35\%) and will be considered equal to 50\% to be conservative and accounting also for errors in the aggregated boiler fuel measurements.
-In addition, the uncertainty on the efficiency can be considered to be around 10\% based on the discrepancy between the considered sources and on the expected variability of the efficiency with load.
-Similarly to the previous case, the combination of these efficiencies lead to a total uncertainty of 51\%, where the main contribution comes from the uncertainty of the model output compared to the actual fuel consumption.
-
-The estimation of the uncertainty of $\dot{Q}_{HTHR}$ can be based on \ref{eqn:HTHR2} having assigned an acceptable variation of 20\% to the effectiveness of the heat exchanger, 20\% on the estimation of the reference mass flow, $\pm 5$ K uncertainty on water temperature measurements and considering negligible the uncertainty on $c_{p,w}$, leading to a 145\% uncertainty on this measurement.
+ADD TABLE WITH LEGEND
+
+The determination of the heat demand is the most challenging in the analysis of ship energy systems, as extensive and appropriate measurements related to this contribution are rarely available.
+In this case the limited data available was fit with assumptions related to the behavior of the energy system and to the daily repartition of the heating demand.
+A more detailed description of this procedure is provided in Appendix REF.
 
@@
 Several algorithms are introduced in the literature to partition these datasets.
-The one selected in this work is the Lloyd’s (k-means) algorithm, which is a non-deterministic method computationally faster than conventional hierarchical tools.
+The one selected in this work is the Lloyd’s (k-means) algorithm \cite{Kanungo2002}, which is a non-deterministic method computationally faster than conventional hierarchical tools.
 The approach applied by Fazlollahi et al.
-was used to select the optimum number of clusters (Nk).
+\cite{Fazlollah2014} was used to select the optimum number of clusters (Nk).
 It (Nk) should be as low as possible for data handling purposes, while preserving a high accuracy of the retrieved data.
@@
 \begin{enumerate}
-	\item the profile deviation (deviation between original and typical period profiles),$\sigma_{profile}$;
-	\item the deviation from the load duration curve of the average values of each period, $\sigma{CDC}$;
+	\item the profile deviation (deviation between original and typical period profiles), $\sigma_{profile}$;
+	\item the deviation from the load duration curve of the average values of each period, $\sigma_{CDC}$;
 	\item the relative error in load duration curve deviation $ELDC$;
@@
 Figure \ref{fig:vship_hist} represents the distribution of the ship speed.
-The ship operates for almost  the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
+The ship operates for almost  the entire year according to a fixed schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed.
 In Figure \ref{fig:opMode_pie} it can be also observed that the ship spends a relatively large amount of time in port.
@@
 
-The ship energy demand is first subdivided among  The evolution of the demands of propulsion, electric power, and heat over time during a typical voyage is shown in Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}, representing a winter and a summer day respectively.
+The ship energy demand is first subdivided among the different types of consumers.
+The evolution of the demands of propulsion, electric power, and heat over time during a typical voyage is shown in Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}, representing a winter and a summer day respectively.
 As expected, the heating demand in winter is higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduced need for compartment heating.
@@
 
-Heat recovery has a large impact on the overall heat demand (see Fig.
+Heat recovery has a large impact on the overall external heat demand (see Fig.
 \ref{fig:Pie_HeatGeneration}).
@@
 The installation of batteries was also considered in a previous publication by the authors \cite{Baldi2017} and proved the potential for yearly savings of 1-2\%.
-In this second case, however, it should be noted that the economic performance was not ensured.
+In this second case, however, it should be noted that the economic performance was not evaluated quantitatively.
 
 The availability of waste heat that is not already recovered by the system suggests that the system could benefit from the installation of a Heat-to-power WHR.
-This possibility was tested by the authors both in the case that no additional retrofit is performed \cite{Ahlgren2016,Mondejar2017}, with estimated savings of 22\% of the auxiliary power demand for a reference voyage, and in the case of a full integration with the rest of the system \cite{Baldi2017}, with estimated savings of up to 6\% of the total fuel consumption for a year of ship operations.
+This possibility was tested by the authors both in the case that no additional retrofit is performed \cite{Mondejar2017,Ahlgren2016}, with estimated savings of 22\% of the auxiliary power demand for a reference voyage, and in the case of a full integration with the rest of the system \cite{Baldi2017}, with estimated savings of up to 6\% of the total fuel consumption for a year of ship operations.
 
@@
 
-\section{Engine modelling, figures and tables} \label{sec:appendix:engines}
+\section{Energy system modelling}
+
+\subsection{Data gathering and pre-processing} \label{sec:met:gathering}
+
+The operational data was collected on board from the ships' machine logging and surveillance system.
+The on board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB.
+The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average.
+The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python \cite{mckinney-proc-scipy-2010}.
+A new structured naming of headers and a consistent time frequency over all data points were created, and 245 selected data points were saved in a HDF5 table time series database which is the base for the analysis.
+
+All data-points were checked individually by creating a descriptive statistic and a histogram of each data-point.
+A filter was applied by setting up a pre-defined maximum and minimum value.
+As we do not know, or had any meaningful way of deriving, each individual sensors placement, sensor type or calibration status all data must be filtered and checked accordingly for outliers or bad data.
+
+Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska H\"{o}garna for the ambient temperature.
+The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska H\"{o}garna lighthouse is along the ship route, in between the Swedish archipelago and Åland.
+The assumption was validated based on  June-December period, for which onboard measurements were available.
+This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
+The fit the SMHI-data with the rest of the data-set the SMHI data was resampled from 1h for the seawater and 3h for the ambient temperature, to 15min frequency using a linear interpolation.
+
+\subsection{Engine modelling, figures and tables} \label{sec:appendix:engines}
+
+The validity of this assumption can be seen by observing Figures \ref{fig:Pme_vs_vship_Neng} and \ref{fig:Pme_vs_TCspeed} where the calculated  power from the main engines is represented against the ship speed and the turbocharger speed, respectively.
+
+\begin{align}
+\dot{m}_{air,comp} & =  \dot{m}_{air,cyl} + \dot{m}_{air,bp} \label{eq:mb:comp} \\ 
+\dot{m}_{eg,turb} & = \dot{m}_{eg,cyl} + \dot{m}_{air,bp} \label{eq:mb:turb} \\
+\dot{m}_{eg,cyl} & = \dot{m}_{air,cyl} + \dot{m}_{fuel,cyl} \label{eq:mb:cyl} \\
+\dot{m}_{air,comp} \Delta h_{comp} & = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech} \label{eq:eb:turbocharger} \\
+\dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0) & = \dot{m}_{air,bp} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) \label{eq:eb:bpmerge}
+\end{align}
+Where equations \ref{eq:mb:comp} to \ref{eq:mb:cyl} represent the mass balances of the bypass split, bypass merge, and cylinder respectively, while equations \ref{eq:eb:turbocharger} and \ref{eq:eb:bpmerge} represent the energy balances of the turbocharger and of the bypass merge, respectively.
+
+The energy balance over the whole engine is presented in Equation \ref{eqn:engineEnergyBalance}.
+\begin{equation}
+\dot{Q}_{fuel} + \dot{Q}_{air,in} = \dot{W}_{mech} + \dot{Q}_{eg} + \dot{Q}_{cooling}  \label{eqn:engineEnergyBalance} \\
+\end{equation}
+With: 
+\begin{align}
+\dot{Q}_{fuel} & =\dot{m}_{fuel} (LHV_{fuel} + c_{p,fuel} (T_{fuel,in} - T_0))\\
+\dot{Q}_{air,in} & = \dot{m}_{air} c_{p,air} (T_{air,comp,in} - T_0) \\ 
+\dot{Q}_{eg} & = \dot{m}_{eg} c_{p,eg} (T_{eg,TC,out} - T_0) \\
+\dot{Q}_{cooling} & = \dot{Q}_{CAC,HT} + \dot{Q}_{CAC,LT} + \dot{Q}_{JWC} + \dot{Q}_{LOC} \label{eq:eb:cooling}
+\end{align}
+
+The value of $\dot{Q}_{cooling}$ is determined based on the balance in Equation \ref{eqn:engineEnergyBalance}.
+The share between the four contributions in Equation \ref{eq:eb:cooling} is calculated based on the data for the heat loss to the HT and LT systems presented in the engines project guides as functions of the engine load.
+The values are interpolated based on a 2$^nd$ degree polynomial and scaled to respect the energy balance.
+In absence of more accurate data, it is then assumed that the contributions of $\dot{Q}_{JWC}$ and $\dot{Q}_{LOC}$ are equal for all loads.
 
@@
 \end{figure}
+
+The most relevant operative values are provided in Tables \ref{tab:ME_values} and \ref{tab:AE_values} and the main energy flows are listed in Tables \ref{tab:ME_flows} and \ref{tab:AE_flows}.
 
@@
 
+\subsection{Electric demand modelling}
+
+\subsubsection{HVAC systems}
+
+The observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a  period, corresponding to the summer months, when the electric power demand is significantly higher then during the remaining months.
+Based on this observation, in this work we assumed that the demand of the HVAC compressors related to the need for space cooling is concentrated during the summer months.
+Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a typical summer and winter day, shows that the daily variation is comparable.
+This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pel_vs_time}
+	\caption[Total electric power demand versus time (Daily average)]{Total electric power demand versus time (Daily average)}
+	\label{fig:pelvstime}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/PelRel_vs_time}
+	\caption{Instantaneous power demand over daily average, winter versus summer day}
+	\label{fig:pelrelvstime}
+\end{figure}
+
+Hence, the HVAC electric power demand was estimated as follows:
+\begin{equation}
+P_{el,HVAC} =
+\begin{cases}
+P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
+0, & \text{otherwise}
+\end{cases}
+\end{equation}
+where $P_{el,ref}(t)$ is calculated as:
+\begin{equation}
+P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
+\end{equation}
+
+\subsubsection{Thrusters}
+
+When entering port areas, the ship needs to use thrusters to maneuver and berth.
+In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.
+In order to isolate the demand from the thrusters, we extracted from ship operations a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average.
+From this daily profile (made of 96 points), we manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).
+
+By comparing the reference profile to the instantaneous one, the points where the former is more than 10\% higher than the latter are identified as "thruster-on" points and treated consequently.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Thruster_selection}
+	\caption{Example of the procedure of selection of thruster power demand}
+	\label{fig:thruster_selection}
+\end{figure}
+
+\clearpage
+
+\subsubsection{Heat balance parameter estimation}
+
+The individual contributions of different consumers were determined by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters.
+The parameter estimation problem is hence written as a minimization problem:
+\begin{eqnarray}
+min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
+\end{eqnarray}
+
+where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections.
+A list of the parameters $\textbf{p}$ is shown in Table \ref{tab:ParameterEstimation}, together with the chosen upper and lower boundaries for the calibration procedure.
+\begin{table}
+	\centering
+	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
+		\hline 
+		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
+		\hline
+		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 & 20.5 \\ 
+		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 & 0 \\
+		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 & 0.40 \\ 
+		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 & 1.00 \\ 
+		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 & 0.72 \\ 
+		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 & 0.68 \\ 
+		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 & 0.21 \\ 
+		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 & 348 \\
+		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 & 0.5 \\
+		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 & 9.5e6 \\ 
+		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 & 5100 \\ 
+		\hline
+	\end{tabular}
+	\caption{Parameters optimized in the parameter estimation for the heat balance}
+	\label{tab:ParameterEstimation} 
+\end{table}
+
+\subsubsection{Heat demand}
+
+The heat demand was calculated as the sum of the contributions of the elements listed in the ship's heat balance documentation.
+The assumptions used in the estimation are summarized in Table \ref{tab:HeatDemand}:
+\begin{table}
+	\centering
+	{\tablinesep=2ex\tabcolsep=10pt
+		\begin{tabular}{p{2.8cm}l}
+			\hline 
+			Heat flow name & Equation \\
+			\hline
+			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
+			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
+			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
+			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
+			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
+			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
+			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
+			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
+			\hline
+	\end{tabular}}
+	\caption{Summary of the heat demand contributions and their calculation}
+	\label{tab:HeatDemand}
+\end{table}
+
+where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
+The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
+The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure \ref{fig:galleyhotwater}.
+
+\begin{figure}[tbp!]
+	\centering
+	\includegraphics[width=0.7\linewidth]{Figures/galleyHotWater.pdf}
+	\caption{Hourly relative demand of hot water heating and galley}
+	\label{fig:galleyhotwater}
+\end{figure}
+
+\subsubsection{Heat generation}
+
+The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty.
+The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
+\begin{equation}
+\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
+\end{equation}\label{eq:egb}
+
+where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance.
+
+It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
+Based on the available measurements, we calculated the heat exchanged in the two HTHRs according to equation \ref{eqn:HTHR2},
+
+\begin{eqnarray}
+\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
+& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
+\end{eqnarray}
+
+where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
+The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
+
+The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems.
+The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit.
+This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value.
+In this calculation, the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
+
+\subsubsection{Parameter estimation and uncertainty quantification}
+
+In this study, we base the estimation of the uncertainty on the production side, as it is the one that has the largest amount of information available.
+In these regards, the uncertainty can be reduced to the contribution of three elements: $U(\dot{Q}_{EGB})$, $U(\dot{Q}_{HTHR})$, $U(\dot{Q}_{AB})$.
+
+The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$.
+In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT].
+The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions and can be estimated of being up to 10\%.
+The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value.
+These assumptions lead to the following estimation of the uncertainty:
+\begin{equation}
+\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
+\end{equation}
+
+That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements.
+
+The uncertainty of the heat generated by the ABs can be reduced to the contribution of two elements: the uncertainty on $\dot{m}_{fuel,AB}$ and that on $\eta_{AB}$.
+The former will be at least as large as the calibration error (35\%) and will be considered equal to 50\% to be conservative and accounting also for errors in the aggregated boiler fuel measurements.
+In addition, the uncertainty on the efficiency can be considered to be around 10\% based on the discrepancy between the considered sources and on the expected variability of the efficiency with load.
+Similarly to the previous case, the combination of these efficiencies lead to a total uncertainty of 51\%, where the main contribution comes from the uncertainty of the model output compared to the actual fuel consumption.
+
+The estimation of the uncertainty of $\dot{Q}_{HTHR}$ can be based on \ref{eqn:HTHR2} having assigned an acceptable variation of 20\% to the effectiveness of the heat exchanger, 20\% on the estimation of the reference mass flow, $\pm 5$ K uncertainty on water temperature measurements and considering negligible the uncertainty on $c_{p,w}$, leading to a 145\% uncertainty on this measurement.
+
+\clearpage
+
 \section{Full tables of energy and exergy flows} \label{sec:appendix:energy}

### 2018-06-05 Manuscript/Manuscript.tex
Updated Appendix

@@
 
-\section{Energy system modelling}
+\section{Extended method description}
+
+In this appendix, a thorough description of the methods of this paper is provided, with particular focus on the data processing from raw measurements and ship technical documentation to the values required to calculate energy and exergy flows.
+These details were omitted in the main part of the paper to improve itse readability.
 
@@
 
-\subsection{Engine modelling, figures and tables} \label{sec:appendix:engines}
+\subsection{Diesel engines modeling} \label{sec:appendix:engines}
+
+In the case of the main engines, the engine power output was not measured and needed to be estimated.
+In this work, we calculated the engine power output based on measurements of engine fuel rack position (used as a proxy of the mass of fuel injected per cycle) and of the engine speed (see equations \ref{eqn:ME:eq1} to \ref{eqn:ME:eq3}).
+In the case of the auxiliary engines, instead, direct measurements of the power generated by each engine were available.
+\begin{eqnarray}
+\dot{m}_{fuel} & = & \dot{m}_{fuel,des} \left(a_0 + a_1 \frac{frp}{frp_{des}}\right) \frac{\omega}{\omega_{des}} \label{eqn:ME:eq1} \\
+\eta_{ME} & = & a_0 + a_1 \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} + a_2 \left( \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} \right)^2 \label{eqn:ME:eq2} \\
+\dot{W}_{ME} & = & \dot{m}_{fuel} \eta_{ME} LHV \label{eqn:ME:eq3}
+\end{eqnarray}
 
 The validity of this assumption can be seen by observing Figures \ref{fig:Pme_vs_vship_Neng} and \ref{fig:Pme_vs_TCspeed} where the calculated  power from the main engines is represented against the ship speed and the turbocharger speed, respectively.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Pme_vs_vship_Neng}
+	\caption{Scatter plot, Propulsion power versus ship speed}
+	\label{fig:Pme_vs_vship_Neng}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/Scatter_Pme_vs_TCspeed}
+	\caption{Scatter plot, Main engine power versus turbocharger speed (ME1)}
+	\label{fig:Pme_vs_TCspeed}
+\end{figure}
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pme_vs_vship_Neng}
-	\caption{Scatter plot, Propulsion power versus ship speed}
-	\label{fig:Pme_vs_vship_Neng}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Scatter_Pme_vs_TCspeed}
-	\caption{Scatter plot, Main engine power versus turbocharger speed (ME1)}
-	\label{fig:Pme_vs_TCspeed}
-\end{figure}
-
 The most relevant operative values are provided in Tables \ref{tab:ME_values} and \ref{tab:AE_values} and the main energy flows are listed in Tables \ref{tab:ME_flows} and \ref{tab:AE_flows}.
@@
 	\begin{tabular}{llccrrrr}
-		\hline
+		\toprule
 		\multicolumn{2}{l}{Variable} & Unit & M/C & 30\% & 50\% & 70\% & 90\% \\
-		\hline
+		\midrule
 		\multicolumn{2}{l}{Compressor} & & & & & & \\ 
@@
 		&	$T_{w,out}$	   		& $^o$C & C  & 360.8 & 360.2 & 359.6 & 358.7 \\ 
-		\hline
+		\bottomrule
 	\end{tabular}
@@
 		&	$T_{w,out}$	   		& $^o$C & C  & 362.5 & 364.2 & 368.0 & 368.2 \\  
-		\hline
+		\bottomrule
 	\end{tabular}
@@
 	\begin{tabular}{lcrrrr}
-		\hline
+		\toprule
 		Energy flow & type & 30\% & 50\% & 70\% & 90\% \\
-		\hline
+		\midrule
 		Power output & Mech &  1767 & 2946 & 4118 & 5293 \\	
@@
 		LOC & Heat & 501 & 507 & 549 & 620 \\
-		\hline
+		\bottomrule
 	\end{tabular}
@@
 	\begin{tabular}{lcrrrr}
-		\hline
+		\toprule
 		Energy flow & type & 30\% & 50\% & 70\% & 90\% \\
-		\hline
+		\midrule
 		Power output & Mech & 828 & 1381 & 1932 & 2482 \\
-
 		Exhaust gas (after turbine) & Heat & 808 & 1144 & 1456 & 1753 \\
@@
 		LOC & Heat & 278 & 366 & 436 & 510 \\ 
-		\hline
+		\bottomrule
 	\end{tabular}
@@
 
-\subsection{Electric demand modelling}
+\subsection{Electric demand modeling}
 
@@
 
-Hence, the HVAC electric power demand was estimated as follows:
+The HVAC electric power demand was estimated as follows:
 \begin{equation}
@@
 \clearpage
+
+\subsubsection{Heat demand}
+
+The heat demand was calculated as the sum of the contributions of the elements listed in the ship's heat balance documentation.
+The assumptions used in the estimation are summarized in Table \ref{tab:HeatDemand}:
+\begin{table}
+	\centering
+	{\tablinesep=2ex\tabcolsep=10pt
+		\begin{tabular}{p{2.8cm}l}
+			\hline 
+			Heat flow name & Equation \\
+			\hline
+			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
+			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
+			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
+			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
+			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
+			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
+			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
+			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
+			\hline
+	\end{tabular}}
+	\caption{Summary of the heat demand contributions and their calculation}
+	\label{tab:HeatDemand}
+\end{table}
+
+where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
+The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
+The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure \ref{fig:galleyhotwater}.
+
+\begin{figure}[tbp!]
+	\centering
+	\includegraphics[width=0.7\linewidth]{Figures/galleyHotWater.pdf}
+	\caption{Hourly relative demand of hot water heating and galley}
+	\label{fig:galleyhotwater}
+\end{figure}
+
+\subsubsection{Heat generation}
+
+The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty.
+The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
+\begin{equation}
+\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
+\end{equation}\label{eq:egb}
+
+where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance.
+
+It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
+Based on the available measurements, we calculated the heat exchanged in the two HTHRs according to equation \ref{eqn:HTHR2},
+
+\begin{eqnarray}
+\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
+& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
+\end{eqnarray}
+
+where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
+The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
+
+The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems.
+The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit.
+This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value.
+In this calculation, the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
 
@@
 
-\subsubsection{Heat demand}
-
-The heat demand was calculated as the sum of the contributions of the elements listed in the ship's heat balance documentation.
-The assumptions used in the estimation are summarized in Table \ref{tab:HeatDemand}:
-\begin{table}
-	\centering
-	{\tablinesep=2ex\tabcolsep=10pt
-		\begin{tabular}{p{2.8cm}l}
-			\hline 
-			Heat flow name & Equation \\
-			\hline
-			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
-			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
-			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
-			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
-			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
-			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
-			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
-			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
-			\hline
-	\end{tabular}}
-	\caption{Summary of the heat demand contributions and their calculation}
-	\label{tab:HeatDemand}
-\end{table}
-
-where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}).
-The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively.
-The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure \ref{fig:galleyhotwater}.
-
-\begin{figure}[tbp!]
-	\centering
-	\includegraphics[width=0.7\linewidth]{Figures/galleyHotWater.pdf}
-	\caption{Hourly relative demand of hot water heating and galley}
-	\label{fig:galleyhotwater}
-\end{figure}
-
-\subsubsection{Heat generation}
-
-The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty.
-The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
-\begin{equation}
-\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
-\end{equation}\label{eq:egb}
-
-where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance.
-
-It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines.
-Based on the available measurements, we calculated the heat exchanged in the two HTHRs according to equation \ref{eqn:HTHR2},
-
-\begin{eqnarray}
-\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
-& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
-\end{eqnarray}
-
-where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}).
-The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
-
-The heat generated by the auxiliary boilers is calculated as to close the heat balance of the ship energy systems.
-The contribution of the boiler heat storage capacity is taken into account by a calibration parameter $Q_{ab,max}$ that determines the maximum heat deficit.
-This corresponds, in practice, to assuming that the boiler is started up when the steam pressure inside the boiler drops below a certain value, and stopped once the pressure has achieved its maximum operative value.
-In this calculation, the calibration parameters are the heat storage capacity ($Q_{ab,max}$) and the fixed heat rate ($\dot{Q}_{ab,des}$) of the auxiliary boilers (see table \ref{tab:ParameterEstimation}).
-
-\subsubsection{Parameter estimation and uncertainty quantification}
+\subsubsection{Uncertainty quantification of heat demand}
 
@@
 \section{Full tables of energy and exergy flows} \label{sec:appendix:energy}
+
+In this appendix, the cumulated yearly values for the main ship energy and exergy flows are provided.
+Apart from yearly values, we also provide the same statistics separated based on the ship operational mode (P/S: Port stay, LSS: Low speed sailing, Man: Maneuvering, HSS, High speed sailing) and based on the season (Win: Winter, Mid: Spring/autumn, Sum: Summer).
 
@@
 
+In this section, the full list of all measurements available in the dataset that was used in this paper is provided. It should be noted that not all the values were used in the analysis.
+
 \clearpage
+
+\section{Energy demand for typical operational days}
+
+In this section, the values for the main energy demands (Propulsion power, electric power, high temperature heat and low temperature heat) are provided for the typical days identified in the demand clustering.
 

### 2018-06-20 Manuscript/Manuscript.tex
Figures updates

@@
 \usepackage{tikz}
+\usepackage{pgfplots}
 
@@
 \dot{Q}_{gen} & = \sum_{i=2,3} \dot{Q}_{EGB,ME,i} +  \sum_{i=1:4} \dot{Q}_{EGB,AE,i} + \sum_{j=ME,AE, i=1:4}\dot{Q}_{HTHR,j,i} + \dot{Q}_{AB} \label{eq:Qgen} \\
-\dot{Q}_{dem} & = \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{HH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \label{eq:Qdem} \\
+\dot{Q}_{dem} & = \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{HH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \label{eq:Qdem}
 \end{align}
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{Figures/segmentation_power_heat.png}
+ 	% \includegraphics[width=0.95\linewidth]{Figures/segmentation_power_heat.png}
+ 	\includegraphics[width=0.95\linewidth]{Figures/segmentation_new.pdf}
 	\caption{Power (left) and heat (right) profiles, original (dotted) and segmented (full), for the 5 typical days and duration curves.}

### 2018-06-21 Manuscript/Figures/dataProcessing.tex
Some updates to the tables

@@
 	text = white,
+	font = \sffamily\Large,
 	fill = coltwo
@@
   fill = colthree,
+  font = \sffamily\Large,
   text centered
@@
 % Entries on the left
-\node [frameSections] (section1) at (0,3) {Data collection} ;
-\node [frameSections] (section2) at (14,2) {Data processing} ;
-\node [frameSections] (section3) at (31,2) {Analysis} ;
+\node [frameSections] (section1) at (1,2) {Data collection} ;
+\node [frameSections] (section2) at (15,2) {Data processing} ;
+\node [frameSections] (section3) at (29.5,2) {Analysis} ;
 
@@
 \path [lineData] (input_data.east) -- (process_filtering.west) ;
-\path [lineInfo] (input_tech.east) -| (process_filtering.south) ;
 
@@
 % LINE INFO
-\path [lineInfo] (11,-6) -| (process_AE.south) ;
-\path [lineInfo] (11,-2) -| (process_ME.south) ;
-\path [lineInfo] (13.5,-10) -| (process_HTHR.south) ;
-\path [lineInfo] (19,-9) -| (process_Q.south) ;
-\path [lineInfo] (19,-6) -| (process_Paux.south) ;
+%\path [lineInfo] (11,-6) -| (process_AE.south) ;
+%\path [lineInfo] (11,-2) -| (process_ME.south) ;
+%\path [lineInfo] (13.5,-10) -| (process_HTHR.south) ;
+%\path [lineInfo] (19,-9) -| (process_Q.south) ;
+%\path [lineInfo] (19,-6) -| (process_Paux.south) ;
+%\path [lineInfo] (input_tech.east) -| (process_filtering.south) ;
 
 % VARIABLES
-\node [frameVar] (aa1) at (25,-2.5) {$\dot{W}, \dot{Q}$} ;
-\node [frameVar] (aa1) at (25,-5.5) {$\dot{B}$} ;
+%\node [frameVar] (aa1) at (25,-2.5) {$\dot{W}, \dot{Q}$} ;
+%\node [frameVar] (aa1) at (25,-5.5) {$\dot{B}$} ;
 % \node [frameVar] (aa1) at (25,-6.5) {$\dot{m}, T$} ;

### 2018-06-21 Manuscript/Manuscript.tex
Some updates to the tables

@@
 
-The main principles applied to the modeling are presented in the following sections and summarized in Figure \ref{fig:method}, while a more thorough description and validation of the different approaches is presented in the Appendix.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.95\linewidth]{Figures/dataProcessing}
+The main principles applied to the modeling are presented in the following sections and summarized in Figure \ref{fig:dataprocessing}, while a more thorough description and validation of the different approaches is presented in the Appendix.
+The data is first processed to the main and auxiliary engines modules to derive the related energy flows.
+These are then used to calculate the electric power demand, the energy flows in the cooling systems and, consequently, the heat recovered in the HTHR systems.
+This information combined is used to determine the heating demand.
+Finally, all flows thus obtained are aggregated to determine energy and exergy flows and related efficiencies.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.99\linewidth]{Figures/dataProcessing}
 	\caption{Overview of the data processing from inputs to outputs}
@@
 	\includegraphics[width=0.999\linewidth]{Figures/Heating_cooling_systems}
-	\caption{General view of the ship heating and cooling systems}
+	\caption{General view of the ship heating and cooling systems. The full list of heat exchangers in the heat exchanger network is provided in the Appendix in Table \ref{tab:appendix:HEN}}
 	\label{fig:ship:general}
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/Pie_EnergyDemand}
+	\includegraphics[width=0.99\linewidth]{Figures/Pie_EnergyDemand}
 	\caption{Energy demand, aggregated data for one year of operations}
@@
 
+\subsubsection{Full list of exchangers in the ship's heat exchanger network}
+
+\begin{table}[hbp!]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{cll}
+\toprule
+& Identifier & Full name \\
+\midrule
+\multirow{ 14}{*}{\rotatebox[origin=c]{90}{Low temperature cooling}} & $HEX_{LT1.1}$ & Charge air cooler, LT stage (ME1)  \\
+& $HEX_{LT1.2}$ & Lubricating oil cooler (ME1) \\
+& $HEX_{LT2.1}$ & Charge air cooler, LT stage (ME3) \\
+& $HEX_{LT2.2}$ & Lubricating oil cooler (ME3) \\
+& $HEX_{LT3.1}$ & Charge air cooler, LT stage (AE1) \\
+& $HEX_{LT3.2}$ & Lubricating oil cooler (AE1) \\
+& $HEX_{LT4.1}$ & Charge air cooler, LT stage (AE3) \\
+& $HEX_{LT4.2}$ & Lubricating oil cooler (AE3) \\
+& $HEX_{LT5}$ & Generator cooler (AE) \\
+& $HEX_{LT6}$ & Propeller oil cooler \\
+& $HEX_{LT7}$ & Shaft bearing cooler (ME) \\
+& $HEX_{LT8}$ & Steam condenser \\
+& $HEX_{LT9}$ & Fin stabiliser cooler \\
+& $HEX_{LT10}$ & Gearbox cooler \\
+\midrule
+\multirow{ 9}{*}{\rotatebox[origin=c]{90}{\parbox[c]{3.5cm}{\centering High temperature cooling}}} & $HEX_{HT1.1}$ & Jacket water cooler (ME1) \\
+& $HEX_{HT1.2}$ & Charge air cooler, HT stage (ME1) \\
+& $HEX_{HT2.1}$ & Jacket water cooler (ME3) \\
+& $HEX_{HT2.2}$ & Charge air cooler, HT stage (ME3) \\
+& $HEX_{HT3.1}$ & Jacket water cooler  (AE1) \\
+& $HEX_{HT3.2}$ & Charge air cooler, HT stage  (AE1) \\
+& $HEX_{HT4.1}$ & Jacket water cooler  (AE3) \\
+& $HEX_{HT4.2}$ & Charge air cooler, HT stage  (AE3) \\
+& $HEX_{HT-HR1}$ & Heat recovery from HT cooling (eng.room1/3) \\
+\midrule
+\multirow{ 7}{*}{\rotatebox[origin=c]{90}{\parbox[c]{3cm}{\centering Hot water heat recovery system}}} & $HEX_{HT-HR1}$ & Heat recovery from HT cooling (eng.room1/3) \\
+& $HEX_{HT-HR2}$ & Heat recovery from HT cooling (eng.room 2/4) \\
+& $HEX_{St-HR}$ & Steam heater \\
+& $HEX_{HR1}$ & AC-Reating \\
+& $HEX_{HR2}$ & AC-Preheating \\
+& $HEX_{HR3}$ & Hot water heater \\
+& $HEX_{HR4}$ & Technical water generator \\
+\midrule 
+\multirow{ 8}{*}{\rotatebox[origin=c]{90}{Steam systems}} & $HEX_{HRSG1}$ & Heat recovery steam generator (AE1) \\
+& $HEX_{HRSG2}$ & Heat recovery steam generator (AE3) \\
+& $HEX_{HRSG3}$ & Heat recovery steam generator (ME3) \\
+& $HEX_{HRSG4}$ & Heat recovery steam generator (AE2) \\
+& $HEX_{HRSG5}$ & Heat recovery steam generator (AE4) \\
+& $HEX_{HRSG6}$ & Heat recovery steam generator (ME2) \\
+& $HEX_{AB1}$ & Auxiliary boiler 1 \\
+& $HEX_{AB2}$ & Auxiliary boiler 2 \\
+\bottomrule
+\end{tabular}}
+\caption{Full list of exchangers in the ship's heat exchanger networks}
+\label{tab:appendix:HEN}
+\end{table}
+
 \clearpage

### 2018-06-21 Python files/plotting.py

@@ -357,7 +357,7 @@ def predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure, filenames
                           processed["Demands:Heat:HFOheater:Edot"].sum(),
                           processed["Demands:Heat:HFOtankHeating:Edot"].sum(),
                           processed["Demands:Heat:Galley:Edot"].sum()]
-            labels = ["Propeller-1" , "Propeller-1" ,
+            labels = ["Propeller-1" , "Propeller-2" ,
                       "HVAC", "Thrusters", "Other users",
                       "HVAC Preheater", "HVAC Reheater", "Hot water heater",
                       "Machinery space heaters", "Other tanks heating", "Fuel tanks heating", "HFO tanks heating", "HFO pre-injection heating", "Galley"]
@@ -368,8 +368,8 @@ def predefinedPlots(processed, dataset_raw, CONSTANTS, dict_structure, filenames
                      "blue", "blue", "blue",
                      "sandybrown","sandybrown","sandybrown",
                      "red","red","red","red","red","red"]
-            ax.pie(quantities, labels=labels, explode=explode, autopct='%1.1f%%', colors=colors)
-
+            patches, texts, autotexts = ax.pie(quantities, labels=labels, explode=explode, autopct='%1.1f%%', colors=colors)
+            [_.set_fontsize(12) for _ in texts]
         if filename == "Pie:GenerationFull":
             quantities = [processed["ME1:Cyl:FuelPh_in:mdot"].sum(),
                           processed["ME2:Cyl:FuelPh_in:mdot"].sum(),

### 2018-07-04 Manuscript/Manuscript.tex
Addition of the abstract and of the nomenclature

@@
 \usepackage{pgfplots}
+% Package for the nomenclature
+\usepackage{nomencl} 
+\usepackage{ifthen} % Required for having groups in the nomenclature
+\makenomenclature
+%Redefines the command "nomgroup" for the nomeclature sot that several subgroups can be included in the nomenclature
+\RequirePackage{ifthen}
+\renewcommand{\nomgroup}[1]{%
+	\ifthenelse{\equal{#1}{A}}{\item[\textbf{Acronyms}]}{%
+		\ifthenelse{\equal{#1}{C}}{\item[\textbf{Constants}]}{%
+			\ifthenelse{\equal{#1}{S}}{\item[\textbf{Subscripts}]}{%
+				\ifthenelse{\equal{#1}{V}}{\item[\textbf{Variables}]}}}}}
 
@@
 \author[Linnaeus]{Fredrik Ahlgren}
-\author[DTU]{Tuong-Van Nguyen}
+\author[DTU,SP]{Tuong-Van Nguyen}
 \author[LTU]{Marcus Thern}
@@
 \address[Linnaeus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
+\address[SP]{Laboratory of Environmental and Thermal Engineering, Polytechnic School - University of S\~{a}o Paulo, S\~{a}o Paulo, Brazil}
 \address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
@@
 \begin{abstract}
-%% Text of abstract
+The shipping sector is today facing numerous challenges.
+Fuel prices are expected to increase in the medium-long term, and the recent decision of the IMO to require a reduction of shipping's greenhouse gas emissions by 50\% with respect to 2009 levels will important effort within this industry.
+In this context, passenger ships represent a small but increasing share of the industry.
+The complexity of the energy system of a ship where the energy required by propulsion is no longer the obvious main contributor to the whole energy use makes this kind of ship of particular interest for the analysis of how energy is converted from its original form to its final use on board.
+To illustrate this, we analyzed the energy and exergy flow rates of a cruise ship sailing in the Baltic Sea based on a combination of available measurements from ship operations and of mechanistic knowledge of the system.
+The energy analysis allows identifying propulsion as the main energy user (46\% of the total) followed by heat (27\%) and electric power (27\%) generation; the exergy analysis allowed instead identifying the main inefficiencies of the system: While exergy is primarily destroyed in all processes involving combustion (76\% of the exergy destruction is generated by the combustion in the Diesel engines and in the oil-fired boilers), the other main causes of exergy destruction are the turbochargers (29\% of the remaining exergy destruction), the heat recovery steam generators (11\%), the steam heaters (11\%) the preheater in the accommodation heating systems (10\%), the sea water coolers (9\%), and the electric generators (6\%); the main exergy losses happen instead in the exhaust gas, mostly from the main engines (67\% of total losses), particularly from those not equipped with heat recovery devices.
+The improved understanding which derives from the results of the energy and exergy analysis can be used as a guidance to identify where improvements of the systems should be directed.
+As a result of the energy flow analysis of the ship, in this paper we also propose a clustering of ship operations based on the concept of typical operational days.
+We found that, based on the operations of the case study, the use of five typical days, clustered based on the sum of mechanical and electrical power, and on heating demand, provides a good approximation of yearly ship operations and can hence be used for the design and optimization of the energy systems of the ship.
 
@@
 
+\printnomenclature % Prints the nomenclature
+
 \section{Introduction} \label{sec:introduction}
@@
 Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios \cite{Smith2014}.
-As an example, global transport demand has increased by 3.4\% in 2014, compared to a global GBP growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}.
+As an example, global transport demand has increased by 3.4\% in 2014, compared to a global gross product growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}.
 The OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3].
@@
 
+\nomenclature[A]{OECD}{Organisation for Economic Co-operation and Development}
+
 In addition to considerations related to GHG emissions, an emission control area (ECA) is enforced in the Baltic Sea  by the International Maritime Organisation since January 2015.
 This ECA stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels.
-More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}
+More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}.
+
+\nomenclature[A]{GHG}{Greenhous gas}
+\nomenclature[A]{ECA}{Emission Controlled Area}
 
@@
 In this context, it has been acknowledged that the world fleet is heterogeneous, and measures need to be evaluated on a ship-to-ship basis \cite{Bouman2017}.
-In this process, a deeper understanding of energy use on board of the specific ship is vital.
+In this process, a deeper understanding of energy use on-board of the specific ship is vital.
 
@@
 
-Most authors focused on the propulsion part of the problem, as this is often the most relevant energy demand on board.
+Most authors focused on the propulsion part of the problem, as this is often the most relevant energy demand on-board.
 Shi et al.
@@
 \cite{Basurko2013} shows the application of energy auditing methods to fishing vessels.
-This approach represents a step forward towards improving the understanding of how energy is used on board during actual ship operations.
+This approach represents a step forward towards improving the understanding of how energy is used on-board during actual ship operations.
 
@@
 % Heat availability
-Heat demand is rarely a subject of concern on board ships, with some notable exceptions.
+Heat demand is rarely a subject of concern on-board ships, with some notable exceptions.
 This is due to a combination of generally low demand and high availability from the waste heat of the engines.
@@
 With reference to different types of technologies, case studies, and designs,the several authors showed the existance of a quite significant potential for energy saving when WHR systems are employed, ranging from around 1\% for single-pressure steam cycles applied to two-stroke engines \cite{Theotokatos2012} to more complex systems based on ORCs (up to 10\%, \cite{Hountalas2012}) or including the cooling systems as a source of waste heat (over 10\%, \cite{Dimopoulos2012}).
-The case of the installation of an ORC on board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2017}, showing a similar potential.
-
-The potential uses for waste heat on board are not limited to improving the efficiency of the power plant.
-Waste heat is commonly used for fulfilling on board heat demand for spatial heating and freshwater generation \cite{Baldi2015a,Molland2011,Mccarthy1990};  Balaji et al.
+The case of the installation of an ORC on-board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2017}, showing a similar potential.
+
+\nomenclature[A]{WHR}{Waste heat recovery}
+\nomenclature[A]{ORC}{Organic Rankine cycle}
+
+The potential uses for waste heat on-board are not limited to improving the efficiency of the power plant.
+Waste heat is commonly used for fulfilling on-board heat demand for spatial heating and freshwater generation \cite{Baldi2015a,Molland2011,Mccarthy1990};  Balaji et al.
 \cite{Balaji2012} proposed the use of waste heat for ballast-water systems ; Salmi et al.
@@
 % Exergy analysis
-Exergy analysis provides a more accurate estimation of the potential for energy recovery on board.
+Exergy analysis provides a more accurate estimation of the potential for energy recovery on-board.
 The application of exergy analysis to the case of ship energy systems is, however, still limited.
@@
 \cite{Dimopoulos2012} showed how the process optimizing the WHR system of a container ship can be more efficient if exergy efficiency, rather than energy efficiency, is used as the target of the optimization; Baldi et al.
-\cite{Baldi2015a} also analyzed the exergy flows on board of a product tanker, showing that this allows having a more accurate understanding of what parts of the system show potential for improvement ; similar results were obtained by Marty et al.
+\cite{Baldi2015a} also analyzed the exergy flows on-board of a product tanker, showing that this allows having a more accurate understanding of what parts of the system show potential for improvement ; similar results were obtained by Marty et al.
 \cite{Marty2016}, who focused on the power plant of a cruise ship.
@@
 In this paper, we present the application of energy and exergy analysis (described in Section \ref{sec:met:energyExergy}) to a cruise ship.
-This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; the specific assumptions and methods used for processing the available information into  energy and exergy flows are specified in Section \ref{sec:met:processing}.
-In these regards, the estimation of the heat demand is treated in a separate section (Section \ref{sec:met:heat}) as it constitutes a particularly challenging task and is one of the central aspects of this article.
+This is done for one specific case study vessel, described in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; the specific assumptions and methods used for processing the available information into  energy and exergy flows are summarized in Section \ref{sec:met:processing}, while a more thorough description is provided in \ref{sec:app:method}.
 
@@
 \begin{align}
-	\sum_{\mathrm{in}} \dot{H}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{H}_{\mathrm{out}} \\
-	\dot{H}_{\mathrm{fuel}} + \dot{H}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}+\dot{W}_{\mathrm{el}}+\dot{Q}_{\mathrm{heating}}
+	\sum_{\mathrm{in}} \dot{H}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{H}_{\mathrm{out}}  \label{eqn:energyBalance}  \\
+	\dot{H}_{\mathrm{fuel}} + \dot{H}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}+\dot{W}_{\mathrm{el}}+\dot{Q}_{\mathrm{heating}} \label{eqn:energyBalance2}
 \end{align}
 
-The left-hand side term represents, on a time rate basis, the energy associated with the fuel consumed in the boilers and engines ($\dot{H}_{\mathrm{fuel}}$) and the air used in the combustion processes ($\dot{H}_{\mathrm{air}}$).
+\nomenclature[V]{H}{Enthalpy}
+\nomenclature[S]{in}{Inlet}
+\nomenclature[S]{out}{Outlet}
+\nomenclature[V]{Q}{Heat}
+\nomenclature[S]{el}{Electric}
+
+In Equation \ref{eqn:energyBalance2}, the left-hand side term represents, on a time rate basis, the energy associated with the fuel consumed in the boilers and engines ($\dot{H}_{\mathrm{fuel}}$) and the air used in the combustion processes ($\dot{H}_{\mathrm{air}}$).
 The right-hand side term denotes the power ($\dot{W}_{\mathrm{el}}$) and heat ($\dot{Q}_{\mathrm{heating}}$) required on-site (e.g.
@@
 The physical energy is taken as the relative enthalpy, as underlined in \cite{Kotas2013}, and the chemical energy is taken as the lower/higher heating value.
-The environmental conditions taken for the present analysis are the ambient pressure (1.01 bar) and seawater temperature (measured) .
+The environmental conditions taken for the present analysis are the ambient pressure (1.01 bar) and seawater temperature (measured).
 
@@
 \begin{align}
-	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \\
-	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d} 
+	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \label{eqn:exergyBalance} \\
+	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d}  \label{eqn:exergyBalance2}
 \end{align}
 
-The left-hand side term represents, on a time rate basis, the exergy associated with the fuel and air.
+\nomenclature[V]{E}{Exergy}
+\nomenclature[S]{d}{Destroyed}
+\nomenclature[S]{l}{Lost}
+\nomenclature[S]{prod}{Product}
+
+In Equation \ref{eqn:exergyBalance2} the left-hand side term represents, on a time rate basis, the exergy associated with the fuel and air.
 The right-hand side term denotes the exergy of the waste streams, the exergy transfers with heat and power, and the exergy destroyed in the ship system.
@@
 The chemical exergy is calculated based on the reference environment of \cite{Szargut1989}, and its value is approximatively equal to the higher heating value for hydrocarbon fuels.
-The exergy destruction can be calculated from the Gouy-Stodola theorem \cite{Kotas2013}.
+The exergy destruction $(\dot{E}_d)$ can be calculated using the Gouy-Stodola theorem \cite{Kotas2013}.
 
@@
 \begin{equation}
-\dot{E}_{\mathrm{p}} = \dot{E}_{\mathrm{f}} - \dot{E}_{\mathrm{d}} - \dot{E}_{\mathrm{l}} 
+\dot{E}_{\mathrm{p}} = \dot{E}_{\mathrm{fuel}} - \dot{E}_{\mathrm{d}} - \dot{E}_{\mathrm{l}} 
 \end{equation}
@@
 where $\dot{E}_{\mathrm{p}}$ is called the exergy product, and corresponds to the desired output of the system, in exergy terms (for example, the power produced in an engine).
-$\dot{E}_{\mathrm{f}}$ denotes the exergy fuel, and represents the resources spent to drive the studied process (for instance, the fuel used in a combustion process).
+$\dot{E}_{\mathrm{fuel}}$ denotes the exergy fuel, and represents the resources spent to drive the studied process (for instance, the fuel used in a combustion process).
 The last term $\dot{E}_{\mathrm{l}}$ corresponds to the losses of a system, such as the heat discharged into the environment with cooling water.
@@
 
+\nomenclature[V]{$\varepsilon$}{Exergy efficiency}
+\nomenclature[V]{$\lambda$}{Efficiency defect}
+\nomenclature[V]{$\delta$}{Irreversibility share}
+\nomenclature[S]{tot}{Total}
+
 \subsection{Case study vessel} \label{sec:met:case}
@@
 The ship systems are summarized in Figure \ref{fig:shipSystems}.
-The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller.
+The propulsion system is composed of two propulsion lines, each made of two engines, a gearbox, and a propeller.
 The main engines are four W\"{a}rtsil\"{a} 4-stroke Diesel engines (ME) rated 5850 kW each.
@@
 Auxiliary power is provided by four auxiliary engines (AE) rated 2760 kW each.
-Auxiliary power is needed on board for a number of alternative functions, from pumps in the engine room to lights, restaurants, ventilation and entertainment for the passengers.
+Auxiliary power is needed on-board for a number of alternative functions, from pumps in the engine room to lights, restaurants, ventilation and entertainment for the passengers.
 
@@
 
-The operational data was collected on board from the ships' machine logging and surveillance system.
-The data sampling rate was selected to 15 minutes, automatically calculated as averages over the respective periods by the on board data logging systems, as an acceptable compromise between accuracy and computational time required for the data processing.
+The operational data was collected on-board from the ship's machine logging and surveillance system.
+The data sampling rate was selected to 15 minutes, automatically calculated as averages over the respective periods by the on-board data logging systems, as an acceptable compromise between accuracy and computational time required for the data processing.
 Data consistency was checked individually by creating a descriptive statistic and a histogram of each measurement.
-A filter was applied by setting up a pre-defined maximum and minimum value.
+A filter was applied by setting up a pre-defined maximum and minimum value and eliminating from the analysis all data points where measurements were located outside these boundaries.
+This led to discarding XXX out of YYY data points.
 
@@
 
-The main principles applied to the modeling are presented in the following sections and summarized in Figure \ref{fig:dataprocessing}, while a more thorough description and validation of the different approaches is presented in the Appendix.
-The data is first processed to the main and auxiliary engines modules to derive the related energy flows.
+The main principles applied to the modeling are presented in the following sections and summarized in Figure \ref{fig:dataprocessing}, while a more thorough description and validation of the different approaches is presented in  \ref{sec:app:method}.
+The data are first processed by the main and auxiliary engines modules to derive the related energy flows.
 These are then used to calculate the electric power demand, the energy flows in the cooling systems and, consequently, the heat recovered in the HTHR systems.
 This information combined is used to determine the heating demand.
-Finally, all flows thus obtained are aggregated to determine energy and exergy flows and related efficiencies.
+Finally, all flows thus obtained are aggregated to determine energy and exergy flows and related performance indicators.
 
@@
 
-The Diesel engines are the only power source of the system, subdivided in four main engines uniquely devoted to propulsion and four auxiliary engines used for on board electric power generation.
+The Diesel engines are the only source of electric and mechanical power in the system, subdivided in four main engines for propulsion and four auxiliary engines for on-board electric power generation.
 The energy and exergy analysis of the engines is not limited to the boundary of the system, but is brought to the detail of the different components that constitute the engine systems, as shown in Figure \ref{fig:DieselEngines}.
@@
 
+\nomenclature[V]{$\omega$}{Rotational speed}
+\nomenclature[V]{$\dot{m}$}{Mass flow}
+\nomenclature[V]{$\eta$}{Energy efficiency}
+\nomenclature[S]{des}{Design}
+\nomenclature[A]{ME}{Main engine}
+\nomenclature[A]{AE}{Auxiliary engine}
+\nomenclature[V]{LHV}{Lower heating value}
+\nomenclature[V]{frp}{Fuel rack position}
+
+In Equations \ref{eqn:ME:eq1} to \ref{eqn:ME:eq3}, $\dot{m}_{fuel}$, $frp$, $\omega$, $\eta_{ME}$, and $LHV$ represent the fuel flow to the engine, the fuel rack position, the engine speed, the engine efficiency and the lower heating value of the fuel, respectively, and the $des$ subscript refers to design conditions.
+
 \begin{figure}
@@
 For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine.
-Values for temperatures, mass flows and energy flows for different engine loads are provided in the \ref{sec:appendix:engines}, for both main and auxiliary engines.
-
-\subsection{Electric power demand} \label{sec:met:electric}
-
-The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy.
+Values for temperatures, mass flows and energy flows for different engine loads are provided in the \ref{sec:app:met:engines}, for both main and auxiliary engines.
+
+\subsubsection{Electric power demand} \label{sec:met:electric}
+
+The total electric power demand of the system ($P_{el,tot}$) is easily estimated as the sum of the power generated by the four auxiliary engines ($P_{AE}$), that is directly measured on the electrical side of the generators and has a rather high accuracy.
 \begin{equation}
@@
 
+\nomenclature[V]{P}{Power}
+
 The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc.
@@
 In this paper only the contributions of the HVAC and the bow thrusters are identified separately from the remaining parts of the system.
-More details about the methods employed in this process are provided in Appendix REF.
-
-\subsection{Heating and cooling systems} \label{sec:met:heat}
-
+More details about how these contributions were separated are provided in \ref{sec:app:met:electric}.
+
+\subsubsection{Heating and cooling systems} \label{sec:met:heat}
+
+The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}.
 The case study ship is not provided with any means to directly measure the heat demand (e.g.
-steam flow meters), and it was hence necessary to provide estimations based on other measurements and assumptions.
-The heat demand is subdivided among the contributions of the HVAC preheater and reheater ($\dot{Q}_{HVAC,PH}$ and $\dot{Q}_{HVAC,RH}$), the hot water heater ($\dot{Q}_{HWH}$), the HFO heaters ($\dot{Q}_{HH}$) the tank heating systems for HFO and other liquids ($\dot{Q}_{HTH}$ and $\dot{Q}_{TH}$), the machinery space heating ($\dot{Q}_{MSH}$) and other tanks ($\dot{Q}_{OT}$).
+steam flow meters) and cooling flows, and it was hence necessary to provide estimations based on other measurements and assumptions.
+
+The ship's cooling systems are designed similarly to most ships and divided in two cooling systems operated at two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C).
+The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on-board.
+
+\nomenclature[A]{HT}{High temperature}
+\nomenclature[A]{LT}{Low temperature}
+
+All values not directly measured in the cooling systems were calculated based on mass and energy balances.
+Mass flows were determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps as provided in the engine project guides.
+
+The heat demand is subdivided among the contributions of the HVAC preheater ($\dot{Q}_{HVAC,PH}$) and reheater ($\dot{Q}_{HVAC,RH}$), the hot water heater ($\dot{Q}_{HWH}$), the HFO heaters ($\dot{Q}_{HH}$) the tank heating systems for HFO ($\dot{Q}_{HTH}$) and other liquids ($\dot{Q}_{TH}$), the machinery space heating ($\dot{Q}_{MSH}$) and other tanks ($\dot{Q}_{OT}$).
+
 The heat generation is instead composed of the contribution of the exhaust gas boilers ($\dot{Q}_{EGB,ME,i}$ and $\dot{Q}_{EGB,AE,i}$), the heat recovery from the high temperature cooling systems ($\dot{Q}_{HTHR,j,i}$) and the boilers ($\dot{Q}_{AB}$).
@@
 
-The ship's cooling systems are designed similarly to most ships and divided in two cooling systems operated at two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C).
-The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board.
-An overview of the ship heating and cooling systems is provided in Figure \ref{fig:generalsystemssimplified}.
-
-All values not directly measured in the cooling systems were calculated based on mass and energy balances.
-Mass flows were determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps as provided in the engine project guides.
-The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}.
-
 \begin{figure}
@@
 	\includegraphics[width=0.999\linewidth]{Figures/Heating_cooling_systems}
-	\caption{General view of the ship heating and cooling systems. The full list of heat exchangers in the heat exchanger network is provided in the Appendix in Table \ref{tab:appendix:HEN}}
+	\caption{General view of the ship heating and cooling systems.
+The full list of heat exchangers in the heat exchanger network is provided in Appendix \ref{sec:app:met:hen} in Table \ref{tab:appendix:HEN}}
 	\label{fig:ship:general}
@@
 
-ADD TABLE WITH LEGEND
-
-The determination of the heat demand is the most challenging in the analysis of ship energy systems, as extensive and appropriate measurements related to this contribution are rarely available.
-In this case the limited data available was fit with assumptions related to the behavior of the energy system and to the daily repartition of the heating demand.
-A more detailed description of this procedure is provided in Appendix REF.
+ As a consequence of the limited amount of available data, a we propose a method based on a combination of operational measurements and assumptions related to the behavior of the energy system for the determination of the heat demand and of its share among different consumers.
+According to this method, the demand and supply of heat on-board are defined using a number of unknown parameters, that are subsequently trained using an optimization procedure in order to minimize the error on the monthly fuel consumption from the boilers, which is available from ship documentation.
+A more detailed description of this procedure is provided in \ref{sec:app:met:heat}.
+
+\nomenclature[A]{AB}{Auxiliar boiler}
+\nomenclature[A]{HVAC}{Heat, ventilation and air conditioning}
+\nomenclature[A]{PH}{Preheater}
+\nomenclature[A]{RH}{Reheater}
+\nomenclature[A]{HWH}{Hot water heater}
+\nomenclature[A]{TH}{Fuel Tank heating}
+\nomenclature[A]{OT}{Other tanks heating}
+\nomenclature[A]{MSH}{Machinery space heating}
+\nomenclature[A]{HH}{HFO heater}
+\nomenclature[A]{HTH}{HFO tank heating}
+\nomenclature[A]{EGB}{Exhaust gas boilers}
+\nomenclature[A]{HTHR}{High temperature heat recovery}
 
@@
 The approach applied by Fazlollahi et al.
-\cite{Fazlollah2014} was used to select the optimum number of clusters (Nk).
-It (Nk) should be as low as possible for data handling purposes, while preserving a high accuracy of the retrieved data.
+\cite{Fazlollah2014} was used to select the optimum number of clusters ($N_k$).
+It ($N_k$) should be as low as possible for data handling purposes, while preserving a high accuracy of the retrieved data.
 It builds on the calculation of three criteria:
@@
 \end{enumerate}
+
+\nomenclature[V]{$\Delta_{prod}$}{Number of periods with relative errors}
+\nomenclature[V]{$\sigma_{profile}$}{Profile deviation}
+\nomenclature[V]{$\sigma_{CDC}$}{Deviation from the load duration curve}
+\nomenclature[V]{$ELDC$}{Relative error in load duration curve deviation}
+\nomenclature[V]{$\Delta_{MLDC}$}{Maximum duration load curve difference}
 
@@
 
+\nomenclature[A]{APCA}{Adaptive piecewise constant approximation}
+\nomenclature[V]{$N_k$}{Number of clusters}
+
 \section{Results and discussion} \label{sec:results}
@@
 	\includegraphics[width=0.99\linewidth]{Figures/Pie_HeatGeneration}
-	\caption{Share of thermal power generated from on board heat systems}
+	\caption{Share of thermal power generated from on-board heat systems}
 	\label{fig:Pie_HeatGeneration}
@@
 	\includegraphics[width=0.99\linewidth]{Figures/TimeSeries_HeatGeneration}
-	\caption{Time series representation of the thermal power generated from on board heat systems}
+	\caption{Time series representation of the thermal power generated from on-board heat systems}
 	\label{fig:TimeSeries_HeatGeneration}
@@
 
-The full Sankey diagram representation of the energy flows on board is provided in Figure \ref{fig:Sankey}.
+The full Sankey diagram representation of the energy flows on-board is provided in Figure \ref{fig:Sankey}.
 
@@
 As expected, the individual largest contribution to exergy destruction in the ship systems is associated to combustion, both in the engines and in the boilers.
-The loss of energy resulting from the conversion of chemical energy to heat contributes to more than 76\% of the total exergy destruction on board.
+The loss of energy resulting from the conversion of chemical energy to heat contributes to more than 76\% of the total exergy destruction on-board.
 Focusing on the remaining parts of the system, the largest sources of exergy destruction are located in the engine turbochargers (28.9\%), in the exhaust gas boilers (10.7\%), in the steam heater of the HTHR system (10.5\%), the HVAC preheater (10.4\%) and the sea water cooler (8.5\%).
@@
 This includes, for instance, the concept of using heat from the LT cooling systems (currently unused) for the HVAC preheater, hence reducing exergy destruction in the HVAC preheater itself and in the steam heater.
-This concept could be extended to other low-temperature heat consumers (such as the machinery space heating), but it would clearly constitute a larger benefit if applied to the HVAC preheater, which is the largest heat demand on board.
-
-The analysis also suggests that the electric power demand resulting from the use of the HVAC on board is relatively low, a consequence of the fact that the case study ship mostly operates in cold climates.
+This concept could be extended to other low-temperature heat consumers (such as the machinery space heating), but it would clearly constitute a larger benefit if applied to the HVAC preheater, which is the largest heat demand on-board.
+
+The analysis also suggests that the electric power demand resulting from the use of the HVAC on-board is relatively low, a consequence of the fact that the case study ship mostly operates in cold climates.
 This suggests that demand-related efforts should be focused on other consumers (e.g.
@@
 % What we did in this paper
-In this paper, we analyzed the energy system of a cruise ship operating in the Baltic Sea, with the aim of providing a better understanding of how energy is used on board, for what purposes, and with what efficiency.
+In this paper, we analyzed the energy system of a cruise ship operating in the Baltic Sea, with the aim of providing a better understanding of how energy is used on-board, for what purposes, and with what efficiency.
 Although these numbers can vary substantially depending on the ship type, size and operating region, the results presented in this paper represent a unique contribution to scientific literature in terms of the detail of the analysis and of the extent of the data available.
@@
 The application of exergy analysis allowed the identification of the main causes for losses of energy quality.
-In addition to the combustion processes in the main engines and in the boilers, that together contribute to 77\% of the exergy destruction on board, the main potential for improvement was identified in the turbochargers, the exhaust gas boilers (HRSGs), the steam heater in the hot water heat recovery systems, in the HVAC pre-heater and in the sea water cooler.
+In addition to the combustion processes in the main engines and in the boilers, that together contribute to 77\% of the exergy destruction on-board, the main potential for improvement was identified in the turbochargers, the exhaust gas boilers (HRSGs), the steam heater in the hot water heat recovery systems, in the HVAC pre-heater and in the sea water cooler.
 The main basic causes of exergy destruction are hence to be considered the low-load operations of the engines and the mismatch in the operating temperatures of the heat exchangers.
@@
 
+\bibliographystyle{unsrt}
+\bibliography{bibliography}
+
 \appendix
 
-\section{Extended method description}
+\section{Full tables of energy and exergy flows} \label{sec:app:flows}
+
+In this appendix, the cumulated yearly values for the main ship energy and exergy flows are provided.
+Apart from yearly values, we also provide the same statistics separated based on the ship operational mode (P/S: Port stay, LSS: Low speed sailing, Man: Maneuvering, HSS, High speed sailing) and based on the season (Win: Winter, Mid: Spring/autumn, Sum: Summer).
+
+\subsection{Energy flows} \label{sec:app:flo:energy}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+			\midrule
+E & Env. & ME & Comp. & Ph & 4.57 & 0.005 & 3.819 & 0.385 & 0.359 & 2.327 & 1.671 & 0.571\\
+ME & Comp. & ME & CAC-HT & Ph & 19.27 & 0.010 & 14.412 & 1.645 & 3.208 & 7.494 & 7.822 & 3.958\\
+ME & CAC-HT & ME & CAC-LT & Ph & 13.98 & 0.010 & 10.900 & 1.190 & 1.885 & 5.717 & 5.728 & 2.541\\
+ME & CAC-LT & ME & Cyl. & Ph & 7.35 & 0.009 & 5.951 & 0.655 & 0.737 & 3.340 & 2.969 & 1.043\\
+ME & Turb. & ME & Comp. & W & 20.16 & 0.006 & 15.287 & 1.659 & 3.207 & 7.539 & 8.523 & 4.098\\
+ME & Cyl. & ME & Turb. & Ph & 103.85 & 0.093 & 82.544 & 9.285 & 11.924 & 41.276 & 43.973 & 18.597\\
+ME & Turb. & E & HRSG & Ph & 44.77 & 0.044 & 36.434 & 3.974 & 4.318 & 18.383 & 18.840 & 7.546\\
+ME & Turb. & E & Losses & Q & 3.32 & 0.003 & 2.652 & 0.289 & 0.378 & 1.307 & 1.425 & 0.590\\
+ME & Comp. & ME & BP valve & Ph & 24.73 & 0.011 & 19.106 & 2.045 & 3.567 & 9.866 & 10.194 & 4.669\\
+ME & Bypass valve & ME & Turb. & Ph & 5.45 & 0.002 & 4.694 & 0.399 & 0.359 & 2.372 & 2.372 & 0.711\\
+ME & CAC-HT & CS & HT Cool. & Q & 5.29 & 0.000 & 3.512 & 0.455 & 1.323 & 1.778 & 2.094 & 1.418\\
+ME & CAC-LT & CS & LT Cool. & Q & 6.63 & 0.001 & 4.949 & 0.535 & 1.147 & 2.377 & 2.758 & 1.497\\
+ME & Cyl. & ME & LOC & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
+ME & Cyl. & ME & JWC & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
+ME & JWC & CS & HT Cool. & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
+ME & LOC & CS & LT Cool. & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
+ME & Cyl. & E & Losses & Q & 2.59 & 0.007 & 2.133 & 0.280 & 0.172 & 1.054 & 1.143 & 0.396\\
+ME & HRSG & S & Boiler & Q & 13.41 & 0.018 & 11.130 & 1.277 & 0.987 & 5.810 & 5.439 & 2.163\\
+ME & HRSG & E & Env. & Ph & 31.36 & 0.025 & 25.304 & 2.697 & 3.331 & 12.573 & 13.402 & 5.383\\
+ME & Cyl. & P & Prop. sh. & W & 107.39 & 0.141 & 85.158 & 10.141 & 11.947 & 42.238 & 45.916 & 19.233\\
+ME & Fuel tank & ME & Cyl. & Chl & 250.62 & 0.360 & 199.949 & 23.977 & 26.331 & 99.059 & 107.585 & 43.973\\
+ME & Fuel tank & ME & Cyl. & Q & 1.09 & 0.002 & 0.875 & 0.102 & 0.109 & 0.455 & 0.458 & 0.174\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of energy flows for the main engines}
+\label{tab:flows:en:ME}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+			\midrule
+E & Env. & AE & Comp. & Ph & 0.42 & 0.13 & 0.23 & 0.05 & 0.01 & 0.25 & 0.09 & 0.08\\
+AE & Comp. & AE & CAC-HT & Ph & 9.96 & 3.64 & 5.08 & 0.89 & 0.35 & 4.34 & 3.85 & 1.76\\
+AE & CAC-HT & AE & CAC-LT & Ph & 8.91 & 3.13 & 4.68 & 0.83 & 0.27 & 3.99 & 3.54 & 1.38\\
+AE & CAC-LT & AE & Cyl. & Ph & 5.62 & 1.76 & 3.13 & 0.59 & 0.14 & 2.75 & 2.20 & 0.67\\
+AE & Turb. & AE & Comp. & W & 9.54 & 3.50 & 4.85 & 0.84 & 0.34 & 4.09 & 3.76 & 1.69\\
+AE & Cyl. & AE & Turb. & Ph & 70.69 & 21.90 & 39.33 & 7.46 & 2.00 & 30.33 & 29.78 & 10.57\\
+AE & Turb. & E & HRSG & Ph & 56.90 & 17.21 & 32.13 & 6.00 & 1.57 & 24.42 & 24.18 & 8.30\\
+AE & Turb. & E & Losses & Q & 1.89 & 0.54 & 1.05 & 0.25 & 0.05 & 0.80 & 0.82 & 0.27\\
+AE & Comp. & AE & BP valve & Ph & 9.96 & 3.64 & 5.08 & 0.89 & 0.35 & 4.34 & 3.85 & 1.76\\
+AE & Bypass valve & AE & Turb. & Ph & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
+AE & CAC-HT & CS & HT Cool. & Q & 1.04 & 0.51 & 0.41 & 0.06 & 0.07 & 0.34 & 0.31 & 0.39\\
+AE & CAC-LT & CS & LT Cool. & Q & 3.29 & 1.37 & 1.55 & 0.24 & 0.13 & 1.24 & 1.34 & 0.70\\
+AE & Cyl. & AE & LOC & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
+AE & Cyl. & AE & JWC & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
+AE & JWC & CS & HT Cool. & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
+AE & LOC & CS & LT Cool. & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
+AE & Cyl. & E & Losses & Q & 1.50 & 0.39 & 0.88 & 0.19 & 0.04 & 0.64 & 0.67 & 0.20\\
+AE & HRSG & S & Boiler & Q & 18.30 & 5.43 & 10.43 & 1.99 & 0.45 & 8.08 & 7.49 & 2.73\\
+AE & HRSG & E & Env. & Ph & 38.60 & 11.78 & 21.69 & 4.01 & 1.11 & 16.34 & 16.68 & 5.58\\
+AE & Cyl. & EL & Switchboard & W & 62.02 & 20.78 & 34.33 & 6.25 & 1.85 & 26.97 & 26.34 & 9.89\\
+AE & Cyl. & E & AG losses & Q & 3.47 & 0.93 & 2.03 & 0.43 & 0.09 & 1.47 & 1.53 & 0.47\\
+AE & Fuel tank & AE & Cyl. & Ch & 162.13 & 51.30 & 89.38 & 16.88 & 4.57 & 68.96 & 68.53 & 24.64\\
+AE & Fuel tank & AE & Cyl. & Q & 0.62 & 0.15 & 0.38 & 0.06 & 0.02 & 0.28 & 0.25 & 0.08\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of energy flows for the auxiliary engines}
+\label{tab:flows:en:AE}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+			\midrule
+CS & HT Cool. & CS & LT Cool. & Q & 31.84 & 3.48 & 21.94 & 3.57 & 2.85 & 12.29 & 13.62 & 5.93\\
+CS & HT Cool. & S & HRHT & Q & 16.33 & 2.39 & 11.60 & 1.59 & 0.75 & 7.15 & 7.01 & 2.17\\
+S & Boiler & S & HRHT & Q & 31.06 & 14.15 & 14.50 & 2.15 & 0.26 & 18.16 & 11.25 & 1.66\\
+S & Aux Boiler & S & Boiler & Ph & 16.22 & 13.56 & 2.32 & 0.43 & 0.09 & 10.50 & 5.42 & 0.30\\
+S & Aux Boiler & E & Env. & Ph & 4.05 & 2.83 & 1.00 & 0.22 & 0.00 & 2.62 & 1.35 & 0.07\\
+S & Fuel Tank & S & Aux Boiler & Ch. & 20.32 & 14.19 & 5.04 & 1.08 & 0.01 & 13.13 & 6.81 & 0.37\\
+S & HRHT & D & Preheater & Q & 34.51 & 12.33 & 19.12 & 2.56 & 0.50 & 21.02 & 13.49 & 0.00\\
+S & HRHT & D & Reheater & Q & 1.83 & 0.60 & 0.86 & 0.19 & 0.18 & 0.00 & 0.00 & 1.83\\
+S & HRHT & D & HWH & Q & 11.05 & 3.60 & 6.12 & 1.00 & 0.33 & 4.29 & 4.76 & 2.00\\
+S & Boiler & D & MSH & Q & 1.99 & 0.69 & 1.07 & 0.17 & 0.06 & 0.81 & 0.90 & 0.28\\
+S & Boiler & D & HFO TH & Q & 1.92 & 0.67 & 1.03 & 0.16 & 0.05 & 0.79 & 0.86 & 0.27 \\ 
+S & Boiler & D & HFO H & Q & 1.29 & 0.16 & 0.90 & 0.13 & 0.10 & 0.52 & 0.55 & 0.21\\
+S & Boiler & D & TH & Q & 1.49 & 0.52 & 0.80 & 0.13 & 0.04 & 0.61 & 0.67 & 0.21\\
+S & Boiler & D & G & Q & 7.11 & 1.96 & 4.33 & 0.64 & 0.17 & 2.97 & 3.19 & 0.95\\
+S & Boiler & D & OT & Q & 1.49 & 0.52 & 0.80 & 0.13 & 0.04 & 0.61 & 0.67 & 0.21\\
+EL & Switchboard & EL & Thrusters & W & 0.94 & 0.00 & 0.00 & 0.94 & 0.00 & 0.36 & 0.43 & 0.15\\
+EL & Switchboard & EL & HVAC & W & 1.79 & 0.59 & 0.85 & 0.18 & 0.18 & 0.00 & 0.00 & 1.79\\
+EL & Switchboard & EL & Other el. dem. & W & 59.29 & 19.99 & 32.66 & 4.99 & 1.64 & 26.06 & 25.45 & 7.78\\
+P & Prop. sh. shaft & P & Prop. sh. & W & 104.19 & 0.14 & 82.62 & 9.84 & 11.59 & 40.98 & 44.55 & 18.66\\
+P & Prop. sh. shaft & E & Losses & Q & 3.20 & 0.00 & 2.54 & 0.30 & 0.36 & 1.26 & 1.37 & 0.57\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of energy flows for the boiler/steam systems}
+\label{tab:flows:en:S}
+\end{table}
+
+\clearpage
+
+\subsection{Exergy flows}  \label{sec:app:flo:exergy}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+\midrule
+E & Env. & ME & Comp. & Ph & 0.16 & 0.000 & 0.141 & 0.013 & 0.010 & 0.102 & 0.048 & 0.014\\
+ME & Comp. & ME & CAC-HT & Ph & 13.64 & 0.004 & 9.932 & 1.148 & 2.555 & 4.976 & 5.626 & 3.037\\
+ME & CAC-HT & ME & CAC-LT & Ph & 12.20 & 0.004 & 9.008 & 1.023 & 2.164 & 4.476 & 5.079 & 2.643\\
+ME & CAC-LT & ME & Cyl. & Ph & 11.04 & 0.004 & 8.144 & 0.932 & 1.960 & 4.019 & 4.619 & 2.401\\
+ME & Turb. & ME & Comp. & W & 20.16 & 0.006 & 15.287 & 1.659 & 3.207 & 7.539 & 8.523 & 4.098\\
+ME & Cyl. & ME & Turb. & Ph & 55.91 & 0.037 & 43.951 & 4.839 & 7.085 & 22.177 & 23.408 & 10.327\\
+ME & Turb. & E & HRSG & Ph & 15.45 & 0.014 & 12.621 & 1.353 & 1.458 & 6.605 & 6.320 & 2.520\\
+ME & Turb. & E & Losses & Q & 0.81 & 0.001 & 0.653 & 0.069 & 0.087 & 0.341 & 0.338 & 0.131\\
+ME & Turb. & E & Env. & Ph & 14.04 & 0.014 & 11.256 & 1.285 & 1.489 & 5.694 & 5.941 & 2.409\\
+ME & Comp. & ME & BP valve & Ph & 17.44 & 0.004 & 13.179 & 1.425 & 2.831 & 6.557 & 7.323 & 3.559\\
+ME & BP valve & ME & Turb. & Ph & 3.80 & 0.001 & 3.246 & 0.277 & 0.275 & 1.581 & 1.697 & 0.522\\
+ME & CAC-HT & CS & HT Cool. & Q & 1.16 & 0.000 & 0.924 & 0.125 & 0.392 & 0.501 & 0.547 & 0.394\\
+ME & CAC-LT & CS & LT Cool. & Q & 0.59 & 0.000 & 0.864 & 0.091 & 0.203 & 0.456 & 0.460 & 0.242\\
+ME & Cyl. & ME & LOC & Q & 4.18 & 0.011 & 3.442 & 0.443 & 0.286 & 1.860 & 1.758 & 0.565\\
+ME & Cyl. & ME & JWC & Q & 8.13 & 0.022 & 6.660 & 0.879 & 0.565 & 3.424 & 3.514 & 1.188\\
+ME & JWC & CS & HT Cool. & Q & 5.30 & 0.022 & 6.660 & 0.879 & 0.565 & 3.424 & 3.514 & 1.188\\
+ME & LOC & CS & LT Cool. & Q & 2.45 & 0.011 & 3.442 & 0.443 & 0.286 & 1.860 & 1.758 & 0.565\\
+ME & Cyl. & E & Losses & Q & 0.63 & 0.002 & 0.525 & 0.067 & 0.040 & 0.275 & 0.271 & 0.088\\
+ME & HRSG & S & Boiler & Q & 4.63 & 0.008 & 5.622 & 0.628 & 0.498 & 3.015 & 2.677 & 1.064\\
+ME & HRSG & E & Env. & Ph & 8.69 & 0.006 & 6.999 & 0.725 & 0.961 & 3.590 & 3.643 & 1.456\\
+ME & Cyl. & P & Prop. sh. & W & 107.39 & 0.141 & 85.158 & 10.141 & 11.947 & 42.238 & 45.916 & 19.233\\
+ME & Fuel tank & ME & Cyl. & Ch & 266.02 & 0.382 & 212.237 & 25.450 & 27.949 & 105.146 & 114.197 & 46.675\\
+ME & Fuel tank & ME & Cyl. & Q & 0.15 & 0.000 & 0.125 & 0.014 & 0.014 & 0.068 & 0.063 & 0.022\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of exergy flows for the main engines}
+\label{tab:flows:ex:ME}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+			\midrule
+E & Env. & AE & Comp. & Ph & 0.01 & 0.002 & 0.003 & 0.001 & 0.000 & 0.003 & 0.001 & 0.001\\
+AE & Comp. & AE & CAC-HT & Ph & 7.91 & 2.960 & 4.008 & 0.659 & 0.288 & 3.392 & 3.091 & 1.432\\
+AE & CAC-HT & AE & CAC-LT & Ph & 7.65 & 2.835 & 3.906 & 0.644 & 0.269 & 3.302 & 3.017 & 1.335\\
+AE & CAC-LT & AE & Cyl. & Ph & 7.09 & 2.589 & 3.650 & 0.606 & 0.248 & 3.066 & 2.802 & 1.227\\
+AE & Turb. & AE & Comp. & W & 9.54 & 3.504 & 4.853 & 0.841 & 0.338 & 4.087 & 3.762 & 1.687\\
+AE & Cyl. & AE & Turb. & Ph & 36.26 & 11.776 & 19.815 & 3.592 & 1.079 & 15.705 & 14.961 & 5.597\\
+AE & Turb. & E & HRSG & Ph & 21.04 & 6.378 & 11.936 & 2.157 & 0.571 & 9.166 & 8.866 & 3.011\\
+AE & Turb. & E & Losses & Q & 0.46 & 0.132 & 0.259 & 0.060 & 0.012 & 0.209 & 0.194 & 0.060\\
+AE & Comp. & AE & BP valve & Ph & 7.91 & 2.960 & 4.008 & 0.659 & 0.288 & 3.392 & 3.091 & 1.432\\
+AE & BP valve & AE & Turb. & Ph & 0.00 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
+AE & CAC-HT & CS & HT Cool. & Q & 0.24 & 0.125 & 0.102 & 0.015 & 0.018 & 0.090 & 0.074 & 0.097\\
+AE & CAC-LT & CS & LT Cool. & Q & 0.36 & 0.245 & 0.256 & 0.037 & 0.021 & 0.236 & 0.215 & 0.108\\
+AE & Cyl. & AE & LOC & Q & 3.14 & 0.962 & 1.757 & 0.344 & 0.075 & 1.476 & 1.275 & 0.388\\
+AE & Cyl. & AE & JWC & Q & 5.84 & 1.772 & 3.270 & 0.656 & 0.144 & 2.608 & 2.451 & 0.783\\
+AE & JWC & CS & HT Cool. & Q & 3.90 & 1.772 & 3.270 & 0.656 & 0.144 & 2.608 & 2.451 & 0.783\\
+AE & LOC & CS & LT Cool. & Q & 2.18 & 0.962 & 1.757 & 0.344 & 0.075 & 1.476 & 1.275 & 0.388\\
+AE & Cyl. & E & Losses & Q & 0.37 & 0.096 & 0.217 & 0.047 & 0.008 & 0.166 & 0.159 & 0.044\\
+AE & HRSG & S & Boiler & Q & 6.32 & 2.869 & 5.504 & 1.011 & 0.237 & 4.309 & 3.901 & 1.412\\
+AE & HRSG & E & Env. & Ph & 11.42 & 3.509 & 6.432 & 1.145 & 0.334 & 4.857 & 4.965 & 1.599\\
+AE & Cyl. & EL & SB & W & 62.02 & 20.781 & 34.328 & 6.249 & 1.851 & 26.971 & 26.344 & 9.894\\
+AE & Cyl. & E & AG losses & Q & 0.85 & 0.228 & 0.499 & 0.105 & 0.020 & 0.384 & 0.363 & 0.104\\
+AE & Fuel tank & AE & Cyl. & Ch & 172.09 & 54.452 & 94.876 & 17.916 & 4.849 & 73.199 & 72.744 & 26.150\\
+AE & Fuel tank & AE & Cyl. & Q & 0.08 & 0.019 & 0.054 & 0.009 & 0.003 & 0.041 & 0.033 & 0.010\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of exergy flows for the auxiliary engines}
+\label{tab:flows:ex:AE}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+			\midrule
+CS & HT Cool. & CS & LT Cool. & Q & 7.36 & 1.46 & 8.63 & 1.36 & 0.97 & 5.10 & 5.25 & 2.08\\
+CS & HT Cool. & S & HRHT & Q & 3.24 & 0.46 & 2.33 & 0.31 & 0.14 & 1.53 & 1.34 & 0.38\\
+S & Boiler & S & HRHT & Q & 10.87 & 4.94 & 5.09 & 0.75 & 0.09 & 6.51 & 3.82 & 0.53\\
+S & Aux Boiler & S & Boiler & Ph & 5.71 & 4.66 & 0.90 & 0.17 & -0.03 & 3.77 & 1.85 & 0.10\\
+S & Aux Boiler & E & Env. & Ph & 1.51 & 1.04 & 0.38 & 0.08 & 0.00 & 0.99 & 0.50 & 0.03\\
+S & Fuel tank & S & Aux Boiler & Ch & 20.12 & 14.37 & 4.73 & 1.02 & 0.01 & 12.99 & 6.76 & 0.37\\
+S & HRHT & D & HVAC-PH & Q & 7.43 & 1.22 & 1.89 & 0.25 & 0.05 & 2.29 & 1.11 & 0.00\\
+S & HRHT & D & HVAC-RH & Q & 0.33 & 0.09 & 0.13 & 0.03 & 0.03 & 0.00 & 0.00 & 0.28\\
+S & HRHT & D & HWH  & Q & 2.27 & 0.64 & 1.09 & 0.17 & 0.05 & 0.84 & 0.81 & 0.31\\
+S & Boiler & D & MSH & Q & 0.68 & 0.06 & 0.10 & 0.01 & 0.00 & 0.09 & 0.07 & 0.02\\
+S & Boiler & D & HFO TH & Q & 0.66 & 0.13 & 0.20 & 0.03 & 0.01 & 0.16 & 0.16 & 0.04\\
+S & Boiler & D & HFO H & Q & 0.44 & 0.04 & 0.21 & 0.03 & 0.02 & 0.13 & 0.12 & 0.04\\
+S & Boiler & D & TH & Q & 0.51 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
+S & Boiler & D & G & Q & 2.45 & 0.44 & 0.98 & 0.14 & 0.04 & 0.71 & 0.69 & 0.19\\
+S & Boiler & D & OT & Q & 0.34 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
+EL & SB & EL & Thrusters & W & 0.94 & 0.00 & 0.00 & 0.94 & 0.00 & 0.36 & 0.43 & 0.15\\
+EL & SB & EL & HVAC & W & 1.79 & 0.59 & 0.85 & 0.18 & 0.18 & 0.00 & 0.00 & 1.79\\
+EL & SB & EL & Other el. Dem. & W & 59.29 & 19.99 & 32.66 & 4.99 & 1.64 & 26.06 & 25.45 & 7.78\\
+P & Prop. shaft & P & Prop. & W & 104.19 & 0.14 & 82.62 & 9.84 & 11.59 & 40.98 & 44.55 & 18.66\\
+P & Prop. shaft & E & Losses & Q & 3.20 & 0.00 & 2.54 & 0.30 & 0.36 & 1.26 & 1.37 & 0.57\\
+S & Boiler & E & Dumped steam & Q & 0.81 & 0.00 & 0.43 & 0.13 & 0.25 & 0.05 & 0.20 & 0.55\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of exergy flows for the boiler and steam systems}
+\label{tab:flows:ex:S}
+\end{table}
+
+\begin{table}[hbp!]
+	\centering
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{clclcrrrrrrrr}
+			\toprule
+			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
+			\midrule
+D& HVAC-PH & D & Dem. & Q & 3.41 & 1.22 & 1.89 & 0.25 & 0.05 & 2.29 & 1.11 & 0.00\\
+D& HVAC-RH & D & Dem. & Q & 0.28 & 0.09 & 0.13 & 0.03 & 0.03 & 0.00 & 0.00 & 0.28\\
+D& HWH  & D & Dem. & Q & 1.96 & 0.64 & 1.09 & 0.17 & 0.05 & 0.84 & 0.81 & 0.31\\
+D& MSH & D & Dem. & Q & 0.18 & 0.06 & 0.10 & 0.01 & 0.00 & 0.09 & 0.07 & 0.02\\
+D& HFO TH & D & Dem. & Q & 0.37 & 0.13 & 0.20 & 0.03 & 0.01 & 0.16 & 0.16 & 0.04\\
+D& HFO H & D & Dem. & Q & 0.29 & 0.04 & 0.21 & 0.03 & 0.02 & 0.13 & 0.12 & 0.04\\
+D& TH & D & Dem. & Q & 0.23 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
+D& G & D & Dem. & Q & 1.59 & 0.44 & 0.98 & 0.14 & 0.04 & 0.71 & 0.69 & 0.19\\
+D& OT & D & Dem. & Q & 0.15 & 0.05 & 0.08 & 0.01 & 0.00 & 0.07 & 0.07 & 0.02\\
+\bottomrule
+\end{tabular}}
+\caption{Summary of exergy flows for the heating demand}
+\label{tab:flows:ex:D}
+\end{table}
+
+\clearpage
+
+\section{Energy demand for typical operational days} \label{sec:app:typicalDays}
+
+In this section, the values for the main energy demands (Propulsion power, electric power, high temperature heat and low temperature heat) are provided for the typical days identified in the demand clustering.
+
+\clearpage
+
+\section{Extended method description} \label{sec:app:method}
 
@@
 
-\subsection{Data gathering and pre-processing} \label{sec:met:gathering}
-
-The operational data was collected on board from the ships' machine logging and surveillance system.
-The on board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB.
+\subsection{Data gathering and pre-processing} \label{sec:app:met:gathering}
+
+The operational data was collected on-board from the ships' machine logging and surveillance system.
+The on-board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB.
 The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average.
@@
 
-\subsection{Diesel engines modeling} \label{sec:appendix:engines}
+\subsection{Diesel engines modeling} \label{sec:app:met:engines}
 
 In the case of the main engines, the engine power output was not measured and needed to be estimated.
-In this work, we calculated the engine power output based on measurements of engine fuel rack position (used as a proxy of the mass of fuel injected per cycle) and of the engine speed (see equations \ref{eqn:ME:eq1} to \ref{eqn:ME:eq3}).
+In this work, we calculated the engine power output based on measurements of engine fuel rack position (used as a proxy of the mass of fuel injected per cycle) and of the engine speed (see equations \ref{eqn:ME:eq1:app} to \ref{eqn:ME:eq3:app}).
 In the case of the auxiliary engines, instead, direct measurements of the power generated by each engine were available.
 \begin{eqnarray}
-\dot{m}_{fuel} & = & \dot{m}_{fuel,des} \left(a_0 + a_1 \frac{frp}{frp_{des}}\right) \frac{\omega}{\omega_{des}} \label{eqn:ME:eq1} \\
-\eta_{ME} & = & a_0 + a_1 \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} + a_2 \left( \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} \right)^2 \label{eqn:ME:eq2} \\
-\dot{W}_{ME} & = & \dot{m}_{fuel} \eta_{ME} LHV \label{eqn:ME:eq3}
+\dot{m}_{fuel} & = & \dot{m}_{fuel,des} \left(a_0 + a_1 \frac{frp}{frp_{des}}\right) \frac{\omega}{\omega_{des}} \label{eqn:ME:eq1:app} \\
+\eta_{ME} & = & a_0 + a_1 \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} + a_2 \left( \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} \right)^2 \label{eqn:ME:eq2:app} \\
+\dot{W}_{ME} & = & \dot{m}_{fuel} \eta_{ME} LHV \label{eqn:ME:eq3:app}
 \end{eqnarray}
@@
 
-\subsection{Electric demand modeling}
+\subsection{Electric demand modeling} \label{sec:app:met:electric}
 
@@
 
-\subsubsection{Heat demand}
+\subsection{Heat demand} \label{sec:app:met:heat}
 
@@
 
-\subsubsection{Full list of exchangers in the ship's heat exchanger network}
+\subsection{Full list of exchangers in the ship's heat exchanger network} \label{sec:app:met:hen}
 
@@
 
-\section{Full tables of energy and exergy flows} \label{sec:appendix:energy}
-
-In this appendix, the cumulated yearly values for the main ship energy and exergy flows are provided.
-Apart from yearly values, we also provide the same statistics separated based on the ship operational mode (P/S: Port stay, LSS: Low speed sailing, Man: Maneuvering, HSS, High speed sailing) and based on the season (Win: Winter, Mid: Spring/autumn, Sum: Summer).
-
-\subsection{Energy flows}
-
-\begin{table}[hbp!]
-	\centering
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{clclcrrrrrrrr}
-			\toprule
-			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
-			\midrule
-E & Env. & ME & Comp. & Ph & 4.57 & 0.005 & 3.819 & 0.385 & 0.359 & 2.327 & 1.671 & 0.571\\
-ME & Comp. & ME & CAC-HT & Ph & 19.27 & 0.010 & 14.412 & 1.645 & 3.208 & 7.494 & 7.822 & 3.958\\
-ME & CAC-HT & ME & CAC-LT & Ph & 13.98 & 0.010 & 10.900 & 1.190 & 1.885 & 5.717 & 5.728 & 2.541\\
-ME & CAC-LT & ME & Cyl. & Ph & 7.35 & 0.009 & 5.951 & 0.655 & 0.737 & 3.340 & 2.969 & 1.043\\
-ME & Turb. & ME & Comp. & W & 20.16 & 0.006 & 15.287 & 1.659 & 3.207 & 7.539 & 8.523 & 4.098\\
-ME & Cyl. & ME & Turb. & Ph & 103.85 & 0.093 & 82.544 & 9.285 & 11.924 & 41.276 & 43.973 & 18.597\\
-ME & Turb. & E & HRSG & Ph & 44.77 & 0.044 & 36.434 & 3.974 & 4.318 & 18.383 & 18.840 & 7.546\\
-ME & Turb. & E & Losses & Q & 3.32 & 0.003 & 2.652 & 0.289 & 0.378 & 1.307 & 1.425 & 0.590\\
-ME & Comp. & ME & BP valve & Ph & 24.73 & 0.011 & 19.106 & 2.045 & 3.567 & 9.866 & 10.194 & 4.669\\
-ME & Bypass valve & ME & Turb. & Ph & 5.45 & 0.002 & 4.694 & 0.399 & 0.359 & 2.372 & 2.372 & 0.711\\
-ME & CAC-HT & CS & HT Cool. & Q & 5.29 & 0.000 & 3.512 & 0.455 & 1.323 & 1.778 & 2.094 & 1.418\\
-ME & CAC-LT & CS & LT Cool. & Q & 6.63 & 0.001 & 4.949 & 0.535 & 1.147 & 2.377 & 2.758 & 1.497\\
-ME & Cyl. & ME & LOC & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
-ME & Cyl. & ME & JWC & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
-ME & JWC & CS & HT Cool. & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
-ME & LOC & CS & LT Cool. & Q & 24.36 & 0.067 & 19.869 & 2.665 & 1.763 & 9.835 & 10.735 & 3.793\\
-ME & Cyl. & E & Losses & Q & 2.59 & 0.007 & 2.133 & 0.280 & 0.172 & 1.054 & 1.143 & 0.396\\
-ME & HRSG & S & Boiler & Q & 13.41 & 0.018 & 11.130 & 1.277 & 0.987 & 5.810 & 5.439 & 2.163\\
-ME & HRSG & E & Env. & Ph & 31.36 & 0.025 & 25.304 & 2.697 & 3.331 & 12.573 & 13.402 & 5.383\\
-ME & Cyl. & P & Prop. sh. & W & 107.39 & 0.141 & 85.158 & 10.141 & 11.947 & 42.238 & 45.916 & 19.233\\
-ME & Fuel tank & ME & Cyl. & Chl & 250.62 & 0.360 & 199.949 & 23.977 & 26.331 & 99.059 & 107.585 & 43.973\\
-ME & Fuel tank & ME & Cyl. & Q & 1.09 & 0.002 & 0.875 & 0.102 & 0.109 & 0.455 & 0.458 & 0.174\\
-\bottomrule
-\end{tabular}}
-\caption{Summary of energy flows for the main engines}
-\label{tab:flows:en:ME}
-\end{table}
-
-\begin{table}[hbp!]
-	\centering
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{clclcrrrrrrrr}
-			\toprule
-			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
-			\midrule
-E & Env. & AE & Comp. & Ph & 0.42 & 0.13 & 0.23 & 0.05 & 0.01 & 0.25 & 0.09 & 0.08\\
-AE & Comp. & AE & CAC-HT & Ph & 9.96 & 3.64 & 5.08 & 0.89 & 0.35 & 4.34 & 3.85 & 1.76\\
-AE & CAC-HT & AE & CAC-LT & Ph & 8.91 & 3.13 & 4.68 & 0.83 & 0.27 & 3.99 & 3.54 & 1.38\\
-AE & CAC-LT & AE & Cyl. & Ph & 5.62 & 1.76 & 3.13 & 0.59 & 0.14 & 2.75 & 2.20 & 0.67\\
-AE & Turb. & AE & Comp. & W & 9.54 & 3.50 & 4.85 & 0.84 & 0.34 & 4.09 & 3.76 & 1.69\\
-AE & Cyl. & AE & Turb. & Ph & 70.69 & 21.90 & 39.33 & 7.46 & 2.00 & 30.33 & 29.78 & 10.57\\
-AE & Turb. & E & HRSG & Ph & 56.90 & 17.21 & 32.13 & 6.00 & 1.57 & 24.42 & 24.18 & 8.30\\
-AE & Turb. & E & Losses & Q & 1.89 & 0.54 & 1.05 & 0.25 & 0.05 & 0.80 & 0.82 & 0.27\\
-AE & Comp. & AE & BP valve & Ph & 9.96 & 3.64 & 5.08 & 0.89 & 0.35 & 4.34 & 3.85 & 1.76\\
-AE & Bypass valve & AE & Turb. & Ph & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
-AE & CAC-HT & CS & HT Cool. & Q & 1.04 & 0.51 & 0.41 & 0.06 & 0.07 & 0.34 & 0.31 & 0.39\\
-AE & CAC-LT & CS & LT Cool. & Q & 3.29 & 1.37 & 1.55 & 0.24 & 0.13 & 1.24 & 1.34 & 0.70\\
-AE & Cyl. & AE & LOC & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
-AE & Cyl. & AE & JWC & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
-AE & JWC & CS & HT Cool. & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
-AE & LOC & CS & LT Cool. & Q & 17.47 & 5.30 & 9.75 & 1.98 & 0.44 & 7.49 & 7.48 & 2.50\\
-AE & Cyl. & E & Losses & Q & 1.50 & 0.39 & 0.88 & 0.19 & 0.04 & 0.64 & 0.67 & 0.20\\
-AE & HRSG & S & Boiler & Q & 18.30 & 5.43 & 10.43 & 1.99 & 0.45 & 8.08 & 7.49 & 2.73\\
-AE & HRSG & E & Env. & Ph & 38.60 & 11.78 & 21.69 & 4.01 & 1.11 & 16.34 & 16.68 & 5.58\\
-AE & Cyl. & EL & Switchboard & W & 62.02 & 20.78 & 34.33 & 6.25 & 1.85 & 26.97 & 26.34 & 9.89\\
-AE & Cyl. & E & AG losses & Q & 3.47 & 0.93 & 2.03 & 0.43 & 0.09 & 1.47 & 1.53 & 0.47\\
-AE & Fuel tank & AE & Cyl. & Ch & 162.13 & 51.30 & 89.38 & 16.88 & 4.57 & 68.96 & 68.53 & 24.64\\
-AE & Fuel tank & AE & Cyl. & Q & 0.62 & 0.15 & 0.38 & 0.06 & 0.02 & 0.28 & 0.25 & 0.08\\
-\bottomrule
-\end{tabular}}
-\caption{Summary of energy flows for the auxiliary engines}
-\label{tab:flows:en:AE}
-\end{table}
-
-\begin{table}[hbp!]
-	\centering
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{clclcrrrrrrrr}
-			\toprule
-			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
-			\midrule
-CS & HT Cool. & CS & LT Cool. & Q & 31.84 & 3.48 & 21.94 & 3.57 & 2.85 & 12.29 & 13.62 & 5.93\\
-CS & HT Cool. & S & HRHT & Q & 16.33 & 2.39 & 11.60 & 1.59 & 0.75 & 7.15 & 7.01 & 2.17\\
-S & Boiler & S & HRHT & Q & 31.06 & 14.15 & 14.50 & 2.15 & 0.26 & 18.16 & 11.25 & 1.66\\
-S & Aux Boiler & S & Boiler & Ph & 16.22 & 13.56 & 2.32 & 0.43 & 0.09 & 10.50 & 5.42 & 0.30\\
-S & Aux Boiler & E & Env. & Ph & 4.05 & 2.83 & 1.00 & 0.22 & 0.00 & 2.62 & 1.35 & 0.07\\
-S & Fuel Tank & S & Aux Boiler & Ch. & 20.32 & 14.19 & 5.04 & 1.08 & 0.01 & 13.13 & 6.81 & 0.37\\
-S & HRHT & D & Preheater & Q & 34.51 & 12.33 & 19.12 & 2.56 & 0.50 & 21.02 & 13.49 & 0.00\\
-S & HRHT & D & Reheater & Q & 1.83 & 0.60 & 0.86 & 0.19 & 0.18 & 0.00 & 0.00 & 1.83\\
-S & HRHT & D & HWH & Q & 11.05 & 3.60 & 6.12 & 1.00 & 0.33 & 4.29 & 4.76 & 2.00\\
-S & Boiler & D & MSH & Q & 1.99 & 0.69 & 1.07 & 0.17 & 0.06 & 0.81 & 0.90 & 0.28\\
-S & Boiler & D & HFO TH & Q & 1.92 & 0.67 & 1.03 & 0.16 & 0.05 & 0.79 & 0.86 & 0.27 \\ 
-S & Boiler & D & HFO H & Q & 1.29 & 0.16 & 0.90 & 0.13 & 0.10 & 0.52 & 0.55 & 0.21\\
-S & Boiler & D & TH & Q & 1.49 & 0.52 & 0.80 & 0.13 & 0.04 & 0.61 & 0.67 & 0.21\\
-S & Boiler & D & G & Q & 7.11 & 1.96 & 4.33 & 0.64 & 0.17 & 2.97 & 3.19 & 0.95\\
-S & Boiler & D & OT & Q & 1.49 & 0.52 & 0.80 & 0.13 & 0.04 & 0.61 & 0.67 & 0.21\\
-EL & Switchboard & EL & Thrusters & W & 0.94 & 0.00 & 0.00 & 0.94 & 0.00 & 0.36 & 0.43 & 0.15\\
-EL & Switchboard & EL & HVAC & W & 1.79 & 0.59 & 0.85 & 0.18 & 0.18 & 0.00 & 0.00 & 1.79\\
-EL & Switchboard & EL & Other el. dem. & W & 59.29 & 19.99 & 32.66 & 4.99 & 1.64 & 26.06 & 25.45 & 7.78\\
-P & Prop. sh. shaft & P & Prop. sh. & W & 104.19 & 0.14 & 82.62 & 9.84 & 11.59 & 40.98 & 44.55 & 18.66\\
-P & Prop. sh. shaft & E & Losses & Q & 3.20 & 0.00 & 2.54 & 0.30 & 0.36 & 1.26 & 1.37 & 0.57\\
-\bottomrule
-\end{tabular}}
-\caption{Summary of energy flows for the boiler/steam systems}
-\label{tab:flows:en:S}
-\end{table}
-
-\clearpage
-
-\subsection{Exergy flows}   \label{sec:appendix:exergy}
-
-\begin{table}[hbp!]
-	\centering
-	\resizebox{\textwidth}{!}{%
-	\begin{tabular}{clclcrrrrrrrr}
-			\toprule
-\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
-\midrule
-E & Env. & ME & Comp. & Ph & 0.16 & 0.000 & 0.141 & 0.013 & 0.010 & 0.102 & 0.048 & 0.014\\
-ME & Comp. & ME & CAC-HT & Ph & 13.64 & 0.004 & 9.932 & 1.148 & 2.555 & 4.976 & 5.626 & 3.037\\
-ME & CAC-HT & ME & CAC-LT & Ph & 12.20 & 0.004 & 9.008 & 1.023 & 2.164 & 4.476 & 5.079 & 2.643\\
-ME & CAC-LT & ME & Cyl. & Ph & 11.04 & 0.004 & 8.144 & 0.932 & 1.960 & 4.019 & 4.619 & 2.401\\
-ME & Turb. & ME & Comp. & W & 20.16 & 0.006 & 15.287 & 1.659 & 3.207 & 7.539 & 8.523 & 4.098\\
-ME & Cyl. & ME & Turb. & Ph & 55.91 & 0.037 & 43.951 & 4.839 & 7.085 & 22.177 & 23.408 & 10.327\\
-ME & Turb. & E & HRSG & Ph & 15.45 & 0.014 & 12.621 & 1.353 & 1.458 & 6.605 & 6.320 & 2.520\\
-ME & Turb. & E & Losses & Q & 0.81 & 0.001 & 0.653 & 0.069 & 0.087 & 0.341 & 0.338 & 0.131\\
-ME & Turb. & E & Env. & Ph & 14.04 & 0.014 & 11.256 & 1.285 & 1.489 & 5.694 & 5.941 & 2.409\\
-ME & Comp. & ME & BP valve & Ph & 17.44 & 0.004 & 13.179 & 1.425 & 2.831 & 6.557 & 7.323 & 3.559\\
-ME & BP valve & ME & Turb. & Ph & 3.80 & 0.001 & 3.246 & 0.277 & 0.275 & 1.581 & 1.697 & 0.522\\
-ME & CAC-HT & CS & HT Cool. & Q & 1.16 & 0.000 & 0.924 & 0.125 & 0.392 & 0.501 & 0.547 & 0.394\\
-ME & CAC-LT & CS & LT Cool. & Q & 0.59 & 0.000 & 0.864 & 0.091 & 0.203 & 0.456 & 0.460 & 0.242\\
-ME & Cyl. & ME & LOC & Q & 4.18 & 0.011 & 3.442 & 0.443 & 0.286 & 1.860 & 1.758 & 0.565\\
-ME & Cyl. & ME & JWC & Q & 8.13 & 0.022 & 6.660 & 0.879 & 0.565 & 3.424 & 3.514 & 1.188\\
-ME & JWC & CS & HT Cool. & Q & 5.30 & 0.022 & 6.660 & 0.879 & 0.565 & 3.424 & 3.514 & 1.188\\
-ME & LOC & CS & LT Cool. & Q & 2.45 & 0.011 & 3.442 & 0.443 & 0.286 & 1.860 & 1.758 & 0.565\\
-ME & Cyl. & E & Losses & Q & 0.63 & 0.002 & 0.525 & 0.067 & 0.040 & 0.275 & 0.271 & 0.088\\
-ME & HRSG & S & Boiler & Q & 4.63 & 0.008 & 5.622 & 0.628 & 0.498 & 3.015 & 2.677 & 1.064\\
-ME & HRSG & E & Env. & Ph & 8.69 & 0.006 & 6.999 & 0.725 & 0.961 & 3.590 & 3.643 & 1.456\\
-ME & Cyl. & P & Prop. sh. & W & 107.39 & 0.141 & 85.158 & 10.141 & 11.947 & 42.238 & 45.916 & 19.233\\
-ME & Fuel tank & ME & Cyl. & Ch & 266.02 & 0.382 & 212.237 & 25.450 & 27.949 & 105.146 & 114.197 & 46.675\\
-ME & Fuel tank & ME & Cyl. & Q & 0.15 & 0.000 & 0.125 & 0.014 & 0.014 & 0.068 & 0.063 & 0.022\\
-\bottomrule
-\end{tabular}}
-\caption{Summary of exergy flows for the main engines}
-\label{tab:flows:ex:ME}
-\end{table}
-
-\begin{table}[hbp!]
-	\centering
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{clclcrrrrrrrr}
-			\toprule
-			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
-			\midrule
-E & Env. & AE & Comp. & Ph & 0.01 & 0.002 & 0.003 & 0.001 & 0.000 & 0.003 & 0.001 & 0.001\\
-AE & Comp. & AE & CAC-HT & Ph & 7.91 & 2.960 & 4.008 & 0.659 & 0.288 & 3.392 & 3.091 & 1.432\\
-AE & CAC-HT & AE & CAC-LT & Ph & 7.65 & 2.835 & 3.906 & 0.644 & 0.269 & 3.302 & 3.017 & 1.335\\
-AE & CAC-LT & AE & Cyl. & Ph & 7.09 & 2.589 & 3.650 & 0.606 & 0.248 & 3.066 & 2.802 & 1.227\\
-AE & Turb. & AE & Comp. & W & 9.54 & 3.504 & 4.853 & 0.841 & 0.338 & 4.087 & 3.762 & 1.687\\
-AE & Cyl. & AE & Turb. & Ph & 36.26 & 11.776 & 19.815 & 3.592 & 1.079 & 15.705 & 14.961 & 5.597\\
-AE & Turb. & E & HRSG & Ph & 21.04 & 6.378 & 11.936 & 2.157 & 0.571 & 9.166 & 8.866 & 3.011\\
-AE & Turb. & E & Losses & Q & 0.46 & 0.132 & 0.259 & 0.060 & 0.012 & 0.209 & 0.194 & 0.060\\
-AE & Comp. & AE & BP valve & Ph & 7.91 & 2.960 & 4.008 & 0.659 & 0.288 & 3.392 & 3.091 & 1.432\\
-AE & BP valve & AE & Turb. & Ph & 0.00 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
-AE & CAC-HT & CS & HT Cool. & Q & 0.24 & 0.125 & 0.102 & 0.015 & 0.018 & 0.090 & 0.074 & 0.097\\
-AE & CAC-LT & CS & LT Cool. & Q & 0.36 & 0.245 & 0.256 & 0.037 & 0.021 & 0.236 & 0.215 & 0.108\\
-AE & Cyl. & AE & LOC & Q & 3.14 & 0.962 & 1.757 & 0.344 & 0.075 & 1.476 & 1.275 & 0.388\\
-AE & Cyl. & AE & JWC & Q & 5.84 & 1.772 & 3.270 & 0.656 & 0.144 & 2.608 & 2.451 & 0.783\\
-AE & JWC & CS & HT Cool. & Q & 3.90 & 1.772 & 3.270 & 0.656 & 0.144 & 2.608 & 2.451 & 0.783\\
-AE & LOC & CS & LT Cool. & Q & 2.18 & 0.962 & 1.757 & 0.344 & 0.075 & 1.476 & 1.275 & 0.388\\
-AE & Cyl. & E & Losses & Q & 0.37 & 0.096 & 0.217 & 0.047 & 0.008 & 0.166 & 0.159 & 0.044\\
-AE & HRSG & S & Boiler & Q & 6.32 & 2.869 & 5.504 & 1.011 & 0.237 & 4.309 & 3.901 & 1.412\\
-AE & HRSG & E & Env. & Ph & 11.42 & 3.509 & 6.432 & 1.145 & 0.334 & 4.857 & 4.965 & 1.599\\
-AE & Cyl. & EL & SB & W & 62.02 & 20.781 & 34.328 & 6.249 & 1.851 & 26.971 & 26.344 & 9.894\\
-AE & Cyl. & E & AG losses & Q & 0.85 & 0.228 & 0.499 & 0.105 & 0.020 & 0.384 & 0.363 & 0.104\\
-AE & Fuel tank & AE & Cyl. & Ch & 172.09 & 54.452 & 94.876 & 17.916 & 4.849 & 73.199 & 72.744 & 26.150\\
-AE & Fuel tank & AE & Cyl. & Q & 0.08 & 0.019 & 0.054 & 0.009 & 0.003 & 0.041 & 0.033 & 0.010\\
-\bottomrule
-\end{tabular}}
-\caption{Summary of exergy flows for the auxiliary engines}
-\label{tab:flows:ex:AE}
-\end{table}
-
-\begin{table}[hbp!]
-	\centering
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{clclcrrrrrrrr}
-			\toprule
-			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
-			\midrule
-CS & HT Cool. & CS & LT Cool. & Q & 7.36 & 1.46 & 8.63 & 1.36 & 0.97 & 5.10 & 5.25 & 2.08\\
-CS & HT Cool. & S & HRHT & Q & 3.24 & 0.46 & 2.33 & 0.31 & 0.14 & 1.53 & 1.34 & 0.38\\
-S & Boiler & S & HRHT & Q & 10.87 & 4.94 & 5.09 & 0.75 & 0.09 & 6.51 & 3.82 & 0.53\\
-S & Aux Boiler & S & Boiler & Ph & 5.71 & 4.66 & 0.90 & 0.17 & -0.03 & 3.77 & 1.85 & 0.10\\
-S & Aux Boiler & E & Env. & Ph & 1.51 & 1.04 & 0.38 & 0.08 & 0.00 & 0.99 & 0.50 & 0.03\\
-S & Fuel tank & S & Aux Boiler & Ch & 20.12 & 14.37 & 4.73 & 1.02 & 0.01 & 12.99 & 6.76 & 0.37\\
-S & HRHT & D & HVAC-PH & Q & 7.43 & 1.22 & 1.89 & 0.25 & 0.05 & 2.29 & 1.11 & 0.00\\
-S & HRHT & D & HVAC-RH & Q & 0.33 & 0.09 & 0.13 & 0.03 & 0.03 & 0.00 & 0.00 & 0.28\\
-S & HRHT & D & HWH  & Q & 2.27 & 0.64 & 1.09 & 0.17 & 0.05 & 0.84 & 0.81 & 0.31\\
-S & Boiler & D & MSH & Q & 0.68 & 0.06 & 0.10 & 0.01 & 0.00 & 0.09 & 0.07 & 0.02\\
-S & Boiler & D & HFO TH & Q & 0.66 & 0.13 & 0.20 & 0.03 & 0.01 & 0.16 & 0.16 & 0.04\\
-S & Boiler & D & HFO H & Q & 0.44 & 0.04 & 0.21 & 0.03 & 0.02 & 0.13 & 0.12 & 0.04\\
-S & Boiler & D & TH & Q & 0.51 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
-S & Boiler & D & G & Q & 2.45 & 0.44 & 0.98 & 0.14 & 0.04 & 0.71 & 0.69 & 0.19\\
-S & Boiler & D & OT & Q & 0.34 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
-EL & SB & EL & Thrusters & W & 0.94 & 0.00 & 0.00 & 0.94 & 0.00 & 0.36 & 0.43 & 0.15\\
-EL & SB & EL & HVAC & W & 1.79 & 0.59 & 0.85 & 0.18 & 0.18 & 0.00 & 0.00 & 1.79\\
-EL & SB & EL & Other el. Dem. & W & 59.29 & 19.99 & 32.66 & 4.99 & 1.64 & 26.06 & 25.45 & 7.78\\
-P & Prop. shaft & P & Prop. & W & 104.19 & 0.14 & 82.62 & 9.84 & 11.59 & 40.98 & 44.55 & 18.66\\
-P & Prop. shaft & E & Losses & Q & 3.20 & 0.00 & 2.54 & 0.30 & 0.36 & 1.26 & 1.37 & 0.57\\
-S & Boiler & E & Dumped steam & Q & 0.81 & 0.00 & 0.43 & 0.13 & 0.25 & 0.05 & 0.20 & 0.55\\
-\bottomrule
-\end{tabular}}
-\caption{Summary of exergy flows for the boiler and steam systems}
-\label{tab:flows:ex:S}
-\end{table}
-
-\begin{table}[hbp!]
-	\centering
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{clclcrrrrrrrr}
-			\toprule
-			\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & P/S & LSS & Man & HSS & Win & Mid & Sum\\
-			\midrule
-D& HVAC-PH & D & Dem. & Q & 3.41 & 1.22 & 1.89 & 0.25 & 0.05 & 2.29 & 1.11 & 0.00\\
-D& HVAC-RH & D & Dem. & Q & 0.28 & 0.09 & 0.13 & 0.03 & 0.03 & 0.00 & 0.00 & 0.28\\
-D& HWH  & D & Dem. & Q & 1.96 & 0.64 & 1.09 & 0.17 & 0.05 & 0.84 & 0.81 & 0.31\\
-D& MSH & D & Dem. & Q & 0.18 & 0.06 & 0.10 & 0.01 & 0.00 & 0.09 & 0.07 & 0.02\\
-D& HFO TH & D & Dem. & Q & 0.37 & 0.13 & 0.20 & 0.03 & 0.01 & 0.16 & 0.16 & 0.04\\
-D& HFO H & D & Dem. & Q & 0.29 & 0.04 & 0.21 & 0.03 & 0.02 & 0.13 & 0.12 & 0.04\\
-D& TH & D & Dem. & Q & 0.23 & 0.08 & 0.12 & 0.02 & 0.01 & 0.11 & 0.10 & 0.03\\
-D& G & D & Dem. & Q & 1.59 & 0.44 & 0.98 & 0.14 & 0.04 & 0.71 & 0.69 & 0.19\\
-D& OT & D & Dem. & Q & 0.15 & 0.05 & 0.08 & 0.01 & 0.00 & 0.07 & 0.07 & 0.02\\
-\bottomrule
-\end{tabular}}
-\caption{Summary of exergy flows for the heating demand}
-\label{tab:flows:ex:D}
-\end{table}
-
-\clearpage
-
-\section{Full list of available measurements}   \label{sec:appendix:measurements}
+\section{Full list of available measurements}   \label{sec:app:measurements}
 
 In this section, the full list of all measurements available in the dataset that was used in this paper is provided. It should be noted that not all the values were used in the analysis.
-
-\clearpage
-
-\section{Energy demand for typical operational days}
-
-In this section, the values for the main energy demands (Propulsion power, electric power, high temperature heat and low temperature heat) are provided for the typical days identified in the demand clustering.
-
-\bibliographystyle{unsrt}
-\bibliography{bibliography}
 
